### 2018-05-10 PaperLatex/frontiers.tex
Add paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.   
%                                              %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.3 Generated 2018/01/26 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
%%%  NB logo1.jpg is required in the path in order to correctly compile front page header %%%

\documentclass[utf8]{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
%\documentclass[utf8]{frontiersHLTH} % for Health articles
%\documentclass[utf8]{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles

%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}
\usepackage{natbib}

\linenumbers

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Hoyland, Gorur-Shandilya, \& Marder} %use et al only if is more than 1 author
\def\Authors{Alec Hoyland, Srinivas Gorur-Shandilya\,$^{*}$, and Eve Marder}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Srinivas Gorur-Shandilya
%
%Volen Center for Complex Systems
%
%Brandeis University
%
%415 South Street
%
%Waltham, MA 02454
%
%USA
}

\def\corrEmail{srinivasgs@brandeis.edu}

\begin{document}
\onecolumn
\firstpage{1}

\title[xolotl: the god of thunder]{Xolotl: A Fast and Comprehensible Neuronal Simulator}

\author[\firstAuthorLast]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}

%%% Leave the Abstract empty if your article does not require one, please see the Summary Table for full details.
\section{}
For full guidelines regarding your manuscript please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines}{Author Guidelines}.

As a primary goal, the abstract should render the general significance and conceptual advance of the work clearly accessible to a broad readership. References should not be cited in the abstract. Leave the Abstract empty if your article does not require one, please see \href{http://www.frontiersin.org/about/AuthorGuidelines#SummaryTable}{Summary Table} for details according to article type.

\tiny
 \keyFont{ \section{Keywords:} simulator, MATLAB, C++, xolotl, conductance-based, computational, keyword, keyword} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\end{abstract}

\section{Introduction}

For Original Research Articles \citep{conference}, Clinical Trial Articles \citep{article}, and Technology Reports \citep{patent}, the introduction should be succinct, with no subheadings \citep{book}. For Case Reports the Introduction should include symptoms at presentation \citep{chapter}, physical exams and lab results \citep{dataset}.

\section{Article types}

For requirements for a specific article type please refer to the Article Types on any Frontiers journal page. Please also refer to  \href{http://home.frontiersin.org/about/author-guidelines#Sections}{Author Guidelines} for further information on how to organize your manuscript in the required sections or their equivalents for your field

% For Original Research articles, please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.

\section{Manuscript Formatting}

\subsection{Heading Levels}

%There are 5 heading levels

\subsection{Level 2}
\subsubsection{Level 3}
\paragraph{Level 4}
\subparagraph{Level 5}

\subsection{Equations}
Equations should be inserted in editable format from the equation editor.

\begin{equation}
\sum x+ y =Z\label{eq:01}
\end{equation}

\subsection{Figures}
Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript. Figures will then be automatically embedded at the bottom of the submitted manuscript. Kindly ensure that each table and figure is mentioned in the text and in numerical order. Figures must be of sufficient resolution for publication \href{http://home.frontiersin.org/about/author-guidelines#ResolutionRequirements}{see here for examples and minimum requirements}. Figures which are not according to the guidelines will cause substantial delay during the production process. Please see \href{http://home.frontiersin.org/about/author-guidelines#GeneralStyleGuidelinesforFigures}{here} for full figure guidelines. Cite figures with subfigures as figure \ref{fig:2}B.

\subsubsection{Permission to reuse and Copyright}
Permission must be obtained for use of copyrighted material from other sources (including the web). Please note that it is compulsory to follow figure instructions.

\subsection{Tables}
Tables should be inserted at the end of the manuscript. Please build your table directly in LaTeX.Tables provided as jpeg/tiff files will not be accepted. Please note that very large tables (covering several pages) cannot be included in the final PDF for reasons of space. These tables will be published as \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} on the online article page at the time of acceptance. The author will be notified during the typesetting of the final article if this is the case.

\section{Nomenclature}

\subsection{Resource Identification Initiative}
To take part in the Resource Identification Initiative, please use the corresponding catalog number and RRID in your current manuscript. For more information about the project and for steps on how to search for an RRID, please click \href{http://www.frontiersin.org/files/pdf/letter_to_author.pdf}{here}.

\subsection{Life Science Identifiers}
Life Science Identifiers (LSIDs) for ZOOBANK registered names or nomenclatural acts should be listed in the manuscript before the keywords. For more information on LSIDs please see \href{http://www.frontiersin.org/about/AuthorGuidelines#InclusionofZoologicalNomenclature}{Inclusion of Zoological Nomenclature} section of the guidelines.

\section{Additional Requirements}

For additional requirements for specific article types and further information please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines#AdditionalRequirements}{Author Guidelines}.

\section*{Conflict of Interest Statement}
%All financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed. If no such relationship exists, authors will be asked to confirm the following statement:

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}

The Author Contributions section is mandatory for all articles, including articles by sole authors. If an appropriate statement is not provided on submission, a standard one will be inserted during the production process. The Author Contributions statement must describe the contributions of individual authors referred to by their initials and, in doing so, all authors agree to be accountable for the content of the work. Please see  \href{http://home.frontiersin.org/about/author-guidelines#AuthorandContributors}{here} for full authorship criteria.

\section*{Funding}
Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.

\section*{Acknowledgments}
This is a short text to acknowledge the contributions of specific colleagues, institutions, or agencies that aided the efforts of the authors.

\section*{Supplemental Data}
 \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure. LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder.

\section*{Data Availability Statement}
The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK].
% Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData

\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
%\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
\bibliography{test}

%%% Make sure to upload the bib file along with the tex file and PDF
%%% Please see the test.bib file for some examples of references

\section*{Figure captions}

%%% Please be aware that for original research articles we only permit a combined number of 15 figures and tables, one figure with multiple subfigures will count as only one figure.
%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.eps and logos.eps) are in the Frontiers LaTeX folder
%%% If using *.tif files convert them to .jpg or .png
%%%  NB logo1.eps is required in the path in order to correctly compile front page header %%%

\begin{figure}[h!]
\begin{center}
\includegraphics[width=10cm]{logo1}% This is a *.eps file
\end{center}
\caption{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures}\label{fig:1}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=15cm]{logos}
\end{center}
\caption{This is a figure with sub figures, \textbf{(A)} is one logo, \textbf{(B)} is a different logo.}\label{fig:2}
\end{figure}

%%% If you are submitting a figure with subfigures please combine these into one image file with part labels integrated.
%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.
%%% Frontiers will add the figures at the end of the provisional pdf automatically
%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.

\end{document}

### 2018-05-17 PaperLatex/frontiers.tex
Begin adding figures

@@
 %%% Leave the Abstract empty if your article does not require one, please see the Summary Table for full details.
-\section{}
-For full guidelines regarding your manuscript please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines}{Author Guidelines}.
-
-As a primary goal, the abstract should render the general significance and conceptual advance of the work clearly accessible to a broad readership.
-References should not be cited in the abstract.
-Leave the Abstract empty if your article does not require one, please see \href{http://www.frontiersin.org/about/AuthorGuidelines#SummaryTable}{Summary Table} for details according to article type.
+\texttt{xolotl} is an open-source neuronal simulator written in \texttt{C++} with \texttt{MATLAB} wrappers.
+Complex models and networks can be designed efficiently using an intuitive language tightly coupled to the object-based architecture of the underlying \texttt{C++} code.
+Models are specified by adding conductances to compartment objects.
+The structure is modular, serialized, and searchable, permitting high-level programmatic control over nearly all features of the models.
+\texttt{C++} templates are provided for developing new conductances, compartments, and integration schemata.
+\texttt{xolotl} readily implements parallel-processing on multicore processors and high-performance computing clusters.
+It also includes a customizable graphical user interface (`puppeteer`) for rapid prototyping and hand-tuning conductances in real-time.
+In addition, \texttt{xolotl} comes packaged with a powerful optimization toolbox (\texttt{procrustes}) for optimizing any model parameters accessible in the \texttt{xolotl} tree.
+The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate interoperability with other specifications (viz.
+\texttt{NeuroML, SBML}), simulators (viz \texttt{NEURON, Brian, NEST}), and web-based applications (viz.
+\texttt{Geppetto}).
+\texttt{xolotl} is freely available at \url{https://github.com/sg-s/xolotl}.
+This tool provides rapid implementation and fast simulation of neuronal models while permitting full control over every aspect of the network and integration.
 
@@
 Cite figures with subfigures as figure \ref{fig:2}B.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_clamp}
+	\caption{\texttt{xolotl} readily implements voltage clamp.
+(A) Cartoon of a cell with potassium conductance with experimentally-fixed voltage.
+(B) Structure of \texttt{xolotl} object in A.
+(C) Code snippet depicting integration under voltage clamp.
+(D-E) Current response to steps in voltage from a holding potential of $V_m = -60$ mV.
+(F) Current-voltage relation of the steady-state current ($t = 400$ ms) indicating a reversal potential of $E = -80$ mV and no inactivation.
+(G) Conductance-voltage relation at steady-state takes the form of a sigmoid.
+(H) Sigmoids of the form $m(V)^n$ fit to the model data indicating that $n=4$ is the best fit.
+(I) $R^2$ correlation of the sigmoid fits at various powers where $n=4$ is an exact fit.}
+	\label{fig:figureclamp}
+\end{figure}
 

### 2018-05-18 PaperLatex/frontiers.tex
Put semi-complete figures in the LaTeX document

@@
 \usepackage{natbib}
+\usepackage{placeins}
 
@@
 	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_HH}
+	\caption{\texttt{xolotl} can quickly set up and simulate conductance-based models.
+(A) Cartoon of a Hodgkin-Huxley single-compartment neuron model with fast sodium, delayed rectifier, and leak currents.
+(B) Code snippet in \texttt{MATLAB} used to implement D, F-I.
+(C) \texttt{xolotl} schematic displayed in the \texttt{MATLAB} command prompt.
+(D) Simulated voltage trace with 0.1 nA applied current.
+Colors indicate the dominant current (gold is fast sodium, blue is delayed rectifier, red is leak).
+(E) Frequency-input relation displaying firing rate as a function of applied current.
+(F-G) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
+Variables not plotted are unity for all voltage.
+(H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables.
+Variables not plotted are unity for all voltage.
+Colors indicate conductance type (gold is fast sodium, blue is delayed rectifier, red is leak).}
+	\label{fig:figurehh}
+\end{figure}
+
+\begin{figure}
+	\centering
 	\includegraphics[width=1.0\linewidth]{gfx/figure_clamp}
@@
 (G) Conductance-voltage relation at steady-state takes the form of a sigmoid.
-(H) Sigmoids of the form $m(V)^n$ fit to the model data indicating that $n=4$ is the best fit.
+(H) Sigmoids of the form $m^n$ fit to the model data indicating that $n=4$ is the best fit.
 (I) $R^2$ correlation of the sigmoid fits at various powers where $n=4$ is an exact fit.}
@@
 \end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
+	\caption{\texttt{xolotl} readily implements conductance-based network models.
+(A) Diagram of a network model of the pyloric rhythm in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004).
+(B) Hierarchical structure of a neuronal network considers compartments as components of the network and conductances and synapses as components of compartments.
+(C) \textit{xolotl} implements conductances as fields of compartments and synapses as connections between compartments.
+(D-F) Simulated voltage trace of a model network for the three compartments.
+(G) Time series of synaptic currents in the simulated network with model cholinergic (chol) and glutamatergic (glut) synapses are outputs of the integration.}
+	\label{fig:figurenetwork}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_integral_control}
+	\caption{\texttt{xolotl} can implement homeostatic tuning rules as integral control.
+(A) Cartoon of a model neuron with integral control.
+(B) Hierarchical structure of a neuronal network considers controllers as components of compartments which act on conductances.
+(C) \texttt{xolotl} implements controllers XYZ.
+(D) Calcium sensors change maximal conductances to move a neuron from quiescence to a bursting state.
+(E) Voltage trace of the controlled neuron shows regular bursting activity.}
+	\label{fig:figureintegralcontrol}
+\end{figure}
+
+\FloatBarrier
 

### 2018-05-18 PaperLatex/frontiers.tex
minor fixes

@@
 
-\title[xolotl: the god of thunder]{Xolotl: A Fast and Comprehensible Neuronal Simulator}
+\title[xolotl: neuronal simulator]{Xolotl: An Intuitive and Comprehensible Neuronal Simulator}
 
@@
 (G) Conductance-voltage relation at steady-state takes the form of a sigmoid.
-(H) Sigmoids of the form $m^n$ fit to the model data indicating that $n=4$ is the best fit.
+(H) Sigmoids $m$ fit to the model as $m^n$ data indicating that $n=4$ is the best fit.
 (I) $R^2$ correlation of the sigmoid fits at various powers where $n=4$ is an exact fit.}

### 2018-05-22 PaperLatex/frontiers.tex
Add descriptions of some figures

@@
 In addition, \texttt{xolotl} comes packaged with a powerful optimization toolbox (\texttt{procrustes}) for optimizing any model parameters accessible in the \texttt{xolotl} tree.
-The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate interoperability with other specifications (viz.
-\texttt{NeuroML, SBML}), simulators (viz \texttt{NEURON, Brian, NEST}), and web-based applications (viz.
-\texttt{Geppetto}).
+The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate rapid construction and assessment of model networks.
 \texttt{xolotl} is freely available at \url{https://github.com/sg-s/xolotl}.
-This tool provides rapid implementation and fast simulation of neuronal models while permitting full control over every aspect of the network and integration.
+This tool provides straightforward implementation and fast simulation of neuronal models while permitting full control over every aspect of the network and integration.
 
@@
 
-For Original Research Articles \citep{conference}, Clinical Trial Articles \citep{article}, and Technology Reports \citep{patent}, the introduction should be succinct, with no subheadings \citep{book}.
-For Case Reports the Introduction should include symptoms at presentation \citep{chapter}, physical exams and lab results \citep{dataset}.
+\texttt{xolotl} (\url{https://github.com/sg-s/xolotl}) is a fast single-compartment and multi-compartment simulator in \texttt{C++} with \texttt{MATLAB} wrappers (\url{https://www.mathworks.com/products/matlab.html}).
+Written with an emphasis on flexibility and speed, \texttt{xolotl} can simulate single-compartment conductance-based models, networks of these, and detailed multi-compartment models.
+\texttt{xolotl} exploits a novel automatic type system, \texttt{cpplab}, which binds \texttt{MATLAB} code to \texttt{C++} header files, creating objects and classes \textit{ad libitum} in \texttt{MATLAB} which reflect the underlying object-oriented code.
+\texttt{xolotl} implements \texttt{cpplab} to represent the nested structure of conductance-based models, and exploits the computational efficiency of the low-level programming language to quickly integrate models.
+For this reason, models can be implemented entirely in \texttt{MATLAB} with few lines of code.
+
+Models are specified in \texttt{MATLAB} by a \texttt{xolotl} object which contains compartment objects which themselves contain conductances.
+Synapses belong to the \texttt{xolotl} object and connect compartments together.
+The high-level specification supports arbitrarily large network and multi-compartment morphologies.
+
+\texttt{xolotl} provides parameter optimization capabilities through the algorithm-agnostic \texttt{procrustes} toolbox.
+Any network parameters accessible through the \texttt{xolotl} structure can be optimized using arbitrary algorithms and objective functions on multi-core computers and high-performance computing clusters.
+
+The software has been implemented in \texttt{MATLAB} due to its ease-of-use and popularity among neuroscientists.
+\texttt{cpplab} provides a powerful backend for specifying and integrating models without relying on the significantly slower and limiting \texttt{MATLAB codegen}.
+Minimal experience with \texttt{MATLAB} is required to use \texttt{xolotl}, and all equations and integration methods are provided transparently to the end user.
+No string parsing of equations is required.
+
+\texttt{xolotl} comes packaged with visualization functions, a graphical user interface for real-time manipulation of model parameters.
+Plotting of voltage, intracellular calcium, conductance gating functions, and time constants is provided by built-in \texttt{xolotl} methods.
+The \texttt{puppeteer} toolbox permits real-time tuning of any network parameters using numerical sliders in a graphical interface which displays the resultant membrane potential and intracellular calcium traces.
+The ease-of-use of these tools lends them to pedagogical applications and rapid exploration of toy models.
+This tool aims to simplify the investigation of dynamics of complex neural network models, facilitate collaborative modeling, and complement other tools being developed in the neuroinformatics community.
 
@@
 % For Original Research articles, please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.
+
+\section{Worked Examples}
+
+Using \texttt{xolotl} in \texttt{MATLAB}, users create a \texttt{xolotl} object and populate it with compartments, synapses, and controllers.
+Each field is a \texttt{cpplab} object constructed by a function call to \texttt{add}.
+The model is integrated with the \texttt{integrate} function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.
+
+\texttt{xolotl} comes packaged with a library of pre-existing conductance and synapse objects which greatly simplify the task of constructing model neurons.
+These objects can be referenced by name and added directly to a compartment.
+Novel conductance dynamics can be easily written by modifying a template header file contained in the \texttt{xolotl} distribution.
+
+\subsection{Simulating a Hodgkin-Huxley Model}
+
+The seminal Hodgkin-Huxley model contains a fast inactivating sodium conductance which promotes spiking, a non-inactivating potassium delayed rectifer, and a passive leak current (\ref{fig:figurehh}A).
+A compartment named \texttt{HH} with compartment properties of membrane capacitance $C_m = 10 ~ \mu\mathrm{F/mm^2}$ and surface area $A = 0.1 ~ \mathrm{mm^2}$ can be specified by \ref{fig:figurehh}B.
+Compartment, conductance, synapse, and controller properties can be specified during the call to the \texttt{add} function, or after construction, using dot-notation in \texttt{MATLAB} (e.g.
+\texttt{x.HH.Cm}).
+\ref{fig:figurehh}C shows the \texttt{MATLAB} command prompt after invoking the \texttt{xolotl} object \texttt{x}, displaying the hierarchical structure inherent in conductance-based treatments of neurodynamics.
+
+This model was constructed using conductances from Liu \textit{et al} 1998.
+In the absence of applied positive current, the model is quiescent and tonically spikes under 0.2 nA of applied current (\ref{fig:figurehh}D).
+The \texttt{integrate} function takes the applied current as an argument (e.g.
+\texttt{x.integrate(AppliedCurrent)}), so that the \texttt{xolotl} object is agnostic to integration-specific perturbations.
+The \texttt{plot} function generates voltage and intracellular calcium traces, where the voltage trace is colored by the dominant current.
+If the membrane potential is increasing, the strongest instantaneous inward current colors the trace.
+Conversely, if the membrane potential is decreasing, the strongest outward current colors the trace instead.
+\ref{fig:figurehh}F-I display the results of the \texttt{show} function.
+Activation and inactivation steady-states and the voltage-dependent time constants of these gating variables describe the conductance dynamics in absence of other channel types.
+
+\subsection{Performing a Voltage Clamp Experiment \textit{in-silico}}
+
+\texttt{xolotl} can recapitulate the results of voltage clamp experiments.
+\ref{fig:figureclamp} displays the procedure to clamp the membrane potential of a cell with a delayed rectifier potassium conductance.
+The second argument of the \texttt{integrate} function determines the clamped voltage (e.g.
+\texttt{x.integrate([], VoltageClamp)}).
+Isolated currents under voltage clamp approach the steady-state (\ref{fig:figureclamp}D-E) so that a current-voltage relation at steady-state can be extracted (\ref{fig:figureclamp}F).
+The derivative of the IV curve is the steady-state conductance (\ref{fig:figureclamp}G).
+Fitting a sigmoid to various powers yields a model for the current dynamics (\ref{fig:figureclamp}H-I).
+These figures describe graphically the theoretical underpinnings of current analysis through voltage clamp and can serve as an effective pedagogical tool for computational and quantitative neuroscience.
+
+\subsection{Simulating Network Models}
+
+Network models in \texttt{xolotl} consist of compartment objects connected by synapses.
+Synapses are stored in a vector array as a field of the \texttt{xolotl} object in \texttt{MATLAB}.
+Presynaptic and postsynaptic labels indicate the connectivity of the synapse.
+Figure \ref{fig:figurenetwork} implements a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans.
+The pyloric model contains three compartments and seven synapses (\ref{fig:figurenetwork}A).
+This structure is reciprocated in the hierarchy of the \texttt{xolotl} object, where conductances are contained within compartments (\ref{fig:figurenetwork}B).
+Synapses are upper-level properties of the network which point between two compartments (\ref{fig:figurenetwork}C).
+This exploits vectorized operations in \texttt{MATLAB} and does not require each synapse to possess a unique name.
+The \texttt{plot} function generates multiple subplots when called for a network with multiple compartments (\ref{fig:figurenetwork}D-F).
 
@@
 Cite figures with subfigures as figure \ref{fig:2}B.
+
+\subsubsection{Permission to reuse and Copyright}
+Permission must be obtained for use of copyrighted material from other sources (including the web). Please note that it is compulsory to follow figure instructions.
+
+\subsection{Tables}
+Tables should be inserted at the end of the manuscript.
+Please build your table directly in LaTeX.Tables provided as jpeg/tiff files will not be accepted.
+Please note that very large tables (covering several pages) cannot be included in the final PDF for reasons of space.
+These tables will be published as \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} on the online article page at the time of acceptance.
+The author will be notified during the typesetting of the final article if this is the case.
+
+\section*{Conflict of Interest Statement}
+%All financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed.
+If no such relationship exists, authors will be asked to confirm the following statement:
+
+The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
+
+\section*{Author Contributions}
+
+AH and SG-S wrote the manuscript and the code. EM provided funding and moral support.
+
+\section*{Funding}
+Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.
+
+AH received funding from National Institute on Drug Abuse (NIDA) through the undergraduate training grant in computational neuroscience (1R90DA033463-01).
+
+\section*{Acknowledgments}
+The authors would like to thank Mara CP Rue and Hillary Rodgers for beta-testing the \texttt{xolotl} software.
+
+\section*{Supplemental Data}
+ \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure.
+LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder.
+
+\section*{Data Availability Statement}
+
+The code to generate all figures can be found in the \texttt{xolotl} repository (\url{https://github.com/sg-s/xolotl}).
+
+\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
+%\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
+\bibliography{test}
+
+%%% Make sure to upload the bib file along with the tex file and PDF
+%%% Please see the test.bib file for some examples of references
+
+\section*{Figure captions}
+
+%%% Please be aware that for original research articles we only permit a combined number of 15 figures and tables, one figure with multiple subfigures will count as only one figure.
+%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
+%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.eps and logos.eps) are in the Frontiers LaTeX folder
+%%% If using *.tif files convert them to .jpg or .png
+%%%  NB logo1.eps is required in the path in order to correctly compile front page header %%%
 
@@
 
-\FloatBarrier
-
-\subsubsection{Permission to reuse and Copyright}
-Permission must be obtained for use of copyrighted material from other sources (including the web). Please note that it is compulsory to follow figure instructions.
-
-\subsection{Tables}
-Tables should be inserted at the end of the manuscript.
-Please build your table directly in LaTeX.Tables provided as jpeg/tiff files will not be accepted.
-Please note that very large tables (covering several pages) cannot be included in the final PDF for reasons of space.
-These tables will be published as \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} on the online article page at the time of acceptance.
-The author will be notified during the typesetting of the final article if this is the case.
-
-\section{Nomenclature}
-
-\subsection{Resource Identification Initiative}
-To take part in the Resource Identification Initiative, please use the corresponding catalog number and RRID in your current manuscript.
-For more information about the project and for steps on how to search for an RRID, please click \href{http://www.frontiersin.org/files/pdf/letter_to_author.pdf}{here}.
-
-\subsection{Life Science Identifiers}
-Life Science Identifiers (LSIDs) for ZOOBANK registered names or nomenclatural acts should be listed in the manuscript before the keywords.
-For more information on LSIDs please see \href{http://www.frontiersin.org/about/AuthorGuidelines#InclusionofZoologicalNomenclature}{Inclusion of Zoological Nomenclature} section of the guidelines.
-
-\section{Additional Requirements}
-
-For additional requirements for specific article types and further information please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines#AdditionalRequirements}{Author Guidelines}.
-
-\section*{Conflict of Interest Statement}
-%All financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed.
-If no such relationship exists, authors will be asked to confirm the following statement:
-
-The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
-
-\section*{Author Contributions}
-
-The Author Contributions section is mandatory for all articles, including articles by sole authors.
-If an appropriate statement is not provided on submission, a standard one will be inserted during the production process.
-The Author Contributions statement must describe the contributions of individual authors referred to by their initials and, in doing so, all authors agree to be accountable for the content of the work.
-Please see  \href{http://home.frontiersin.org/about/author-guidelines#AuthorandContributors}{here} for full authorship criteria.
-
-\section*{Funding}
-Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.
-
-\section*{Acknowledgments}
-This is a short text to acknowledge the contributions of specific colleagues, institutions, or agencies that aided the efforts of the authors.
-
-\section*{Supplemental Data}
- \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure.
-LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder.
-
-\section*{Data Availability Statement}
-The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK].
-% Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData
-
-\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
-%\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
-\bibliography{test}
-
-%%% Make sure to upload the bib file along with the tex file and PDF
-%%% Please see the test.bib file for some examples of references
-
-\section*{Figure captions}
-
-%%% Please be aware that for original research articles we only permit a combined number of 15 figures and tables, one figure with multiple subfigures will count as only one figure.
-%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
-%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.eps and logos.eps) are in the Frontiers LaTeX folder
-%%% If using *.tif files convert them to .jpg or .png
-%%%  NB logo1.eps is required in the path in order to correctly compile front page header %%%
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=10cm]{logo1}% This is a *.eps file
-\end{center}
-\caption{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures}\label{fig:1}
-\end{figure}
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=15cm]{logos}
-\end{center}
-\caption{This is a figure with sub figures, \textbf{(A)} is one logo, \textbf{(B)} is a different logo.}\label{fig:2}
-\end{figure}
-
-%%% If you are submitting a figure with subfigures please combine these into one image file with part labels integrated.
-%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.
-%%% Frontiers will add the figures at the end of the provisional pdf automatically
-%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.
-
 \end{document}

### 2018-05-22 PaperLatex/frontiers.tex
Add new section to worked examples

@@
 \def\firstAuthorLast{Hoyland, Gorur-Shandilya, \& Marder} %use et al only if is more than 1 author
-\def\Authors{Alec Hoyland, Srinivas Gorur-Shandilya\,$^{*}$, and Eve Marder}
-% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
-% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
+\def\Authors{Alec Hoyland\,$^{\dagger}$, Srinivas Gorur-Shandilya\,$^{\dagger}$, and Eve Marder\,$^{*}$}
+
 \def\Address{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
-% The Corresponding Author should be marked with an asterisk
-% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
-\def\corrAuthor{Srinivas Gorur-Shandilya
-%
-%Volen Center for Complex Systems
-%
-%Brandeis University
-%
-%415 South Street
-%
-%Waltham, MA 02454
-%
-%USA
+
+\def\corrAuthor{Eve Marder
+
+Volen Center for Complex Systems
+
+Brandeis University
+
+415 South Street
+
+Waltham, MA 02454
+
+USA
 }
 
-\def\corrEmail{srinivasgs@brandeis.edu}
+\def\corrEmail{marder@brandeis.edu}
 
@@
 
-\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
-%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
+%\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
+\extraAuth{$^{\dagger}$ Co-first authors}
 
@@
 Complex models and networks can be designed efficiently using an intuitive language tightly coupled to the object-based architecture of the underlying \texttt{C++} code.
-Models are specified by adding conductances to compartment objects.
+Models can be specified by adding conductances to compartment objects.
 The structure is modular, serialized, and searchable, permitting high-level programmatic control over nearly all features of the models.
 \texttt{C++} templates are provided for developing new conductances, compartments, and integration schemata.
-\texttt{xolotl} readily implements parallel-processing on multicore processors and high-performance computing clusters.
 It also includes a customizable graphical user interface (`puppeteer`) for rapid prototyping and hand-tuning conductances in real-time.
-In addition, \texttt{xolotl} comes packaged with a powerful optimization toolbox (\texttt{procrustes}) for optimizing any model parameters accessible in the \texttt{xolotl} tree.
 The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate rapid construction and assessment of model networks.
-\texttt{xolotl} is freely available at \url{https://github.com/sg-s/xolotl}.
+\texttt{xolotl} is freely available at \url{https://github.com/marderlab/xolotl}.
 This tool provides straightforward implementation and fast simulation of neuronal models while permitting full control over every aspect of the network and integration.
@@
 The \texttt{plot} function generates multiple subplots when called for a network with multiple compartments (\ref{fig:figurenetwork}D-F).
+
+\subsection{Simulating Integral Control}
+
+\texttt{xolotl} can implement homeostatic tuning rules as integral control.
+Figure \ref{fig:figureintegralcontrol} depicts generation of a bursting neuron model from quiescent conditions.
+Calcium sensors supervene on maximal conductance density (\ref{fig:figureintegralcontrol}).
+In \texttt{xolotl}, integral controllers are properties of conductances (\ref{fig:figureintegralcontrol}B-C).
+They modify properties of the conductance in response to an error signal.
+In a demonstration adapted from O'Leary \textit{et al.} 2013, integral control changes maximal conductances to bring a neuron from quiescence into a bursting regime.
+Maximal conductances increase from random initial conditions to a set which elicits the desired network output by minimizing the error signal (\ref{fig:figureintegralcontrol}D-F).
 

### 2018-05-22 PaperLatex/frontiers.tex
Begin technical documentation

@@
 
-\section{Manuscript Formatting}
+\section{Technical Details}
 
-\subsection{Heading Levels}
+\subsection{Modeling}
 
-%There are 5 heading levels
+Models are specified by adding compartments and synapses to the \texttt{xolotl} object.
+Conductances are added to compartments and controllers can be added to conductances.
+This modular structure recapitulates the biophysics of the Hodgkin-Huxley formalism and obviates the need to explicitly write out equations, which in \texttt{xolotl} are contained within the conductance header files.
 
-\subsection{Level 2}
-\subsubsection{Level 3}
-\paragraph{Level 4}
-\subparagraph{Level 5}
+\texttt{xolotl} relies on \texttt{cpplab} constructions, which allow the user to exploit the efficiency of low-level \texttt{C++} code.
+\texttt{MATLAB} treats \texttt{cpplab} objects as standard variables allowing for symbolic manipulation using only the high-level programming language and graphical interfaces.
+\texttt{xolotl} is fast specifically because all time-intensive code is written in native \texttt{C++}.
+While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{coder} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
 
-\subsection{Equations}
-Equations should be inserted in editable format from the equation editor.
+Native \texttt{C++} provides drastic speed improvements beyond the benefits of translating \texttt{MATLAB} features into low-level code.
+For this reason, \texttt{cpplab} has been designed to provide an interface for constructing, transpiling, and compiling \texttt{C++} code to be called from within \texttt{MATLAB}.
+\texttt{xolotl} simulations can be run entirely from \texttt{C++} executables.
+To facilitate use in \texttt{MATLAB}, \texttt{xolotl} uses the MD5 algorithm to automatically hash the network and compile a new binary and \texttt{MEX} bridge file only if needed.
+\texttt{MATLAB} provides a high level programmatic and graphical interface for implementing, manipulating, and visualizing models without sacrificing the enhancements of native \texttt{C++} code.
 
-\begin{equation}
-\sum x+ y =Z\label{eq:01}
-\end{equation}
+\subsubsection{Compartments and Synapses}
 
-\subsection{Figures}
-Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript.
-Figures will then be automatically embedded at the bottom of the submitted manuscript.
-Kindly ensure that each table and figure is mentioned in the text and in numerical order.
-Figures must be of sufficient resolution for publication \href{http://home.frontiersin.org/about/author-guidelines#ResolutionRequirements}{see here for examples and minimum requirements}.
-Figures which are not according to the guidelines will cause substantial delay during the production process.
-Please see \href{http://home.frontiersin.org/about/author-guidelines#GeneralStyleGuidelinesforFigures}{here} for full figure guidelines.
-Cite figures with subfigures as figure \ref{fig:2}B.
+Once the \texttt{xolotl} object is instantiated in \texttt{MATLAB}, the \texttt{add} function constructs new \texttt{cpplab} objects.
+A model neuron consists of one or more compartments, each representing a section of membrane with capacitance and surface area.
+Isopotential models require one compartment.
+Models with multiple neurons, units, or non-trivial morphology require multiple compartments.
+All specifiable properties of compartments are shown in Supplementary Table 1.
 
-\subsubsection{Permission to reuse and Copyright}
-Permission must be obtained for use of copyrighted material from other sources (including the web). Please note that it is compulsory to follow figure instructions.
-
-\subsection{Tables}
-Tables should be inserted at the end of the manuscript.
-Please build your table directly in LaTeX.Tables provided as jpeg/tiff files will not be accepted.
-Please note that very large tables (covering several pages) cannot be included in the final PDF for reasons of space.
-These tables will be published as \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} on the online article page at the time of acceptance.
-The author will be notified during the typesetting of the final article if this is the case.
+\texttt{xolotl} provides some features for generating complex models.
+Synapses can be added with the \texttt{connect} function.
+At minimum possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses.
+All specifiable properties of synapses are shown in Supplementary Table 2.
+To create axons or transport chains, the \texttt{slice} function splits a compartment into $n$ discrete segments and adds these compartments to the network connected by electrical synapses.
 
@@
 \section*{Supplemental Data}
- \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure.
-LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder.
+ Table including all conductances packaged with \texttt{xolotl} should be put in the supplementary material.
 
@@
 
-The code to generate all figures can be found in the \texttt{xolotl} repository (\url{https://github.com/sg-s/xolotl}).
+The code to generate all figures can be found in the \texttt{xolotl} repository (\url{https://github.com/marderlab/xolotl}).
 
@@
 %%%  NB logo1.eps is required in the path in order to correctly compile front page header %%%
+
+\FloatBarrier
 

### 2018-05-23 PaperLatex/frontiers.tex
Begin simulation section

@@
 \usepackage{placeins}
-
+\usepackage{matlab-prettifier}
 \linenumbers
@@
 \texttt{C++} templates are provided for developing new conductances, compartments, and integration schemata.
-It also includes a customizable graphical user interface (`puppeteer`) for rapid prototyping and hand-tuning conductances in real-time.
+It also includes a customizable graphical user interface (GUI, `puppeteer`) for rapid prototyping and hand-tuning conductances in real-time.
 The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate rapid construction and assessment of model networks.
@@
 
-\texttt{xolotl} comes packaged with visualization functions, a graphical user interface for real-time manipulation of model parameters.
+\texttt{xolotl} comes packaged with visualization functions, a graphical user interface (GUI) for real-time manipulation of model parameters.
 Plotting of voltage, intracellular calcium, conductance gating functions, and time constants is provided by built-in \texttt{xolotl} methods.
@@
 
+\subsubsection{Using the \texttt{cpplab} Framework}
+
+The \texttt{add} function will construct a \texttt{cpplab} object and affix it to as a field in the \texttt{xolotl} structure.
+All compartments, conductances, synapses, and controllers are \texttt{cpplab}.
+Compartments add to the \texttt{xolotl} object and conductances add to compartments.
+Specific properties can be specified using key-value pair arguments (e.g.
+\ref{fig:figurehh}A).
+
+\texttt{cpplab} comes with several features which simplify the handling of high-dimensional models.
+The \texttt{find} function acquires a cell array of all properties of the network which satisfy a search condition.
+
 \subsubsection{Compartments and Synapses}
 
-Once the \texttt{xolotl} object is instantiated in \texttt{MATLAB}, the \texttt{add} function constructs new \texttt{cpplab} objects.
 A model neuron consists of one or more compartments, each representing a section of membrane with capacitance and surface area.
-Isopotential models require one compartment.
-Models with multiple neurons, units, or non-trivial morphology require multiple compartments.
+Isopotential models require one compartment, whereas models with multiple neurons, units, or non-trivial morphology require multiple compartments.
 All specifiable properties of compartments are shown in Supplementary Table 1.
@@
 Synapses can be added with the \texttt{connect} function.
-At minimum possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses.
+At minimum synapses possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses.
 All specifiable properties of synapses are shown in Supplementary Table 2.
 To create axons or transport chains, the \texttt{slice} function splits a compartment into $n$ discrete segments and adds these compartments to the network connected by electrical synapses.
+
+\subsubsection{Conductances and Controllers}
+
+All conductances contain fields for maximal conductance and reversal potential.
+Conductances with activation and inactivation variables include them as $m$ and $h$ respectively.
+Gating functions and their respective time constants are contained within the conductance header file.
+\texttt{xolotl} comes packaged with conductances from several dozen papers (Supplementary Table 3).
+
+\subsubsection{Creating Custom \texttt{cpplab} Objects}
+
+\texttt{xolotl} contains template header files for producing custom conductances. The template contains instructions on how to design novel conductances with arbitrary specifications.
+
+\subsection{Simulation}
+
+Models are simulated in \texttt{xolotl} with the \texttt{integrate} function which outputs as time series the membrane potentials, intracellular calcium concentrations, controller states, intrinsic currents, and synaptic currents.
+The \texttt{integrate} function also accepts an argument which specifies injected current or clamped voltage.
+
+\texttt{xolotl} uses the exponential Euler method (Dayan \& Abbott 2001) for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments.
+The simulation time-resolution can be specified to target arbitrary precision, and an output time step can be selected to support automatic down-sampling for memory considerations.
+
+Simulations can be run in "open-loop" mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or "closed-loop" mode which begins simulation with the current network state.
+
+\subsection{Using the \texttt{puppeteer} graphical interface}
+
+\texttt{xolotl} comes packaged with a graphical user interface for visualizing parameter changes in real-time. The \texttt{manipulate} function begins the
 

### 2018-05-23 PaperLatex/frontiers.tex
Add to Simulation section

@@
 \usepackage{placeins}
+
+\usepackage{xcolor}
+\usepackage{listings}
 \usepackage{matlab-prettifier}
+\lstMakeShortInline[style=Matlab-editor]"
+
 \linenumbers
@@
 The \texttt{integrate} function takes the applied current as an argument (e.g.
-\texttt{x.integrate(AppliedCurrent)}), so that the \texttt{xolotl} object is agnostic to integration-specific perturbations.
+"x.integrate(AppliedCurrent)"), so that the \texttt{xolotl} object is agnostic to integration-specific perturbations.
 The \texttt{plot} function generates voltage and intracellular calcium traces, where the voltage trace is colored by the dominant current.
@@
 The \texttt{find} function acquires a cell array of all properties of the network which satisfy a search condition.
+For example, one can find all paths to maximal conductances within the "'HH'" compartment in two ways:
+
+\begin{lstlisting}[style=Matlab-editor]
+	x.find('HH*gbar');
+	x.HH.find('gbar');
+\end{lstlisting}
+
+To extract a vector of the maximal conductances:
+
+\begin{lstlisting}[style=Matlab-editor]
+	gbars = x.get(x.find('HH*gbar'));
+\end{lstlisting}
+
+To set the maximal conductances all at once:
+
+\begin{lstlisting}[style=Matlab-editor]
+	x.set(x.find('HH*gbar'), gbars)
+\end{lstlisting}
+
+"MATLAB" can easily control the "cpplab" objects using the standard, flexible data structure notation popular in high-level scripting languages.
 
@@
 
-Simulations can be run in "open-loop" mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or "closed-loop" mode which begins simulation with the current network state.
+Simulations can be run in 'open-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or 'closed-loop' mode which begins simulation with the current network state.
 
@@
 
-\texttt{xolotl} comes packaged with a graphical user interface for visualizing parameter changes in real-time. The \texttt{manipulate} function begins the
+\texttt{xolotl} comes packaged with a graphical user interface for visualizing parameter changes in real-time.
+The \texttt{manipulate} function opens the GUI, which displays a figure plotting the membrane potential and intracellular calcium concentration of all compartments as time series, and a dialog box with customizable sliders for all parameters of the model.
+Moving the sliders integrates the model in `open-loop` mode with the new parameters.
+The parameters available in the sliders can be customized by passing a cell array to manipulate.
+For example, to only see sliders for maximal conductances of the "HH" compartment, call "x.manipulate(x.find('HH*gbar'))".
+Closing the GUI saves the network state of the model to the "xolotl" object.
+This is particularly helpful for rapid prototyping of models.
 
@@
 \section*{Supplemental Data}
- Table including all conductances packaged with \texttt{xolotl} should be put in the supplementary material.
+ Tables including all conductances packaged with \texttt{xolotl} should be put in the supplementary material.
 

### 2018-05-23 PaperLatex/paper.tex
Overhaul of template and add bibliography

--- 2018-05-23 PaperLatex/frontiers.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
-% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.   
-%                                              %
-%                                                                                                                                                         %
-% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass{frontiersSCNS} % for Science articles
 
-%%% Version 3.3 Generated 2018/01/26 %%%
-%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
-%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
-%%%  NB logo1.jpg is required in the path in order to correctly compile front page header %%%
+\usepackage{graphicx}
+%
+\usepackage{amsmath} 
+\usepackage{mathptmx}      % use Times fonts if available on your TeX system
+%
+\usepackage{color, soul}
+\usepackage{url}
+\usepackage{multirow}
+\usepackage{array}
+\usepackage{fixltx2e}
+\usepackage{textcomp}
+\usepackage{booktabs}
 
-\documentclass[utf8]{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
-%\documentclass[utf8]{frontiersHLTH} % for Health articles
-%\documentclass[utf8]{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{float}
+\floatstyle{plaintop}
 
-%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
-\usepackage{url,hyperref,lineno,microtype,subcaption}
-\usepackage[onehalfspacing]{setspace}
-\usepackage{natbib}
+\usepackage{paralist} % inparaenum
+
+\usepackage[bookmarks]{hyperref}
+
+% bibliography
+\usepackage[english]{babel}
+\usepackage[backend=biber, bibencoding=utf8, style=authoryear, citestyle=authoryear-comp, autocite=inline]{biblatex}
+\addbibresource{Bibliography.bib}
+
+\usepackage[draft, nomargin, marginclue, footnote]{fixme}
+\fxsetup{targetlayout=color}
+
+% new commands and shortcuts
+\usepackage{xspace}
+\newcommand{\eg}{\textit{e.g.}\xspace}
+\newcommand{\etal}{\textit{et al.}\xspace}
+\newcommand{\ie}{\textit{i.e.}\xspace}
+\newcommand{\etc}{\textit{etc.}\xspace}
+\newcommand{\vs}{\textit{vs.}\xspace}
+
+% shortcut for float placing
 \usepackage{placeins}
 
-\usepackage{xcolor}
+% make pretty MATLAB code
 \usepackage{listings}
@@
 
-\linenumbers
+\usepackage{xifthen}% provides \isempty test
 
-% Leave a blank line between paragraphs instead of using \\
+\copyrightyear{}
+\pubyear{}
 
+\def\journal{Neuroinformatics}
+\def\DOI{}
+\def\articleType{Technology Report}
 \def\keyFont{\fontsize{8}{11}\helveticabold }
-\def\firstAuthorLast{Hoyland, Gorur-Shandilya, \& Marder} %use et al only if is more than 1 author
-\def\Authors{Alec Hoyland\,$^{\dagger}$, Srinivas Gorur-Shandilya\,$^{\dagger}$, and Eve Marder\,$^{*}$}
+\def\firstAuthorLast{Hoyland {et~al.}} 
+\def\Authors{Alec Hoyland\,$^{1\dagger}$, Srinivas Gorur-Shandilya\,$^{1\dagger}$, and Eve Marder$^{1*}$}
+\def\Address{$^{1}$Volen Center for Complex Systems and Biology Department \\ Brandeis University \\ Waltham, MA \\
+\vspace{0.25cm}
 
-\def\Address{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
-
-\def\corrAuthor{Eve Marder
-
-Volen Center for Complex Systems
-
-Brandeis University
-
-415 South Street
-
-Waltham, MA 02454
-
-USA
+\vspace{0.5em}\tiny{$\dagger$ These authors have equally contributed to this article.}
 }
-
+\def\corrAuthor{Eve Marder \vspace{1mm}}
+\def\corrAddress{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
 \def\corrEmail{marder@brandeis.edu}
@@
 
-\title[xolotl: neuronal simulator]{Xolotl: An Intuitive and Comprehensible Neuronal Simulator}
-
-\author[\firstAuthorLast]{\Authors} %This field will be automatically populated
-\address{} %This field will be automatically populated
-\correspondance{} %This field will be automatically populated
-
-%\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
-\extraAuth{$^{\dagger}$ Co-first authors}
+\title[xolotl: neuronal simulator]{Xolotl: An Intuitive and Definitively Sexy Neuronal Simulator}
+\author[\firstAuthorLast ]{\Authors}
+\address{}
+\correspondance{}
+\extraAuth{}
+\topic{An Intuitive Neuronal Simulator}
 
@@
 
-%%% Leave the Abstract empty if your article does not require one, please see the Summary Table for full details.
 	\texttt{xolotl} is an open-source neuronal simulator written in \texttt{C++} with \texttt{MATLAB} wrappers.
@@
 \texttt{C++} templates are provided for developing new conductances, compartments, and integration schemata.
-It also includes a customizable graphical user interface (GUI, `puppeteer`) for rapid prototyping and hand-tuning conductances in real-time.
+It also includes a customizable graphical user interface (GUI, "puppeteer") for rapid prototyping and hand-tuning conductances in real-time.
 The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate rapid construction and assessment of model networks.
@@
 	\keyFont{ \section{Keywords:} simulator, MATLAB, C++, xolotl, conductance-based, computational, keyword, keyword} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
+
 \end{abstract}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		INTRODUCTION & OVERVIEW
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Introduction}
+\label{sec:intro}
 
@@
 
-\section{Article types}
-
-For requirements for a specific article type please refer to the Article Types on any Frontiers journal page.
-Please also refer to  \href{http://home.frontiersin.org/about/author-guidelines#Sections}{Author Guidelines} for further information on how to organize your manuscript in the required sections or their equivalents for your field
-
-% For Original Research articles, please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		WORKED EXAMPLES
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \section{Worked Examples}
+\label{sec:examples}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		SIMULATING A HODGKIN-HUXLEY MODEL
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Simulating a Hodgkin-Huxley Model}
 
-The seminal Hodgkin-Huxley model contains a fast inactivating sodium conductance which promotes spiking, a non-inactivating potassium delayed rectifer, and a passive leak current (\ref{fig:figurehh}A).
-A compartment named \texttt{HH} with compartment properties of membrane capacitance $C_m = 10 ~ \mu\mathrm{F/mm^2}$ and surface area $A = 0.1 ~ \mathrm{mm^2}$ can be specified by \ref{fig:figurehh}B.
+The seminal Hodgkin-Huxley model \autocite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952} contains a fast inactivating sodium conductance which promotes spiking, a non-inactivating potassium delayed rectifer, and a passive leak current (\ref{fig:figurehh}A).
+A compartment, \texttt{HH}, with membrane capacitance $C_m = 10 ~ \mu\mathrm{F/mm^2}$ and surface area $A = 0.1 ~ \mathrm{mm^2}$ can be specified by \ref{fig:figurehh}B.
 Compartment, conductance, synapse, and controller properties can be specified during the call to the \texttt{add} function, or after construction, using dot-notation in \texttt{MATLAB} (e.g.
@@
 
-This model was constructed using conductances from Liu \textit{et al} 1998.
+This model was constructed using conductances from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion \autocite{turrigianoSelectiveRegulationCurrent1995}.
 In the absence of applied positive current, the model is quiescent and tonically spikes under 0.2 nA of applied current (\ref{fig:figurehh}D).
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		VOLTAGE CLAMP
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Performing a Voltage Clamp Experiment \textit{in-silico}}
 
-\texttt{xolotl} can recapitulate the results of voltage clamp experiments.
+\texttt{xolotl} can recapitulate the results of voltage clamp experiments \autocite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}.
 \ref{fig:figureclamp} displays the procedure to clamp the membrane potential of a cell with a delayed rectifier potassium conductance.
@@
 These figures describe graphically the theoretical underpinnings of current analysis through voltage clamp and can serve as an effective pedagogical tool for computational and quantitative neuroscience.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		NETWORK MODELS
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		INTEGRAL CONTROL
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Simulating Integral Control}
@@
 They modify properties of the conductance in response to an error signal.
-In a demonstration adapted from O'Leary \textit{et al.} 2013, integral control changes maximal conductances to bring a neuron from quiescence into a bursting regime.
+In a demonstration adapted from \cite{olearyCorrelationsIonChannel2013}, integral control changes maximal conductances to bring a neuron from quiescence into a bursting regime.
 Maximal conductances increase from random initial conditions to a set which elicits the desired network output by minimizing the error signal (\ref{fig:figureintegralcontrol}D-F).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		TECHNICAL DETAILS
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\texttt{xolotl} uses the exponential Euler method (Dayan \& Abbott 2001) for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments.
+\texttt{xolotl} uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments.
 The simulation time-resolution can be specified to target arbitrary precision, and an output time step can be selected to support automatic down-sampling for memory considerations.
@@
 
-\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
-%\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
-\bibliography{test}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		BIBLIOGRAPHY
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%% Make sure to upload the bib file along with the tex file and PDF
-%%% Please see the test.bib file for some examples of references
+%\bibliographystyle{frontiersinSCNS&ENG} % for Science and Engineering articles
+%\bibliography{Bibliography}   % name your BibTeX data base
+\printbibliography
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		FIGURES
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \section*{Figure captions}
-
-%%% Please be aware that for original research articles we only permit a combined number of 15 figures and tables, one figure with multiple subfigures will count as only one figure.
-%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
-%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.eps and logos.eps) are in the Frontiers LaTeX folder
-%%% If using *.tif files convert them to .jpg or .png
-%%%  NB logo1.eps is required in the path in order to correctly compile front page header %%%
 

### 2018-05-25 PaperLatex/paper.tex
Add benchmarked figure, minor fixes

@@
 
+\subsection{Benchmarking}
+
+\texttt{xolotl} performs well in benchmarking tests.
+To assess speed and accuracy, "xolotl", "DynaSim" \autocite{sherfeyDynaSimMATLABToolbox2018}, and "NEURON" \autocite{hinesNEURONSimulationEnvironment1997} were compared in simulations over varied time-resolution and simulation time (\ref{fig:figurebenchmark}).
+A single-compartment Hodgkin-Huxley-like model was generated using conductance dynamics from \cite{liuModelNeuronActivitydependent1998} in "xolotl" and "DynaSim".
+To assess speed and accuracy over time-resolution, these models were simulated over varied time-steps for a total real-time of 5 s.
+The speed factor was defined as the ratio between the real-time (5 s) and the runtime of the simulation (simulation-time).
+Therefore, the speed factor represents how many times faster the simulation is than a real-time observation.
+Accuracy was computed by establishing a canonical voltage trace at $dt = 0.001$ ms and performing $r^2$ correlations between voltage traces with greater time-steps and the canonical trace.
+
+"xolotl" uses the exponential euler method for integrating membrane potential \autocite{dayanTheoreticalNeuroscience2001}.
+"DynaSim" was implemented with a 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance in the documentation.
+
+"xolotl" performed well in time-step benchmarking, with comparable accuracy at high time-resolution. At low time-resolution, "xolotl" significantly outperforms "DynaSim" in both accuracy and speed.
+
+To test whether transient overhead effects had a significant effect on performance, "xolotl" and "DynaSim" were tested with time-step $dt = 0.1$ ms for varied lengths of time.
+"xolotl" and "DynaSim" both performed best during longer simulations, approaching maximal performance at $>10^5$ time steps.
+"xolotl" is about 20 times faster for short simulation times and 3.5 times faster for arbitrarily large ones.
+
 \section*{Conflict of Interest Statement}
@@
 (D-F) Simulated voltage trace of a model network for the three compartments.
-(G) Time series of synaptic currents in the simulated network with model cholinergic (chol) and glutamatergic (glut) synapses are outputs of the integration.}
+(G) Time series of synaptic currents in the simulated network are outputs of the integration.
+Two cholinergic (chol) current traces are plotted.}
 	\label{fig:figurenetwork}
@@
 (D) Calcium sensors change maximal conductances to move a neuron from quiescence to a bursting state.
-(E) Voltage trace of the controlled neuron shows regular bursting activity.}
+(E) Voltage trace shows regular bursting activity after integral control.}
 	\label{fig:figureintegralcontrol}
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_benchmark}
+	\caption{\texttt{xolotl} benchmarked against \texttt{DynaSim}.}
+	\label{fig:figurebenchmark}
+\end{figure}
+
 \end{document}

### 2018-05-25 figures/figure_benchmark.m

@@ -161,7 +161,7 @@ for ii = 1:length(t_end)
   V = x.integrate(0.2);
   t_sim = toc;
   % compute the speed as real-time / simulation-time
-  Qfactor(ii, 1) = t_end(ii) / t_sim;
+  Qfactor(ii, 1) = t_end(ii) / 1e3 / t_sim;
 end
 
 % test DynaSim
@@ -173,7 +173,7 @@ for ii = 1:length(t_end)
   data = dsSimulate(equations, 'solver', 'rk2', 'tspan', [0 t_end(ii)], 'dt', 0.1, 'compile_flag', 0);
   t_sim = toc;
   % compute the speed as real-time / simulation-time
-  Qfactor(ii, 2) = t_end(ii) / t_sim;
+  Qfactor(ii, 2) = t_end(ii) / 1e3 / t_sim;
 end
 
 % plot benchmark test #3

### 2018-05-25 figures/figure_network.m

@@ -108,17 +108,18 @@ end
 
 % plot the synaptic currents
 c = lines(10);
-plot(ax(7), time, synaptic_currents);
-for ii = 1:size(synaptic_currents, 2)
+plot(ax(7), time, synaptic_currents(:,1:2));
+for ii = 1:2
 	hplot(ii) = plot(NaN, NaN, 'o', 'MarkerFaceColor', c(ii, :), 'MarkerEdgeColor', c(ii, :), 'MarkerSize', 12);
 end
 xlabel(ax(7), 'time (s)')
 ylabel(ax(7), 'I_{syn} (nA)')
 set(ax(7), 'YScale', 'log')
-ylim(ax(7), [0.1 5000])
+% ylim(ax(7), [0.1 5000])
 xlim(ax(7), [0 max(time)]);
-legend(hplot, {'AB→LP (Chol)', 'AB→PY (Chol)', 'AB→LP (Glut)', 'AB→PY (Glut)', ...
-	'LP→PY (Glut)', 'PY→LP (Glut)', 'LP→AB (Glut)'}, 'Location', 'EastOutside')
+% legend(hplot, {'AB→LP (Chol)', 'AB→PY (Chol)', 'AB→LP (Glut)', 'AB→PY (Glut)', ...
+% 	'LP→PY (Glut)', 'PY→LP (Glut)', 'LP→AB (Glut)'}, 'Location', 'EastOutside')
+legend(hplot, {'AB→LP (Chol)', 'AB→PY (Chol)'}, 'Location', 'EastOutside')
 
 %% Post-Processing
 

### 2018-05-25 PaperLatex/paper.tex
Aesthetic fixes and some clarifications

@@
 
-\title[xolotl: neuronal simulator]{Xolotl: An Intuitive and Definitively Sexy Neuronal Simulator}
+\title[xolotl: neuronal simulator]{Xolotl: An Intuitive and Flexible Neuronal Simulator}
 \author[\firstAuthorLast ]{\Authors}
@@
 
-	\texttt{xolotl} is an open-source neuronal simulator written in \texttt{C++} with \texttt{MATLAB} wrappers.
-Complex models and networks can be designed efficiently using an intuitive language tightly coupled to the object-based architecture of the underlying \texttt{C++} code.
+	\texttt{xolotl} is a free and open-source neuronal simulator written in \texttt{C++} with \texttt{MATLAB} wrappers.
+Biophysically-detailed models of networks can be designed efficiently using an intuitive language tightly coupled to the object-based architecture of the underlying \texttt{C++} code.
 Models can be specified by adding conductances to compartment objects.
@@
 \texttt{C++} templates are provided for developing new conductances, compartments, and integration schemata.
-It also includes a customizable graphical user interface (GUI, "puppeteer") for rapid prototyping and hand-tuning conductances in real-time.
+It also includes a customizable graphical user interface (GUI) for rapid prototyping and hand-tuning conductances in real-time.
 The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate rapid construction and assessment of model networks.
@@
 	\tiny
-	\keyFont{ \section{Keywords:} simulator, MATLAB, C++, xolotl, conductance-based, computational, keyword, keyword} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
+	\keyFont{ \section{Keywords:} simulator, MATLAB, C++, conductance-based, neuron, network, keyword} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
 
@@
 
-\texttt{xolotl} (\url{https://github.com/sg-s/xolotl}) is a fast single-compartment and multi-compartment simulator in \texttt{C++} with \texttt{MATLAB} wrappers (\url{https://www.mathworks.com/products/matlab.html}).
-Written with an emphasis on flexibility and speed, \texttt{xolotl} can simulate single-compartment conductance-based models, networks of these, and detailed multi-compartment models.
+\texttt{xolotl} (\url{https://github.com/marderlab/xolotl}) is a fast single-compartment and multi-compartment simulator in \texttt{C++} with \texttt{MATLAB} wrappers.
+Written with an emphasis on ease-of-use, \texttt{xolotl} can simulate single-compartment conductance-based models, networks of these, and detailed multi-compartment models.
 \texttt{xolotl} exploits a novel automatic type system, \texttt{cpplab}, which binds \texttt{MATLAB} code to \texttt{C++} header files, creating objects and classes \textit{ad libitum} in \texttt{MATLAB} which reflect the underlying object-oriented code.
 \texttt{xolotl} implements \texttt{cpplab} to represent the nested structure of conductance-based models, and exploits the computational efficiency of the low-level programming language to quickly integrate models.
-For this reason, models can be implemented entirely in \texttt{MATLAB} with few lines of code.
+For this reason, models can be implemented entirely in \texttt{MATLAB} with a few lines of code.
 
-Models are specified in \texttt{MATLAB} by a \texttt{xolotl} object which contains compartment objects which themselves contain conductances.
+Models are specified in \texttt{MATLAB} by a nested structure.
+The \texttt{xolotl} object contains compartments which themselves contain conductances.
 Synapses belong to the \texttt{xolotl} object and connect compartments together.
@@
 
-\texttt{xolotl} provides parameter optimization capabilities through the algorithm-agnostic \texttt{procrustes} toolbox.
+%\texttt{xolotl} provides parameter optimization capabilities through the algorithm-agnostic \texttt{procrustes} toolbox.
 Any network parameters accessible through the \texttt{xolotl} structure can be optimized using arbitrary algorithms and objective functions on multi-core computers and high-performance computing clusters.
@@
 \texttt{cpplab} provides a powerful backend for specifying and integrating models without relying on the significantly slower and limiting \texttt{MATLAB codegen}.
+While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{coder} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
 Minimal experience with \texttt{MATLAB} is required to use \texttt{xolotl}, and all equations and integration methods are provided transparently to the end user.
@@
 Plotting of voltage, intracellular calcium, conductance gating functions, and time constants is provided by built-in \texttt{xolotl} methods.
-The \texttt{puppeteer} toolbox permits real-time tuning of any network parameters using numerical sliders in a graphical interface which displays the resultant membrane potential and intracellular calcium traces.
+The GUI permits real-time tuning of any network parameters using numerical sliders in a graphical interface which displays the resultant membrane potential and intracellular calcium traces.
 The ease-of-use of these tools lends them to pedagogical applications and rapid exploration of toy models.
@@
 
-Using \texttt{xolotl} in \texttt{MATLAB}, users create a \texttt{xolotl} object and populate it with compartments, synapses, and controllers.
-Each field is a \texttt{cpplab} object constructed by a function call to \texttt{add}.
+Using \texttt{xolotl} in \texttt{MATLAB}, users create a \texttt{xolotl} object and populate it "cpplab" objects which describe compartments, conductances, synapses, and controllers.
 The model is integrated with the \texttt{integrate} function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.
@@
 These objects can be referenced by name and added directly to a compartment.
-Novel conductance dynamics can be easily written by modifying a template header file contained in the \texttt{xolotl} distribution.
+Novel conductance dynamics can be easily written by modifying a template header file contained in the \texttt{xolotl} distribution, or designed entirely from scratch.
 
@@
 
-The seminal Hodgkin-Huxley model \autocite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952} contains a fast inactivating sodium conductance which promotes spiking, a non-inactivating potassium delayed rectifer, and a passive leak current (\ref{fig:figurehh}A).
-A compartment, \texttt{HH}, with membrane capacitance $C_m = 10 ~ \mu\mathrm{F/mm^2}$ and surface area $A = 0.1 ~ \mathrm{mm^2}$ can be specified by \ref{fig:figurehh}B.
-Compartment, conductance, synapse, and controller properties can be specified during the call to the \texttt{add} function, or after construction, using dot-notation in \texttt{MATLAB} (e.g.
+The seminal Hodgkin-Huxley model of action potentials in the squid giant axon \autocite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952} contains a fast inactivating sodium conductance ("NaV"), a non-inactivating delayed rectifier ("Kd"), and a passive leak current (\ref{fig:figurehh}A).
+A compartment, \texttt{HH}, with membrane capacitance ("Cm") and surface area ("A") can be specified by \ref{fig:figurehh}B.
+Network properties can be set during construction or afterwards using dot-notation in \texttt{MATLAB} (e.g.
 \texttt{x.HH.Cm}).
@@
 This model was constructed using conductances from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion \autocite{turrigianoSelectiveRegulationCurrent1995}.
-In the absence of applied positive current, the model is quiescent and tonically spikes under 0.2 nA of applied current (\ref{fig:figurehh}D).
+In the absence of applied positive current, the model is quiescent.
+When 0.2 nA is injected, the model tonically spikes (\ref{fig:figurehh}D).
 The \texttt{integrate} function takes the applied current as an argument (e.g.

### 2018-05-25 figures/figure_HH.m

@@ -78,11 +78,11 @@ a           = z + 1;
 curr_index  = x.contributingCurrents(this_V, this_I);
 
 % plot the voltage
-plot(ax(4), time0, V0, 'k', 'LineWidth', 3)
+plot(ax(4), time0, V0, 'k', 'LineWidth', 1.5)
 for qq = 1:size(this_I, 2)
   Vplot = this_V;
   Vplot(curr_index ~= qq) = NaN;
-  plot(ax(4), time + max(time0), Vplot, 'Color', c(qq,:), 'LineWidth', 3);
+  plot(ax(4), time + max(time0), Vplot, 'Color', c(qq,:), 'LineWidth', 1.5);
 end
 
 % xlabel(ax(4), 'time (s)')
@@ -112,7 +112,7 @@ end
 % plot on the correct axes
 plot(ax(6), all_I_ext, all_f, '-k')
 xlabel(ax(6), 'applied current (nA)')
-ylabel(ax(6), 'frequency (Hz)')
+ylabel(ax(6), 'firing rate (Hz)')
 set(ax(6), 'XLim', [min(all_I_ext)*1.1 max(all_I_ext)*1.05], 'XTick', [0 0.5 1])
 
 % set up tags
@@ -166,19 +166,19 @@ ax(9).Tag   = 'τ_m';
 ax(10).Tag  = 'τ_h';
 
 % set the xlabels and ylabels
-title(ax(7), 'm_∞')
+ylabel(ax(7), 'm_∞')
 xlabel(ax(7), 'V (mV)')
 set(ax(7), 'YLim', [0 1]);
 
 xlabel(ax(8), 'V (mV)')
-title(ax(8), 'h_∞')
+ylabel(ax(8), 'h_∞')
 set(ax(8), 'YLim', [0 1]);
 
-title(ax(9), 'τ_m (ms)')
+ylabel(ax(9), 'τ_m (ms)')
 xlabel(ax(9), 'V (mV)')
 set(ax(9),    'YScale','log', 'YTick', [10^-2 10^0 10^2], 'YLim', [10^-2 10^2])
 
-title(ax(10),'τ_h (ms)')
+ylabel(ax(10),'τ_h (ms)')
 xlabel(ax(10),'V (mV)')
 set(ax(10),   'YScale','log', 'YTick', [10^-2 10^0 10^2], 'YLim', [10^-2 10^2])
 
@@ -198,16 +198,16 @@ end
 
[…]

### 2018-05-25 figures/figure_clamp.m

@@ -116,15 +116,15 @@ set(ax(5), 'XLim', [90 121], 'YLim', [-90 60], 'YTick', [-80 -40 0 40]);
 %% Plot Current vs. Voltage
 
 plot(ax(6), Vsteps, current(end,:), 'k')
-xlabel(ax(6), 'voltage clamp (mV)')
+xlabel(ax(6), 'clamped voltage (mV)')
 ylabel(ax(6), 'current (nA)')
 set(ax(6), 'XLim', [min(Vsteps) max(Vsteps)], 'XTick', [-80 -40 0 40])
 
 %% Plot Conductance vs. Voltage
 conductance = current(end, :) ./ (Vsteps - x.AB.Kd.E) / x.AB.A;
 plot(ax(7), Vsteps, conductance, 'k');
-xlabel(ax(7), 'voltage clamp (mV)')
-ylabel(ax(7), 'ḡ (μS/mm^2)')
+xlabel(ax(7), 'clamped voltage (mV)')
+ylabel(ax(7), 'conductance (\muS/mm^2)')
 set(ax(7), 'XLim', [min(Vsteps) max(Vsteps)], 'XTick', [-80 -40 0 40])
 
 %% Plot Steady-State Kd Activation Gating Variable
@@ -147,12 +147,14 @@ end
 plot(ax(8), Vsteps, conductance/conductance(end), 'ok')
 hplot(end+1) = plot(ax(8), NaN, NaN, 'o', 'MarkerEdgeColor', [0 0 0], 'MarkerSize', 12);
 xlabel(ax(8), 'voltage clamp (mV)')
+ylabel(ax(8), 'm_\infty')
 set(ax(8), 'XLim', [min(Vsteps) max(Vsteps)], 'XTick', [-80 -40 0 40])
 legend(hplot, {'n = 1', 'n = 2', 'n = 3', 'n = 4', 'data'}, 'Location', 'EastOutside')
 %% Plot R^2 value
 
 warning off
 for j = 1:6
+	all_n(j) = j;
 	temp = conductance.^(1/j);
 	rm_this = isnan(temp) | isinf(temp);
 	all_r2(j) = rsquare(temp(~rm_this),minf(~rm_this));
@@ -164,6 +166,7 @@ warning on
 plot(ax(9), all_n, 1-all_r2, 'k')
 xlabel(ax(9), 'exponent')
 ylabel(ax(9), '1 - r^2')
[…]

### 2018-05-25 figures/figure_integral_control.m

@@ -89,10 +89,10 @@ time = x.dt*(1:length(C))*1e-3;
 Cplot = C(:,2:2:end);
 plot(ax(4), time, Cplot);
 for ii = 1:size(Cplot,2)
-	hplot(ii) = plot(ax(4), NaN, NaN, 'o', 'MarkerFaceColor', c(ii, :), 'MarkerEdgeColor', c(ii, :), 'MarkerSize', 12);
+	hplot(ii) = plot(ax(4), NaN, NaN, 'o', 'MarkerFaceColor', c(ii, :), 'MarkerEdgeColor', c(ii, :), 'MarkerSize', 8);
 end
-set(ax(4), 'XScale', 'log', 'YScale','log', 'YTick', [1e-2 1e0 1e2 1e4], 'XLim', [0 1.1e3])
-ylabel(ax(4), '$\bar{g}~\mu S/mm^2$', 'Interpreter', 'Latex')
+set(ax(4), 'XScale', 'log', 'YScale','log', 'YTick', [1e-2 1e0 1e2 1e4], 'XLim', [0 510], 'XTick', [0 1e0 1e1 1e2 1e3])
+ylabel(ax(4), {'maximal conductance'; '(\muS/mm^2)'})
 xlabel('time (s)')
 leg(1) = legend(hplot, x.AB.find('conductance'), 'Location', 'EastOutside');
 
@@ -102,14 +102,14 @@ x.dt = .1;
 x.t_end = 1e3;
 V = x.integrate;
 time = x.dt*(1:length(V))*1e-3;
-hplot(1) = plot(ax(5), NaN, NaN, 'o', 'MarkerFaceColor', [1 0 0], 'MarkerEdgeColor', [1 0 0], 'MarkerSize', 12);
-hplot(2) = plot(ax(5), NaN, NaN, 'o', 'MarkerFaceColor', [0 0 0], 'MarkerEdgeColor', [0 0 0], 'MarkerSize', 12);
+hplot(1) = plot(ax(5), NaN, NaN, 'o', 'MarkerFaceColor', [1 0 0], 'MarkerEdgeColor', [1 0 0], 'MarkerSize', 8);
+hplot(2) = plot(ax(5), NaN, NaN, 'o', 'MarkerFaceColor', [0 0 0], 'MarkerEdgeColor', [0 0 0], 'MarkerSize', 8);
 plot(ax(5), time, V_init, 'r', 'LineWidth', 1)
[…]

### 2018-05-25 getFigPosition.m

@@ -1,0 +1,10 @@
+function pos = getFigPosition(ax)
+  % creates a matrix of positions for a vector of axes
+
+  pos = zeros(length(ax), 4);
+
+  for ii = 1:length(ax)
+    pos(ii,:) = ax(ii).Position;
+  end
+
+end

### 2018-05-25 PaperLatex/paper.tex
Fixes to figure captions

@@
 
-AH and SG-S wrote the manuscript and the code. EM provided funding and moral support.
+SG-S designed and implemented the core of the "xolotl" toolbox.
+AH contributed to the code base, created the online user documentation, and wrote the manuscript.
+EM supervised the project.
+All authors reviewed the paper.
 
 \section*{Funding}
-Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.
 
@@
 (C) \texttt{xolotl} schematic displayed in the \texttt{MATLAB} command prompt.
-(D) Simulated voltage trace with 0.1 nA applied current.
+(D) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current.
 Colors indicate the dominant current (gold is fast sodium, blue is delayed rectifier, red is leak).
-(E) Frequency-input relation displaying firing rate as a function of applied current.
+(E) Firing rate-input relation displaying firing rate as a function of injected current current.
 (F-G) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
-Variables not plotted are unity for all voltage.
 (H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables.
-Variables not plotted are unity for all voltage.
-Colors indicate conductance type (gold is fast sodium, blue is delayed rectifier, red is leak).}
+Variables not plotted are unity for all voltage.}
 	\label{fig:figurehh}
@@
 	\includegraphics[width=1.0\linewidth]{gfx/figure_clamp}
-	\caption{\texttt{xolotl} readily implements voltage clamp.
-(A) Cartoon of a cell with potassium conductance with experimentally-fixed voltage.
+	\caption{Simulating a voltage-clamp experiment.
+(A) Cartoon of a cell with delayed rectifier potassium conductance \autocite{liuModelNeuronActivitydependent1998} with experimentally-fixed voltage.
 (B) Structure of \texttt{xolotl} object in A.
@@
 (H) Sigmoids $m$ fit to the model as $m^n$ data indicating that $n=4$ is the best fit.
-(I) $R^2$ correlation of the sigmoid fits at various powers where $n=4$ is an exact fit.}
+(I) $r^2$ correlation of the sigmoid fits at various powers where $n=4$ is an exact fit.}
 	\label{fig:figureclamp}
@@
 	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
-	\caption{\texttt{xolotl} readily implements conductance-based network models.
+	\caption{Simulating a network of conductance-based model neurons.
 (A) Diagram of a network model of the pyloric rhythm in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004).
@@
 	\includegraphics[width=1.0\linewidth]{gfx/figure_integral_control}
-	\caption{\texttt{xolotl} can implement homeostatic tuning rules as integral control.
-(A) Cartoon of a model neuron with integral control.
+	\caption{Simulating neurons under homeostatic regulation.
+(A) Cartoon of a model neuron \autocite{liuModelNeuronActivitydependent1998} with integral control \autocite{olearyCorrelationsIonChannel2013}.
 (B) Hierarchical structure of a neuronal network considers controllers as components of compartments which act on conductances.

### 2018-05-26 PaperLatex/paper.tex
Add more detail to the worked examples

@@
 The \texttt{integrate} function takes the applied current as an argument (e.g.
-"x.integrate(AppliedCurrent)"), so that the \texttt{xolotl} object is agnostic to integration-specific perturbations.
+"x.integrate(Iapp)"), so that the \texttt{xolotl} object is agnostic to integration-specific perturbations.
 The \texttt{plot} function generates voltage and intracellular calcium traces, where the voltage trace is colored by the dominant current.
@@
 \texttt{xolotl} can recapitulate the results of voltage clamp experiments \autocite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}.
-\ref{fig:figureclamp} displays the procedure to clamp the membrane potential of a cell with a delayed rectifier potassium conductance.
-The second argument of the \texttt{integrate} function determines the clamped voltage (e.g.
-\texttt{x.integrate([], VoltageClamp)}).
-Isolated currents under voltage clamp approach the steady-state (\ref{fig:figureclamp}D-E) so that a current-voltage relation at steady-state can be extracted (\ref{fig:figureclamp}F).
-The derivative of the IV curve is the steady-state conductance (\ref{fig:figureclamp}G).
+Figure \ref{fig:figureclamp} displays steps in the procedure to clamp the membrane potential of a cell with delayed rectifier potassium conductance.
+During an \textit{in-vitro} experiment, confounding currents would be pharmacologically-blocked and two-electrode voltage clamp used to record tail currents at fixed membrane potential \autocite{connorInwardDelayedOutward1971, connorVoltageClampStudies1971}.
+
+A single-compartment model with a delayed-rectifier conductance is simulated at stepped membrane potentials.
+The model is simulated using the "integrate" function.
+The second argument determines the clamped voltage and the fourth output is the current trace.
+
+\begin{lstlisting}[style=Matlab-editor]
+	[~, ~, ~, current_trace] = x.integrate([], clamped_voltage)
+\end{lstlisting}
+
+Currents under voltage clamp approach the steady-state holding current (\ref{fig:figureclamp}D-E).
+The current-voltage relation is the steady-state current over the clamped voltage, and the effective conductance is the derivative of that relation (\ref{fig:figureclamp}F-G).
 Fitting a sigmoid to various powers yields a model for the current dynamics (\ref{fig:figureclamp}H-I).
@@
 This structure is reciprocated in the hierarchy of the \texttt{xolotl} object, where conductances are contained within compartments (\ref{fig:figurenetwork}B).
+
+Representing the network in "xolotl" requires constructing three compartments and eight conductances in each using the "add" function.
+
+\begin{lstlisting}[style=Matlab-editor]
+	x.add('AB', 'compartment', 'Cm', 10, 'A', 0.628, ...)
+	x.AB.add('prinz/NaV', 'gbar', 1000, 'E', 50)
+	...
+\end{lstlisting}
+
 Synapses are upper-level properties of the network which point between two compartments (\ref{fig:figurenetwork}C).
 This exploits vectorized operations in \texttt{MATLAB} and does not require each synapse to possess a unique name.
-The \texttt{plot} function generates multiple subplots when called for a network with multiple compartments (\ref{fig:figurenetwork}D-F).
+The "connect" function quickly adds synapses to the network.
+
+\begin{lstlisting}[style=Matlab-editor]
+	x.connect('AB', 'LP', 'Chol', 'gbar', 30)
+\end{lstlisting}
+
+The "plot" method generates multiple subplots when called for a network with multiple compartments (\ref{fig:figurenetwork}D-F), providing a quick way to visualize a simulation.
 
@@
 
-\texttt{xolotl} uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments.
+"xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments.
+These defaults provide a mix of speed, accuracy, and stability, and are built into the "cpplab" header files.
+Custom "cpplab" header files can be customized with any iterative integration method.
 The simulation time-resolution can be specified to target arbitrary precision, and an output time step can be selected to support automatic down-sampling for memory considerations.
@@
 
-\subsection{Using the \texttt{puppeteer} graphical interface}
+\subsection{Using the graphical interface to manipulate parameters}
 
 \texttt{xolotl} comes packaged with a graphical user interface for visualizing parameter changes in real-time.
-The \texttt{manipulate} function opens the GUI, which displays a figure plotting the membrane potential and intracellular calcium concentration of all compartments as time series, and a dialog box with customizable sliders for all parameters of the model.
+The \texttt{manipulate} function opens the GUI, which displays a figure plotting the membrane potential and intracellular calcium concentration of all compartments as time series, and a dialog box with customizable sliders for all parameters of the model, much like the "Manipulate" function in Wolfram "Mathematica".
 Moving the sliders integrates the model in `open-loop` mode with the new parameters.
@@
 
-\texttt{xolotl} performs well in benchmarking tests.
 To assess speed and accuracy, "xolotl", "DynaSim" \autocite{sherfeyDynaSimMATLABToolbox2018}, and "NEURON" \autocite{hinesNEURONSimulationEnvironment1997} were compared in simulations over varied time-resolution and simulation time (\ref{fig:figurebenchmark}).
@@
 
-"xolotl" performed well in time-step benchmarking, with comparable accuracy at high time-resolution. At low time-resolution, "xolotl" significantly outperforms "DynaSim" in both accuracy and speed.
+"xolotl" and "DynaSim" performed with comparable accuracy at high time-resolution. At low time-resolution, "xolotl" significantly outperforms "DynaSim" in both accuracy and speed.
 
@@
 
-The code to generate all figures can be found in the \texttt{xolotl} repository (\url{https://github.com/marderlab/xolotl}).
+The code to generate all figures can be found in the \texttt{xolotl-paper} repository (\url{https://github.com/marderlab/xolotl-paper}).
+"xolotl" can be downloaded from its repository (\url{https://github.com/marderlab/xolotl}).
 

### 2018-05-29 PaperLatex/paper.tex
minor fixes

@@
 
-\title[xolotl: neuronal simulator]{Xolotl: An Intuitive and Flexible Neuronal Simulator}
+\title[xolotl: neuronal simulator]{Xolotl: An Intuitive and Approachable Neuron \& Network Simulator in MATLAB}
 \author[\firstAuthorLast ]{\Authors}
@@
 Minimal experience with \texttt{MATLAB} is required to use \texttt{xolotl}, and all equations and integration methods are provided transparently to the end user.
-No string parsing of equations is required.
+No string parsing of equations is required \autocite{sherfeyDynaSimMATLABToolbox2018, stimbergBrianSecondComing2013, stimbergEquationorientedSpecificationNeural2014}.
 
-\texttt{xolotl} comes packaged with visualization functions, a graphical user interface (GUI) for real-time manipulation of model parameters.
+\texttt{xolotl} comes packaged with visualization functions and a graphical user interface (GUI) for real-time manipulation of model parameters.
 Plotting of voltage, intracellular calcium, conductance gating functions, and time constants is provided by built-in \texttt{xolotl} methods.
@@
 
-Using \texttt{xolotl} in \texttt{MATLAB}, users create a \texttt{xolotl} object and populate it "cpplab" objects which describe compartments, conductances, synapses, and controllers.
+Using \texttt{xolotl} in \texttt{MATLAB}, users create a \texttt{xolotl} object and populate it with "cpplab" objects which describe compartments, conductances, synapses, and controllers.
 The model is integrated with the \texttt{integrate} function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.
@@
 \begin{lstlisting}[style=Matlab-editor]
-	[~, ~, ~, current_trace] = x.integrate([], clamped_voltage)
+	[V, Ca, ~, I] = x.integrate([], clamped_voltage)
 \end{lstlisting}
@@
 The current-voltage relation is the steady-state current over the clamped voltage, and the effective conductance is the derivative of that relation (\ref{fig:figureclamp}F-G).
+Since the effective conductance is the product of the maximal conductance and the gating variables \autocite{dayanTheoreticalNeuroscience2001, turrigianoSelectiveRegulationCurrent1995} and the tail current is monotonically-increasing with time under voltage clamp, the current can be represented as non-inactivating.
 Fitting a sigmoid to various powers yields a model for the current dynamics (\ref{fig:figureclamp}H-I).
@@
 This exploits vectorized operations in \texttt{MATLAB} and does not require each synapse to possess a unique name.
-The "connect" function quickly adds synapses to the network.
+The "connect" function adds synapses to the network.
 
@@
 \end{lstlisting}
-
-The "plot" method generates multiple subplots when called for a network with multiple compartments (\ref{fig:figurenetwork}D-F), providing a quick way to visualize a simulation.
 
@@
 \texttt{xolotl} can implement homeostatic tuning rules as integral control.
-Figure \ref{fig:figureintegralcontrol} depicts generation of a bursting neuron model from quiescent conditions.
-Calcium sensors supervene on maximal conductance density (\ref{fig:figureintegralcontrol}).
-In \texttt{xolotl}, integral controllers are properties of conductances (\ref{fig:figureintegralcontrol}B-C).
-They modify properties of the conductance in response to an error signal.
+The controller computes an error signal (typically a function of intracellular calcium concentration), and adjusts the conductance or synapse it controls accordingly \autocite{olearyCorrelationsIonChannel2013}.
+In "xolotl", integral controllers are "cpplab" objects added to the conductance or synapse they regulate.
+
 In a demonstration adapted from \cite{olearyCorrelationsIonChannel2013}, integral control changes maximal conductances to bring a neuron from quiescence into a bursting regime.
+Calcium sensors supervene on maximal conductance density (\ref{fig:figureintegralcontrol}) to change neuronal activity.
+Each conductance in the "xolotl" structure contains a calcium-sensitive controller (\ref{fig:figureintegralcontrol}B-C).
 Maximal conductances increase from random initial conditions to a set which elicits the desired network output by minimizing the error signal (\ref{fig:figureintegralcontrol}D-F).
@@
 \texttt{xolotl} relies on \texttt{cpplab} constructions, which allow the user to exploit the efficiency of low-level \texttt{C++} code.
-\texttt{MATLAB} treats \texttt{cpplab} objects as standard variables allowing for symbolic manipulation using only the high-level programming language and graphical interfaces.
-\texttt{xolotl} is fast specifically because all time-intensive code is written in native \texttt{C++}.
-While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{coder} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
+\texttt{MATLAB} treats \texttt{cpplab} objects as fully-typed variables allowing for symbolic manipulation using only the high-level programming language and graphical interfaces.
+\texttt{xolotl} is fast because all time-intensive code is written in \texttt{C++}.
+While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
 
-Native \texttt{C++} provides drastic speed improvements beyond the benefits of translating \texttt{MATLAB} features into low-level code.
+\texttt{C++} provides speed improvements beyond the benefits of translating \texttt{MATLAB} features into low-level code.
 For this reason, \texttt{cpplab} has been designed to provide an interface for constructing, transpiling, and compiling \texttt{C++} code to be called from within \texttt{MATLAB}.
-\texttt{xolotl} simulations can be run entirely from \texttt{C++} executables.
-To facilitate use in \texttt{MATLAB}, \texttt{xolotl} uses the MD5 algorithm to automatically hash the network and compile a new binary and \texttt{MEX} bridge file only if needed.
-\texttt{MATLAB} provides a high level programmatic and graphical interface for implementing, manipulating, and visualizing models without sacrificing the enhancements of native \texttt{C++} code.
+\texttt{xolotl} simulations are run entirely from \texttt{C++} executables.
+
+\texttt{xolotl} automatically uses the MD5 algorithm to hash the network and compile a new binary and \texttt{MEX} bridge file only if needed.
+\texttt{MATLAB} provides a high level programmatic and graphical interface for implementing, manipulating, and visualizing models without sacrificing the enhancements of the underlying \texttt{C++} code.
 
@@
 
-\texttt{cpplab} comes with several features which simplify the handling of high-dimensional models.
+\texttt{cpplab} comes with several features which simplify the handling of complexly-nested models.
 The \texttt{find} function acquires a cell array of all properties of the network which satisfy a search condition.
-For example, one can find all paths to maximal conductances within the "'HH'" compartment in two ways:
+For example, one can find all paths to maximal conductances within the "'HH'" compartment by:
 
@@
 	x.find('HH*gbar');
-x.HH.find('gbar');
 \end{lstlisting}

### 2018-05-29 PaperLatex/paper.tex
minor fixes

@@
 \texttt{cpplab} provides a powerful backend for specifying and integrating models without relying on the significantly slower and limiting \texttt{MATLAB codegen}.
-While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{coder} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
+While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
 Minimal experience with \texttt{MATLAB} is required to use \texttt{xolotl}, and all equations and integration methods are provided transparently to the end user.
@@
 
-Simulations can be run in 'open-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or 'closed-loop' mode which begins simulation with the current network state.
+Simulations can be run in 'closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or 'open-loop' mode which begins simulation with the current network state.
 
@@
 
-The code to generate all figures can be found in the \texttt{xolotl-paper} repository (\url{https://github.com/marderlab/xolotl-paper}).
-"xolotl" can be downloaded from its repository (\url{https://github.com/marderlab/xolotl}).
+The code to generate all figures is available at (\url{https://github.com/marderlab/xolotl-paper}). "xolotl" is frely available at (\url{https://github.com/marderlab/xolotl}).
 

### 2018-05-29 PaperLatex/paper.tex
Add citations and clarify language

@@
 
-"xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments.
+"xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments \autocite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001, butcherNumericalDifferentialEquation2016}.
 These defaults provide a mix of speed, accuracy, and stability, and are built into the "cpplab" header files.
@@
 (H) Sigmoids $m$ fit to the model as $m^n$ data indicating that $n=4$ is the best fit.
-(I) $r^2$ correlation of the sigmoid fits at various powers where $n=4$ is an exact fit.}
+(I) Goodness of fit vs.
+exponent $n$, suggesting $n=4$ as the best fit to the data.}
 	\label{fig:figureclamp}

### 2018-05-29 PaperLatex/paper.tex
Clarify language in figure captions

@@
 (A) Diagram of a network model of the pyloric rhythm in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004).
-(B) Hierarchical structure of a neuronal network considers compartments as components of the network and conductances and synapses as components of compartments.
+(B) Each neuron is modeled as a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
 (C) \textit{xolotl} implements conductances as fields of compartments and synapses as connections between compartments.
 (D-F) Simulated voltage trace of a model network for the three compartments.
-(G) Time series of synaptic currents in the simulated network are outputs of the integration.
-Two cholinergic (chol) current traces are plotted.}
+(G) Time series of synaptic currents in the simulated network can be obtained from the integration.}
 	\label{fig:figurenetwork}

### 2018-05-29 figures/figure_benchmark.m
Set up simulations in NEURON

@@ -42,104 +42,6 @@ equations = { ...
   'taun(v)=7.2-6.4/(1.0+exp((v+28.3)/-19.2))'};
 
 %% Benchmark Test #1
-% simulate a hodgkin-huxley model over a series of time-resolutions
-
-% set up time-resolution vector
-max_dt    = 1000;
-k         = 1:max_dt;
-dt        = k(rem(max_dt, k) == 0) / 1e3;
-
-% set up outputs
-sim_time          = NaN(length(dt), 3); % duration of runtime
-r2                = NaN(length(dt), 3); % correlation coefficient
-
-% perform benchmark for xolotl
-
-x.closed_loop = false;
-
-x.sim_dt = dt(1);
-x.dt = max_dt*1e-3;
-tic
-V0 = x.integrate(0.2);
-V0_diff = diff(V0);
-sim_time(1,1) = toc;
-
-
-for i = 2:length(dt)
-  textbar(i, length(dt))
-	x.sim_dt = dt(i);
-	tic
-	V = x.integrate(0.2);
-	sim_time(i,1) = toc;
-
-	% measure distance b/w diff-embedded attractors
-	V_diff = diff(V);
-
-	this_cost = 0;
-	for j = 2:10:length(V0_diff)
-		for k = 1:size(V,2)
-			this_cost = this_cost + min(sqrt((V0_diff(j,k) - V_diff(:,k)).^2 + (V0(2:end,k) - V(2:end,k)).^2));
-		end
-	end
-
-	r2(i,1) = this_cost;
-
-end
-
-% perform benchmark for DynaSim
-sim_dt            = dt(1);
-sampling          = max_dt / 1e3;
-
-% precompile
-data              = dsSimulate(equations, 'solver', 'rk2', 'tspan', [0 t_end], 'dt', sim_dt, 'compile_flag', 0);
-
-% set up "best-case" for accuracy
-tic
-data0             = dsSimulate(equations, 'solver', 'rk2', 'tspan', [0 t_end], 'dt', dt(1), 'compile_flag', 0, 'downsample_factor', 1/dt(1));
-sim_time(1,2)     = toc;
-V0                = data0.(data0.labels{1});
-V0_diff           = diff(V0);
-
-for ii = 2:length(dt)
-  textbar(ii, length(dt))
-  sim_dt          = dt(ii);
-  tic
[…]

### 2018-05-29 neuron/benchmark1.py

@@ -1,18 +1,21 @@
 ## Benchmark Test #1
 # single-compartment, Hodgkin-Huxley
-# test speed and accuracy with increasing time-step
+# test speed with increasing simulation time
+
+# execute the main neuron file
+exec("neuron_HH.py")
 
 # set up vectors to hold outputs
-dt              = np.array([0.0010, 0.0020, 0.0040, 0.0050, 0.0080, 0.0100,
-                0.0200, 0.0250, 0.0400, 0.0500, 0.1000, 0.1250, 0.2000 0.2500,
-                0.5000, 1.0000])
-sim_time        = np.zeros((len(dt), 1))
-r2              = np.zeros((len(dt), 1))
+t_end           = np.round(np.logspace(1, 1e6, num=20))
+sim_time        = np.zeros((len(t_end),1))
 
 # perform the simulation
-for ii in range(1, len(dt)):
-    h.dt = dt[ii]
+for ii in range(1,len(t_end)):
+    h.tspan = t_end[ii]
     tic = time.clock()
     h.run()
     toc = time.clock() - tic
     sim_dt[ii] = toc;
+
+# save the results
+np.savetxt("neuron_benchmark2.csv", sim_dt, delimiter=",")

### 2018-05-29 neuron/benchmark2.py

@@ -1,21 +1,0 @@
-## Benchmark Test #2
-# single-compartment, Hodgkin-Huxley
-# test speed with increasing simulation time
-
-# execute the main neuron file
-exec("neuron_HH.py")
-
-# set up vectors to hold outputs
-t_end           = np.round(np.logspace(1, 1e6, num=20))
-sim_time        = np.zeros((len(t_end),1))
-
-# perform the simulation
-for ii in range(1,len(t_end)):
-    h.tspan = t_end[ii]
-    tic = time.clock()
-    h.run()
-    toc = time.clock() - tic
-    sim_dt[ii] = toc;
-
-# save the results
-np.savetxt("neuron_benchmark2.csv", sim_dt, delimiter=",")

### 2018-05-29 neuron/neuron_HH.py

@@ -6,7 +6,7 @@ from matplotlib import pyplot
 
 # create the neuron
 soma        = h.Section(name='soma');
-h.psection(sec=soma)
+# h.psection(sec=soma)
 
 # set the size of the soma
 soma.L      = 28.209 # microns
@@ -18,13 +18,14 @@ soma.cm     = 1 # uF/cm^2
 
 # add conductances from Liu et al. 1998
 soma.insert('pas')
-soma.insert('liu_Kd')
-soma.insert('liu_NaV')
 
-# set up maximal conductances
-soma.pas.gbar       = 1 / 1e4 # S/cm^2
-soma.liu_NaV.gbar   = 1000 / 1e4 # S/cm^2
-soma.liu_Kd.gbar    = 300 / 1e4 # S/cm^2
+soma.insert('na')
+soma.insert('kd')
+
+# set maximal conductances
+soma(0.5).pas.g = 0.0001
+
+h.psection(sec=soma)
 
 # set up injected current
 stim = h.IClamp(soma(0.5))
@@ -38,6 +39,15 @@ v_vec.record(soma(0.5)._ref_v)
 t_vec.record(h._ref_t)
 
 # set up simulation
-h.dt    = 0.05/1000 # s
-h.tstop = 100.0 # s
-h.secondorder = 2 # use Crank-Nicholson
+h.dt    = 0.05
+h.tstop = 1000.0 # ms
+# h.secondorder = 2 # use Crank-Nicholson
+
+h.run()
+
+# visualize the results
+pyplot.figure(figsize=(8,4)) # Default figsize is (8,6)
+pyplot.plot(t_vec, v_vec)
+pyplot.xlabel('time (ms)')
+pyplot.ylabel('mV')
+pyplot.show()

### 2018-05-29 neuron/neuron_python_HH.py

@@ -1,14 +1,0 @@
-
-
-# run simulation
-aaa = time.clock()
-h.run()
-print(time.clock()-aaa)
-
-# visualize
-from matplotlib import pyplot
-pyplot.figure(figsize=(8,4)) # Default figsize is (8,6)
-pyplot.plot(t_vec, v_vec)
-pyplot.xlabel('time (s)')
-pyplot.ylabel('mV')
-pyplot.show()

### 2018-05-29 neuron/test.py

@@ -1,0 +1,25 @@
+from neuron import h, gui
+
+soma = h.Section(name='soma')
+
+# soma.insert('pas')
+soma.insert('CaT')
+
+# asyn = h.AlphaSynapse(soma(0.5))
+# asyn.onset = 20
+# asyn.gmax = 1
+
+v_vec = h.Vector()             # Membrane potential vector
+t_vec = h.Vector()             # Time stamp vector
+v_vec.record(soma(0.5)._ref_v)
+t_vec.record(h._ref_t)
+
+h.tstop = 40000.0
+h.run()
+
+from matplotlib import pyplot
+pyplot.figure(figsize=(8,4)) # Default figsize is (8,6)
+pyplot.plot(t_vec, v_vec)
+pyplot.xlabel('time (ms)')
+pyplot.ylabel('mV')
+pyplot.show()

### 2018-05-29 neuron/x86_64/CaT.c

@@ -1,0 +1,580 @@
+/* Created by Language version: 6.2.0 */
+/* VECTORIZED */
+#define NRN_VECTORIZED 1
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include "scoplib_ansi.h"
+#undef PI
+#define nil 0
+#include "md1redef.h"
+#include "section.h"
+#include "nrniv_mf.h"
+#include "md2redef.h"
+ 
+#if METHOD3
+extern int _method3;
+#endif
+
+#if !NRNGPU
+#undef exp
+#define exp hoc_Exp
+extern double hoc_Exp(double);
+#endif
+ 
+#define nrn_init _nrn_init__CaT
+#define _nrn_initial _nrn_initial__CaT
+#define nrn_cur _nrn_cur__CaT
+#define _nrn_current _nrn_current__CaT
+#define nrn_jacob _nrn_jacob__CaT
+#define nrn_state _nrn_state__CaT
+#define _net_receive _net_receive__CaT 
+#define _f_settables _f_settables__CaT 
+#define settables settables__CaT 
+#define states states__CaT 
+ 
+#define _threadargscomma_ _p, _ppvar, _thread, _nt,
+#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
+#define _threadargs_ _p, _ppvar, _thread, _nt
+#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
+ 	/*SUPPRESS 761*/
+	/*SUPPRESS 762*/
+	/*SUPPRESS 763*/
+	/*SUPPRESS 765*/
+	 extern double *getarg();
+ /* Thread safe. No static _p or _ppvar. */
+ 
+#define t _nt->_t
+#define dt _nt->_dt
+#define gmax _p[0]
+#define r _p[1]
+#define s _p[2]
+#define d _p[3]
+#define eca _p[4]
+#define ica _p[5]
+#define ralpha _p[6]
+#define rbeta _p[7]
+#define salpha _p[8]
+#define sbeta _p[9]
+#define dalpha _p[10]
+#define dbeta _p[11]
+#define Dr _p[12]
+#define Ds _p[13]
+#define Dd _p[14]
+#define v _p[15]
[…]

### 2018-05-29 neuron/x86_64/Kd.c

@@ -1,0 +1,407 @@
+/* Created by Language version: 6.2.0 */
+/* VECTORIZED */
+#define NRN_VECTORIZED 1
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include "scoplib_ansi.h"
+#undef PI
+#define nil 0
+#include "md1redef.h"
+#include "section.h"
+#include "nrniv_mf.h"
+#include "md2redef.h"
+ 
+#if METHOD3
+extern int _method3;
+#endif
+
+#if !NRNGPU
+#undef exp
+#define exp hoc_Exp
+extern double hoc_Exp(double);
+#endif
+ 
+#define nrn_init _nrn_init__Kd
+#define _nrn_initial _nrn_initial__Kd
+#define nrn_cur _nrn_cur__Kd
+#define _nrn_current _nrn_current__Kd
+#define nrn_jacob _nrn_jacob__Kd
+#define nrn_state _nrn_state__Kd
+#define _net_receive _net_receive__Kd 
+#define settables settables__Kd 
+#define states states__Kd 
+ 
+#define _threadargscomma_ _p, _ppvar, _thread, _nt,
+#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
+#define _threadargs_ _p, _ppvar, _thread, _nt
+#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
+ 	/*SUPPRESS 761*/
+	/*SUPPRESS 762*/
+	/*SUPPRESS 763*/
+	/*SUPPRESS 765*/
+	 extern double *getarg();
+ /* Thread safe. No static _p or _ppvar. */
+ 
+#define t _nt->_t
+#define dt _nt->_dt
+#define gbar _p[0]
+#define m _p[1]
+#define i _p[2]
+#define minf _p[3]
+#define taum _p[4]
+#define Dm _p[5]
+#define v _p[6]
+#define _g _p[7]
+ 
+#if MAC
+#if !defined(v)
+#define v _mlhv
+#endif
+#if !defined(h)
+#define h _mlhh
+#endif
+#endif
+ 
+#if defined(__cplusplus)
+extern "C" {
+#endif
+ static int hoc_nrnpointerindex =  -1;
+ static Datum* _extcall_thread;
+ static Prop* _extcall_prop;
[…]

### 2018-05-29 neuron/x86_64/NaV.c

@@ -1,0 +1,419 @@
+/* Created by Language version: 6.2.0 */
+/* VECTORIZED */
+#define NRN_VECTORIZED 1
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include "scoplib_ansi.h"
+#undef PI
+#define nil 0
+#include "md1redef.h"
+#include "section.h"
+#include "nrniv_mf.h"
+#include "md2redef.h"
+ 
+#if METHOD3
+extern int _method3;
+#endif
+
+#if !NRNGPU
+#undef exp
+#define exp hoc_Exp
+extern double hoc_Exp(double);
+#endif
+ 
+#define nrn_init _nrn_init__NaV
+#define _nrn_initial _nrn_initial__NaV
+#define nrn_cur _nrn_cur__NaV
+#define _nrn_current _nrn_current__NaV
+#define nrn_jacob _nrn_jacob__NaV
+#define nrn_state _nrn_state__NaV
+#define _net_receive _net_receive__NaV 
+#define settables settables__NaV 
+#define states states__NaV 
+ 
+#define _threadargscomma_ _p, _ppvar, _thread, _nt,
+#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
+#define _threadargs_ _p, _ppvar, _thread, _nt
+#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
+ 	/*SUPPRESS 761*/
+	/*SUPPRESS 762*/
+	/*SUPPRESS 763*/
+	/*SUPPRESS 765*/
+	 extern double *getarg();
+ /* Thread safe. No static _p or _ppvar. */
+ 
+#define t _nt->_t
+#define dt _nt->_dt
+#define gbar _p[0]
+#define m _p[1]
+#define h _p[2]
+#define i _p[3]
+#define minf _p[4]
+#define hinf _p[5]
+#define taum _p[6]
+#define tauh _p[7]
+#define Dm _p[8]
+#define Dh _p[9]
+#define v _p[10]
+#define _g _p[11]
+ 
+#if MAC
+#if !defined(v)
+#define v _mlhv
+#endif
+#if !defined(h)
+#define h _mlhh
+#endif
+#endif
+ 
+#if defined(__cplusplus)
+extern "C" {
+#endif
[…]

### 2018-05-29 neuron/x86_64/kd.c

@@ -1,0 +1,443 @@
+/* Created by Language version: 6.2.0 */
+/* VECTORIZED */
+#define NRN_VECTORIZED 1
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include "scoplib_ansi.h"
+#undef PI
+#define nil 0
+#include "md1redef.h"
+#include "section.h"
+#include "nrniv_mf.h"
+#include "md2redef.h"
+ 
+#if METHOD3
+extern int _method3;
+#endif
+
+#if !NRNGPU
+#undef exp
+#define exp hoc_Exp
+extern double hoc_Exp(double);
+#endif
+ 
+#define nrn_init _nrn_init__kd
+#define _nrn_initial _nrn_initial__kd
+#define nrn_cur _nrn_cur__kd
+#define _nrn_current _nrn_current__kd
+#define nrn_jacob _nrn_jacob__kd
+#define nrn_state _nrn_state__kd
+#define _net_receive _net_receive__kd 
+#define rates rates__kd 
+#define states states__kd 
+ 
+#define _threadargscomma_ _p, _ppvar, _thread, _nt,
+#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
+#define _threadargs_ _p, _ppvar, _thread, _nt
+#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
+ 	/*SUPPRESS 761*/
+	/*SUPPRESS 762*/
+	/*SUPPRESS 763*/
+	/*SUPPRESS 765*/
+	 extern double *getarg();
+ /* Thread safe. No static _p or _ppvar. */
+ 
+#define t _nt->_t
+#define dt _nt->_dt
+#define Erev _p[0]
+#define i _p[1]
+#define m _p[2]
+#define g _p[3]
+#define minf _p[4]
+#define tau_m _p[5]
+#define Dm _p[6]
+#define v _p[7]
+#define _g _p[8]
+ 
+#if MAC
+#if !defined(v)
+#define v _mlhv
+#endif
+#if !defined(h)
+#define h _mlhh
+#endif
+#endif
+ 
+#if defined(__cplusplus)
+extern "C" {
+#endif
+ static int hoc_nrnpointerindex =  -1;
+ static Datum* _extcall_thread;
[…]

### 2018-05-29 neuron/x86_64/mod_func.c

@@ -1,0 +1,19 @@
+#include <stdio.h>
+#include "hocdec.h"
+extern int nrnmpi_myid;
+extern int nrn_nobanner_;
+
+extern void _na_reg(void);
+extern void _kd_reg(void);
+
+void modl_reg(){
+  if (!nrn_nobanner_) if (nrnmpi_myid < 1) {
+    fprintf(stderr, "Additional mechanisms from files\n");
+
+    fprintf(stderr," na.mod");
+    fprintf(stderr," kd.mod");
+    fprintf(stderr, "\n");
+  }
+  _na_reg();
+  _kd_reg();
+}

### 2018-05-29 neuron/x86_64/na.c

@@ -1,0 +1,478 @@
+/* Created by Language version: 6.2.0 */
+/* VECTORIZED */
+#define NRN_VECTORIZED 1
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+#include "scoplib_ansi.h"
+#undef PI
+#define nil 0
+#include "md1redef.h"
+#include "section.h"
+#include "nrniv_mf.h"
+#include "md2redef.h"
+ 
+#if METHOD3
+extern int _method3;
+#endif
+
+#if !NRNGPU
+#undef exp
+#define exp hoc_Exp
+extern double hoc_Exp(double);
+#endif
+ 
+#define nrn_init _nrn_init__na
+#define _nrn_initial _nrn_initial__na
+#define nrn_cur _nrn_cur__na
+#define _nrn_current _nrn_current__na
+#define nrn_jacob _nrn_jacob__na
+#define nrn_state _nrn_state__na
+#define _net_receive _net_receive__na 
+#define rates rates__na 
+#define states states__na 
+ 
+#define _threadargscomma_ _p, _ppvar, _thread, _nt,
+#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
+#define _threadargs_ _p, _ppvar, _thread, _nt
+#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
+ 	/*SUPPRESS 761*/
+	/*SUPPRESS 762*/
+	/*SUPPRESS 763*/
+	/*SUPPRESS 765*/
+	 extern double *getarg();
+ /* Thread safe. No static _p or _ppvar. */
+ 
+#define t _nt->_t
+#define dt _nt->_dt
+#define Erev _p[0]
+#define i _p[1]
+#define m _p[2]
+#define h _p[3]
+#define g _p[4]
+#define minf _p[5]
+#define hinf _p[6]
+#define tau_h _p[7]
+#define tau_m _p[8]
+#define Dm _p[9]
+#define Dh _p[10]
+#define v _p[11]
+#define _g _p[12]
+ 
+#if MAC
+#if !defined(v)
+#define v _mlhv
+#endif
+#if !defined(h)
+#define h _mlhh
+#endif
+#endif
+ 
+#if defined(__cplusplus)
+extern "C" {
+#endif
[…]

### 2018-05-30 PaperLatex/paper.tex
Set up simulations in NEURON
Restructure text

@@
 %
-%		WORKED EXAMPLES
+%		DESIGN GOALS
 %
@@
 
-\section{Worked Examples}
-\label{sec:examples}
+\section{Design Goals}
+\label{design}
+
+"xolotl" is designed to be easy-to-use without sacrificing speed.
+The software has been designed in "MATLAB" due to its popularity among neuroscientists for pedagogy and research.
+"xolotl" capitalizes on "MATLAB"'s straightforward structure array syntax to permit rapid prototyping and experimentation, especially for small neuronal networks of complex models.
+Parameters of conductances, neuronal compartments, and simulations may all be edited in the structure before any calls to integration functions.
+The underlying code is written in "C++" for speed and memory optimization, and while models can indeed be integrated using the compiled binary, symbolic manipulations can be readily performed in "MATLAB" without ever touching the foundational code.
+
+\subsection{Features}
+\label{features}
+
+Models are specified by adding compartments and synapses to the \texttt{xolotl} object.
+Conductances are added to compartments and controllers can be added to conductances.
+This modular structure recapitulates the biophysics of the Hodgkin-Huxley formalism and obviates the need to explicitly write out equations, which in \texttt{xolotl} are contained within the conductance header files.
+
+\texttt{xolotl} relies on \texttt{cpplab} constructions, which allow the user to exploit the efficiency of low-level \texttt{C++} code.
+\texttt{MATLAB} treats \texttt{cpplab} objects as fully-typed variables allowing for symbolic manipulation using only the high-level programming language and graphical interfaces.
+\texttt{xolotl} is fast because all time-intensive code is written in \texttt{C++}.
+While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
+
+\texttt{C++} provides speed improvements beyond the benefits of translating \texttt{MATLAB} features into low-level code.
+For this reason, \texttt{cpplab} has been designed to provide an interface for constructing, transpiling, and compiling \texttt{C++} code to be called from within \texttt{MATLAB}.
+\texttt{xolotl} simulations are run entirely from \texttt{C++} executables.
+
+\texttt{xolotl} automatically uses the MD5 algorithm to hash the network and compile a new binary and \texttt{MEX} bridge file only if needed.
+\texttt{MATLAB} provides a high level programmatic and graphical interface for implementing, manipulating, and visualizing models without sacrificing the enhancements of the underlying \texttt{C++} code.
+
+\paragraph{Using the \texttt{cpplab} framework.}
+
+The \texttt{add} function will construct a \texttt{cpplab} object and affix it to as a field in the \texttt{xolotl} structure.
+All compartments, conductances, synapses, and controllers are \texttt{cpplab}.
+Compartments add to the \texttt{xolotl} object and conductances add to compartments.
+Specific properties can be specified using key-value pair arguments (e.g.
+\ref{fig:figurehh}A).
+
+\texttt{cpplab} comes with several features which simplify the handling of complexly-nested models.
+The \texttt{find} function acquires a cell array of all properties of the network which satisfy a search condition.
+For example, one can find all paths to maximal conductances within the "'HH'" compartment by:
+
+\begin{lstlisting}[style=Matlab-editor]
+x.find('HH*gbar');
+\end{lstlisting}
+
+To extract a vector of the maximal conductances:
+
+\begin{lstlisting}[style=Matlab-editor]
+gbars = x.get(x.find('HH*gbar'));
+\end{lstlisting}
+
+To set the maximal conductances all at once:
+
+\begin{lstlisting}[style=Matlab-editor]
+x.set('HH*gbar', gbars)
+\end{lstlisting}
+
+"MATLAB" can easily control the "cpplab" objects using the standard, flexible data structure notation popular in high-level scripting languages.
+
+\paragraph{Compartments and synapses.}
+
+A model neuron consists of one or more compartments, each representing a section of membrane with capacitance and surface area.
+Isopotential models require one compartment, whereas models with multiple neurons, units, or non-trivial morphology require multiple compartments.
+All specifiable properties of compartments are shown in Supplementary Table 1.
+
+\texttt{xolotl} provides some features for generating complex models.
+Synapses can be added with the \texttt{connect} function.
+At minimum synapses possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses.
+All specifiable properties of synapses are shown in Supplementary Table 2.
+To create axons or transport chains, the \texttt{slice} function splits a compartment into $n$ discrete segments and adds these compartments to the network connected by electrical synapses.
+
+\paragraph{Conductances and controllers.}
+
+All conductances contain fields for maximal conductance and reversal potential.
+Conductances with activation and inactivation variables include them as $m$ and $h$ respectively.
+Gating functions and their respective time constants are contained within the conductance header file.
+\texttt{xolotl} comes packaged with conductances from several dozen papers (Supplementary Table 3).
+
+\paragraph{Creating custom \texttt{cpplab} objects.}
+
+\texttt{xolotl} contains template header files for producing custom conductances. The template contains instructions on how to design novel conductances with arbitrary specifications.
+
+\paragraph{Simulation.}
+
+Models are simulated in \texttt{xolotl} with the \texttt{integrate} function which outputs as time series the membrane potentials, intracellular calcium concentrations, controller states, intrinsic currents, and synaptic currents.
+The \texttt{integrate} function also accepts an argument which specifies injected current or clamped voltage.
+
+"xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments \autocite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001, butcherNumericalDifferentialEquation2016}.
+These defaults provide a mix of speed, accuracy, and stability, and are built into the "cpplab" header files.
+Custom "cpplab" header files can be customized with any iterative integration method.
+The simulation time-resolution can be specified to target arbitrary precision, and an output time step can be selected to support automatic down-sampling for memory considerations.
+
+Simulations can be run in 'closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or 'open-loop' mode which begins simulation with the current network state.
+
+\paragraph{Using the graphical interface to manipulate parameters.}
+
+\texttt{xolotl} comes packaged with a graphical user interface for visualizing parameter changes in real-time.
+The \texttt{manipulate} function opens the GUI, which displays a figure plotting the membrane potential and intracellular calcium concentration of all compartments as time series, and a dialog box with customizable sliders for all parameters of the model, much like the "Manipulate" function in Wolfram "Mathematica".
+Moving the sliders integrates the model in `open-loop` mode with the new parameters.
+The parameters available in the sliders can be customized by passing a cell array to manipulate.
+For example, to only see sliders for maximal conductances of the "HH" compartment, call "x.manipulate(x.find('HH*gbar'))".
+Closing the GUI saves the network state of the model to the "xolotl" object.
+This is particularly helpful for rapid prototyping of models.
+
+\paragraph{Optimizing parameters.}
+"xolotl" can use the Global Optimization and Parallel Computing toolboxes for "MATLAB" to optimize network parameters.
+
+\subsection{Limitations}
+\label{limitations}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		USAGE EXAMPLES
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Usage Examples}
+\label{usage}
 
@@
 %
+%		BENCHMARKS
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Benchmarks}
+\label{benchmarks}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		DISCUSSION
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Discussion}
+\label{discussion}
+
+\subsection{Reproducibility}
+
+\subsection{Circumventing Language Tradeoffs}
+
+\subsection{Applications of \texttt{cpplab}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
 %		TECHNICAL DETAILS
@@
 \section{Technical Details}
-
-\subsection{Modeling}
-
-Models are specified by adding compartments and synapses to the \texttt{xolotl} object.
-Conductances are added to compartments and controllers can be added to conductances.
-This modular structure recapitulates the biophysics of the Hodgkin-Huxley formalism and obviates the need to explicitly write out equations, which in \texttt{xolotl} are contained within the conductance header files.
-
-\texttt{xolotl} relies on \texttt{cpplab} constructions, which allow the user to exploit the efficiency of low-level \texttt{C++} code.
-\texttt{MATLAB} treats \texttt{cpplab} objects as fully-typed variables allowing for symbolic manipulation using only the high-level programming language and graphical interfaces.
-\texttt{xolotl} is fast because all time-intensive code is written in \texttt{C++}.
-While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
-
-\texttt{C++} provides speed improvements beyond the benefits of translating \texttt{MATLAB} features into low-level code.
-For this reason, \texttt{cpplab} has been designed to provide an interface for constructing, transpiling, and compiling \texttt{C++} code to be called from within \texttt{MATLAB}.
-\texttt{xolotl} simulations are run entirely from \texttt{C++} executables.
-
-\texttt{xolotl} automatically uses the MD5 algorithm to hash the network and compile a new binary and \texttt{MEX} bridge file only if needed.
-\texttt{MATLAB} provides a high level programmatic and graphical interface for implementing, manipulating, and visualizing models without sacrificing the enhancements of the underlying \texttt{C++} code.
-
-\subsubsection{Using the \texttt{cpplab} Framework}
-
-The \texttt{add} function will construct a \texttt{cpplab} object and affix it to as a field in the \texttt{xolotl} structure.
-All compartments, conductances, synapses, and controllers are \texttt{cpplab}.
-Compartments add to the \texttt{xolotl} object and conductances add to compartments.
-Specific properties can be specified using key-value pair arguments (e.g.
-\ref{fig:figurehh}A).
-
-\texttt{cpplab} comes with several features which simplify the handling of complexly-nested models.
-The \texttt{find} function acquires a cell array of all properties of the network which satisfy a search condition.
-For example, one can find all paths to maximal conductances within the "'HH'" compartment by:
-
-\begin{lstlisting}[style=Matlab-editor]
-	x.find('HH*gbar');
-\end{lstlisting}
-
-To extract a vector of the maximal conductances:
-
-\begin{lstlisting}[style=Matlab-editor]
-	gbars = x.get(x.find('HH*gbar'));
-\end{lstlisting}
-
-To set the maximal conductances all at once:
-
-\begin{lstlisting}[style=Matlab-editor]
-	x.set(x.find('HH*gbar'), gbars)
-\end{lstlisting}
-
-"MATLAB" can easily control the "cpplab" objects using the standard, flexible data structure notation popular in high-level scripting languages.
-
-\subsubsection{Compartments and Synapses}
-
-A model neuron consists of one or more compartments, each representing a section of membrane with capacitance and surface area.
-Isopotential models require one compartment, whereas models with multiple neurons, units, or non-trivial morphology require multiple compartments.
-All specifiable properties of compartments are shown in Supplementary Table 1.
-
-\texttt{xolotl} provides some features for generating complex models.
-Synapses can be added with the \texttt{connect} function.
-At minimum synapses possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses.
-All specifiable properties of synapses are shown in Supplementary Table 2.
-To create axons or transport chains, the \texttt{slice} function splits a compartment into $n$ discrete segments and adds these compartments to the network connected by electrical synapses.
-
-\subsubsection{Conductances and Controllers}
-
-All conductances contain fields for maximal conductance and reversal potential.
-Conductances with activation and inactivation variables include them as $m$ and $h$ respectively.
-Gating functions and their respective time constants are contained within the conductance header file.
-\texttt{xolotl} comes packaged with conductances from several dozen papers (Supplementary Table 3).
-
-\subsubsection{Creating Custom \texttt{cpplab} Objects}
-
-\texttt{xolotl} contains template header files for producing custom conductances. The template contains instructions on how to design novel conductances with arbitrary specifications.
-
-\subsection{Simulation}
-
-Models are simulated in \texttt{xolotl} with the \texttt{integrate} function which outputs as time series the membrane potentials, intracellular calcium concentrations, controller states, intrinsic currents, and synaptic currents.
-The \texttt{integrate} function also accepts an argument which specifies injected current or clamped voltage.
-
-"xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments \autocite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001, butcherNumericalDifferentialEquation2016}.
-These defaults provide a mix of speed, accuracy, and stability, and are built into the "cpplab" header files.
-Custom "cpplab" header files can be customized with any iterative integration method.
-The simulation time-resolution can be specified to target arbitrary precision, and an output time step can be selected to support automatic down-sampling for memory considerations.
-
-Simulations can be run in 'closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or 'open-loop' mode which begins simulation with the current network state.
-
-\subsection{Using the graphical interface to manipulate parameters}
-
-\texttt{xolotl} comes packaged with a graphical user interface for visualizing parameter changes in real-time.
-The \texttt{manipulate} function opens the GUI, which displays a figure plotting the membrane potential and intracellular calcium concentration of all compartments as time series, and a dialog box with customizable sliders for all parameters of the model, much like the "Manipulate" function in Wolfram "Mathematica".
-Moving the sliders integrates the model in `open-loop` mode with the new parameters.
-The parameters available in the sliders can be customized by passing a cell array to manipulate.
-For example, to only see sliders for maximal conductances of the "HH" compartment, call "x.manipulate(x.find('HH*gbar'))".
-Closing the GUI saves the network state of the model to the "xolotl" object.
-This is particularly helpful for rapid prototyping of models.
 

### 2018-05-30 neuron/benchmark1.py

@@ -25,7 +25,7 @@ for ii in range(1,len(t_end)):
     tic = time.perf_counter() # s
     h.run()
     toc = time.perf_counter() - tic # s
-    sim_time[ii] = toc/1000; # ms
+    sim_time[ii] = toc*1000; # ms
     speed_factor[ii] = t_end[ii] / sim_time[ii] # unitless
 
 # save the results

### 2018-05-30 figures/figure_benchmark.m
minor fixes

@@ -3,15 +3,15 @@
 % create figure
 fig = figure('outerposition',[0 0 1200 600],'PaperUnits','points','PaperSize',[1200 600]); hold on;
 
-% speed versus time-step & accuracy vs. time-step
+% speed versus time-step
 ax(1) = subplot(1,3,1); hold on
 ax(1).Tag = 'Q vs t_end';
-% % accuracy vs. time-step
-% ax(2) = subplot(1,3,2); hold on
-% ax(2).Tag = 'Q vs. dt & r2 vs. dt';
-% % speed versus time span
-% ax(3) = subplot(1,3,3); hold on
-% ax(3).Tag = 'Q vs. t_end';
+% speed versus network size
+ax(2) = subplot(1,3,2); hold on
+ax(2).Tag = 'Q vs. nComps';
+% accuracy vs. dt
+ax(3) = subplot(1,3,3); hold on
+ax(3).Tag = 'Accuracy';
 
 % set up xolotl object
 x = xolotl;
@@ -43,7 +43,7 @@ equations = { ...
 
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Benchmark Test #1
+%% Benchmark Test #2
 % simulate a hodgkin-huxley model neuron over a series of simulation times
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@ -87,25 +87,25 @@ end
 % recover benchmark for NEURON
 NEURON_data = csvread('~/code/simulation-environment-paper/neuron/neuron_benchmark1.csv');
 
-% plot benchmark 1
-Qfactor(:,3) = vectorise(BRIAN_data);
+% plot benchmark 2
+% Qfactor(:,3) = vectorise(BRIAN_data);
 Qfactor(:,4) = vectorise(NEURON_data);
 
-plot(ax(1), t_end, Qfactor, '-o')
-xlabel(ax(1), 'simulation time (ms)')
-set(ax(1), 'XScale','log','YScale','log', 'XLim', [0 1.01e7], 'XTick', [1e1 1e4 1e7])
-ylabel(ax(1), 'speed factor')
[…]

### 2018-05-30 PaperLatex/paper.tex
Working on benchmarks, minor edits to paper

@@
 \paragraph{Optimizing parameters.}
-"xolotl" can use the Global Optimization and Parallel Computing toolboxes for "MATLAB" to optimize network parameters.
+"xolotl" can use the Global Optimization toolbox for "MATLAB" to optimize any accessible "xolotl" parameters.
+The toolbox is algorithm-agnostic and accepts any function in "MATLAB" with a scalar first output as the objective function.
+Simulations run on multi-core processors or high-performance computing clusters using the Parallel Computing toolbox.
 
@@
 \label{limitations}
+
+The focus on ease-of-use and speed means some features were elided in the streamlining process.
+
+\paragraph{Reliance on compiled \texttt{C++} code.} While "MATLAB" comes with robust features for compiling "C" and "C++" code, "xolotl" cannot run without "C++" compilation.
+For users, this necessitates the additional step of setting up the "mex" compiler which can be problematical, especially for nonstandard (e.g.
+Arch-based Linux).
+Secondly, compilation adds a small amount to total processing time.
+Longer simulations ($>1000$ time-steps) minimizes this effect.
+Adding new conductances also requires writing some "C++" code.
+For model conductances in the Hodgkin-Huxley formalism \autocite{hodgkinMeasurementCurrentvoltageRelations1952, dayanTheoreticalNeuroscience2001}, adjustments consist of changing default values in a template "C++" header file.
+Implementing a new integration scheme requires much more in-depth usage of "C++".
+
+\paragraph{Limited to conductance-based models.} "xolotl" has been developed specifically for conductance-based models. It does not currently support current-based models.
+
+\paragraph{Limited numerical integration strategies.} While the exponential Euler method performs well in neuronal models \autocite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}, it may be desirable to use other methods under certain conditions.
+"xolotl" does not currently support other integration schemes for its build-in conductances, nor does the software support error-sensitive variable step-sizes.
+
+\paragraph{Inefficient tools for handling large networks.} While "xolotl" can integrate large networks ($>1000$ compartments), the tools used to index, label, hash, and search large networks become prohibitively slow.
+"xolotl" uses string-based comprehension for labeling compartments, which is suited to descriptively-named compartments, but computationally slow for searching and indexing.
 
@@
 
+To assess speed and accuracy, "xolotl", "DynaSim" \autocite{sherfeyDynaSimMATLABToolbox2018}, and "NEURON" \autocite{hinesNEURONSimulationEnvironment1997} were compared in simulations over varied time-resolution and simulation time (\ref{fig:figurebenchmark}).
+A single-compartment Hodgkin-Huxley-like model was generated using conductance dynamics from \cite{liuModelNeuronActivitydependent1998} in "xolotl" and "DynaSim".
+To assess speed and accuracy over time-resolution, these models were simulated over varied time-steps for a total real-time of 5 s.
+The speed factor was defined as the ratio between the real-time (5 s) and the runtime of the simulation (simulation-time).
+Therefore, the speed factor represents how many times faster the simulation is than a real-time observation.
+Accuracy was computed by establishing a canonical voltage trace at $dt = 0.001$ ms and performing $r^2$ correlations between voltage traces with greater time-steps and the canonical trace.
+
+"xolotl" uses the exponential euler method for integrating membrane potential \autocite{dayanTheoreticalNeuroscience2001}.
+"DynaSim" was implemented with a 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance in the documentation.
+
+"xolotl" and "DynaSim" performed with comparable accuracy at high time-resolution. At low time-resolution, "xolotl" significantly outperforms "DynaSim" in both accuracy and speed.
+
+To test whether transient overhead effects had a significant effect on performance, "xolotl" and "DynaSim" were tested with time-step $dt = 0.1$ ms for varied lengths of time.
+"xolotl" and "DynaSim" both performed best during longer simulations, approaching maximal performance at $>10^5$ time steps.
+"xolotl" is about 20 times faster for short simulation times and 3.5 times faster for arbitrarily large ones.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+"xolotl" fosters reproducibility in science. While the availability of hosting sites with version control such as GitHub and GitLab
+
 \subsection{Circumventing Language Tradeoffs}
@@
 %
-%		TECHNICAL DETAILS
+%		PUBLICATION DETAILS
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Technical Details}
-
-\subsection{Benchmarking}
-
-To assess speed and accuracy, "xolotl", "DynaSim" \autocite{sherfeyDynaSimMATLABToolbox2018}, and "NEURON" \autocite{hinesNEURONSimulationEnvironment1997} were compared in simulations over varied time-resolution and simulation time (\ref{fig:figurebenchmark}).
-A single-compartment Hodgkin-Huxley-like model was generated using conductance dynamics from \cite{liuModelNeuronActivitydependent1998} in "xolotl" and "DynaSim".
-To assess speed and accuracy over time-resolution, these models were simulated over varied time-steps for a total real-time of 5 s.
-The speed factor was defined as the ratio between the real-time (5 s) and the runtime of the simulation (simulation-time).
-Therefore, the speed factor represents how many times faster the simulation is than a real-time observation.
-Accuracy was computed by establishing a canonical voltage trace at $dt = 0.001$ ms and performing $r^2$ correlations between voltage traces with greater time-steps and the canonical trace.
-
-"xolotl" uses the exponential euler method for integrating membrane potential \autocite{dayanTheoreticalNeuroscience2001}.
-"DynaSim" was implemented with a 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance in the documentation.
-
-"xolotl" and "DynaSim" performed with comparable accuracy at high time-resolution. At low time-resolution, "xolotl" significantly outperforms "DynaSim" in both accuracy and speed.
-
-To test whether transient overhead effects had a significant effect on performance, "xolotl" and "DynaSim" were tested with time-step $dt = 0.1$ ms for varied lengths of time.
-"xolotl" and "DynaSim" both performed best during longer simulations, approaching maximal performance at $>10^5$ time steps.
-"xolotl" is about 20 times faster for short simulation times and 3.5 times faster for arbitrarily large ones.
 

### 2018-05-30 figures/figure_benchmark.m

@@ -105,7 +105,7 @@ ylabel(ax(2), 'speed factor')
 % set up general simulation parameters
 t_end     = 5e3; % ms
 dt        = 0.1; % ms
-nComps    = [1, 2, 4, 8, 16, 32, 64, 128 250 500 1000 2000 4000 10000];
+nComps    = [1, 2, 4, 8, 16, 32, 64, 128 250 500 1000];% 2000 4000 10000];
 Qfactor_nComps = zeros(length(nComps),4);
 
 % test xolotl
@@ -132,6 +132,7 @@ for ii = 1:length(nComps)
   end
   x.skip_hash = false;
   x.md5hash
+  x.transpile; x.compile;
   Iext = 0.2 * ones(nComps(ii), 1);
 
   % begin timing
@@ -142,13 +143,32 @@ for ii = 1:length(nComps)
   Qfactor_nComps(ii, 1) = t_end / 1e3 / t_sim;
 end
 
+% test DynaSim
+
+for ii = 1:length(nComps)
+  textbar(ii, length(nComps))
+  % set up the DynaSim 'specification'
+  clear S
+  S = struct; % holds the DynaSim population information
+  S.population.name       = 'test';
+  S.population.size       = nComps(ii);
+  S.population.equations  = equations;
+
+  % begin timing
+  tic;
+  dsSimulate(equations, 'solver', 'rk2', 'tspan', [0 5e3], 'dt', 0.1, 'compile_flag', 0);
+  t_sim = toc;
+  % compute the speed as real-time / simulation-time
+  Qfactor_nComps(ii, 2) = t_end / 1e3 / t_sim;
+end
+
 % plot benchmark 3
 % Qfactor(:,3) = vectorise(BRIAN_data);
 % Qfactor(:,4) = vectorise(NEURON_data);
 
 plot(ax(3), nComps, Qfactor_nComps, '-o')
 xlabel(ax(3), 'simulation time (ms)')
-set(ax(3), 'XScale','log','YScale','log', 'XLim', [0 32010], 'XTick', [1e1 1e2 1e3 1e4])
+set(ax(3), 'XScale','log','YScale','log', 'XLim', [0 1010], 'XTick', [1e1 1e2 1e3])
 ylabel(ax(3), 'speed factor')
 leg = legend(ax(3), {'xolotl', 'DynaSim', 'BRIAN 2', 'NEURON'}, 'Location', 'EastOutside');
 

### 2018-05-30 figures/figure_transpile.m

@@ -1,0 +1,60 @@
+% test transpile and compile speed in xolotl
+
+nComps      = round(logspace(1, 3, 11));
+init_time   = zeros(length(nComps), 1);
+trans_time  = zeros(length(nComps), 1);
+comp_time   = zeros(length(nComps), 1);
+sim_time    = zeros(length(nComps), 1);
+
+fig = figure; hold on;
+
+% perform benchmarking
+for ii = 1:length(nComps)
+  textbar(ii, length(nComps))
+
+  % set up the xolotl object
+  clear x
+
+  % benchmark instantiation
+  tic;
+  x = xolotl;
+  x.cleanup
+  x.skip_hash = true;
+  for qq = 1:nComps(ii)
+    compName = ['HH' mat2str(qq)];
+    x.add(compName, 'compartment', 'Cm', 10, 'A', 0.01);
+    x.(compName).add('liu/NaV', 'gbar', 1000, 'E', 50);
+    x.(compName).add('liu/Kd', 'gbar', 300, 'E', -80);
+    x.(compName).add('Leak', 'gbar', 1, 'E', -50);
+  end
+  x.skip_hash = false;
+  x.md5hash
+  init_time(ii) = toc;
+
+  % benchmark transpiling
+  tic;
+  x.transpile;
+  trans_time(ii) = toc;
+
+  % benchmark compiling
+  tic;
+  x.compile;
+  comp_time(ii) = toc;
+
+  % benchmark simulation
+  x.sim_dt = 0.1;
+  x.dt = 0.1;
+  x.t_end = 1000;
+  Iext = 0.2 * ones(nComps(ii), 1);
+  tic;
+  x.integrate;
+  sim_time(ii) = toc;
+
+  % plot now
+  plot(nComps(ii), init_time(ii), 'ko')
+  plot(nComps(ii), trans_time(ii), 'ro')
+  plot(nComps(ii), comp_time(ii), 'bo')
+  plot(nComps(ii), sim_time(ii), 'go')
+
+  drawnow
+end

### 2018-05-30 neuron/benchmark2.py

@@ -1,4 +1,4 @@
-## Benchmark Test #1
+## Benchmark Test #2
 # single-compartment, Hodgkin-Huxley
 # test speed with increasing simulation time
 

### 2018-05-31 PaperLatex/paper.tex
General improvements to the paper

@@
 \usepackage{amsmath} 
-\usepackage{mathptmx}      % use Times fonts if available on your TeX system
+%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
 %
@@
 \lstMakeShortInline[style=Matlab-editor]"
+
+% other nice features
+\usepackage{lineno}
+\linenumbers
+\usepackage{setspace}
+\onehalfspacing
 
@@
 
-%\texttt{xolotl} provides parameter optimization capabilities through the algorithm-agnostic \texttt{procrustes} toolbox.
-Any network parameters accessible through the \texttt{xolotl} structure can be optimized using arbitrary algorithms and objective functions on multi-core computers and high-performance computing clusters.
-
 The software has been implemented in \texttt{MATLAB} due to its ease-of-use and popularity among neuroscientists.
@@
 "xolotl" is designed to be easy-to-use without sacrificing speed.
+
 The software has been designed in "MATLAB" due to its popularity among neuroscientists for pedagogy and research.
@@
 
+\paragraph{Modular structure.}
+
 Models are specified by adding compartments and synapses to the \texttt{xolotl} object.
@@
 This modular structure recapitulates the biophysics of the Hodgkin-Huxley formalism and obviates the need to explicitly write out equations, which in \texttt{xolotl} are contained within the conductance header files.
+
+\paragraph{Interface between \texttt{C++} and \texttt{MATLAB}.}
 
@@
 While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
-
 \texttt{C++} provides speed improvements beyond the benefits of translating \texttt{MATLAB} features into low-level code.
@@
 
+\paragraph{Automatic and efficient compiling.}
 \texttt{xolotl} automatically uses the MD5 algorithm to hash the network and compile a new binary and \texttt{MEX} bridge file only if needed.
@@
 
-\paragraph{Using the \texttt{cpplab} framework.}
+\subsection{Syntax}
 
+"MATLAB" can easily control the "cpplab" objects using the standard, flexible data structure notation popular in high-level scripting languages.
+
+\paragraph{Adding features.}
 The \texttt{add} function will construct a \texttt{cpplab} object and affix it to as a field in the \texttt{xolotl} structure.
@@
 Specific properties can be specified using key-value pair arguments (e.g.
-\ref{fig:figurehh}A).
+Figure \ref{fig:figurehh}A).
 
+\paragraph{Finding features.}
 \texttt{cpplab} comes with several features which simplify the handling of complexly-nested models.
@@
 \begin{lstlisting}[style=Matlab-editor]
-gbars = x.get(x.find('HH*gbar'));
+	gbars = x.get('HH*gbar');
 \end{lstlisting}
@@
 
-"MATLAB" can easily control the "cpplab" objects using the standard, flexible data structure notation popular in high-level scripting languages.
-
-\paragraph{Compartments and synapses.}
-
+\paragraph{Compartments.}
 A model neuron consists of one or more compartments, each representing a section of membrane with capacitance and surface area.
@@
 
+\paragraph{Synapses.}
 \texttt{xolotl} provides some features for generating complex models.
@@
 
+\paragraph{Numerical integration.}
 "xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments \autocite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001, butcherNumericalDifferentialEquation2016}.
@@
 
+\paragraph{'Closed-loop' vs. 'open-loop.'}
 Simulations can be run in 'closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or 'open-loop' mode which begins simulation with the current network state.
@@
 
-The seminal Hodgkin-Huxley model of action potentials in the squid giant axon \autocite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952} contains a fast inactivating sodium conductance ("NaV"), a non-inactivating delayed rectifier ("Kd"), and a passive leak current (\ref{fig:figurehh}A).
-A compartment, \texttt{HH}, with membrane capacitance ("Cm") and surface area ("A") can be specified by \ref{fig:figurehh}B.
+The seminal Hodgkin-Huxley model of action potentials in the squid giant axon \autocite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952} contains a fast inactivating sodium conductance ("NaV"), a non-inactivating delayed rectifier ("Kd"), and a passive leak current (Figure \ref{fig:figurehh}A).
+A compartment, \texttt{HH}, with membrane capacitance ("Cm") and surface area ("A") can be specified by Figure \ref{fig:figurehh}B.
 Network properties can be set during construction or afterwards using dot-notation in \texttt{MATLAB} (e.g.
 \texttt{x.HH.Cm}).
-\ref{fig:figurehh}C shows the \texttt{MATLAB} command prompt after invoking the \texttt{xolotl} object \texttt{x}, displaying the hierarchical structure inherent in conductance-based treatments of neurodynamics.
+Figure \ref{fig:figurehh}C shows the \texttt{MATLAB} command prompt after invoking the \texttt{xolotl} object \texttt{x}, displaying the hierarchical structure inherent in conductance-based treatments of neurodynamics.
 
@@
 In the absence of applied positive current, the model is quiescent.
-When 0.2 nA is injected, the model tonically spikes (\ref{fig:figurehh}D).
+When 0.2 nA is injected, the model tonically spikes (Figure \ref{fig:figurehh}D).
 The \texttt{integrate} function takes the applied current as an argument (e.g.
@@
 Conversely, if the membrane potential is decreasing, the strongest outward current colors the trace instead.
-\ref{fig:figurehh}F-I display the results of the \texttt{show} function.
+Figure \ref{fig:figurehh}F-I display the results of the \texttt{show} function.
 Activation and inactivation steady-states and the voltage-dependent time constants of these gating variables describe the conductance dynamics in absence of other channel types.
@@
 \texttt{xolotl} can recapitulate the results of voltage clamp experiments \autocite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}.
-Figure \ref{fig:figureclamp} displays steps in the procedure to clamp the membrane potential of a cell with delayed rectifier potassium conductance.
+Figure Figure \ref{fig:figureclamp} displays steps in the procedure to clamp the membrane potential of a cell with delayed rectifier potassium conductance.
 During an \textit{in-vitro} experiment, confounding currents would be pharmacologically-blocked and two-electrode voltage clamp used to record tail currents at fixed membrane potential \autocite{connorInwardDelayedOutward1971, connorVoltageClampStudies1971}.
@@
 
-Currents under voltage clamp approach the steady-state holding current (\ref{fig:figureclamp}D-E).
-The current-voltage relation is the steady-state current over the clamped voltage, and the effective conductance is the derivative of that relation (\ref{fig:figureclamp}F-G).
+Currents under voltage clamp approach the steady-state holding current (Figure \ref{fig:figureclamp}D-E).
+The current-voltage relation is the steady-state current over the clamped voltage, and the effective conductance is the derivative of that relation (Figure \ref{fig:figureclamp}F-G).
 Since the effective conductance is the product of the maximal conductance and the gating variables \autocite{dayanTheoreticalNeuroscience2001, turrigianoSelectiveRegulationCurrent1995} and the tail current is monotonically-increasing with time under voltage clamp, the current can be represented as non-inactivating.
-Fitting a sigmoid to various powers yields a model for the current dynamics (\ref{fig:figureclamp}H-I).
+Fitting a sigmoid to various powers yields a model for the current dynamics (Figure \ref{fig:figureclamp}H-I).
 These figures describe graphically the theoretical underpinnings of current analysis through voltage clamp and can serve as an effective pedagogical tool for computational and quantitative neuroscience.
@@
 Presynaptic and postsynaptic labels indicate the connectivity of the synapse.
-Figure \ref{fig:figurenetwork} implements a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans.
-The pyloric model contains three compartments and seven synapses (\ref{fig:figurenetwork}A).
-This structure is reciprocated in the hierarchy of the \texttt{xolotl} object, where conductances are contained within compartments (\ref{fig:figurenetwork}B).
+Figure Figure \ref{fig:figurenetwork} implements a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans.
+The pyloric model contains three compartments and seven synapses (Figure \ref{fig:figurenetwork}A).
+This structure is reciprocated in the hierarchy of the \texttt{xolotl} object, where conductances are contained within compartments (Figure \ref{fig:figurenetwork}B).
 
@@
 
-Synapses are upper-level properties of the network which point between two compartments (\ref{fig:figurenetwork}C).
+Synapses are upper-level properties of the network which point between two compartments (Figure \ref{fig:figurenetwork}C).
 This exploits vectorized operations in \texttt{MATLAB} and does not require each synapse to possess a unique name.
@@
 In a demonstration adapted from \cite{olearyCorrelationsIonChannel2013}, integral control changes maximal conductances to bring a neuron from quiescence into a bursting regime.
-Calcium sensors supervene on maximal conductance density (\ref{fig:figureintegralcontrol}) to change neuronal activity.
-Each conductance in the "xolotl" structure contains a calcium-sensitive controller (\ref{fig:figureintegralcontrol}B-C).
-Maximal conductances increase from random initial conditions to a set which elicits the desired network output by minimizing the error signal (\ref{fig:figureintegralcontrol}D-F).
+Calcium sensors supervene on maximal conductance density (Figure \ref{fig:figureintegralcontrol}) to change neuronal activity.
+Each conductance in the "xolotl" structure contains a calcium-sensitive controller (Figure \ref{fig:figureintegralcontrol}B-C).
+Maximal conductances increase from random initial conditions to a set which elicits the desired network output by minimizing the error signal (Figure \ref{fig:figureintegralcontrol}D-F).
 
@@
 
-To assess speed and accuracy, "xolotl", "DynaSim" \autocite{sherfeyDynaSimMATLABToolbox2018}, and "NEURON" \autocite{hinesNEURONSimulationEnvironment1997} were compared in simulations over varied time-resolution and simulation time (\ref{fig:figurebenchmark}).
-A single-compartment Hodgkin-Huxley-like model was generated using conductance dynamics from \cite{liuModelNeuronActivitydependent1998} in "xolotl" and "DynaSim".
-To assess speed and accuracy over time-resolution, these models were simulated over varied time-steps for a total real-time of 5 s.
-The speed factor was defined as the ratio between the real-time (5 s) and the runtime of the simulation (simulation-time).
+To assess speed and accuracy, "xolotl", "DynaSim" \autocite{sherfeyDynaSimMATLABToolbox2018}, and "NEURON" \autocite{hinesNEURONSimulationEnvironment1997} were compared in simulations over varied simulation time and number of compartments (Figure \ref{fig:figurebenchmark}).
+
+Single-compartment Hodgkin-Huxley-like models were generated using conductance dynamics from \cite{liuModelNeuronActivitydependent1998} in the simulation environments.
+Models were simulated with a time-resolution of $0.1$ ms over increasing simulation time (Figure \ref{fig:figurebenchmark}B).
+The speed factor was defined as the ratio between time represented in the simulation and actual runtime (simulation-time).
 Therefore, the speed factor represents how many times faster the simulation is than a real-time observation.
-Accuracy was computed by establishing a canonical voltage trace at $dt = 0.001$ ms and performing $r^2$ correlations between voltage traces with greater time-steps and the canonical trace.
 
@@
 
-"xolotl" fosters reproducibility in science. While the availability of hosting sites with version control such as GitHub and GitLab
+"xolotl" fosters reproducibility in science.
+While the availability of hosting sites with version control (viz.
+GitHub (\url{https://github.com}), GitLab (\url{https://gitlab.com/}), and Open Science Framework (\url{https://osf.io/})) and the push for reproducibility in computational science \autocite{eklundClusterFailureWhy2016, stoddenEnhancingReproducibilityComputational2016, bakerWhyScientistsMust2016} has resulted in the availability of source code, much of this code base is bespoke and difficult to implement \autocite{sedanoCodeReadabilityTesting2016, xuMeasurementSourceCode2017}.
+
+To this end, "xolotl" provides an environment with readability and reproducibility in mind.
+Each network is hashed to provide a unique alphanumeric identifier.
+Conductance header files are easily viewed in the "xolotl" source files; conductances in "MATLAB" contain links to the full path of the generating file.
 
@@
 
+Executing "C/C++" code in higher-level languages such as "MATLAB" or "Python" often provides speed improvements for iterative code in algorithms.
+
+"C" is statically-typed, with procedural syntax that provides low-level access to memory \autocite{kernighanProgrammingLanguage1978}, providing significant advantages for time-intensive computations.
+Unfortunately, automatic code-generation is limited by the supervening language.
+"MATLAB", for instance, cannot use pointers or pass by reference, which limits the efficiency of "C" code automatically generated from "MATLAB".
+Conversely, custom "C/C++" code provides significant increases in performance and memory conservation, but lacks the ease-of-use and flexibility of scripting languages.
+
+"xolotl" handles this problem through symbolic manipulation of "C++" objects in "MATLAB".
+Built from the ground up in "C++", "xolotl" maintains all the advantages of custom compiled code, but can run in "MATLAB" without the user having to touch the "C++" code.
+"xolotl" represents compartment, conductance, synapse, and controllers as "cpplab" objects, which map to underlying "C++" header files.
+In this way, properties of the "xolotl" network can be examined and changed using object-oriented paradigms.
+The object specifies the "integrate" function, not the other way around.
+
 \subsection{Applications of \texttt{cpplab}}
+
+NEED TO WRITE THIS
 
@@
 \section*{Acknowledgments}
-The authors would like to thank Mara CP Rue and Hillary Rodgers for beta-testing the \texttt{xolotl} software.
+The authors would like to thank Mara CP Rue and Hillary Rodgers for beta-testing the \texttt{xolotl} software. Janis Li helped to prepare some conductance header files.
 

### 2018-05-31 figures/figure_clamp.m

@@ -142,10 +142,10 @@ end
 
 for ii = 1:length(all_n)
 	plot(ax(8), Vsteps, minf.^all_n(ii), 'Color', c(ii, :))
-	hplot(ii) = plot(ax(8), NaN, NaN, 'o', 'MarkerFaceColor', c(ii, :), 'MarkerEdgeColor', c(ii, :), 'MarkerSize', 12);
+	hplot(ii) = plot(ax(8), NaN, NaN, 'o', 'MarkerFaceColor', c(ii, :), 'MarkerEdgeColor', c(ii, :), 'MarkerSize', 8);
 end
 plot(ax(8), Vsteps, conductance/conductance(end), 'ok')
-hplot(end+1) = plot(ax(8), NaN, NaN, 'o', 'MarkerEdgeColor', [0 0 0], 'MarkerSize', 12);
+hplot(end+1) = plot(ax(8), NaN, NaN, 'o', 'MarkerEdgeColor', [0 0 0], 'MarkerSize', 8);
 xlabel(ax(8), 'clamped voltage (mV)')
 ylabel(ax(8), 'm_\infty')
 set(ax(8), 'XLim', [min(Vsteps) max(Vsteps)], 'XTick', [-80 -40 0 40])

### 2018-06-01 PaperLatex/paper.tex
General clarifications

@@
 
-	\texttt{xolotl} is a free and open-source neuronal simulator written in \texttt{C++} with \texttt{MATLAB} wrappers.
+	"xolotl" is a free and open-source neuronal simulator written in \texttt{C++} with \texttt{MATLAB} wrappers.
 Biophysically-detailed models of networks can be designed efficiently using an intuitive language tightly coupled to the object-based architecture of the underlying \texttt{C++} code.
@@
 The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate rapid construction and assessment of model networks.
-\texttt{xolotl} is freely available at \url{https://github.com/marderlab/xolotl}.
+"xolotl" is freely available at \url{https://github.com/marderlab/xolotl}.
 This tool provides straightforward implementation and fast simulation of neuronal models while permitting full control over every aspect of the network and integration.
@@
 
-\texttt{xolotl} (\url{https://github.com/marderlab/xolotl}) is a fast single-compartment and multi-compartment simulator in \texttt{C++} with \texttt{MATLAB} wrappers.
-Written with an emphasis on ease-of-use, \texttt{xolotl} can simulate single-compartment conductance-based models, networks of these, and detailed multi-compartment models.
-\texttt{xolotl} exploits a novel automatic type system, \texttt{cpplab}, which binds \texttt{MATLAB} code to \texttt{C++} header files, creating objects and classes \textit{ad libitum} in \texttt{MATLAB} which reflect the underlying object-oriented code.
-\texttt{xolotl} implements \texttt{cpplab} to represent the nested structure of conductance-based models, and exploits the computational efficiency of the low-level programming language to quickly integrate models.
+"xolotl" (\url{https://github.com/marderlab/xolotl}) is a fast single-compartment and multi-compartment simulator in \texttt{C++} with \texttt{MATLAB} wrappers.
+Written with an emphasis on ease-of-use, "xolotl" can simulate single-compartment conductance-based models, networks of these, and detailed multi-compartment models.
+"xolotl" exploits a novel automatic type system, \texttt{cpplab}, which binds \texttt{MATLAB} code to \texttt{C++} header files, creating objects and classes \textit{ad libitum} in \texttt{MATLAB} which reflect the underlying object-oriented code.
+"xolotl" implements \texttt{cpplab} to represent the nested structure of conductance-based models, and exploits the computational efficiency of the low-level programming language to quickly integrate models.
 For this reason, models can be implemented entirely in \texttt{MATLAB} with a few lines of code.
@@
 Models are specified in \texttt{MATLAB} by a nested structure.
-The \texttt{xolotl} object contains compartments which themselves contain conductances.
-Synapses belong to the \texttt{xolotl} object and connect compartments together.
+The "xolotl" object contains compartments which themselves contain conductances.
+Synapses belong to the "xolotl" object and connect compartments together.
 The high-level specification supports arbitrarily large network and multi-compartment morphologies.
@@
 While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
-Minimal experience with \texttt{MATLAB} is required to use \texttt{xolotl}, and all equations and integration methods are provided transparently to the end user.
+Minimal experience with \texttt{MATLAB} is required to use "xolotl", and all equations and integration methods are provided transparently to the end user.
 No string parsing of equations is required \autocite{sherfeyDynaSimMATLABToolbox2018, stimbergBrianSecondComing2013, stimbergEquationorientedSpecificationNeural2014}.
 
-\texttt{xolotl} comes packaged with visualization functions and a graphical user interface (GUI) for real-time manipulation of model parameters.
-Plotting of voltage, intracellular calcium, conductance gating functions, and time constants is provided by built-in \texttt{xolotl} methods.
+"xolotl" comes packaged with visualization functions and a graphical user interface (GUI) for real-time manipulation of model parameters.
+Plotting of voltage, intracellular calcium, conductance gating functions, and time constants is provided by built-in "xolotl" methods.
 The GUI permits real-time tuning of any network parameters using numerical sliders in a graphical interface which displays the resultant membrane potential and intracellular calcium traces.
@@
 
+\texttt{MATLAB} provides a high level programmatic and graphical interface for implementing, manipulating, and visualizing models without sacrificing the enhancements of the underlying \texttt{C++} code.
+
 \paragraph{Modular structure.}
 
-Models are specified by adding compartments and synapses to the \texttt{xolotl} object.
+Models are specified by adding compartments and synapses to the "xolotl" object.
 Conductances are added to compartments and controllers can be added to conductances.
-This modular structure recapitulates the biophysics of the Hodgkin-Huxley formalism and obviates the need to explicitly write out equations, which in \texttt{xolotl} are contained within the conductance header files.
+This modular structure recapitulates the biophysics of the Hodgkin-Huxley formalism and obviates the need to explicitly write out equations, which in "xolotl" are contained within the conductance header files.
 
@@
 
-\texttt{xolotl} relies on \texttt{cpplab} constructions, which allow the user to exploit the efficiency of low-level \texttt{C++} code.
+"xolotl" relies on \texttt{cpplab} constructions, which allow the user to exploit the efficiency of low-level \texttt{C++} code.
 \texttt{MATLAB} treats \texttt{cpplab} objects as fully-typed variables allowing for symbolic manipulation using only the high-level programming language and graphical interfaces.
-\texttt{xolotl} is fast because all time-intensive code is written in \texttt{C++}.
+"xolotl" is fast because all time-intensive code is written in \texttt{C++}.
 While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
@@
 For this reason, \texttt{cpplab} has been designed to provide an interface for constructing, transpiling, and compiling \texttt{C++} code to be called from within \texttt{MATLAB}.
-\texttt{xolotl} simulations are run entirely from \texttt{C++} executables.
+"xolotl" simulations are run entirely from \texttt{C++} executables.
 
 \paragraph{Automatic and efficient compiling.}
-\texttt{xolotl} automatically uses the MD5 algorithm to hash the network and compile a new binary and \texttt{MEX} bridge file only if needed.
-\texttt{MATLAB} provides a high level programmatic and graphical interface for implementing, manipulating, and visualizing models without sacrificing the enhancements of the underlying \texttt{C++} code.
+"xolotl" automatically handles transpiling and compiling "MATLAB" code into "C++".
+The MD5 algorithm hashes the network to compile a new binary and "MEX" bridge file only if needed and to confirm that the correct binary fetched during execution.
 
@@
 \paragraph{Adding features.}
-The \texttt{add} function will construct a \texttt{cpplab} object and affix it to as a field in the \texttt{xolotl} structure.
-All compartments, conductances, synapses, and controllers are \texttt{cpplab}.
-Compartments add to the \texttt{xolotl} object and conductances add to compartments.
+The "add" function creates a compartment, conductance, or controller and affixes it as a field in the "xolotl" network structure.
+This function generates a "MATLAB struct" that faithfully represents the underlying "C++" code.
+Compartments add to the "xolotl" object and conductances add to compartments.
 Specific properties can be specified using key-value pair arguments (e.g.
@@
 \texttt{cpplab} comes with several features which simplify the handling of complexly-nested models.
-The \texttt{find} function acquires a cell array of all properties of the network which satisfy a search condition.
+The "find" function acquires a cell array of all properties of the network which satisfy a search condition.
 For example, one can find all paths to maximal conductances within the "'HH'" compartment by:
@@
 \paragraph{Synapses.}
-\texttt{xolotl} provides some features for generating complex models.
+"xolotl" provides some features for generating complex models.
 Synapses can be added with the \texttt{connect} function.
@@
 Gating functions and their respective time constants are contained within the conductance header file.
-\texttt{xolotl} comes packaged with conductances from several dozen papers (Supplementary Table 3).
+"xolotl" comes packaged with conductances from several dozen papers (Supplementary Table 3).
 
@@
 
-\texttt{xolotl} contains template header files for producing custom conductances. The template contains instructions on how to design novel conductances with arbitrary specifications.
+"xolotl" contains template header files for producing custom conductances. The template contains instructions on how to design novel conductances with arbitrary specifications.
 
@@
 
-Models are simulated in \texttt{xolotl} with the \texttt{integrate} function which outputs as time series the membrane potentials, intracellular calcium concentrations, controller states, intrinsic currents, and synaptic currents.
+Models are simulated in "xolotl" with the \texttt{integrate} function which outputs as time series the membrane potentials, intracellular calcium concentrations, controller states, intrinsic currents, and synaptic currents.
 The \texttt{integrate} function also accepts an argument which specifies injected current or clamped voltage.
@@
 
-\paragraph{'Closed-loop' vs. 'open-loop.'}
-Simulations can be run in 'closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or 'open-loop' mode which begins simulation with the current network state.
+\paragraph{`Closed-loop' vs. `open-loop.'}
+Simulations can be run in `closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or `open-loop' mode which begins simulation with the current network state.
 
@@
 
-\texttt{xolotl} comes packaged with a graphical user interface for visualizing parameter changes in real-time.
+"xolotl" comes packaged with a graphical user interface for visualizing parameter changes in real-time.
 The \texttt{manipulate} function opens the GUI, which displays a figure plotting the membrane potential and intracellular calcium concentration of all compartments as time series, and a dialog box with customizable sliders for all parameters of the model, much like the "Manipulate" function in Wolfram "Mathematica".
@@
 The parameters available in the sliders can be customized by passing a cell array to manipulate.
-For example, to only see sliders for maximal conductances of the "HH" compartment, call "x.manipulate(x.find('HH*gbar'))".
+For example, to only see sliders for maximal conductances of the "HH" compartment, call "x.manipulate('HH*gbar')".
 Closing the GUI saves the network state of the model to the "xolotl" object.
@@
 
-\paragraph{Limited to conductance-based models.} "xolotl" has been developed specifically for conductance-based models. It does not currently support current-based models.
+\paragraph{Limited to conductance-based models.} "xolotl" has been developed specifically for conductance-based models. It does not currently support rate- or current-based models.
 
 \paragraph{Limited numerical integration strategies.} While the exponential Euler method performs well in neuronal models \autocite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}, it may be desirable to use other methods under certain conditions.
-"xolotl" does not currently support other integration schemes for its build-in conductances, nor does the software support error-sensitive variable step-sizes.
+"xolotl" does not currently support other integration schemes for its built-in conductances, nor does the software support error-sensitive variable step-sizes `out-of-the-box.'
 
-\paragraph{Inefficient tools for handling large networks.} While "xolotl" can integrate large networks ($>1000$ compartments), the tools used to index, label, hash, and search large networks become prohibitively slow.
-"xolotl" uses string-based comprehension for labeling compartments, which is suited to descriptively-named compartments, but computationally slow for searching and indexing.
+\paragraph{Inefficient tools for handling large networks.} While "xolotl" can integrate large networks ($>1000$ compartments), "xolotl" uses string-based comprehension for labeling compartments which is suited to descriptive naming, but prohibits vector operations over compartments.
 
@@
 
-Using \texttt{xolotl} in \texttt{MATLAB}, users create a \texttt{xolotl} object and populate it with "cpplab" objects which describe compartments, conductances, synapses, and controllers.
-The model is integrated with the \texttt{integrate} function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.
+In "MATLAB", users create a "xolotl" object and populate it with "cpplab"-generated objects which describe compartments, conductances, synapses, and controllers.
+The model is integrated with the "integrate" function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.
 
-\texttt{xolotl} comes packaged with a library of pre-existing conductance and synapse objects which greatly simplify the task of constructing model neurons.
+"xolotl" comes packaged with a library of pre-existing conductance and synapse objects which greatly simplify the task of constructing model neurons.
 These objects can be referenced by name and added directly to a compartment.
-Novel conductance dynamics can be easily written by modifying a template header file contained in the \texttt{xolotl} distribution, or designed entirely from scratch.
+Novel conductance dynamics can be easily written by modifying a template header file contained in the "xolotl" distribution, or designed entirely from scratch.
 
@@
 \texttt{x.HH.Cm}).
-Figure \ref{fig:figurehh}C shows the \texttt{MATLAB} command prompt after invoking the \texttt{xolotl} object \texttt{x}, displaying the hierarchical structure inherent in conductance-based treatments of neurodynamics.
+Figure \ref{fig:figurehh}C shows the \texttt{MATLAB} command prompt after invoking the "xolotl" object \texttt{x}, displaying the hierarchical structure inherent in conductance-based treatments of neurodynamics.
 
@@
 The \texttt{integrate} function takes the applied current as an argument (e.g.
-"x.integrate(Iapp)"), so that the \texttt{xolotl} object is agnostic to integration-specific perturbations.
+"x.integrate(Iapp)"), so that the "xolotl" object is agnostic to integration-specific perturbations.
 The \texttt{plot} function generates voltage and intracellular calcium traces, where the voltage trace is colored by the dominant current.
@@
 
-\texttt{xolotl} can recapitulate the results of voltage clamp experiments \autocite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}.
+"xolotl" can recapitulate the results of voltage clamp experiments \autocite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}.
 Figure Figure \ref{fig:figureclamp} displays steps in the procedure to clamp the membrane potential of a cell with delayed rectifier potassium conductance.
@@
 
-Network models in \texttt{xolotl} consist of compartment objects connected by synapses.
-Synapses are stored in a vector array as a field of the \texttt{xolotl} object in \texttt{MATLAB}.
+Network models in "xolotl" consist of compartment objects connected by synapses.
+Synapses are stored in a vector array as a field of the "xolotl" object in \texttt{MATLAB}.
 Presynaptic and postsynaptic labels indicate the connectivity of the synapse.
@@
 The pyloric model contains three compartments and seven synapses (Figure \ref{fig:figurenetwork}A).
-This structure is reciprocated in the hierarchy of the \texttt{xolotl} object, where conductances are contained within compartments (Figure \ref{fig:figurenetwork}B).
+This structure is reciprocated in the hierarchy of the "xolotl" object, where conductances are contained within compartments (Figure \ref{fig:figurenetwork}B).
 
@@
 
-\texttt{xolotl} can implement homeostatic tuning rules as integral control.
+"xolotl" can implement homeostatic tuning rules as integral control.
 The controller computes an error signal (typically a function of intracellular calcium concentration), and adjusts the conductance or synapse it controls accordingly \autocite{olearyCorrelationsIonChannel2013}.
@@
 \section*{Acknowledgments}
-The authors would like to thank Mara CP Rue and Hillary Rodgers for beta-testing the \texttt{xolotl} software. Janis Li helped to prepare some conductance header files.
+The authors would like to thank Mara CP Rue and Hillary Rodgers for beta-testing the "xolotl" software. Janis Li helped to prepare some conductance header files.
 
 \section*{Supplemental Data}
-Tables including all conductances packaged with \texttt{xolotl} should be put in the supplementary material.
+Tables including all conductances packaged with "xolotl" should be put in the supplementary material.
 
@@
 (F-G) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
-(H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables.
-Variables not plotted are unity for all voltage.}
+(H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables.}
 	\label{fig:figurehh}
@@
 (B) Each neuron is modeled as a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
-(C) \textit{xolotl} implements conductances as fields of compartments and synapses as connections between compartments.
+(C) \texttt{xolotl} implements conductances as fields of compartments and synapses as connections between compartments.
 (D-F) Simulated voltage trace of a model network for the three compartments.

### 2018-06-03 PaperLatex/paper.tex
Align formatting to Frontiers specifications
Rewrite benchmarks section

@@
 \usepackage{amsmath} 
-%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
+\usepackage{mathptmx}      % use Times fonts if available on your TeX system
 %
@@
 \linenumbers
-\usepackage{setspace}
-\onehalfspacing
 
@@
 \def\corrAddress{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
-\def\corrEmail{marder@brandeis.edu}
+\def\corrEmail{\texttt{marder@brandeis.edu}}
 
@@
 
-To assess speed and accuracy, "xolotl", "DynaSim" \autocite{sherfeyDynaSimMATLABToolbox2018}, and "NEURON" \autocite{hinesNEURONSimulationEnvironment1997} were compared in simulations over varied simulation time and number of compartments (Figure \ref{fig:figurebenchmark}).
+To assess speed and accuracy, "xolotl", "DynaSim" \autocite{sherfeyDynaSimMATLABToolbox2018}, and "NEURON" \autocite{hinesNEURONSimulationEnvironment1997} were compared in simulations over varied time resolution,  simulation time, and number of compartments (Figure \ref{fig:figurebenchmark}).
 
 Single-compartment Hodgkin-Huxley-like models were generated using conductance dynamics from \cite{liuModelNeuronActivitydependent1998} in the simulation environments.
-Models were simulated with a time-resolution of $0.1$ ms over increasing simulation time (Figure \ref{fig:figurebenchmark}B).
+"xolotl" uses the exponential Euler method for integrating membrane potential \autocite{dayanTheoreticalNeuroscience2001}.
+"DynaSim" was implemented with a 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance in the documentation.
+"NEURON" used an implicit Euler regime \autocite{hinesNEURONSimulationEnvironment1997}.
+
+To compare the integration methods, models were simulated for 5 s at varying time-resolution (Figure \ref{fig:figurebenchmark}A).
+The ratio between 'simulated' time and actual runtime was defined as the speed factor.
+Higher values indicate faster simulations.
+The coincidence factor determines the correlative overlap between two spike trains \autocite{jolivetBenchmarkTestQuantitative2008}.
+To assess accuracy over decreasing time-resolution for the three simulation environments, spike trains at each resolution were compared to a 'canonical' spike train (exponential Euler at a time-step of $dt = 0.001$ ms).
+
+To assess the performance of the simulators in absence of set-up overhead, models were simulated with a time-resolution of $0.1$ ms over increasing simulation time (Figure \ref{fig:figurebenchmark}B).
 The speed factor was defined as the ratio between time represented in the simulation and actual runtime (simulation-time).
@@
 
-"xolotl" uses the exponential euler method for integrating membrane potential \autocite{dayanTheoreticalNeuroscience2001}.
-"DynaSim" was implemented with a 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance in the documentation.
+Many simulators perform well in simulations of many compartments \autocite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}.
+Networks of up to 1,000 Hodgkin-Huxley cells were simulated for 5 s at a time-resolution of 0.1 ms (Figure \ref{fig:figurebenchmark}C).
 
@@
 (B) Hierarchical structure of a neuronal network considers controllers as components of compartments which act on conductances.
-(C) \texttt{xolotl} implements controllers XYZ.
+(C) \texttt{xolotl} implements controllers as properties of conductances and synapses.
 (D) Calcium sensors change maximal conductances to move a neuron from quiescence to a bursting state.
@@
 	\includegraphics[width=1.0\linewidth]{gfx/figure_benchmark}
-	\caption{\texttt{xolotl} benchmarked against \texttt{DynaSim}.}
+	\caption{\texttt{xolotl} benchmarked against \texttt{DynaSim} and \texttt{NEURON}.}
 	\label{fig:figurebenchmark}

### 2018-06-03 PaperLatex/paper.tex
minor fixes

@@
 The "connect" function adds synapses to the network.
+The first two arguments specify the presynaptic and postsynaptic compartments.
+The third dictates the type of synapse.
+All others follow the "'keyword', value" paradigm and preset parameters of the synapse.
 
@@
 Maximal conductances increase from random initial conditions to a set which elicits the desired network output by minimizing the error signal (Figure \ref{fig:figureintegralcontrol}D-F).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		GRAPHICAL INTERFACE
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Using the GUI to manipulate parameters}
+
+The "manipulate" function opens the GUI which permits visualization of changing parameters in real-time.
+Moving sliders representing the values of network parameters updates a plot (Figure \ref{fig:puppeteerscreenshot}B).
+By default, the function opens a figure displaying the results of the "plot" function, which shows the voltage and intracellular calcium traces for each compartment (Figure \ref{fig:puppeteerscreenshot}A).
+"manipulate" grants slider control over all "xolotl" parameters by default, but specific ones can be selected by passing them as arguments.
+For example, to manipulate only the maximal conductances and visualize using the "myPlot" function
+
+\begin{lstlisting}[style=Matlab-editor]
+	x.manipulate('*gbar', @myplot)
+\end{lstlisting}
 
@@
 
-"xolotl" and "DynaSim" performed with comparable accuracy at high time-resolution. At low time-resolution, "xolotl" significantly outperforms "DynaSim" in both accuracy and speed.
-
-To test whether transient overhead effects had a significant effect on performance, "xolotl" and "DynaSim" were tested with time-step $dt = 0.1$ ms for varied lengths of time.
-"xolotl" and "DynaSim" both performed best during longer simulations, approaching maximal performance at $>10^5$ time steps.
-"xolotl" is about 20 times faster for short simulation times and 3.5 times faster for arbitrarily large ones.
+DESCRIBE BENCHMARK RESULTS
 
@@
 	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/puppeteer_screenshot}
+	\caption{Using the GUI to manipulate neuron parameters.
+(A) Real-time output of the \texttt{plot} function displaying voltage (colored) and intracellular calcium (black) traces of a bursting neuron model \autocite{prinzAlternativeHandtuningConductancebased2003, prinzSimilarNetworkActivity2004}.
+Colors indicate the dominant current.
+(B) Sliders control the maximal conductances, which updates on the figure.}
+	\label{fig:puppeteerscreenshot}
+\end{figure}
+
+\begin{figure}
+	\centering
 	\includegraphics[width=1.0\linewidth]{gfx/figure_benchmark}
-	\caption{\texttt{xolotl} benchmarked against \texttt{DynaSim} and \texttt{NEURON}.}
+	\caption{\texttt{xolotl} benchmarked against \texttt{DynaSim} and \texttt{NEURON}.
+(A) Ratio of 'simulated' time to runtime (speed factor) and accuracy, measured by spike train coincidence plotted against decreasing time-resolution.
+(B) Speed factor for models at increasing simulation times.
+(C) Speed factor over number of compartments.}
 	\label{fig:figurebenchmark}

### 2018-06-03 PaperLatex/paper.tex
Beautify bibliography

@@
 \usepackage[english]{babel}
-\usepackage[backend=biber, bibencoding=utf8, style=authoryear, citestyle=authoryear-comp, autocite=inline]{biblatex}
+\usepackage[backend=biber, bibencoding=utf8, style=authoryear, citestyle=authoryear-comp, date=year, maxbibnames=6, minbibnames=6, url=false, isbn=false, autocite=inline]{biblatex}
 \addbibresource{Bibliography.bib}
@@
 \def\Authors{Alec Hoyland\,$^{1\dagger}$, Srinivas Gorur-Shandilya\,$^{1\dagger}$, and Eve Marder$^{1*}$}
-\def\Address{$^{1}$Volen Center for Complex Systems and Biology Department \\ Brandeis University \\ Waltham, MA \\
+\def\Address{$^{1}$Volen Center and Biology Department \\ Brandeis University \\ Waltham, MA 02454 \\ USA
 \vspace{0.25cm}
@@
 }
-\def\corrAuthor{Eve Marder \vspace{1mm}}
-\def\corrAddress{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
-\def\corrEmail{\texttt{marder@brandeis.edu}}
+\def\corrAuthor{Srinivas Gorur-Shandilya \vspace{1mm}}
+%\def\corrAddress{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
+\def\corrAddress{Marder Lab}
+\def\corrEmail{\texttt{srinivas@brandeis.edu}}
 
@@
 	\tiny
-	\keyFont{ \section{Keywords:} simulator, MATLAB, C++, conductance-based, neuron, network, keyword} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
+	\keyFont{ \section{Keywords:} simulator, MATLAB, C++, conductance-based, neuron, network, pedagogy}
 
@@
 
-"xolotl" (\url{https://github.com/marderlab/xolotl}) is a fast single-compartment and multi-compartment simulator in \texttt{C++} with \texttt{MATLAB} wrappers.
+Conductance-based models are the simplest biophysical representation of an electrically-excitable cell \autocite{hodgkinQuantitativeDescriptionMembrane1952}.
+Modeling studies now contribute significantly to mainstream research in small-circuit networks \autocite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}.
+Additionally, conductance-based models provide an approachable formalism based in electrophysiology for understanding salient principles of neuroscience.
+To this end, we have developed "xolotl" (\url{https://github.com/marderlab/xolotl}), a fast single-compartment and multi-compartment simulator in \texttt{C++} with \texttt{MATLAB} wrappers.
 Written with an emphasis on ease-of-use, "xolotl" can simulate single-compartment conductance-based models, networks of these, and detailed multi-compartment models.
@@
 \paragraph{Finding features.}
-\texttt{cpplab} comes with several features which simplify the handling of complexly-nested models.
+\texttt{cpplab} comes with several features which simplify the handling of complex, nested models.
 The "find" function acquires a cell array of all properties of the network which satisfy a search condition.
@@
 Isopotential models require one compartment, whereas models with multiple neurons, units, or non-trivial morphology require multiple compartments.
-All specifiable properties of compartments are shown in Supplementary Table 1.
+All specifiable properties of compartments are shown in Table 1.
 
@@
 Synapses can be added with the \texttt{connect} function.
-At minimum synapses possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses.
-All specifiable properties of synapses are shown in Supplementary Table 2.
+Synapses possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses.
+All specifiable properties of synapses are shown in Table 2.
 To create axons or transport chains, the \texttt{slice} function splits a compartment into $n$ discrete segments and adds these compartments to the network connected by electrical synapses.
@@
 Gating functions and their respective time constants are contained within the conductance header file.
-"xolotl" comes packaged with conductances from several dozen papers (Supplementary Table 3).
+"xolotl" comes packaged with conductances from several dozen papers (Table 3).
 
@@
 
-The focus on ease-of-use and speed means some features were elided in the streamlining process.
+The focus on ease-of-use and speed means some features were intentionally neglected in the streamlining process.
 
@@
 
+We envision "xolotl" will be helpful in teaching students how to
+
 \subsection{Reproducibility}

### 2018-06-03 PaperLatex/paper.tex
Add more citations

@@
 \usepackage{color, soul}
-\usepackage{url}
-\usepackage{multirow}
-\usepackage{array}
-\usepackage{fixltx2e}
-\usepackage{textcomp}
-\usepackage{booktabs}
 
 \usepackage{caption}
-\usepackage{subcaption}
 \usepackage{float}
@@
 
-\usepackage{paralist} % inparaenum
-
-\usepackage[bookmarks]{hyperref}
+\usepackage{hyperref}
 
@@
 \subsection{Syntax}
+\label{sec:syntax}
 
@@
 Isopotential models require one compartment, whereas models with multiple neurons, units, or non-trivial morphology require multiple compartments.
-All specifiable properties of compartments are shown in Table 1.
+
+%\begin{table}
+%	\begin{center}
+%		\begin{tabular}{ l | l | c }
+%			\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
+%			Membrane potential & "V" & $mV$ \\
+%			Intracellular calcium concentration & "Ca" & $\mu M$ \\
+%			Membrane capacitance & "Cm" & $nF/mm^2$ \\
+%			Volume & "vol" & $mm^3$ \\
+%			Surface area & "A" & $mm^2$ \\
+%			Radius of compartment & "radius" & $mm$ \\
+%			Length of compartment & "len" & $mm$ \\
+%			Extracellular calcium steady-state & "Ca_out" & $\mu M$ \\
+%			Intracellular calcium steady-state & "Ca_in" & $\mu M$ \\
+%			Calcium buffering shell thickness & "shell_thickness" & $\mu M$ \\
+%			Calcium current to concentration factor & "phi" & ??? \\
+%			Calcium buffering time constant & "tau_Ca" & $ms$ \\
+%			Average calcium concentration & "Ca_average" & $\mu M$ \\
+%			Target calcium concentration & "Ca_target" & $\mu M$ \\
+%		\end{tabular}
+%	\end{center}
+%	\label{tab:compartmentprops}
+%\end{table}
 
@@
 Synapses possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses.
-All specifiable properties of synapses are shown in Table 2.
+All specifiable properties of synapses are shown in Table \ref{tab:synapseprops}.
 To create axons or transport chains, the \texttt{slice} function splits a compartment into $n$ discrete segments and adds these compartments to the network connected by electrical synapses.
+
+%\begin{center}
+%	\begin{tabular}{ l | l | c }
+%		\label{tab:synapseprops}
+%		\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
+%		Maximal conductance & "gbar" & $\mu S /mm^2$ \\
+%		Reversal potential & "E" & $mV$ \\
+%		Synaptic gating variable & "s" &  \\
+%	\end{tabular}
+%\end{center}
 
@@
 Gating functions and their respective time constants are contained within the conductance header file.
-"xolotl" comes packaged with conductances from several dozen papers (Table 3).
+"xolotl" comes packaged with conductances from several dozen papers \autocite{dethierPositiveFeedbackCellular2015}.
+
+%\begin{center}
+%	\begin{tabular}{ l | l | c }
+%		\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
+%		Maximal conductance & "gbar" & $\mu S /mm^2$ \\
+%		Reversal potential & "E" & $mV$ \\
+%		Activation gating variable & "m" &  \\
+%		Inactivation gating variable & "h" &  \\
+%	\end{tabular}
+%\end{center}
 
@@
 Many simulators perform well in simulations of many compartments \autocite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}.
-Networks of up to 1,000 Hodgkin-Huxley cells were simulated for 5 s at a time-resolution of 0.1 ms (Figure \ref{fig:figurebenchmark}C).
+To assess how "xolotl" performs in these conditions, networks of up to 1,000 Hodgkin-Huxley cells were simulated for 5 s at a time-resolution of 0.1 ms (Figure \ref{fig:figurebenchmark}C).
 
@@
 
-We envision "xolotl" will be helpful in teaching students how to
+We envision that "xolotl" will be helpful in teaching students how to interpret cellular biophysics.
+The modular structure of "cpplab" and the graphical interface simplifies the process of manipulating and analyzing the properties of electrical excitability.
 
@@
 
-The code to generate all figures is available at (\url{https://github.com/marderlab/xolotl-paper}). "xolotl" is frely available at (\url{https://github.com/marderlab/xolotl}).
+The code to generate all figures is available at (\url{https://github.com/marderlab/xolotl-paper}). "xolotl" is freely available at (\url{https://github.com/marderlab/xolotl}).
 

### 2018-06-04 PaperLatex/paper.tex
Begin paper rewrite

@@
 
-	"xolotl" is a free and open-source neuronal simulator written in \texttt{C++} with \texttt{MATLAB} wrappers.
-Biophysically-detailed models of networks can be designed efficiently using an intuitive language tightly coupled to the object-based architecture of the underlying \texttt{C++} code.
-Models can be specified by adding conductances to compartment objects.
-The structure is modular, serialized, and searchable, permitting high-level programmatic control over nearly all features of the models.
-\texttt{C++} templates are provided for developing new conductances, compartments, and integration schemata.
-It also includes a customizable graphical user interface (GUI) for rapid prototyping and hand-tuning conductances in real-time.
-The modular structure and accessibility to all parameters, variables, and dynamics of the model network in \texttt{MATLAB} facilitate rapid construction and assessment of model networks.
-"xolotl" is freely available at \url{https://github.com/marderlab/xolotl}.
-This tool provides straightforward implementation and fast simulation of neuronal models while permitting full control over every aspect of the network and integration.
+	Information processing by neurons relies on the transmission and interaction of electrical signals that arise from the biophysics of ion channels and and synapses.
+Electrophysiological characterization of these low-level mechanisms have  allowed for the construction of conductance-based models that can reproduce many features of neuronal and circuit behavior.
+However, working with conductance-based models continues to be a challenge due to their high dimensionality, hindering intuition of their dynamical features.
+Here, we present a neuron and circuit simulator using a novel automatic type system that binds class templates written in "C++" to object-oriented code in "MATLAB".
+This approach combines the speed of "C++" code with the ease-of-use of scientific programming languages like "MATLAB".
+Neuron models are hierarchical, named and searchable, permitting high-level programmatic control over all parameters.
+The simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter on model behavior.
+The simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be easily extended to include arbitrary mechanisms and dynamics.
+Finally, the simulator is written in a modular fashion and has been released under a permissive free software license, enabling it to be integrated easily in third party applications.
 
@@
 Additionally, conductance-based models provide an approachable formalism based in electrophysiology for understanding salient principles of neuroscience.
+
 To this end, we have developed "xolotl" (\url{https://github.com/marderlab/xolotl}), a fast single-compartment and multi-compartment simulator in \texttt{C++} with \texttt{MATLAB} wrappers.

### 2018-06-04 figures/figure_benchmark.m

@@ -139,7 +139,7 @@ end
 xlabel(ax(2), 'time step (ms)')
 ylabel(ax(2), 'coincidence factor')
 set(ax(2), 'XScale', 'log', 'YLim', [0 1], 'XLim', [0 1.01])
-
+return
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% Benchmark Test #2
 % simulate a hodgkin-huxley model neuron over a series of simulation times
@@ -337,7 +337,7 @@ function Gamma = coincidence(canonSpikes, modelSpikes, dt, delta)
   % count the number of coincidences
   % between the canonical trace and the model trace
   for ii = 1:length(canonSpikes)-spikeRange
-    if sum(modelSpikes(ii + spikeRange)) > 0
+    if sum(modelSpikes(ii + spikeRange)) > 0 & sum(canonSpikes(ii+spikeRange)) > 0
       nCoincidences = nCoincidences + 1;
     end
   end

### 2018-06-04 PaperLatex/paper.tex
Rewrite introduction

@@
 
-Conductance-based models are the simplest biophysical representation of an electrically-excitable cell \autocite{hodgkinQuantitativeDescriptionMembrane1952}.
-Modeling studies now contribute significantly to mainstream research in small-circuit networks \autocite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}.
-Additionally, conductance-based models provide an approachable formalism based in electrophysiology for understanding salient principles of neuroscience.
+Nervous systems process and transmit information through electrically-excitable membranes.
+Conductance-based models are the simplest biophysical representation of an electrically-excitable cell (\cite{hodgkinQuantitativeDescriptionMembrane1952}).
+Studied based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small-circuit networks (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}).
+Additionally, these models provide an approachable framework for understanding salient principles of neuroscience.
+However, challenges remain in simulating biophysically-realistic neuron models.
+Conductance-based models are typically high-dimensional with many coupled nonlinear differential equations.
+Conductances are coupled through the membrane potential, and in multicompartment models, all membrane potentials are coupled.
+Simulators written in languages like "C", "C++", or "hoc" integrate equations quickly, but these simulators often lack the ease-of-use and interoperability of those written in scientific programming languages (e.g.
+"Python", "Julia", "MATLAB").
 
-To this end, we have developed "xolotl" (\url{https://github.com/marderlab/xolotl}), a fast single-compartment and multi-compartment simulator in \texttt{C++} with \texttt{MATLAB} wrappers.
-Written with an emphasis on ease-of-use, "xolotl" can simulate single-compartment conductance-based models, networks of these, and detailed multi-compartment models.
-"xolotl" exploits a novel automatic type system, \texttt{cpplab}, which binds \texttt{MATLAB} code to \texttt{C++} header files, creating objects and classes \textit{ad libitum} in \texttt{MATLAB} which reflect the underlying object-oriented code.
-"xolotl" implements \texttt{cpplab} to represent the nested structure of conductance-based models, and exploits the computational efficiency of the low-level programming language to quickly integrate models.
+Two major approaches have dominated the design of neuron simulators.
+Some simulators, such as "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}) are specified in a compiled language with pre-specified network components.
+These simulators tend to perform very fast computations with little overhead, but suffer from a steep learning curve.
+Implementations in more approachable languages and use of graphical interfaces mitigate these drawbacks, but obfuscate the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
+In contrast, many simulators have been designed in popular scientific programming languages to emphasize ease-of-use and flexibility.
+"DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}) parse strings of equations that are specified in the scientific programming language.
+The equations can be translated into a fast implementation language (\cite{stimbergEquationorientedSpecificationNeural2014}).
+This approach permits tremendous flexibility for simulating systems of differential equations, but the syntax tends to be verbose and the hierarchical nature of conductance-based models is not generally reflected in the user-generated code.
+There is no simulator that combines the advantages of both approaches without sacrificing efficiency, ease-of-use, or clarity.
+
+%Simulating conductance-based neuron models requires numerically integrating coupled nonlinear differential equations, for which several specialized methods have been developed.
+The exponential Euler method can efficiently integrate conductance-based models (\cite{dayanTheoreticalNeuroscience2001, ohErrorAnalysisSpecialized2006}).
+Exponential Euler also has the advantage of asymptotic stability at low time-resolution.
+Multicompartment models with spatial discretization require an integration scheme more suited to stiff equations.
+The implicit Crank-Nicolson method is both asymptotically stable and much faster than higher-order Runge-Kutta methods (\cite{dayanTheoreticalNeuroscience2001}).
+To expedite integrating large multi-compartment networks, "NEURON" implements a modified Gaussian elimination strategy (\cite{hinesEfficientComputationBranched1984, vooturiParallelizingHinesMatrix2017}).
+While these numerical and algebraic methods are powerful, we consider ease-of-use and interoperability with extant scientific programming tools to be just as important to the efficiency and flexibility of a neuronal simulator.
+
+We have developed a novel neuronal simulator, "xolotl", written in "C++" with a "MATLAB" interface.
+Designed with an emphasis on ease-of-use, "xolotl" can simulate single-compartment conductance-based models, networks of these, and detailed multi-compartment models.
+"xolotl" exploits a novel automatic type system, \texttt{cpplab}, which binds \texttt{MATLAB} code to \texttt{C++} header files, creating objects in \texttt{MATLAB} which reflect the underlying object-oriented code.
+"xolotl" implements \texttt{cpplab} to represent the nested structure of conductance-based models, and exploits the computational efficiency of the "C++" to quickly integrate models.
 For this reason, models can be implemented entirely in \texttt{MATLAB} with a few lines of code.
-
-Models are specified in \texttt{MATLAB} by a nested structure.
-The "xolotl" object contains compartments which themselves contain conductances.
-Synapses belong to the "xolotl" object and connect compartments together.
-The high-level specification supports arbitrarily large network and multi-compartment morphologies.
-
-The software has been implemented in \texttt{MATLAB} due to its ease-of-use and popularity among neuroscientists.
-\texttt{cpplab} provides a powerful backend for specifying and integrating models without relying on the significantly slower and limiting \texttt{MATLAB codegen}.
-While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
+The software has been implemented in \texttt{MATLAB} due to its ease-of-use and popularity among neuroscientists while \texttt{cpplab} provides a powerful backend for specifying and integrating models without relying on the significantly slower and limiting \texttt{MATLAB codegen}.
+While automated \texttt{C++} transpiling using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of features not accessible in "MATLAB", such as passing by reference and strongly-typed object-oriented programming.
 Minimal experience with \texttt{MATLAB} is required to use "xolotl", and all equations and integration methods are provided transparently to the end user.
-No string parsing of equations is required \autocite{sherfeyDynaSimMATLABToolbox2018, stimbergBrianSecondComing2013, stimbergEquationorientedSpecificationNeural2014}.
+No string parsing of equations is required (\cite{sherfeyDynaSimMATLABToolbox2018, stimbergBrianSecondComing2013, stimbergEquationorientedSpecificationNeural2014}.
 
@@
 Gating functions and their respective time constants are contained within the conductance header file.
-"xolotl" comes packaged with conductances from several dozen papers \autocite{dethierPositiveFeedbackCellular2015}.
+"xolotl" comes packaged with conductances from several dozen papers (\cite{dethierPositiveFeedbackCellular2015}.
 
@@
 \paragraph{Numerical integration.}
-"xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments \autocite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001, butcherNumericalDifferentialEquation2016}.
+"xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001, butcherNumericalDifferentialEquation2016}.
 These defaults provide a mix of speed, accuracy, and stability, and are built into the "cpplab" header files.
@@
 Adding new conductances also requires writing some "C++" code.
-For model conductances in the Hodgkin-Huxley formalism \autocite{hodgkinMeasurementCurrentvoltageRelations1952, dayanTheoreticalNeuroscience2001}, adjustments consist of changing default values in a template "C++" header file.
+For model conductances in the Hodgkin-Huxley formalism (\cite{hodgkinMeasurementCurrentvoltageRelations1952, dayanTheoreticalNeuroscience2001}, adjustments consist of changing default values in a template "C++" header file.
 Implementing a new integration scheme requires much more in-depth usage of "C++".
@@
 
-\paragraph{Limited numerical integration strategies.} While the exponential Euler method performs well in neuronal models \autocite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}, it may be desirable to use other methods under certain conditions.
+\paragraph{Limited numerical integration strategies.} While the exponential Euler method performs well in neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}, it may be desirable to use other methods under certain conditions.
 "xolotl" does not currently support other integration schemes for its built-in conductances, nor does the software support error-sensitive variable step-sizes `out-of-the-box.'
@@
 
-The seminal Hodgkin-Huxley model of action potentials in the squid giant axon \autocite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952} contains a fast inactivating sodium conductance ("NaV"), a non-inactivating delayed rectifier ("Kd"), and a passive leak current (Figure \ref{fig:figurehh}A).
+The seminal Hodgkin-Huxley model of action potentials in the squid giant axon (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952} contains a fast inactivating sodium conductance ("NaV"), a non-inactivating delayed rectifier ("Kd"), and a passive leak current (Figure \ref{fig:figurehh}A).
 A compartment, \texttt{HH}, with membrane capacitance ("Cm") and surface area ("A") can be specified by Figure \ref{fig:figurehh}B.
@@
 
-This model was constructed using conductances from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion \autocite{turrigianoSelectiveRegulationCurrent1995}.
+This model was constructed using conductances from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}.
 In the absence of applied positive current, the model is quiescent.
@@
 
-"xolotl" can recapitulate the results of voltage clamp experiments \autocite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}.
+"xolotl" can recapitulate the results of voltage clamp experiments (\cite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}.
 Figure Figure \ref{fig:figureclamp} displays steps in the procedure to clamp the membrane potential of a cell with delayed rectifier potassium conductance.
-During an \textit{in-vitro} experiment, confounding currents would be pharmacologically-blocked and two-electrode voltage clamp used to record tail currents at fixed membrane potential \autocite{connorInwardDelayedOutward1971, connorVoltageClampStudies1971}.
+During an \textit{in-vitro} experiment, confounding currents would be pharmacologically-blocked and two-electrode voltage clamp used to record tail currents at fixed membrane potential (\cite{connorInwardDelayedOutward1971, connorVoltageClampStudies1971}.
 
@@
 The current-voltage relation is the steady-state current over the clamped voltage, and the effective conductance is the derivative of that relation (Figure \ref{fig:figureclamp}F-G).
-Since the effective conductance is the product of the maximal conductance and the gating variables \autocite{dayanTheoreticalNeuroscience2001, turrigianoSelectiveRegulationCurrent1995} and the tail current is monotonically-increasing with time under voltage clamp, the current can be represented as non-inactivating.
+Since the effective conductance is the product of the maximal conductance and the gating variables (\cite{dayanTheoreticalNeuroscience2001, turrigianoSelectiveRegulationCurrent1995} and the tail current is monotonically-increasing with time under voltage clamp, the current can be represented as non-inactivating.
 Fitting a sigmoid to various powers yields a model for the current dynamics (Figure \ref{fig:figureclamp}H-I).
@@
 "xolotl" can implement homeostatic tuning rules as integral control.
-The controller computes an error signal (typically a function of intracellular calcium concentration), and adjusts the conductance or synapse it controls accordingly \autocite{olearyCorrelationsIonChannel2013}.
+The controller computes an error signal (typically a function of intracellular calcium concentration), and adjusts the conductance or synapse it controls accordingly (\cite{olearyCorrelationsIonChannel2013}.
 In "xolotl", integral controllers are "cpplab" objects added to the conductance or synapse they regulate.
@@
 
-To assess speed and accuracy, "xolotl", "DynaSim" \autocite{sherfeyDynaSimMATLABToolbox2018}, and "NEURON" \autocite{hinesNEURONSimulationEnvironment1997} were compared in simulations over varied time resolution,  simulation time, and number of compartments (Figure \ref{fig:figurebenchmark}).
+To assess speed and accuracy, "xolotl", "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}, and "NEURON" (\cite{hinesNEURONSimulationEnvironment1997} were compared in simulations over varied time resolution,  simulation time, and number of compartments (Figure \ref{fig:figurebenchmark}).
 
 Single-compartment Hodgkin-Huxley-like models were generated using conductance dynamics from \cite{liuModelNeuronActivitydependent1998} in the simulation environments.
-"xolotl" uses the exponential Euler method for integrating membrane potential \autocite{dayanTheoreticalNeuroscience2001}.
+"xolotl" uses the exponential Euler method for integrating membrane potential (\cite{dayanTheoreticalNeuroscience2001}.
 "DynaSim" was implemented with a 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance in the documentation.
-"NEURON" used an implicit Euler regime \autocite{hinesNEURONSimulationEnvironment1997}.
+"NEURON" used an implicit Euler regime (\cite{hinesNEURONSimulationEnvironment1997}.
 
@@
 Higher values indicate faster simulations.
-The coincidence factor determines the correlative overlap between two spike trains \autocite{jolivetBenchmarkTestQuantitative2008}.
+The coincidence factor determines the correlative overlap between two spike trains (\cite{jolivetBenchmarkTestQuantitative2008}.
 To assess accuracy over decreasing time-resolution for the three simulation environments, spike trains at each resolution were compared to a 'canonical' spike train (exponential Euler at a time-step of $dt = 0.001$ ms).
@@
 
-Many simulators perform well in simulations of many compartments \autocite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}.
+Many simulators perform well in simulations of many compartments (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}.
 To assess how "xolotl" performs in these conditions, networks of up to 1,000 Hodgkin-Huxley cells were simulated for 5 s at a time-resolution of 0.1 ms (Figure \ref{fig:figurebenchmark}C).
@@
 While the availability of hosting sites with version control (viz.
-GitHub (\url{https://github.com}), GitLab (\url{https://gitlab.com/}), and Open Science Framework (\url{https://osf.io/})) and the push for reproducibility in computational science \autocite{eklundClusterFailureWhy2016, stoddenEnhancingReproducibilityComputational2016, bakerWhyScientistsMust2016} has resulted in the availability of source code, much of this code base is bespoke and difficult to implement \autocite{sedanoCodeReadabilityTesting2016, xuMeasurementSourceCode2017}.
+GitHub (\url{https://github.com}), GitLab (\url{https://gitlab.com/}), and Open Science Framework (\url{https://osf.io/})) and the push for reproducibility in computational science (\cite{eklundClusterFailureWhy2016, stoddenEnhancingReproducibilityComputational2016, bakerWhyScientistsMust2016} has resulted in the availability of source code, much of this code base is bespoke and difficult to implement (\cite{sedanoCodeReadabilityTesting2016, xuMeasurementSourceCode2017}.
 
@@
 
-"C" is statically-typed, with procedural syntax that provides low-level access to memory \autocite{kernighanProgrammingLanguage1978}, providing significant advantages for time-intensive computations.
+"C" is statically-typed, with procedural syntax that provides low-level access to memory (\cite{kernighanProgrammingLanguage1978}, providing significant advantages for time-intensive computations.
 Unfortunately, automatic code-generation is limited by the supervening language.
@@
 	\caption{Simulating a voltage-clamp experiment.
-(A) Cartoon of a cell with delayed rectifier potassium conductance \autocite{liuModelNeuronActivitydependent1998} with experimentally-fixed voltage.
+(A) Cartoon of a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998} with experimentally-fixed voltage.
 (B) Structure of \texttt{xolotl} object in A.
@@
 	\caption{Simulating neurons under homeostatic regulation.
-(A) Cartoon of a model neuron \autocite{liuModelNeuronActivitydependent1998} with integral control \autocite{olearyCorrelationsIonChannel2013}.
+(A) Cartoon of a model neuron (\cite{liuModelNeuronActivitydependent1998} with integral control (\cite{olearyCorrelationsIonChannel2013}.
 (B) Hierarchical structure of a neuronal network considers controllers as components of compartments which act on conductances.
@@
 	\caption{Using the GUI to manipulate neuron parameters.
-(A) Real-time output of the \texttt{plot} function displaying voltage (colored) and intracellular calcium (black) traces of a bursting neuron model \autocite{prinzAlternativeHandtuningConductancebased2003, prinzSimilarNetworkActivity2004}.
+(A) Real-time output of the \texttt{plot} function displaying voltage (colored) and intracellular calcium (black) traces of a bursting neuron model (\cite{prinzAlternativeHandtuningConductancebased2003, prinzSimilarNetworkActivity2004}.
 Colors indicate the dominant current.

### 2018-06-04 neuron/benchmark1.py

@@ -41,7 +41,7 @@ soma(0.5).pas.g = 0.0001
 h.psection(sec=soma)
 
 # set up injected current
-stim = h.IClamp(soma(0.5))
+stim = h.IClamp(soma(0.2))
 stim.amp = 0.2 # nA
 stim.dur = 5000 # ms
 

### 2018-06-04 neuron/benchmark2.py

@@ -7,7 +7,62 @@ from neuron import h, gui
 import time
 
 # execute the main NEURON file
-import neuron_HH
+
+################################################################################
+################################################################################
+################################################################################
+
+# import the graphical interface
+from neuron import h, gui
+# import arrays and graphics
+import numpy as np
+from matplotlib import pyplot
+
+# create the neuron
+soma        = h.Section(name='soma');
+# h.psection(sec=soma)
+
+# set the size of the soma
+soma.L      = 28.209 # microns
+soma.diam   = 28.209 # microns
+print('surface area of the soma = {}'.format(soma(0.5).area()))
+
+# set up the capacitance
+soma.cm     = 1 # uF/cm^2
+
+# add conductances from Liu et al. 1998
+soma.insert('pas')
+
+soma.insert('na')
+soma.insert('kd')
+
+# set maximal conductances
+soma(0.5).pas.g = 0.0001
+
+h.psection(sec=soma)
+
+# set up injected current
+stim = h.IClamp(soma(0.2))
+stim.amp = 0.2 # nA
+stim.dur = 5000 # ms
+
+# set up recording variables
+v_vec       = h.Vector()
+t_vec       = h.Vector()
+v_vec.record(soma(0.5)._ref_v)
+t_vec.record(h._ref_t)
+
+# set up simulation
+h.dt    = 0.1 # ms
+h.tstop = 1000.0 # ms
+# h.secondorder = 2 # use Crank-Nicholson
+
+h.run()
+
+################################################################################
+################################################################################
+################################################################################
+
 
 # set up vectors to hold outputs
 t_end = np.array([10, 18, 34, 62, 113,  207,  379,  695, 1274, 2336, 4281, 7848,

### 2018-06-04 neuron/neuron_HH.py

@@ -3,6 +3,7 @@ from neuron import h, gui
 # import arrays and graphics
 import numpy as np
 from matplotlib import pyplot
+import time
 
 # create the neuron
 soma        = h.Section(name='soma');
@@ -30,7 +31,7 @@ h.psection(sec=soma)
 # set up injected current
 stim = h.IClamp(soma(0.5))
 stim.amp = 0.2 # nA
-stim.dur = 5000 # ms
+stim.dur = 10000 # ms
 
 # set up recording variables
 v_vec       = h.Vector()
@@ -40,14 +41,16 @@ t_vec.record(h._ref_t)
 
 # set up simulation
 h.dt    = 0.1 # ms
-h.tstop = 1000.0 # ms
-# h.secondorder = 2 # use Crank-Nicholson
+h.tstop = 50000 # ms
+h.secondorder = 2 # use Crank-Nicholson
 
+tic = time.perf_counter() # s
 h.run()
+toc = time.perf_counter() - tic # s
 
 # visualize the results
-# pyplot.figure(figsize=(8,4)) # Default figsize is (8,6)
-# pyplot.plot(t_vec, v_vec)
-# pyplot.xlabel('time (ms)')
-# pyplot.ylabel('mV')
-# pyplot.show()
+pyplot.figure(figsize=(8,4)) # Default figsize is (8,6)
+pyplot.plot(t_vec, v_vec)
+pyplot.xlabel('time (ms)')
+pyplot.ylabel('mV')
+pyplot.show()

### 2018-06-04 PaperLatex/paper.tex
delete extraneous

@@
 Conductance-based models are the simplest biophysical representation of an electrically-excitable cell (\cite{hodgkinQuantitativeDescriptionMembrane1952}).
-Studied based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small-circuit networks (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}).
+Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small-circuit networks (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}).
 Additionally, these models provide an approachable framework for understanding salient principles of neuroscience.
@@
 Two major approaches have dominated the design of neuron simulators.
-Some simulators, such as "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}) are specified in a compiled language with pre-specified network components.
+First, some simulators, such as "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}) are specified in a compiled language with pre-specified network components.
 These simulators tend to perform very fast computations with little overhead, but suffer from a steep learning curve.
@@
 This approach permits tremendous flexibility for simulating systems of differential equations, but the syntax tends to be verbose and the hierarchical nature of conductance-based models is not generally reflected in the user-generated code.
-There is no simulator that combines the advantages of both approaches without sacrificing efficiency, ease-of-use, or clarity.
+Neither approach can maintain efficiency, ease-of-use, or clarity without sacrificing one of these aspects.
 

### 2018-06-04 neuron/x86_64/CaT.c

@@ -1,580 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* VECTORIZED */
-#define NRN_VECTORIZED 1
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__CaT
-#define _nrn_initial _nrn_initial__CaT
-#define nrn_cur _nrn_cur__CaT
-#define _nrn_current _nrn_current__CaT
-#define nrn_jacob _nrn_jacob__CaT
-#define nrn_state _nrn_state__CaT
-#define _net_receive _net_receive__CaT 
-#define _f_settables _f_settables__CaT 
-#define settables settables__CaT 
-#define states states__CaT 
- 
-#define _threadargscomma_ _p, _ppvar, _thread, _nt,
-#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
-#define _threadargs_ _p, _ppvar, _thread, _nt
-#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- /* Thread safe. No static _p or _ppvar. */
- 
-#define t _nt->_t
-#define dt _nt->_dt
-#define gmax _p[0]
-#define r _p[1]
-#define s _p[2]
-#define d _p[3]
-#define eca _p[4]
-#define ica _p[5]
-#define ralpha _p[6]
-#define rbeta _p[7]
-#define salpha _p[8]
-#define sbeta _p[9]
-#define dalpha _p[10]
-#define dbeta _p[11]
-#define Dr _p[12]
-#define Ds _p[13]
-#define Dd _p[14]
-#define v _p[15]
-#define _g _p[16]
[…]

### 2018-06-04 neuron/x86_64/Kd.c

@@ -1,407 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* VECTORIZED */
-#define NRN_VECTORIZED 1
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__Kd
-#define _nrn_initial _nrn_initial__Kd
-#define nrn_cur _nrn_cur__Kd
-#define _nrn_current _nrn_current__Kd
-#define nrn_jacob _nrn_jacob__Kd
-#define nrn_state _nrn_state__Kd
-#define _net_receive _net_receive__Kd 
-#define settables settables__Kd 
-#define states states__Kd 
- 
-#define _threadargscomma_ _p, _ppvar, _thread, _nt,
-#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
-#define _threadargs_ _p, _ppvar, _thread, _nt
-#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- /* Thread safe. No static _p or _ppvar. */
- 
-#define t _nt->_t
-#define dt _nt->_dt
-#define gbar _p[0]
-#define m _p[1]
-#define i _p[2]
-#define minf _p[3]
-#define taum _p[4]
-#define Dm _p[5]
-#define v _p[6]
-#define _g _p[7]
- 
-#if MAC
-#if !defined(v)
-#define v _mlhv
-#endif
-#if !defined(h)
-#define h _mlhh
-#endif
-#endif
- 
-#if defined(__cplusplus)
-extern "C" {
-#endif
- static int hoc_nrnpointerindex =  -1;
- static Datum* _extcall_thread;
- static Prop* _extcall_prop;
- /* external NEURON variables */
[…]

### 2018-06-04 neuron/x86_64/NaV.c

@@ -1,419 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* VECTORIZED */
-#define NRN_VECTORIZED 1
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__NaV
-#define _nrn_initial _nrn_initial__NaV
-#define nrn_cur _nrn_cur__NaV
-#define _nrn_current _nrn_current__NaV
-#define nrn_jacob _nrn_jacob__NaV
-#define nrn_state _nrn_state__NaV
-#define _net_receive _net_receive__NaV 
-#define settables settables__NaV 
-#define states states__NaV 
- 
-#define _threadargscomma_ _p, _ppvar, _thread, _nt,
-#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
-#define _threadargs_ _p, _ppvar, _thread, _nt
-#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- /* Thread safe. No static _p or _ppvar. */
- 
-#define t _nt->_t
-#define dt _nt->_dt
-#define gbar _p[0]
-#define m _p[1]
-#define h _p[2]
-#define i _p[3]
-#define minf _p[4]
-#define hinf _p[5]
-#define taum _p[6]
-#define tauh _p[7]
-#define Dm _p[8]
-#define Dh _p[9]
-#define v _p[10]
-#define _g _p[11]
- 
-#if MAC
-#if !defined(v)
-#define v _mlhv
-#endif
-#if !defined(h)
-#define h _mlhh
-#endif
-#endif
- 
-#if defined(__cplusplus)
-extern "C" {
-#endif
- static int hoc_nrnpointerindex =  -1;
[…]

### 2018-06-04 neuron/x86_64/kd.c

@@ -1,443 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* VECTORIZED */
-#define NRN_VECTORIZED 1
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__kd
-#define _nrn_initial _nrn_initial__kd
-#define nrn_cur _nrn_cur__kd
-#define _nrn_current _nrn_current__kd
-#define nrn_jacob _nrn_jacob__kd
-#define nrn_state _nrn_state__kd
-#define _net_receive _net_receive__kd 
-#define rates rates__kd 
-#define states states__kd 
- 
-#define _threadargscomma_ _p, _ppvar, _thread, _nt,
-#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
-#define _threadargs_ _p, _ppvar, _thread, _nt
-#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- /* Thread safe. No static _p or _ppvar. */
- 
-#define t _nt->_t
-#define dt _nt->_dt
-#define Erev _p[0]
-#define i _p[1]
-#define m _p[2]
-#define g _p[3]
-#define minf _p[4]
-#define tau_m _p[5]
-#define Dm _p[6]
-#define v _p[7]
-#define _g _p[8]
- 
-#if MAC
-#if !defined(v)
-#define v _mlhv
-#endif
-#if !defined(h)
-#define h _mlhh
-#endif
-#endif
- 
-#if defined(__cplusplus)
-extern "C" {
-#endif
- static int hoc_nrnpointerindex =  -1;
- static Datum* _extcall_thread;
- static Prop* _extcall_prop;
[…]

### 2018-06-04 neuron/x86_64/mod_func.c

@@ -1,19 +1,0 @@
-#include <stdio.h>
-#include "hocdec.h"
-extern int nrnmpi_myid;
-extern int nrn_nobanner_;
-
-extern void _na_reg(void);
-extern void _kd_reg(void);
-
-void modl_reg(){
-  if (!nrn_nobanner_) if (nrnmpi_myid < 1) {
-    fprintf(stderr, "Additional mechanisms from files\n");
-
-    fprintf(stderr," na.mod");
-    fprintf(stderr," kd.mod");
-    fprintf(stderr, "\n");
-  }
-  _na_reg();
-  _kd_reg();
-}

### 2018-06-04 neuron/x86_64/na.c

@@ -1,478 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* VECTORIZED */
-#define NRN_VECTORIZED 1
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__na
-#define _nrn_initial _nrn_initial__na
-#define nrn_cur _nrn_cur__na
-#define _nrn_current _nrn_current__na
-#define nrn_jacob _nrn_jacob__na
-#define nrn_state _nrn_state__na
-#define _net_receive _net_receive__na 
-#define rates rates__na 
-#define states states__na 
- 
-#define _threadargscomma_ _p, _ppvar, _thread, _nt,
-#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
-#define _threadargs_ _p, _ppvar, _thread, _nt
-#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- /* Thread safe. No static _p or _ppvar. */
- 
-#define t _nt->_t
-#define dt _nt->_dt
-#define Erev _p[0]
-#define i _p[1]
-#define m _p[2]
-#define h _p[3]
-#define g _p[4]
-#define minf _p[5]
-#define hinf _p[6]
-#define tau_h _p[7]
-#define tau_m _p[8]
-#define Dm _p[9]
-#define Dh _p[10]
-#define v _p[11]
-#define _g _p[12]
- 
-#if MAC
-#if !defined(v)
-#define v _mlhv
-#endif
-#if !defined(h)
-#define h _mlhh
-#endif
-#endif
- 
-#if defined(__cplusplus)
-extern "C" {
-#endif
- static int hoc_nrnpointerindex =  -1;
[…]

### 2018-06-04 PaperLatex/paper.tex
Edits to introduction

@@
 Two major approaches have dominated the design of neuron simulators.
-First, some simulators, such as "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}) are specified in a compiled language with pre-specified network components.
-These simulators tend to perform very fast computations with little overhead, but suffer from a steep learning curve.
-Implementations in more approachable languages and use of graphical interfaces mitigate these drawbacks, but obfuscate the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
-In contrast, many simulators have been designed in popular scientific programming languages to emphasize ease-of-use and flexibility.
-"DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}) parse strings of equations that are specified in the scientific programming language.
-The equations can be translated into a fast implementation language (\cite{stimbergEquationorientedSpecificationNeural2014}).
-This approach permits tremendous flexibility for simulating systems of differential equations, but the syntax tends to be verbose and the hierarchical nature of conductance-based models is not generally reflected in the user-generated code.
-Neither approach can maintain efficiency, ease-of-use, or clarity without sacrificing one of these aspects.
+First, some simulators, such as "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}) are specified in a custom language with modular network components.
+These simulators tend to perform fast computations with little overhead, but suffer from a steep learning curve.
+Wrapping "NEURON" in a more approachable language like "Python" and use of graphical interfaces mitigate these drawbacks,at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
+In contrast, many simulators have been designed to be used entirely in popular scientific programming languages.
+This has the benefit of ease-of-use and intercompatibility with other tools.
+"DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}) translate strings of equations that are specified in the scientific programming language, that can be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
+This approach permits considerable flexibility for simulating systems of differential equations, but the syntax tends to be verbose and the hierarchical nature of conductance-based neuron models is not naturally reflected in the user-generated code.
+Neither approach can maintain efficiency, ease-of-use, and clarity without sacrificing one of these aspects.
 
@@
 
-We have developed a novel neuronal simulator, "xolotl", written in "C++" with a "MATLAB" interface.
-Designed with an emphasis on ease-of-use, "xolotl" can simulate single-compartment conductance-based models, networks of these, and detailed multi-compartment models.
-"xolotl" exploits a novel automatic type system, \texttt{cpplab}, which binds \texttt{MATLAB} code to \texttt{C++} header files, creating objects in \texttt{MATLAB} which reflect the underlying object-oriented code.
-"xolotl" implements \texttt{cpplab} to represent the nested structure of conductance-based models, and exploits the computational efficiency of the "C++" to quickly integrate models.
-For this reason, models can be implemented entirely in \texttt{MATLAB} with a few lines of code.
-The software has been implemented in \texttt{MATLAB} due to its ease-of-use and popularity among neuroscientists while \texttt{cpplab} provides a powerful backend for specifying and integrating models without relying on the significantly slower and limiting \texttt{MATLAB codegen}.
-While automated \texttt{C++} transpiling using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of features not accessible in "MATLAB", such as passing by reference and strongly-typed object-oriented programming.
-Minimal experience with \texttt{MATLAB} is required to use "xolotl", and all equations and integration methods are provided transparently to the end user.
-No string parsing of equations is required (\cite{sherfeyDynaSimMATLABToolbox2018, stimbergBrianSecondComing2013, stimbergEquationorientedSpecificationNeural2014}.
-
-"xolotl" comes packaged with visualization functions and a graphical user interface (GUI) for real-time manipulation of model parameters.
+To overcome these design limitations, we have developed a novel automatic type system, "cpplab", which binds "MATLAB" code to "C++" header files.
+This architecture automatically creates objects in "MATLAB" that reflect the underlying object-oriented "C++" code.
+"xolotl" is an implementation of the "cpplab" system specialized for integrating conductance-based neuron and network models.
+Models can be easily constructed from network building blocks in in a few lines of "MATLAB" code using hierarchical, intuitive syntax.
+Since objects in "MATLAB" are linked to the "C++" implementation, changes made in "MATLAB" are reflected in the underlying "C++" objects.
+"xolotl" comes packaged with hundreds of conductances and synapses, built-in visualization functions, and a graphical user interface (GUI) for real-time manipulation of model parameters.
 Plotting of voltage, intracellular calcium, conductance gating functions, and time constants is provided by built-in "xolotl" methods.
@@
 The ease-of-use of these tools lends them to pedagogical applications and rapid exploration of toy models.
-This tool aims to simplify the investigation of dynamics of complex neural network models, facilitate collaborative modeling, and complement other tools being developed in the neuroinformatics community.
+This tool aims to simplify the investigation of dynamics of biophysically-realistic network models, facilitate collaborative modeling, and complement other tools being developed in the neuroinformatics community.
 
@@
 
-The software has been designed in "MATLAB" due to its popularity among neuroscientists for pedagogy and research.
+The software is built in "MATLAB" due to its popularity among neuroscientists for pedagogy and research.
 "xolotl" capitalizes on "MATLAB"'s straightforward structure array syntax to permit rapid prototyping and experimentation, especially for small neuronal networks of complex models.
-Parameters of conductances, neuronal compartments, and simulations may all be edited in the structure before any calls to integration functions.
+Parameters of conductances, cellular compartments, and simulations may all be edited in the structure before any calls to integration functions.
 The underlying code is written in "C++" for speed and memory optimization, and while models can indeed be integrated using the compiled binary, symbolic manipulations can be readily performed in "MATLAB" without ever touching the foundational code.
@@
 
-\texttt{MATLAB} provides a high level programmatic and graphical interface for implementing, manipulating, and visualizing models without sacrificing the enhancements of the underlying \texttt{C++} code.
+"MATLAB" provides a high level programmatic and graphical interface for implementing, manipulating, and visualizing models without sacrificing the enhancements of the underlying "C++" code.
 
-\paragraph{Modular structure.}
-
+\paragraph{Modular structure. }
 Models are specified by adding compartments and synapses to the "xolotl" object.
 Conductances are added to compartments and controllers can be added to conductances.
-This modular structure recapitulates the biophysics of the Hodgkin-Huxley formalism and obviates the need to explicitly write out equations, which in "xolotl" are contained within the conductance header files.
+This modular structure recapitulates the biophysics of the Hodgkin-Huxley formalism and obviates the need to explicitly write out equations, which in "xolotl" are contained within "C++" header files.
 
-\paragraph{Interface between \texttt{C++} and \texttt{MATLAB}.}
+\paragraph{Automatic type system. }
+When a network component is added in "MATLAB", the "xolotl" structure is updated, and "cpplab" automatically constructs and compiles the "C++" from the requisite header files.
+"MATLAB" treats "cpplab" constructions as fully-typed first-class objects allowing for symbolic manipulation using only the high-level programming language and graphical interfaces.
+"xolotl" simulations are run entirely from "C++" executables; the inputs and outputs can be represented entirely in "MATLAB".
 
-"xolotl" relies on \texttt{cpplab} constructions, which allow the user to exploit the efficiency of low-level \texttt{C++} code.
-\texttt{MATLAB} treats \texttt{cpplab} objects as fully-typed variables allowing for symbolic manipulation using only the high-level programming language and graphical interfaces.
-"xolotl" is fast because all time-intensive code is written in \texttt{C++}.
-While automated \texttt{C++} transpiling from \texttt{MATLAB} using the proprietary \texttt{codegen} can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of \texttt{MATLAB} over \texttt{C++} prevents efficient use of low-level features, such as passing by reference and object-oriented programming.
-\texttt{C++} provides speed improvements beyond the benefits of translating \texttt{MATLAB} features into low-level code.
-For this reason, \texttt{cpplab} has been designed to provide an interface for constructing, transpiling, and compiling \texttt{C++} code to be called from within \texttt{MATLAB}.
-"xolotl" simulations are run entirely from \texttt{C++} executables.
+\paragraph{Built in \texttt{C++}. }
+"xolotl" is fast because all time-intensive code is written in "C++".
+While automated "C++" transpiling from "MATLAB" using the proprietary "codegen" can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of "MATLAB" over "C++" prevents efficient use of features, such as passing by reference and object-oriented programming.
+"C++" provides speed improvements beyond the benefits of translating "MATLAB" features into statically-typed code.
+For this reason, "cpplab" has been designed to provide an interface for constructing, transpiling, and compiling "C++" code to be called from within "MATLAB".
 
-\paragraph{Automatic and efficient compiling.}
-"xolotl" automatically handles transpiling and compiling "MATLAB" code into "C++".
-The MD5 algorithm hashes the network to compile a new binary and "MEX" bridge file only if needed and to confirm that the correct binary fetched during execution.
+\paragraph{Automatic compiling. }
+The user experience of "xolotl" can stay entirely in "MATLAB".
+Transpiling and compiling of "C++" is performed automatically.
+Cryptographic hashing of the "xolotl" object confirms that compiling occurs only when necessary, and that the correct binary is used during integration.
 
-\subsection{Syntax}
-\label{sec:syntax}
+\subsection{Limitations}
+\label{limitations}
 
-"MATLAB" can easily control the "cpplab" objects using the standard, flexible data structure notation popular in high-level scripting languages.
+The focus on ease-of-use and fast simulations means some features were intentionally neglected in the streamlining process.
+
+\paragraph{Reliance on compiled \texttt{C++} code.} While "MATLAB" comes with robust features for compiling "C" and "C++" code, "xolotl" cannot run without "C++" compilation.
+For users, this necessitates the additional step of setting up the "mex" compiler which can be problematic.
+Secondly, compilation adds a small amount to total processing time.
+Longer simulations ($>1000$ time-steps) minimizes this effect.
+In addition, adding new network components also requires writing some "C++" code.
+For model conductances in the Hodgkin-Huxley formalism (\cite{hodgkinMeasurementCurrentvoltageRelations1952, dayanTheoreticalNeuroscience2001}, adjustments consist of changing default values in a template "C++" header file.
+Implementing a new integration scheme or "cpplab" class requires much more in-depth usage of "C++".
+
+\paragraph{Limited to conductance-based models.} "xolotl" has been developed specifically for conductance-based models. It does not currently support rate- or current-based models.
+
+\paragraph{Limited numerical integration strategies.} While the exponential Euler method performs well in neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}, it may be desirable to use other methods under certain conditions.
+"xolotl" does not currently support other integration schemes for its built-in conductances, nor does the software support error-sensitive variable step-sizes `out-of-the-box.'
+
+\paragraph{Inefficient tools for handling large networks.} While "xolotl" can integrate large networks ($>1000$ compartments), "xolotl" uses string-based comprehension for labeling compartments which is suited to descriptive naming, but prohibits vector operations over compartments.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		USAGE EXAMPLES
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Usage Examples}
+\label{usage}
+
+In "MATLAB", users create a "xolotl" object and populate it with "cpplab"-generated objects which describe compartments, conductances, synapses, and controllers.
+The model is integrated with the "integrate" function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.
+
+"xolotl" comes packaged with a library of pre-existing conductance and synapse objects which greatly simplify the task of constructing model neurons.
+These objects can be referenced by name and added directly to a compartment.
+Novel conductance dynamics can be easily written by modifying a template header file contained in the "xolotl" distribution, or designed entirely from scratch.
 
@@
 \paragraph{Finding features.}
-\texttt{cpplab} comes with several features which simplify the handling of complex, nested models.
+"cpplab" comes with several features which simplify the handling of complex, nested models.
 The "find" function acquires a cell array of all properties of the network which satisfy a search condition.
@@
 Simulations run on multi-core processors or high-performance computing clusters using the Parallel Computing toolbox.
-
-\subsection{Limitations}
-\label{limitations}
-
-The focus on ease-of-use and speed means some features were intentionally neglected in the streamlining process.
-
-\paragraph{Reliance on compiled \texttt{C++} code.} While "MATLAB" comes with robust features for compiling "C" and "C++" code, "xolotl" cannot run without "C++" compilation.
-For users, this necessitates the additional step of setting up the "mex" compiler which can be problematical, especially for nonstandard (e.g.
-Arch-based Linux).
-Secondly, compilation adds a small amount to total processing time.
-Longer simulations ($>1000$ time-steps) minimizes this effect.
-Adding new conductances also requires writing some "C++" code.
-For model conductances in the Hodgkin-Huxley formalism (\cite{hodgkinMeasurementCurrentvoltageRelations1952, dayanTheoreticalNeuroscience2001}, adjustments consist of changing default values in a template "C++" header file.
-Implementing a new integration scheme requires much more in-depth usage of "C++".
-
-\paragraph{Limited to conductance-based models.} "xolotl" has been developed specifically for conductance-based models. It does not currently support rate- or current-based models.
-
-\paragraph{Limited numerical integration strategies.} While the exponential Euler method performs well in neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}, it may be desirable to use other methods under certain conditions.
-"xolotl" does not currently support other integration schemes for its built-in conductances, nor does the software support error-sensitive variable step-sizes `out-of-the-box.'
-
-\paragraph{Inefficient tools for handling large networks.} While "xolotl" can integrate large networks ($>1000$ compartments), "xolotl" uses string-based comprehension for labeling compartments which is suited to descriptive naming, but prohibits vector operations over compartments.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%
-%		USAGE EXAMPLES
-%
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Usage Examples}
-\label{usage}
-
-In "MATLAB", users create a "xolotl" object and populate it with "cpplab"-generated objects which describe compartments, conductances, synapses, and controllers.
-The model is integrated with the "integrate" function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.
-
-"xolotl" comes packaged with a library of pre-existing conductance and synapse objects which greatly simplify the task of constructing model neurons.
-These objects can be referenced by name and added directly to a compartment.
-Novel conductance dynamics can be easily written by modifying a template header file contained in the "xolotl" distribution, or designed entirely from scratch.
 

### 2018-06-04 PaperLatex/paper.tex
Rewrite main body

@@
 \newcommand{\vs}{\textit{vs.}\xspace}
+\newcommand{\viz}{\textit{viz.}\xspace}
 
@@
 
-	Information processing by neurons relies on the transmission and interaction of electrical signals that arise from the biophysics of ion channels and and synapses.
-Electrophysiological characterization of these low-level mechanisms have  allowed for the construction of conductance-based models that can reproduce many features of neuronal and circuit behavior.
+	Information processing by neurons relies on the transmission and interaction of electrical signals that arise from the biophysics of ion channels and synapses.
+Electrophysiological characterization of these mechanisms has allowed for the construction of conductance-based models that can reproduce many features of neuronal and circuit behavior.
 However, working with conductance-based models continues to be a challenge due to their high dimensionality, hindering intuition of their dynamical features.
-Here, we present a neuron and circuit simulator using a novel automatic type system that binds class templates written in "C++" to object-oriented code in "MATLAB".
+Here, we present a neuron and network simulator using a novel automatic type system that binds class templates written in "C++" to object-oriented code in "MATLAB".
 This approach combines the speed of "C++" code with the ease-of-use of scientific programming languages like "MATLAB".
@@
 Conductances are coupled through the membrane potential, and in multicompartment models, all membrane potentials are coupled.
-Simulators written in languages like "C", "C++", or "hoc" integrate equations quickly, but these simulators often lack the ease-of-use and interoperability of those written in scientific programming languages (e.g.
-"Python", "Julia", "MATLAB").
+Simulators written in languages like "C", "C++", or "hoc" integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages (\eg "Python", "Julia", "MATLAB").
 
@@
 Neither approach can maintain efficiency, ease-of-use, and clarity without sacrificing one of these aspects.
-
-%Simulating conductance-based neuron models requires numerically integrating coupled nonlinear differential equations, for which several specialized methods have been developed.
-The exponential Euler method can efficiently integrate conductance-based models (\cite{dayanTheoreticalNeuroscience2001, ohErrorAnalysisSpecialized2006}).
-Exponential Euler also has the advantage of asymptotic stability at low time-resolution.
-Multicompartment models with spatial discretization require an integration scheme more suited to stiff equations.
-The implicit Crank-Nicolson method is both asymptotically stable and much faster than higher-order Runge-Kutta methods (\cite{dayanTheoreticalNeuroscience2001}).
-To expedite integrating large multi-compartment networks, "NEURON" implements a modified Gaussian elimination strategy (\cite{hinesEfficientComputationBranched1984, vooturiParallelizingHinesMatrix2017}).
-While these numerical and algebraic methods are powerful, we consider ease-of-use and interoperability with extant scientific programming tools to be just as important to the efficiency and flexibility of a neuronal simulator.
 
@@
 "xolotl" is designed to be easy-to-use without sacrificing speed.
-
 The software is built in "MATLAB" due to its popularity among neuroscientists for pedagogy and research.
@@
 Parameters of conductances, cellular compartments, and simulations may all be edited in the structure before any calls to integration functions.
-The underlying code is written in "C++" for speed and memory optimization, and while models can indeed be integrated using the compiled binary, symbolic manipulations can be readily performed in "MATLAB" without ever touching the foundational code.
+While the underlying code is written in "C++" for speed and memory optimization, symbolic manipulations can be readily performed in "MATLAB" without ever touching the foundational code.
 
@@
 
-\paragraph{Modular structure. }
+\paragraph{Library of network components.} "xolotl" comes packaged with hundreds of pre-existing components (viz.
+compartments, synapses, conductances, and controllers) which greatly simplify the task of constructing model neurons.
+Compartments represent cylindrical sections of membrane with a membrane potential.
+Synapses connect two compartments together by introducing a current in the post-synaptic neuron that depends on the state of the presynaptic neuron.
+Conductances represent populations of ion channels in a compartment that produce transmembrane currents.
+Controllers integrate a signal from a conductance or synapse and change the properties of the conductance or synapse in response.
+
+\paragraph{Object-oriented.}
 Models are specified by adding compartments and synapses to the "xolotl" object.
-Conductances are added to compartments and controllers can be added to conductances.
+Conductances are added to compartments and controllers can be added to conductances or synapses.
 This modular structure recapitulates the biophysics of the Hodgkin-Huxley formalism and obviates the need to explicitly write out equations, which in "xolotl" are contained within "C++" header files.
 
-\paragraph{Automatic type system. }
+\paragraph{Automatic type system.}
 When a network component is added in "MATLAB", the "xolotl" structure is updated, and "cpplab" automatically constructs and compiles the "C++" from the requisite header files.
@@
 
-\paragraph{Built in \texttt{C++}. }
+\paragraph{Fast.}
 "xolotl" is fast because all time-intensive code is written in "C++".
@@
 
-\paragraph{Automatic compiling. }
+\paragraph{Automatic compiling.}
 The user experience of "xolotl" can stay entirely in "MATLAB".
@@
 
-\paragraph{Reliance on compiled \texttt{C++} code.} While "MATLAB" comes with robust features for compiling "C" and "C++" code, "xolotl" cannot run without "C++" compilation.
-For users, this necessitates the additional step of setting up the "mex" compiler which can be problematic.
-Secondly, compilation adds a small amount to total processing time.
-Longer simulations ($>1000$ time-steps) minimizes this effect.
-In addition, adding new network components also requires writing some "C++" code.
-For model conductances in the Hodgkin-Huxley formalism (\cite{hodgkinMeasurementCurrentvoltageRelations1952, dayanTheoreticalNeuroscience2001}, adjustments consist of changing default values in a template "C++" header file.
-Implementing a new integration scheme or "cpplab" class requires much more in-depth usage of "C++".
-
 \paragraph{Limited to conductance-based models.} "xolotl" has been developed specifically for conductance-based models. It does not currently support rate- or current-based models.
@@
 \paragraph{Inefficient tools for handling large networks.} While "xolotl" can integrate large networks ($>1000$ compartments), "xolotl" uses string-based comprehension for labeling compartments which is suited to descriptive naming, but prohibits vector operations over compartments.
+
+\paragraph{New mechanisms require new \texttt{C++} code.} Adding new network components also requires writing some "C++" code.
+Adding a new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinMeasurementCurrentvoltageRelations1952, dayanTheoreticalNeuroscience2001}), requires creating a new "C++" header file, though this is generally trivial.
+Implementing a new integration scheme or "cpplab" class requires much more in-depth knowledge of "C++" and modification of the core code.
 
@@
 
-In "MATLAB", users create a "xolotl" object and populate it with "cpplab"-generated objects which describe compartments, conductances, synapses, and controllers.
+In this section, we describe several uses of "xolotl" for common situations.
+
+%In "MATLAB", users create a "xolotl" object and populate it with "cpplab"-generated objects which describe compartments, conductances, synapses, and controllers.
 The model is integrated with the "integrate" function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.
-
-"xolotl" comes packaged with a library of pre-existing conductance and synapse objects which greatly simplify the task of constructing model neurons.
-These objects can be referenced by name and added directly to a compartment.
-Novel conductance dynamics can be easily written by modifying a template header file contained in the "xolotl" distribution, or designed entirely from scratch.
-
-\paragraph{Adding features.}
-The "add" function creates a compartment, conductance, or controller and affixes it as a field in the "xolotl" network structure.
+%
+%
+%
+%\subsection{Constructing Networks}
+%
+%The "add" function creates a compartment, conductance, or controller and affixes it as a field in the "xolotl" network structure.
 This function generates a "MATLAB struct" that faithfully represents the underlying "C++" code.
 Compartments add to the "xolotl" object and conductances add to compartments.
-Specific properties can be specified using key-value pair arguments (e.g.
-Figure \ref{fig:figurehh}A).
-
-\paragraph{Finding features.}
-"cpplab" comes with several features which simplify the handling of complex, nested models.
-The "find" function acquires a cell array of all properties of the network which satisfy a search condition.
-For example, one can find all paths to maximal conductances within the "'HH'" compartment by:
-
-\begin{lstlisting}[style=Matlab-editor]
-	x.find('HH*gbar');
-\end{lstlisting}
-
-To extract a vector of the maximal conductances:
-
-\begin{lstlisting}[style=Matlab-editor]
-	gbars = x.get('HH*gbar');
-\end{lstlisting}
-
-To set the maximal conductances all at once:
-
-\begin{lstlisting}[style=Matlab-editor]
-	x.set('HH*gbar', gbars)
-\end{lstlisting}
-
-\paragraph{Compartments.}
-A model neuron consists of one or more compartments, each representing a section of membrane with capacitance and surface area.
+Specific properties can be specified using key-value pair arguments (\eg Figure \ref{fig:figurehh}A).
+%
+%\paragraph{Compartments.}
+%A model neuron consists of one or more compartments, each representing a section of membrane with capacitance and surface area.
 Isopotential models require one compartment, whereas models with multiple neurons, units, or non-trivial morphology require multiple compartments.
-
-%\begin{table}
-%	\begin{center}
-%		\begin{tabular}{ l | l | c }
-%			\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
-%			Membrane potential & "V" & $mV$ \\
-%			Intracellular calcium concentration & "Ca" & $\mu M$ \\
-%			Membrane capacitance & "Cm" & $nF/mm^2$ \\
-%			Volume & "vol" & $mm^3$ \\
-%			Surface area & "A" & $mm^2$ \\
-%			Radius of compartment & "radius" & $mm$ \\
-%			Length of compartment & "len" & $mm$ \\
-%			Extracellular calcium steady-state & "Ca_out" & $\mu M$ \\
-%			Intracellular calcium steady-state & "Ca_in" & $\mu M$ \\
-%			Calcium buffering shell thickness & "shell_thickness" & $\mu M$ \\
-%			Calcium current to concentration factor & "phi" & ??? \\
-%			Calcium buffering time constant & "tau_Ca" & $ms$ \\
-%			Average calcium concentration & "Ca_average" & $\mu M$ \\
-%			Target calcium concentration & "Ca_target" & $\mu M$ \\
-%		\end{tabular}
-%	\end{center}
-%	\label{tab:compartmentprops}
-%\end{table}
-
-\paragraph{Synapses.}
-"xolotl" provides some features for generating complex models.
+%
+%%\begin{table}
+%%	\begin{center}
+%%		\begin{tabular}{ l | l | c }
+%%			\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
+%%			Membrane potential & "V" & $mV$ \\
+%%			Intracellular calcium concentration & "Ca" & $\mu M$ \\
+%%			Membrane capacitance & "Cm" & $nF/mm^2$ \\
+%%			Volume & "vol" & $mm^3$ \\
+%%			Surface area & "A" & $mm^2$ \\
+%%			Radius of compartment & "radius" & $mm$ \\
+%%			Length of compartment & "len" & $mm$ \\
+%%			Extracellular calcium steady-state & "Ca_out" & $\mu M$ \\
+%%			Intracellular calcium steady-state & "Ca_in" & $\mu M$ \\
+%%			Calcium buffering shell thickness & "shell_thickness" & $\mu M$ \\
+%%			Calcium current to concentration factor & "phi" & ??? \\
+%%			Calcium buffering time constant & "tau_Ca" & $ms$ \\
+%%			Average calcium concentration & "Ca_average" & $\mu M$ \\
+%%			Target calcium concentration & "Ca_target" & $\mu M$ \\
+%%		\end{tabular}
+%%	\end{center}
+%%	\label{tab:compartmentprops}
+%%\end{table}
+%
+%\paragraph{Synapses.}
+%"xolotl" provides some features for generating complex models.
 Synapses can be added with the \texttt{connect} function.
 Synapses possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses.
-All specifiable properties of synapses are shown in Table \ref{tab:synapseprops}.
 To create axons or transport chains, the \texttt{slice} function splits a compartment into $n$ discrete segments and adds these compartments to the network connected by electrical synapses.
-
-%\begin{center}
-%	\begin{tabular}{ l | l | c }
-%		\label{tab:synapseprops}
-%		\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
-%		Maximal conductance & "gbar" & $\mu S /mm^2$ \\
-%		Reversal potential & "E" & $mV$ \\
-%		Synaptic gating variable & "s" &  \\
-%	\end{tabular}
-%\end{center}
-
-\paragraph{Conductances and controllers.}
-
-All conductances contain fields for maximal conductance and reversal potential.
+%
+%%\begin{center}
+%%	\begin{tabular}{ l | l | c }
+%%		\label{tab:synapseprops}
+%%		\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
+%%		Maximal conductance & "gbar" & $\mu S /mm^2$ \\
+%%		Reversal potential & "E" & $mV$ \\
+%%		Synaptic gating variable & "s" &  \\
+%%	\end{tabular}
+%%\end{center}
+%
+%\paragraph{Conductances and controllers.}
+%
+%All conductances contain fields for maximal conductance and reversal potential.
 Conductances with activation and inactivation variables include them as $m$ and $h$ respectively.
@@
 "xolotl" comes packaged with conductances from several dozen papers (\cite{dethierPositiveFeedbackCellular2015}.
-
-%\begin{center}
-%	\begin{tabular}{ l | l | c }
-%		\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
-%		Maximal conductance & "gbar" & $\mu S /mm^2$ \\
-%		Reversal potential & "E" & $mV$ \\
-%		Activation gating variable & "m" &  \\
-%		Inactivation gating variable & "h" &  \\
-%	\end{tabular}
-%\end{center}
-
-\paragraph{Creating custom \texttt{cpplab} objects.}
-
-"xolotl" contains template header files for producing custom conductances. The template contains instructions on how to design novel conductances with arbitrary specifications.
-
-\paragraph{Simulation.}
-
-Models are simulated in "xolotl" with the \texttt{integrate} function which outputs as time series the membrane potentials, intracellular calcium concentrations, controller states, intrinsic currents, and synaptic currents.
+%
+%%\begin{center}
+%%	\begin{tabular}{ l | l | c }
+%%		\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
+%%		Maximal conductance & "gbar" & $\mu S /mm^2$ \\
+%%		Reversal potential & "E" & $mV$ \\
+%%		Activation gating variable & "m" &  \\
+%%		Inactivation gating variable & "h" &  \\
+%%	\end{tabular}
+%%\end{center}
+%
+%\paragraph{Creating custom \texttt{cpplab} objects.}
+%
+%"xolotl" contains template header files for producing custom conductances. The template contains instructions on how to design novel conductances with arbitrary specifications.
+%
+%\subsection{Simulation}
+%
+%Models are simulated in "xolotl" with the \texttt{integrate} function which outputs as time series the membrane potentials, intracellular calcium concentrations, controller states, intrinsic currents, and synaptic currents.
 The \texttt{integrate} function also accepts an argument which specifies injected current or clamped voltage.
-
-\paragraph{Numerical integration.}
-"xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001, butcherNumericalDifferentialEquation2016}.
+%
+%\paragraph{Numerical integration.}
+%"xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001, butcherNumericalDifferentialEquation2016}).
 These defaults provide a mix of speed, accuracy, and stability, and are built into the "cpplab" header files.
@@
 The simulation time-resolution can be specified to target arbitrary precision, and an output time step can be selected to support automatic down-sampling for memory considerations.
-
-\paragraph{`Closed-loop' vs. `open-loop.'}
-Simulations can be run in `closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or `open-loop' mode which begins simulation with the current network state.
-
-\paragraph{Using the graphical interface to manipulate parameters.}
-
-"xolotl" comes packaged with a graphical user interface for visualizing parameter changes in real-time.
+%
+%\paragraph{`Closed-loop' vs. `open-loop.'}
+%Simulations can be run in `closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or `open-loop' mode which begins simulation with the current network state.
+%
+%\subsection{Additional Features}
+%
+%\paragraph{Using the graphical interface to manipulate parameters.}
+%
+%"xolotl" comes packaged with a graphical user interface for visualizing parameter changes in real-time.
 The \texttt{manipulate} function opens the GUI, which displays a figure plotting the membrane potential and intracellular calcium concentration of all compartments as time series, and a dialog box with customizable sliders for all parameters of the model, much like the "Manipulate" function in Wolfram "Mathematica".
 Moving the sliders integrates the model in `open-loop` mode with the new parameters.
-The parameters available in the sliders can be customized by passing a cell array to manipulate.
-For example, to only see sliders for maximal conductances of the "HH" compartment, call "x.manipulate('HH*gbar')".
 Closing the GUI saves the network state of the model to the "xolotl" object.
 This is particularly helpful for rapid prototyping of models.
-
-\paragraph{Optimizing parameters.}
-"xolotl" can use the Global Optimization toolbox for "MATLAB" to optimize any accessible "xolotl" parameters.
+%
+%\paragraph{Optimizing parameters.}
+%"xolotl" can use the Global Optimization toolbox for "MATLAB" to optimize any accessible "xolotl" parameters.
 The toolbox is algorithm-agnostic and accepts any function in "MATLAB" with a scalar first output as the objective function.
@@
 
-The seminal Hodgkin-Huxley model of action potentials in the squid giant axon (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952} contains a fast inactivating sodium conductance ("NaV"), a non-inactivating delayed rectifier ("Kd"), and a passive leak current (Figure \ref{fig:figurehh}A).
-A compartment, \texttt{HH}, with membrane capacitance ("Cm") and surface area ("A") can be specified by Figure \ref{fig:figurehh}B.
-Network properties can be set during construction or afterwards using dot-notation in \texttt{MATLAB} (e.g.
-\texttt{x.HH.Cm}).
-Figure \ref{fig:figurehh}C shows the \texttt{MATLAB} command prompt after invoking the "xolotl" object \texttt{x}, displaying the hierarchical structure inherent in conductance-based treatments of neurodynamics.
+The seminal Hodgkin-Huxley model of action potentials in the squid giant axon (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}) contains a fast inactivating sodium conductance ("NaV"), a non-inactivating delayed rectifier ("Kd"), and a passive leak current (Figure \ref{fig:figurehh}A).
+This model was constructed using conductances from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}).
+A compartment, "HH", with membrane capacitance ("Cm") and surface area ("A") is specified by Figure \ref{fig:figurehh}B.
+Figure \ref{fig:figurehh}C shows the "MATLAB" command window after invoking the "xolotl" object \texttt{x}, displaying the hierarchical structure inherent in conductance-based treatments of neurodynamics.
 
-This model was constructed using conductances from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}.
 In the absence of applied positive current, the model is quiescent.
 When 0.2 nA is injected, the model tonically spikes (Figure \ref{fig:figurehh}D).
-The \texttt{integrate} function takes the applied current as an argument (e.g.
-"x.integrate(Iapp)"), so that the "xolotl" object is agnostic to integration-specific perturbations.
-The \texttt{plot} function generates voltage and intracellular calcium traces, where the voltage trace is colored by the dominant current.
+The \texttt{integrate} function takes the applied current as an argument.
+The "plot" function generates voltage and intracellular calcium traces, where the voltage trace is colored by the dominant current (Figure \ref{fig:figurehh}E).
 If the membrane potential is increasing, the strongest instantaneous inward current colors the trace.
 Conversely, if the membrane potential is decreasing, the strongest outward current colors the trace instead.
-Figure \ref{fig:figurehh}F-I display the results of the \texttt{show} function.
-Activation and inactivation steady-states and the voltage-dependent time constants of these gating variables describe the conductance dynamics in absence of other channel types.
+Figure \ref{fig:figurehh}F-I display the results of the \texttt{show} function which plots the activation and inactivation steady-states and the voltage-dependent time constants of conductance gating variables.
 
@@
 
-"xolotl" can recapitulate the results of voltage clamp experiments (\cite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}.
-Figure Figure \ref{fig:figureclamp} displays steps in the procedure to clamp the membrane potential of a cell with delayed rectifier potassium conductance.
-During an \textit{in-vitro} experiment, confounding currents would be pharmacologically-blocked and two-electrode voltage clamp used to record tail currents at fixed membrane potential (\cite{connorInwardDelayedOutward1971, connorVoltageClampStudies1971}.
+"xolotl" can recapitulate the results of voltage clamp experiments (\cite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}).
+Figure \ref{fig:figureclamp} displays steps in the procedure to clamp the membrane potential of a cell with delayed rectifier potassium conductance.
+During an \textit{in-vitro} experiment, confounding currents would be pharmacologically-blocked and two-electrode voltage clamp used to record tail currents at fixed membrane potential (\cite{connorInwardDelayedOutward1971, connorVoltageClampStudies1971}).
 
-A single-compartment model with a delayed-rectifier conductance is simulated at stepped membrane potentials.
-The model is simulated using the "integrate" function.
-The second argument determines the clamped voltage and the fourth output is the current trace.
-
-\begin{lstlisting}[style=Matlab-editor]
-	[V, Ca, ~, I] = x.integrate([], clamped_voltage)
-\end{lstlisting}
-
+A single-compartment model with a delayed-rectifier conductance (\cite{liuModelNeuronActivitydependent1998}) is simulated at stepped membrane potentials.
+The "integrate" function outputs the membrane potential and current as time series.
 Currents under voltage clamp approach the steady-state holding current (Figure \ref{fig:figureclamp}D-E).
 The current-voltage relation is the steady-state current over the clamped voltage, and the effective conductance is the derivative of that relation (Figure \ref{fig:figureclamp}F-G).
-Since the effective conductance is the product of the maximal conductance and the gating variables (\cite{dayanTheoreticalNeuroscience2001, turrigianoSelectiveRegulationCurrent1995} and the tail current is monotonically-increasing with time under voltage clamp, the current can be represented as non-inactivating.
+Since the effective conductance is the product of the maximal conductance and the gating variables (\cite{dayanTheoreticalNeuroscience2001, turrigianoSelectiveRegulationCurrent1995}) and the tail current is monotonically-increasing with time under voltage clamp, the current can be represented as non-inactivating.
 Fitting a sigmoid to various powers yields a model for the current dynamics (Figure \ref{fig:figureclamp}H-I).
@@
 Network models in "xolotl" consist of compartment objects connected by synapses.
-Synapses are stored in a vector array as a field of the "xolotl" object in \texttt{MATLAB}.
 Presynaptic and postsynaptic labels indicate the connectivity of the synapse.
-Figure Figure \ref{fig:figurenetwork} implements a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans.
+Figure \ref{fig:figurenetwork} implements a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans (\cite{prinzSimilarNetworkActivity2004}).
 The pyloric model contains three compartments and seven synapses (Figure \ref{fig:figurenetwork}A).
-This structure is reciprocated in the hierarchy of the "xolotl" object, where conductances are contained within compartments (Figure \ref{fig:figurenetwork}B).
-
-Representing the network in "xolotl" requires constructing three compartments and eight conductances in each using the "add" function.
-
-\begin{lstlisting}[style=Matlab-editor]
-	x.add('AB', 'compartment', 'Cm', 10, 'A', 0.628, ...)
-	x.AB.add('prinz/NaV', 'gbar', 1000, 'E', 50)
-	...
-\end{lstlisting}
-
-Synapses are upper-level properties of the network which point between two compartments (Figure \ref{fig:figurenetwork}C).
-This exploits vectorized operations in \texttt{MATLAB} and does not require each synapse to possess a unique name.
-The "connect" function adds synapses to the network.
-The first two arguments specify the presynaptic and postsynaptic compartments.
-The third dictates the type of synapse.
-All others follow the "'keyword', value" paradigm and preset parameters of the synapse.
-
-\begin{lstlisting}[style=Matlab-editor]
-	x.connect('AB', 'LP', 'Chol', 'gbar', 30)
-\end{lstlisting}
+This structure is recapitulated in the hierarchy of the "xolotl" object, where conductances are contained within compartments (Figure \ref{fig:figurenetwork}B).
+Synapses are stored in a vector array as a field of the "xolotl" object in "MATLAB" (Figure \ref{fig:figurenetwork}C).
 
@@
 
-\subsection{Simulating Integral Control}
+\subsection{Simulating Homeostasis}
 
 "xolotl" can implement homeostatic tuning rules as integral control.
-The controller computes an error signal (typically a function of intracellular calcium concentration), and adjusts the conductance or synapse it controls accordingly (\cite{olearyCorrelationsIonChannel2013}.
+The controller computes an error signal (typically a function of intracellular calcium concentration), and adjusts the conductance or synapse it controls accordingly (\cite{olearyCorrelationsIonChannel2013}).
 In "xolotl", integral controllers are "cpplab" objects added to the conductance or synapse they regulate.
@@
 "manipulate" grants slider control over all "xolotl" parameters by default, but specific ones can be selected by passing them as arguments.
-For example, to manipulate only the maximal conductances and visualize using the "myPlot" function
-
-\begin{lstlisting}[style=Matlab-editor]
-	x.manipulate('*gbar', @myplot)
-\end{lstlisting}
 
@@
 
-To assess speed and accuracy, "xolotl", "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}, and "NEURON" (\cite{hinesNEURONSimulationEnvironment1997} were compared in simulations over varied time resolution,  simulation time, and number of compartments (Figure \ref{fig:figurebenchmark}).
+To assess speed and accuracy, "xolotl", "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), and "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}) were compared in simulations over varied time resolution,  simulation time, and number of compartments (Figure \ref{fig:figurebenchmark}).
 
-Single-compartment Hodgkin-Huxley-like models were generated using conductance dynamics from \cite{liuModelNeuronActivitydependent1998} in the simulation environments.
-"xolotl" uses the exponential Euler method for integrating membrane potential (\cite{dayanTheoreticalNeuroscience2001}.
+Single-compartment Hodgkin-Huxley-like models were generated using conductance dynamics from \cite{liuModelNeuronActivitydependent1998}) in the simulation environments.
+"xolotl" uses the exponential Euler method for integrating membrane potential (\cite{dayanTheoreticalNeuroscience2001}).
 "DynaSim" was implemented with a 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance in the documentation.
-"NEURON" used an implicit Euler regime (\cite{hinesNEURONSimulationEnvironment1997}.
+"NEURON" used an implicit Euler regime (\cite{hinesNEURONSimulationEnvironment1997}).
 
@@
 Higher values indicate faster simulations.
-The coincidence factor determines the correlative overlap between two spike trains (\cite{jolivetBenchmarkTestQuantitative2008}.
-To assess accuracy over decreasing time-resolution for the three simulation environments, spike trains at each resolution were compared to a 'canonical' spike train (exponential Euler at a time-step of $dt = 0.001$ ms).
+The coincidence factor determines the correlative overlap between two spike trains (\cite{jolivetBenchmarkTestQuantitative2008}).
+To assess accuracy over decreasing time-resolution for the three simulation environments, spike trains at each resolution were compared to a `canonical' spike train (exponential Euler at a time-step of $dt = 0.001$ ms).
 
@@
 
-Many simulators perform well in simulations of many compartments (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}.
+Many simulators perform well in simulations of many compartments (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}).
 To assess how "xolotl" performs in these conditions, networks of up to 1,000 Hodgkin-Huxley cells were simulated for 5 s at a time-resolution of 0.1 ms (Figure \ref{fig:figurebenchmark}C).
@@
 "xolotl" fosters reproducibility in science.
-While the availability of hosting sites with version control (viz.
-GitHub (\url{https://github.com}), GitLab (\url{https://gitlab.com/}), and Open Science Framework (\url{https://osf.io/})) and the push for reproducibility in computational science (\cite{eklundClusterFailureWhy2016, stoddenEnhancingReproducibilityComputational2016, bakerWhyScientistsMust2016} has resulted in the availability of source code, much of this code base is bespoke and difficult to implement (\cite{sedanoCodeReadabilityTesting2016, xuMeasurementSourceCode2017}.
+While the availability of hosting sites with version control (\viz GitHub (\url{https://github.com}), GitLab (\url{https://gitlab.com/}), and Open Science Framework (\url{https://osf.io/})) and the push for reproducibility in computational science (\cite{eklundClusterFailureWhy2016, stoddenEnhancingReproducibilityComputational2016, bakerWhyScientistsMust2016}) has resulted in the availability of source code, much of this code base is bespoke and difficult to implement (\cite{sedanoCodeReadabilityTesting2016, xuMeasurementSourceCode2017}).
 

### 2018-06-05 PaperLatex/paper.tex
Rewrite main body
Fix citation style

@@
 Built from the ground up in "C++", "xolotl" maintains all the advantages of custom compiled code, but can run in "MATLAB" without the user having to touch the "C++" code.
-"xolotl" represents compartment, conductance, synapse, and controllers as "cpplab" objects, which map to underlying "C++" header files.
+"xolotl" represents compartment, conductance, synapse, and controllers as "cpplab" objects, which template from underlying "C++" header files.
 In this way, properties of the "xolotl" network can be examined and changed using object-oriented paradigms.
@@
 SG-S designed and implemented the core of the "xolotl" toolbox.
-AH contributed to the code base, created the online user documentation, and wrote the manuscript.
+AH contributed to the code base, created the online user documentation.
+SG-S and AH wrote the manuscript.
 EM supervised the project.
@@
 \section*{Acknowledgments}
-The authors would like to thank Mara CP Rue and Hillary Rodgers for beta-testing the "xolotl" software. Janis Li helped to prepare some conductance header files.
+The authors would like to thank Mara CP Rue and Hillary Rodgers for beta-testing the "xolotl" software. Janis Li helped to prepare many conductance header files.
 

### 2018-06-05 PaperLatex/paper.tex
minor fixes

@@
 \def\firstAuthorLast{Hoyland {et~al.}} 
-\def\Authors{Alec Hoyland\,$^{1\dagger}$, Srinivas Gorur-Shandilya\,$^{1\dagger}$, and Eve Marder$^{1*}$}
+\def\Authors{Srinivas Gorur-Shandilya\,$^{1*\dagger}$, Alec Hoyland\,$^{1\dagger}$, and Eve Marder$^{1}$}
 \def\Address{$^{1}$Volen Center and Biology Department \\ Brandeis University \\ Waltham, MA 02454 \\ USA
@@
 %\def\corrAddress{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
-\def\corrAddress{Marder Lab}
-\def\corrEmail{\texttt{srinivas@brandeis.edu}}
+\def\corrAddress{..}
+\def\corrEmail{\texttt{srinivasgs@brandeis.edu}}
 
@@
 To compare the integration methods, models were simulated for 5 s at varying time-resolution (Figure \ref{fig:figurebenchmark}A).
-The ratio between 'simulated' time and actual runtime was defined as the speed factor.
-Higher values indicate faster simulations.
+The speed factor was defined as the ratio between time represented in the simulation and actual runtime (simulation-time).
+Therefore, the speed factor represents how many times faster the simulation is than a real-time observation.
 The coincidence factor determines the correlative overlap between two spike trains (\cite{jolivetBenchmarkTestQuantitative2008}).
+In this context, it is a measure of accuracy in the numerical simulation.
+Unity indicates perfect correlation.
 To assess accuracy over decreasing time-resolution for the three simulation environments, spike trains at each resolution were compared to a `canonical' spike train (exponential Euler at a time-step of $dt = 0.001$ ms).
@@
 To assess the performance of the simulators in absence of set-up overhead, models were simulated with a time-resolution of $0.1$ ms over increasing simulation time (Figure \ref{fig:figurebenchmark}B).
-The speed factor was defined as the ratio between time represented in the simulation and actual runtime (simulation-time).
-Therefore, the speed factor represents how many times faster the simulation is than a real-time observation.
 
 Many simulators perform well in simulations of many compartments (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}).
-To assess how "xolotl" performs in these conditions, networks of up to 1,000 Hodgkin-Huxley cells were simulated for 5 s at a time-resolution of 0.1 ms (Figure \ref{fig:figurebenchmark}C).
+To assess how "xolotl" under these conditions, networks of up to 1,000 Hodgkin-Huxley cells were simulated for 5 s at a time-resolution of 0.1 ms (Figure \ref{fig:figurebenchmark}C).
 

### 2018-06-05 paper/paper.tex
rewrote goals

--- 2018-06-05 PaperLatex/paper.tex
@@
+% !BIB TS-program = biber
 \documentclass{frontiersSCNS} % for Science articles
@@
 %\def\corrAddress{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
-\def\corrAddress{..}
+\def\corrAddress{Volen Center and Biology Department}
 \def\corrEmail{\texttt{srinivasgs@brandeis.edu}}
@@
 However, working with conductance-based models continues to be a challenge due to their high dimensionality, hindering intuition of their dynamical features.
-Here, we present a neuron and network simulator using a novel automatic type system that binds class templates written in "C++" to object-oriented code in "MATLAB".
-This approach combines the speed of "C++" code with the ease-of-use of scientific programming languages like "MATLAB".
-Neuron models are hierarchical, named and searchable, permitting high-level programmatic control over all parameters.
+Here, we present a neuron and network simulator built on a novel automatic type system that binds class templates written in "C++" to object-oriented code in "MATLAB".
+Our approach combines the speed and powerful features of "C++" code with the ease-of-use and rich libraries of scientific programming languages like "MATLAB".
+In our software, models are hierarchical, and components are named and searchable, permitting high-level programmatic control over all parameters.
 The simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter on model behavior.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Nervous systems process and transmit information through electrically-excitable membranes.
-Conductance-based models are the simplest biophysical representation of an electrically-excitable cell (\cite{hodgkinQuantitativeDescriptionMembrane1952}).
+Nervous systems process and transmit information using electrically excitable membranes.
+Conductance-based models are powerful biophysical simplification of an electrically excitable cell (\cite{hodgkinQuantitativeDescriptionMembrane1952}).
 Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small-circuit networks (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}).
 Additionally, these models provide an approachable framework for understanding salient principles of neuroscience.
-However, challenges remain in simulating biophysically-realistic neuron models.
-Conductance-based models are typically high-dimensional with many coupled nonlinear differential equations.
-Conductances are coupled through the membrane potential, and in multicompartment models, all membrane potentials are coupled.
-Simulators written in languages like "C", "C++", or "hoc" integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages (\eg "Python", "Julia", "MATLAB").
+However, several challenges remain in simulating biophysically-detailed conductance-based neuron models.
+First, these models can be high-dimensional with many nonlinear differential equations, with several parameters.
+Second, many or all equations are strongly coupled through the variables like the membrane potential, and in multi-compartment models, all membrane potentials and their equations are coupled to each other.
+Finally, the choice of programming language imposes tradeoffs in neuron and network simulators: simulators written in languages like "C", "C++", or "hoc" can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like "Python", "Julia" or "MATLAB".
 
 Two major approaches have dominated the design of neuron simulators.
-First, some simulators, such as "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}) are specified in a custom language with modular network components.
+First, some simulators, such as "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}) use a custom language to specify modular network components.
 These simulators tend to perform fast computations with little overhead, but suffer from a steep learning curve.
-Wrapping "NEURON" in a more approachable language like "Python" and use of graphical interfaces mitigate these drawbacks,at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
+Wrapping "NEURON" in a more approachable language like "Python" and use of graphical interfaces mitigate these drawbacks, at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
 In contrast, many simulators have been designed to be used entirely in popular scientific programming languages.
-This has the benefit of ease-of-use and intercompatibility with other tools.
-"DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}) translate strings of equations that are specified in the scientific programming language, that can be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
-This approach permits considerable flexibility for simulating systems of differential equations, but the syntax tends to be verbose and the hierarchical nature of conductance-based neuron models is not naturally reflected in the user-generated code.
-Neither approach can maintain efficiency, ease-of-use, and clarity without sacrificing one of these aspects.
+These simulators can be easier to use, and can benefit from inter-compatibility with other tools.
+Programs like "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}) allow the user to specify models using strings of equations that are specified in the scientific programming language, that can be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
+This approach permits considerable flexibility for simulating systems of differential equations, but the syntax tends to be verbose and the hierarchical nature of conductance-based neuron models is not naturally reflected in the constructed model.
+Neither approach permits tools that simultaneously maintain efficiency, ease-of-use, and clarity.
 
 To overcome these design limitations, we have developed a novel automatic type system, "cpplab", which binds "MATLAB" code to "C++" header files.
-This architecture automatically creates objects in "MATLAB" that reflect the underlying object-oriented "C++" code.
-"xolotl" is an implementation of the "cpplab" system specialized for integrating conductance-based neuron and network models.
-Models can be easily constructed from network building blocks in in a few lines of "MATLAB" code using hierarchical, intuitive syntax.
-Since objects in "MATLAB" are linked to the "C++" implementation, changes made in "MATLAB" are reflected in the underlying "C++" objects.
+This architecture automatically creates objects in "MATLAB" that represent the underlying object-oriented "C++" code, without having to re-specify these objects in "MATLAB".
+In this paper, we introduce "xolotl", an implementation of the "cpplab" system specialized in integrating conductance-based neuron and network models.
+Models can be easily constructed from network building blocks in a few lines of "MATLAB" code using a hierarchical and intuitive syntax.
+Since model objects in the "MATLAB" workspace are automatically linked to the "C++" specification, configuring these objects in "MATLAB" transparently configures the underlying "C++" objects.
 "xolotl" comes packaged with hundreds of conductances and synapses, built-in visualization functions, and a graphical user interface (GUI) for real-time manipulation of model parameters.
@@
 The ease-of-use of these tools lends them to pedagogical applications and rapid exploration of toy models.
-This tool aims to simplify the investigation of dynamics of biophysically-realistic network models, facilitate collaborative modeling, and complement other tools being developed in the neuroinformatics community.
+This tool aims to simplify the investigation of dynamics of biophysically-realistic network models, facilitate collaborative modeling, and complement other tools being developed in the computational neuroscience community.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % 
@@
 % 
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-"xolotl" is designed to be easy-to-use without sacrificing speed.
-The software is built in "MATLAB" due to its popularity among neuroscientists for pedagogy and research.
-"xolotl" capitalizes on "MATLAB"'s straightforward structure array syntax to permit rapid prototyping and experimentation, especially for small neuronal networks of complex models.
-Parameters of conductances, cellular compartments, and simulations may all be edited in the structure before any calls to integration functions.
-While the underlying code is written in "C++" for speed and memory optimization, symbolic manipulations can be readily performed in "MATLAB" without ever touching the foundational code.
+"xolotl" is designed to be easy-to-use and richly featured and that is fast enough to use in everyday research.
+Our focus was on designing a simulator of conductance-based neurons and networks of these neurons; and simulating arbitrary dynamical systems [cite DynaSim and others] is therefore beyond the scope of this software.
+The software is designed to be used from within "MATLAB", a scientific programming language popular amongst neuroscientists and engineers for pedagogy and research.
+Our goal is to make "xolotl" completely usable entirely within MATLAB, and all parameters of models can be be inspected and edited before running the model.
+Several features of the simulator are designed to be easily extensible, and the simulator appears in the "MATLAB" workspace as a first-class native object, and is fully compatible with the large library of toolboxes that a scientific programming environment like "MATLAB" provides.
 
@@
 (F-G) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
-(H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables.}
-	\label{fig:figurehh}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_clamp}
-	\caption{Simulating a voltage-clamp experiment.
-(A) Cartoon of a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998} with experimentally-fixed voltage.
-(B) Structure of \texttt{xolotl} object in A.
-(C) Code snippet depicting integration under voltage clamp.
-(D-E) Current response to steps in voltage from a holding potential of $V_m = -60$ mV.
-(F) Current-voltage relation of the steady-state current ($t = 400$ ms) indicating a reversal potential of $E = -80$ mV and no inactivation.
-(G) Conductance-voltage relation at steady-state takes the form of a sigmoid.
-(H) Sigmoids $m$ fit to the model as $m^n$ data indicating that $n=4$ is the best fit.
-(I) Goodness of fit vs.
-exponent $n$, suggesting $n=4$ as the best fit to the data.}
-	\label{fig:figureclamp}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
-	\caption{Simulating a network of conductance-based model neurons.
-(A) Diagram of a network model of the pyloric rhythm in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004).
-(B) Each neuron is modeled as a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
-(C) \texttt{xolotl} implements conductances as fields of compartments and synapses as connections between compartments.
-(D-F) Simulated voltage trace of a model network for the three compartments.
-(G) Time series of synaptic currents in the simulated network can be obtained from the integration.}
-	\label{fig:figurenetwork}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_integral_control}
-	\caption{Simulating neurons under homeostatic regulation.
-(A) Cartoon of a model neuron (\cite{liuModelNeuronActivitydependent1998} with integral control (\cite{olearyCorrelationsIonChannel2013}.
-(B) Hierarchical structure of a neuronal network considers controllers as components of compartments which act on conductances.
-(C) \texttt{xolotl} implements controllers as properties of conductances and synapses.
-(D) Calcium sensors change maximal conductances to move a neuron from quiescence to a bursting state.
-(E) Voltage trace shows regular bursting activity after integral control.}
-	\label{fig:figureintegralcontrol}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/puppeteer_screenshot}
-	\caption{Using the GUI to manipulate neuron parameters.
-(A) Real-time output of the \texttt{plot} function displaying voltage (colored) and intracellular calcium (black) traces of a bursting neuron model (\cite{prinzAlternativeHandtuningConductancebased2003, prinzSimilarNetworkActivity2004}.
-Colors indicate the dominant current.
-(B) Sliders control the maximal conductances, which updates on the figure.}
-	\label{fig:puppeteerscreenshot}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_benchmark}
-	\caption{\texttt{xolotl} benchmarked against \texttt{DynaSim} and \texttt{NEURON}.
-(A) Ratio of 'simulated' time to runtime (speed factor) and accuracy, measured by spike train coincidence plotted against decreasing time-resolution.
-(B) Speed factor for models at increasing simulation times.
-(C) Speed factor over number of compartments.}
-	\label{fig:figurebenchmark}
-\end{figure}
-
-\end{document}
+(H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating var

### 2018-06-05 paper/paper.tex
done with paper text

@@
 %\def\corrAddress{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
-\def\corrAddress{Volen Center and Biology Department}
+\def\corrAddress{..}
 \def\corrEmail{\texttt{srinivasgs@brandeis.edu}}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
 %
@@
 %
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 "xolotl" is designed to be easy-to-use and richly featured and that is fast enough to use in everyday research.
-Our focus was on designing a simulator of conductance-based neurons and networks of these neurons; and simulating arbitrary dynamical systems [cite DynaSim and others] is therefore beyond the scope of this software.
+Our focus was on designing a simulator of conductance-based neurons and networks of these neurons; and simulating arbitrary dynamical systems is therefore beyond the scope of this software.
 The software is designed to be used from within "MATLAB", a scientific programming language popular amongst neuroscientists and engineers for pedagogy and research.
@@
 
-"MATLAB" provides a high level programmatic and graphical interface for implementing, manipulating, and visualizing models without sacrificing the enhancements of the underlying "C++" code.
-
-\paragraph{Library of network components.} "xolotl" comes packaged with hundreds of pre-existing components (viz.
-compartments, synapses, conductances, and controllers) which greatly simplify the task of constructing model neurons.
-Compartments represent cylindrical sections of membrane with a membrane potential.
-Synapses connect two compartments together by introducing a current in the post-synaptic neuron that depends on the state of the presynaptic neuron.
+\paragraph{A rich library of network components.} "xolotl" comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and controllers) which are used to construct model neurons and networks.
+Compartments represent cylindrical sections of membrane with a single membrane potential, and can represent either entire cells or parts of cells.
+Synapses connect two compartments together by introducing a current in the post-synaptic neuron that depends on the voltage of the presynaptic neuron.
 Conductances represent populations of ion channels in a compartment that produce transmembrane currents.
-Controllers integrate a signal from a conductance or synapse and change the properties of the conductance or synapse in response.
+Controllers represent homeostatic mechanisms that integrate a Calcium-dependent signal from the compartment they are in and  change the properties of a conductance or synapse in response.
 
 \paragraph{Object-oriented.}
-Models are specified by adding compartments and synapses to the "xolotl" object.
-Conductances are added to compartments and controllers can be added to conductances or synapses.
-This modular structure recapitulates the biophysics of the Hodgkin-Huxley formalism and obviates the need to explicitly write out equations, which in "xolotl" are contained within "C++" header files.
+"xolotl" is designed to mirror the nested and hierarchical structure of networks and neurons.
+Biological neuronal networks consist of neurons that are connected to each other with synapses.
+Each of these neurons contains within it a set of conductances and synaptic currents that affect its electrical behavior.
+Similarly, a "xolotl" model can contain a set of compartments that can represent individual neurons.
+Each compartment can contain a set of conductances.
+All objects in the "xolotl" model tree (compartments, conductances, etc.) are bonafide "MATLAB" objects that have their own type, properties and associated methods.
+This object-oriented programming paradigm naturally represents the hierarchical structure of biological neurons and networks.
 
@@
 "xolotl" simulations are run entirely from "C++" executables; the inputs and outputs can be represented entirely in "MATLAB".
-
-\paragraph{Fast.}
-"xolotl" is fast because all time-intensive code is written in "C++".
-While automated "C++" transpiling from "MATLAB" using the proprietary "codegen" can drastically improve performance over loops through strong typing and memory pre-allocation, supervenience of "MATLAB" over "C++" prevents efficient use of features, such as passing by reference and object-oriented programming.
-"C++" provides speed improvements beyond the benefits of translating "MATLAB" features into statically-typed code.
-For this reason, "cpplab" has been designed to provide an interface for constructing, transpiling, and compiling "C++" code to be called from within "MATLAB".
 
@@
 
-The focus on ease-of-use and fast simulations means some features were intentionally neglected in the streamlining process.
+The focus on ease-of-use and fast simulations means some features were intentionally neglected in the design process.
 
-\paragraph{Limited to conductance-based models.} "xolotl" has been developed specifically for conductance-based models. It does not currently support rate- or current-based models.
+\paragraph{Limited to conductance-based models.} "xolotl" has been developed specifically for conductance-based models.
+It does not currently support rate- or current-based models, or arbitrary dynamical systems.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-In this section, we describe several uses of "xolotl" for common situations.
-
-%In "MATLAB", users create a "xolotl" object and populate it with "cpplab"-generated objects which describe compartments, conductances, synapses, and controllers.
-The model is integrated with the "integrate" function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.
-%
-%
-%
-%\subsection{Constructing Networks}
-%
-%The "add" function creates a compartment, conductance, or controller and affixes it as a field in the "xolotl" network structure.
-This function generates a "MATLAB struct" that faithfully represents the underlying "C++" code.
-Compartments add to the "xolotl" object and conductances add to compartments.
-Specific properties can be specified using key-value pair arguments (\eg Figure \ref{fig:figurehh}A).
-%
-%\paragraph{Compartments.}
-%A model neuron consists of one or more compartments, each representing a section of membrane with capacitance and surface area.
-Isopotential models require one compartment, whereas models with multiple neurons, units, or non-trivial morphology require multiple compartments.
-%
-%%\begin{table}
-%%	\begin{center}
-%%		\begin{tabular}{ l | l | c }
-%%			\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
-%%			Membrane potential & "V" & $mV$ \\
-%%			Intracellular calcium concentration & "Ca" & $\mu M$ \\
-%%			Membrane capacitance & "Cm" & $nF/mm^2$ \\
-%%			Volume & "vol" & $mm^3$ \\
-%%			Surface area & "A" & $mm^2$ \\
-%%			Radius of compartment & "radius" & $mm$ \\
-%%			Length of compartment & "len" & $mm$ \\
-%%			Extracellular calcium steady-state & "Ca_out" & $\mu M$ \\
-%%			Intracellular calcium steady-state & "Ca_in" & $\mu M$ \\
-%%			Calcium buffering shell thickness & "shell_thickness" & $\mu M$ \\
-%%			Calcium current to concentration factor & "phi" & ??? \\
-%%			Calcium buffering time constant & "tau_Ca" & $ms$ \\
-%%			Average calcium concentration & "Ca_average" & $\mu M$ \\
-%%			Target calcium concentration & "Ca_target" & $\mu M$ \\
-%%		\end{tabular}
-%%	\end{center}
-%%	\label{tab:compartmentprops}
-%%\end{table}
-%
-%\paragraph{Synapses.}
-%"xolotl" provides some features for generating complex models.
-Synapses can be added with the \texttt{connect} function.
-Synapses possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses.
-To create axons or transport chains, the \texttt{slice} function splits a compartment into $n$ discrete segments and adds these compartments to the network connected by electrical synapses.
-%
-%%\begin{center}
-%%	\begin{tabular}{ l | l | c }
-%%		\label{tab:synapseprops}
-%%		\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
-%%		Maximal conductance & "gbar" & $\mu S /mm^2$ \\
-%%		Reversal potential & "E" & $mV$ \\
-%%		Synaptic gating variable & "s" &  \\
-%%	\end{tabular}
-%%\end{center}
-%
-%\paragraph{Conductances and controllers.}
-%
-%All conductances contain fields for maximal conductance and reversal potential.
-Conductances with activation and inactivation variables include them as $m$ and $h$ respectively.
-Gating functions and their respective time constants are contained within the conductance header file.
-"xolotl" comes packaged with conductances from several dozen papers (\cite{dethierPositiveFeedbackCellular2015}.
-%
-%%\begin{center}
-%%	\begin{tabular}{ l | l | c }
-%%		\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
-%%		Maximal conductance & "gbar" & $\mu S /mm^2$ \\
-%%		Reversal potential & "E" & $mV$ \\
-%%		Activation gating variable & "m" &  \\
-%%		Inactivation gating variable & "h" &  \\
-%%	\end{tabular}
-%%\end{center}
-%
-%\paragraph{Creating custom \texttt{cpplab} objects.}
-%
-%"xolotl" contains template header files for producing custom conductances. The template contains instructions on how to design novel conductances with arbitrary specifications.
-%
-%\subsection{Simulation}
-%
-%Models are simulated in "xolotl" with the \texttt{integrate} function which outputs as time series the membrane potentials, intracellular calcium concentrations, controller states, intrinsic currents, and synaptic currents.
-The \texttt{integrate} function also accepts an argument which specifies injected current or clamped voltage.
-%
-%\paragraph{Numerical integration.}
-%"xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001, butcherNumericalDifferentialEquation2016}).
-These defaults provide a mix of speed, accuracy, and stability, and are built into the "cpplab" header files.
-Custom "cpplab" header files can be customized with any iterative integration method.
-The simulation time-resolution can be specified to target arbitrary precision, and an output time step can be selected to support automatic down-sampling for memory considerations.
-%
-%\paragraph{`Closed-loop' vs. `open-loop.'}
-%Simulations can be run in `closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or `open-loop' mode which begins simulation with the current network state.
-%
-%\subsection{Additional Features}
-%
-%\paragraph{Using the graphical interface to manipulate parameters.}
-%
-%"xolotl" comes packaged with a graphical user interface for visualizing parameter changes in real-time.
-The \texttt{manipulate} function opens the GUI, which displays a figure plotting the membrane potential and intracellular calcium concentration of all compartments as time series, and a dialog box with customizable sliders for all parameters of the model, much like the "Manipulate" function in Wolfram "Mathematica".
-Moving the sliders integrates the model in `open-loop` mode with the new parameters.
-Closing the GUI saves the network state of the model to the "xolotl" object.
-This is particularly helpful for rapid prototyping of models.
-%
-%\paragraph{Optimizing parameters.}
-%"xolotl" can use the Global Optimization toolbox for "MATLAB" to optimize any accessible "xolotl" parameters.
-The toolbox is algorithm-agnostic and accepts any function in "MATLAB" with a scalar first output as the objective function.
-Simulations run on multi-core processors or high-performance computing clusters using the Parallel Computing toolbox.
+In this section, we illustrate how "xolotl" can be used to generate a variety of models and simulate them.
+These examples have been chosen to illustrate various features of "xolotl", and are intended to serve as a template that researchers and educators can use for their own needs.
 
@@
 "manipulate" grants slider control over all "xolotl" parameters by default, but specific ones can be selected by passing them as arguments.
+
+%In "MATLAB", users create a "xolotl" object and populate it with "cpplab"-generated objects which describe compartments, conductances, synapses, and controllers.
+The model is integrated with the "integrate" function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.
+%
+%
+%
+%\subsection{Constructing Networks}
+%
+%The "add" function creates a compartment, conductance, or controller and affixes it as a field in the "xolotl" network structure.
+This function generates a "MATLAB struct" that faithfully represents the underlying "C++" code.
+Compartments add to the "xolotl" object and conductances add to compartments.
+Specific properties can be specified using key-value pair arguments (\eg Figure \ref{fig:figurehh}A).
+%
+%\paragraph{Compartments.}
+%A model neuron consists of one or more compartments, each representing a section of membrane with capacitance and surface area.
+Isopotential models require one compartment, whereas models with multiple neurons, units, or non-trivial morphology require multiple compartments.
+%
+%%\begin{table}
+%%	\begin{center}
+%%		\begin{tabular}{ l | l | c }
+%%			\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
+%%			Membrane potential & "V" & $mV$ \\
+%%			Intracellular calcium concentration & "Ca" & $\mu M$ \\
+%%			Membrane capacitance & "Cm" & $nF/mm^2$ \\
+%%			Volume & "vol" & $mm^3$ \\
+%%			Surface area & "A" & $mm^2$ \\
+%%			Radius of compartment & "radius" & $mm$ \\
+%%			Length of compartment & "len" & $mm$ \\
+%%			Extracellular calcium steady-state & "Ca_out" & $\mu M$ \\
+%%			Intracellular calcium steady-state & "Ca_in" & $\mu M$ \\
+%%			Calcium buffering shell thickness & "shell_thickness" & $\mu M$ \\
+%%			Calcium current to concentration factor & "phi" & ??? \\
+%%			Calcium buffering time constant & "tau_Ca" & $ms$ \\
+%%			Average calcium concentration & "Ca_average" & $\mu M$ \\
+%%			Target calcium concentration & "Ca_target" & $\mu M$ \\
+%%		\end{tabular}
+%%	\end{center}
+%%	\label{tab:compartmentprops}
+%%\end{table}
+%
+%\paragraph{Synapses.}
+%"xolotl" provides some features for generating complex models.
+Synapses can be added with the \texttt{connect} function.
+Synapses possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses.
+To create axons or transport chains, the \texttt{slice} function splits a compartment into $n$ discrete segments and adds these compartments to the network connected by electrical synapses.
+%
+%%\begin{center}
+%%	\begin{tabular}{ l | l | c }
+%%		\label{tab:synapseprops}
+%%		\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
+%%		Maximal conductance & "gbar" & $\mu S /mm^2$ \\
+%%		Reversal potential & "E" & $mV$ \\
+%%		Synaptic gating variable & "s" &  \\
+%%	\end{tabular}
+%%\end{center}
+%
+%\paragraph{Conductances and controllers.}
+%
+%All conductances contain fields for maximal conductance and reversal potential.
+Conductances with activation and inactivation variables include them as $m$ and $h$ respectively.
+Gating functions and their respective time constants are contained within the conductance header file.
+"xolotl" comes packaged with conductances from several dozen papers (\cite{dethierPositiveFeedbackCellular2015}.
+%
+%%\begin{center}
+%%	\begin{tabular}{ l | l | c }
+%%		\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
+%%		Maximal conductance & "gbar" & $\mu S /mm^2$ \\
+%%		Reversal potential & "E" & $mV$ \\
+%%		Activation gating variable & "m" &  \\
+%%		Inactivation gating variable & "h" &  \\
+%%	\end{tabular}
+%%\end{center}
+%
+%\paragraph{Creating custom \texttt{cpplab} objects.}
+%
+%"xolotl" contains template header files for producing custom conductances. The template contains instructions on how to design novel conductances with arbitrary specifications.
+%
+%\subsection{Simulation}
+%
+%Models are simulated in "xolotl" with the \texttt{integrate} function which outputs as time series the membrane potentials, intracellular calcium concentrations, controller states, intrinsic currents, and synaptic currents.
+The \texttt{integrate} function also accepts an argument which specifies injected current or clamped voltage.
+%
+%\paragraph{Numerical integration.}
+%"xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001, butcherNumericalDifferentialEquation2016}).
+These defaults provide a mix of speed, accuracy, and stability, and are built into the "cpplab" header files.
+Custom "cpplab" header files can be customized with any iterative integration method.
+The simulation time-resolution can be specified to target arbitrary precision, and an output time step can be selected to support automatic down-sampling for memory considerations.
+%
+%\paragraph{`Closed-loop' vs. `open-loop.'}
+%Simulations can be run in `closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or `open-loop' mode which begins simulation with the current network state.
+%
+%\subsection{Additional Features}
+%
+%\paragraph{Using the graphical interface to manipulate parameters.}
+%
+%"xolotl" comes packaged with a graphical user interface for visualizing parameter changes in real-time.
+The \texttt{manipulate} function opens the GUI, which displays a figure plotting the membrane potential and intracellular calcium concentration of all compartments as time series, and a dialog box with customizable sliders for all parameters of the model, much like the "Manipulate" function in Wolfram "Mathematica".
+Moving the sliders integrates the model in `open-loop` mode with the new parameters.
+Closing the GUI saves the network state of the model to the "xolotl" object.
+This is particularly helpful for rapid prototyping of models.
+%
+%\paragraph{Optimizing parameters.}
+%"xolotl" can use the Global Optimization toolbox for "MATLAB" to optimize any accessible "xolotl" parameters.
+The toolbox is algorithm-agnostic and accepts any function in "MATLAB" with a scalar first output as the objective function.
+Simulations run on multi-core processors or high-performance computing clusters using the Parallel Computing toolbox.
 
@@
 (F-G) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
-(H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating var
+(H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables.}
+	\label{fig:figurehh}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_clamp}
+	\caption{Simulating a voltage-clamp experiment.
+(A) Cartoon of a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998} with experimentally-fixed voltage.
+(B) Structure of \texttt{xolotl} object in A.
+(C) Code snippet depicting integration under voltage clamp.
+(D-E) Current response to steps in voltage from a holding potential of $V_m = -60$ mV.
+(F) Current-voltage relation of the steady-state current ($t = 400$ ms) indicating a reversal potential of $E = -80$ mV and no inactivation.
+(G) Conductance-voltage relation at steady-state takes the form of a sigmoid.
+(H) Sigmoids $m$ fit to the model as $m^n$ data indicating that $n=4$ is the best fit.
+(I) Goodness of fit vs.
+exponent $n$, suggesting $n=4$ as the best fit to the data.}
+	\label{fig:figureclamp}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
+	\caption{Simulating a network of conductance-based model neurons.
+(A) Diagram of a network model of the pyloric rhythm in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004).
+(B) Each neuron is modeled as a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
+(C) \texttt{xolotl} implements conductances as fields of compartments and synapses as connections between compartments.
+(D-F) Simulated voltage trace of a model network for the three compartments.
+(G) Time series of synaptic currents in the simulated network can be obtained from the integration.}
+	\label{fig:figurenetwork}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_integral_control}
+	\caption{Simulating neurons under homeostatic regulation.
+(A) Cartoon of a model neuron (\cite{liuModelNeuronActivitydependent1998} with integral control (\cite{olearyCorrelationsIonChannel2013}.
+(B) Hierarchical structure of a neuronal network considers controllers as components of compartments which act on conductances.
+(C) \texttt{xolotl} implements controllers as properties of conductances and synapses.
+(D) Calcium sensors change maximal conductances to move a neuron from quiescence to a bursting state.
+(E) Voltage trace shows regular bursting activity after integral control.}
+	\label{fig:figureintegralcontrol}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/puppeteer_screenshot}
+	\caption{Using the GUI to manipulate neuron parameters.
+(A) Real-time output of the \texttt{plot} function displaying voltage (colored) and intracellular calcium (black) traces of a bursting neuron model (\cite{prinzAlternativeHandtuningConductancebased2003, prinzSimilarNetworkActivity2004}.
+Colors indicate the dominant current.
+(B) Sliders control the maximal conductances, which updates on the figure.}
+	\label{fig:puppeteerscreenshot}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_benchmark}
+	\caption{\texttt{xolotl} benchmarked against \texttt{DynaSim} and \texttt{NEURON}.
+(A) Ratio of 'simulated' time to runtime (speed factor) and accuracy, measured by spike train coincidence plotted against decreasing time-resolution.
+(B) Speed factor for models at increasing simulation times.
+(C) Speed factor over number of compartments.}
+	\label{fig:figurebenchmark}
+\end{figure}
+
+\end{document}

### 2018-06-05 paper/paper.tex
Fix bibliography error

@@
 \usepackage[english]{babel}
-\usepackage[backend=biber, bibencoding=utf8, style=authoryear, citestyle=authoryear-comp, date=year, maxbibnames=6, minbibnames=6, url=false, isbn=false, autocite=inline]{biblatex}
+\usepackage{csquotes}
+\usepackage[
+backend=biber, 
+bibencoding=utf8, 
+style=authoryear, 
+date=year, 
+maxbibnames=6, 
+minbibnames=6, 
+url=false, 
+isbn=false, 
+eprint=false,
+uniquename = false, 
+autocite=inline]{biblatex}
 \addbibresource{Bibliography.bib}

### 2018-06-08 paper/paper.tex
fig1 done

@@
 	\caption{\texttt{xolotl} can quickly set up and simulate conductance-based models.
-(A) Cartoon of a Hodgkin-Huxley single-compartment neuron model with fast sodium, delayed rectifier, and leak currents.
-(B) Code snippet in \texttt{MATLAB} used to implement D, F-I.
-(C) \texttt{xolotl} schematic displayed in the \texttt{MATLAB} command prompt.
-(D) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current.
-Colors indicate the dominant current (gold is fast sodium, blue is delayed rectifier, red is leak).
-(E) Firing rate-input relation displaying firing rate as a function of injected current current.
-(F-G) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
-(H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables.}
+Schematic representation of a electrotonically-neuron with three populations of ion channels (colored rectangles).
+In the simulator, the soma is represented using ab object of class $compartment$ and the populations of ion channels are represented by $conductance$ objects contained within the compartment.
+Code snippet used to set up this neuron model, inject current, integrate and plot the voltage, and show activation functions demonstrates how various several steps can be run in a few lines of code.
+(A) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current.
+Colors indicate the dominant current (gold is fast sodium (NaV), blue is delayed rectifier (Kd), red is Leak).
+(C) f-I curve of this neuron.
+(D-E) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
+(F-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables}
 	\label{fig:figurehh}

### 2018-06-08 figures/figure_HH.m

@@ -16,10 +16,11 @@ fig = figure('outerposition',[0 0 1200 600],'PaperUnits','points','PaperSize',[1
 c = lines;
 clear ax;
 
-% cartoon cell
+
 ax(1) = subplot(3,5,1);
-% xolotl structure
+ax(1).Visible = 'off';
 ax(2) = subplot(3,5,6);
+ax(2).Visible = 'off';
 % xolotl readout from MATLAB
 ax(3) = subplot(3,5,11);
 % voltage trace
@@ -35,21 +36,21 @@ end
 
 %% Make Cartoon Cell
 
-image(ax(1), imread('paper-figs-0.png'))
-axis(ax(1), 'off');
-ax(1).Tag = 'cartoon';
+image(ax(3), imread('fig_HH.png'))
+axis(ax(3), 'off');
+ax(3).Tag = 'cartoon';
 
-%% Make Xolotl Structure
+% %% Make Xolotl Structure
 
-image(ax(2), imread('figure_HH_code_snippet.png'))
-axis(ax(2), 'off')
-ax(2).Tag = 'code_snippet';
+% image(ax(2), imread('figure_HH_code_snippet.png'))
+% axis(ax(2), 'off')
+% ax(2).Tag = 'code_snippet';
 
-%% Make Xolotl Readout from MATLAB
+% %% Make Xolotl Readout from MATLAB
 
-image(ax(3), imread('figure_HH_xolotl_printout.png'))
-axis(ax(3), 'off')
-ax(3).Tag = 'xolotl_printout';
+% image(ax(3), imread('figure_HH_xolotl_printout.png'))
+% axis(ax(3), 'off')
+% ax(3).Tag = 'xolotl_printout';
 
 %% Make Voltage Trac
 
@@ -98,8 +99,8 @@ end
 
 % plot on the correct axes
 plot(ax(6), all_I_ext, all_f, '-k')
-xlabel(ax(6), 'applied current (nA)')
-ylabel(ax(6), 'firing rate (Hz)')
+xlabel(ax(6), 'Injected current (nA)')
+ylabel(ax(6), 'Firing rate (Hz)')
 set(ax(6), 'XLim', [min(all_I_ext)*1.1 max(all_I_ext)*1.05], 'XTick', [0 0.5 1])
 
 % set up tags
@@ -187,11 +188,11 @@ end
 pos = [...
   0.0100    0.7093    0.1174    0.2157;
   0.0100    0.4096    0.1174    0.2157;
[…]

### 2018-06-08 paper/paper.tex
changes to result 1 text

@@
 \def\Authors{Srinivas Gorur-Shandilya\,$^{1*\dagger}$, Alec Hoyland\,$^{1\dagger}$, and Eve Marder$^{1}$}
-\def\Address{$^{1}$Volen Center and Biology Department \\ Brandeis University \\ Waltham, MA 02454 \\ USA
+\def\Address{$^{1}$Volen Center and Biology Department \\ Brandeis University \\ Waltham, MA 02453 \\ USA
 \vspace{0.25cm}
@@
 %\def\corrAddress{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
-\def\corrAddress{..}
+\def\corrAddress{Volen Center for Complex Systems, Brandeis University, Waltham MA 02453}
 \def\corrEmail{\texttt{srinivasgs@brandeis.edu}}
@@
 
-\title[xolotl: neuronal simulator]{Xolotl: An Intuitive and Approachable Neuron \& Network Simulator in MATLAB}
+\title[xolotl: neuron \& network simulator]{Xolotl: An Intuitive and Approachable Neuron \& Network Simulator in MATLAB}
 \author[\firstAuthorLast ]{\Authors}
@@
 
-The seminal Hodgkin-Huxley model of action potentials in the squid giant axon (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}) contains a fast inactivating sodium conductance ("NaV"), a non-inactivating delayed rectifier ("Kd"), and a passive leak current (Figure \ref{fig:figurehh}A).
-This model was constructed using conductances from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}).
-A compartment, "HH", with membrane capacitance ("Cm") and surface area ("A") is specified by Figure \ref{fig:figurehh}B.
-Figure \ref{fig:figurehh}C shows the "MATLAB" command window after invoking the "xolotl" object \texttt{x}, displaying the hierarchical structure inherent in conductance-based treatments of neurodynamics.
+The axon of the giant squid contains contains a fast inactivating sodium conductance ("NaV"), a slower non-inactivating potassium conductance ("Kd"), and a passive leak current.
+Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}).
+The Hodgkin-Huxley model generates action potentials, and, as one of the simplest models of excitable neural mebranes, often serves as a the first model introduced in pedagogical literature (cite some textbooks).
 
-In the absence of applied positive current, the model is quiescent.
-When 0.2 nA is injected, the model tonically spikes (Figure \ref{fig:figurehh}D).
-The \texttt{integrate} function takes the applied current as an argument.
-The "plot" function generates voltage and intracellular calcium traces, where the voltage trace is colored by the dominant current (Figure \ref{fig:figurehh}E).
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_HH}
+	\caption{\texttt{xolotl} can quickly set up and simulate conductance-based models.
+Schematic representation of a electrotonically-neuron with three populations of ion channels (colored rectangles).
+In the simulator, the soma is represented using ab object of class $compartment$ and the populations of ion channels are represented by $conductance$ objects contained within the compartment.
+Code snippet used to set up this neuron model, inject current, integrate and plot the voltage, and show activation functions demonstrates how various several steps can be run in a few lines of code.
+(A) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current.
+Colors indicate the dominant current (gold is fast sodium (NaV), blue is delayed rectifier (Kd), red is Leak).
+(C) f-I curve of this neuron.
+(D-E) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
+(F-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables}
+	\label{fig:figurehh}
+\end{figure}
+
+In this section, we demonstrate how to simulate a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model, and gain insight into the underlying biophysical mechanism.
+The schematic in Figure \ref{fig:figurehh} shows how the hierarchical structure of the ion channel populations within the soma is represented in the simulator, and how to create a compartment (here, labelled "HH") and add three populations of ion channels to it.
+These models of ion channels were obtained from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator.
+Adding an injected current and calling the built-in "plot" function plots the time series of membrane voltage.
+In the absence of injected current, the model is quiescent.
+When 0.2 nA is injected, the model tonically spikes (Figure \ref{fig:figurehh}A-B).
+The \texttt{plot} function generates a voltage trace, which is colored by the dominant current (Figure \ref{fig:figurehh}A).
 If the membrane potential is increasing, the strongest instantaneous inward current colors the trace.
 Conversely, if the membrane potential is decreasing, the strongest outward current colors the trace instead.
-Figure \ref{fig:figurehh}F-I display the results of the \texttt{show} function which plots the activation and inactivation steady-states and the voltage-dependent time constants of conductance gating variables.
+These built-in features reveal that the dominant current during the upswing of every action potential is the sodium current, and the dominant current immediately after the peak of the action potential is the potassium current.
+This feature could be useful in quickly understanding the contributions of a number of ion channel types in a complex voltage trace from a more complicated neuron model.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The model can be integrated for various amplitudes of injected current to determine its f-I curve, or the relationship between injected current and its firing rate (cite).
+Figure \ref{fig:figurehh}C shows the f-I curve of this model, obtained by repeated integration of the model.
+Finally, the built-in \texttt{show} function can plot activation (m) and inactivation (h) curves and voltage-dependent timescales of any channel type in the simulator (Figure \ref{fig:figurehh}D-G).
+These plots revleal that activation kinetics of the "NaV" channels are much faster than that of the "Kd" channels (Figure \ref{fig:figurehh}F), which facilitates the transient depolarization in an action potential.
+In summary, the simulator allows the user to construct and integrate this model in a few lines of code, and provides a rich visualization of the dynamics of the model.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 
-SG-S designed and implemented the core of the "xolotl" toolbox.
-AH contributed to the code base, created the online user documentation.
-SG-S and AH wrote the manuscript.
-EM supervised the project.
-All authors reviewed the paper.
+SG-S & AH designed and wrote the software, and wrote the manuscript. AH created the online user documentation. EM supervised the project. All authors reviewed the paper.
 
@@
 %		FIGURES
+%   (because it's 2018 and we still have this dumb system where all the figures are at the end. why? who knows
+%   why can't they be in a more sensible place? it's not like we're not typesetting this ourselves. why do people
+%   go out of their way to make papers unreadable? who actually likes the figures and the captions at the end, 
+%   separated into their own pages?)
 %
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\section*{Figure captions}
+%\section*{Figure captions}
 
 \FloatBarrier
-
-\begin{figure}
-	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_HH}
-	\caption{\texttt{xolotl} can quickly set up and simulate conductance-based models.
-Schematic representation of a electrotonically-neuron with three populations of ion channels (colored rectangles).
-In the simulator, the soma is represented using ab object of class $compartment$ and the populations of ion channels are represented by $conductance$ objects contained within the compartment.
-Code snippet used to set up this neuron model, inject current, integrate and plot the voltage, and show activation functions demonstrates how various several steps can be run in a few lines of code.
-(A) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current.
-Colors indicate the dominant current (gold is fast sodium (NaV), blue is delayed rectifier (Kd), red is Leak).
-(C) f-I curve of this neuron.
-(D-E) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
-(F-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables}
-	\label{fig:figurehh}
-\end{figure}
 

### 2018-06-08 paper/paper.tex
started work on figure_network

@@
 
+- failure of averaging
+- complex interrelationships b/w parameters and dynamics
+- not clear what parameter does what
+- traditional approach: vary parameters, but for more than 2 D, how do you even visualize this? also slow. expensive to set up. 
+- need something where parameter variation can be linked to dynamics, easy to set up, easy to use.
+
 The "manipulate" function opens the GUI which permits visualization of changing parameters in real-time.
@@
 
-We envision that "xolotl" will be helpful in teaching students how to interpret cellular biophysics.
-The modular structure of "cpplab" and the graphical interface simplifies the process of manipulating and analyzing the properties of electrical excitability.
+We set out to design a neuron and network simulator that could be useful in the classroom setting, especially for students of computational students, while also being powerful, fast and extensible enough to be used for research.
+By using a novel architecture that .., we showed how to create and simulate (1) single-compartment models of Hogkin-Huxley like neurons (Figure \ref{fig:figurehh}), (2) voltage clamp experiments and recover activation functions of single channel types (Figure 2), (3) a network of neurons with multiple synapse types (Figure 3), (4) a multi-compartment neuron model where the voltage of every compartment is coupled, and (5) a neuron model where homeostatic mechanisms control the maximal conductance of multiple channel types.
 
-\subsection{Reproducibility}
+\subsection{Software use and reproducibility of research}
+
+- reproducibility crisis in computational science 
+- multiple sources: software being used is closed-source, researcher's code is not made available; obscure architecture makes improper use possible. 
+- how xolotl conributes to reoridubile science: (1) freely available, code for this paper available too for others to inspect and replicate (2) intutitive syntax, solid documentation.
+(3) architectural choices that ensure accountability: object hashing, parameter hashing possible.
 
@@
 
-\subsection{Applications of \texttt{cpplab}}
+\subsection{Applications of the \texttt{cpplab} architecture}
 
@@
 
-SG-S & AH designed and wrote the software, and wrote the manuscript. AH created the online user documentation. EM supervised the project. All authors reviewed the paper.
+SG-S \& AH designed and wrote the software, and wrote the manuscript. AH created the online user documentation. EM supervised the project. All authors reviewed the paper.
 

### 2018-06-08 figures/figure_network.m

@@ -2,35 +2,32 @@
 
 % set up xolotl object
 % conversion from Prinz to phi
-A = 0.0628;
-vol = A; % this can be anything, doesn't matter
-f = 14.96; % uM/nA
-tau_Ca = 200;
-phi = (2*f*96485*vol)/tau_Ca;
 
-channels = {'NaV','CaT','CaS','ACurrent','KCa','Kd','HCurrent'};
-prefix = 'prinz/';
-gbar(:,1) = [1000 25  60 500  50  1000 .1];
-gbar(:,2) = [1000 0   40 200  0   250  .5];
-gbar(:,3) = [1000 24  20 500  0   1250 .5];
-E =         [50   30  30 -80 -80 -80   -20];
+clear x
+
 
 x = xolotl;
 
-x.add('AB','compartment','Cm',10,'A',A,'vol',vol,'phi',phi,'Ca_out',3000,'Ca_in',0.05,'tau_Ca',tau_Ca);
-x.add('LP','compartment','Cm',10,'A',0.0628,'vol',vol,'phi',phi,'Ca_out',3000,'Ca_in',0.05,'tau_Ca',tau_Ca);
-x.add('PY','compartment','Cm',10,'A',A,'vol',vol,'phi',phi,'Ca_out',3000,'Ca_in',0.05,'tau_Ca',tau_Ca);
+C = {'NaV','CaT','CaS','ACurrent','KCa'...
+    ,'Kd','HCurrent'};
+g(:,3) = [1000 24  20 500  0   1250 .5];
+g(:,2) = [1000 0   40 200  0   250  .5];
+g(:,1) = [1000 25  60 500  50  1000 .1];
+
+x.add('AB','compartment','vol',.0628,'phi',906);
+x.add('LP','compartment','vol',.0628,'phi',906);
+x.add('PY','compartment','vol',.0628,'phi',906);
 
 compartments = x.find('compartment');
 for j = 1:length(compartments)
-	for i = 1:length(channels)
-		x.(compartments{j}).add([prefix channels{i}],'gbar',gbar(i,j),'E',E(i));
+	for i = 1:length(C)
+		x.(compartments{j}).add(...
+      ['prinz/' C{i}],'gbar',g(i,j));
 	end
 end
 
 x.LP.add('Leak','gbar',.3,'E',-50);
 x.PY.add('Leak','gbar',.1,'E',-50);
-
 
 % set up synapses as in Fig. 2e
[…]

### 2018-06-11 figures/figure_network.m
added network snippet

@@ -22,7 +22,7 @@ compartments = x.find('compartment');
 for j = 1:length(compartments)
 	for i = 1:length(C)
 		x.(compartments{j}).add(...
-      ['prinz/' C{i}],'gbar',g(i,j));
+	    ['prinz/' C{i}],'gbar',g(i,j));
 	end
 end
 
@@ -52,8 +52,10 @@ clear ax
 
 % cartoon cell
 ax(1) = subplot(3,5,1);
+ax(1).Visible = 'off';
 % xolotl structure
 ax(2) = subplot(3,5,6);
+ax(2).Visible = 'off';
 % xolotl printout
 ax(3) = subplot(3,5,11);
 % voltage trace
@@ -89,31 +91,31 @@ nComps      = length(nameComps);
 
 % integrate and obtain the current traces
 x.closed_loop = true;
+x.t_end = 4e3;
 x.integrate;
-[V, Ca, ~, currents, synaptic_currents]  = x.integrate;
+[V, Ca, ~, ~, synaptic_currents]  = x.integrate;
 time        = 1e-3 * x.dt * (1:length(V));
 
 % plot the voltage
 for ii = 1:nComps
-  plot(ax(ii+3), time, V(:,ii), 'k', 'LineWidth', 1)
-	set(ax(ii+3), 'XTickLabel', [])
+  plot(ax(ii+3), time, V(:,ii), 'k')
+	set(ax(ii+3), 'XTickLabel', [],'XLim',[0 max(time)],'YLim',[-80 50])
 	ylabel(ax(ii+3), ['V_{' nameComps{ii} '} (mV)'])
-	xlim(ax(ii+3), [0 max(time)]);
 end
 
 % plot the synaptic currents
 c = lines(10);
-plot(ax(7), time, synaptic_currents(:,1));
+plot(ax(7), time, synaptic_currents(:,1),'k');
 
 xlabel(ax(7), 'time (s)')
 ylabel(ax(7), 'I_{syn} (nA)')
-set(ax(7), 'YScale', 'log','YLim',[1e-1 1e2])
+set(ax(7), 'YScale', 'linear','YLim',[1e-1 1e2])
 xlim(ax(7), [0 max(time)]);
 
 
 %% Post-Processing
 
-prettyFig('fs', 12, 'lw', 1)
+prettyFig('fs', 16, 'plw', 1.5,'lw',1.5)
 
 % set the positions of the axes
 pos = [ ...

### 2018-06-11 figures/network_snippet.m

@@ -1,0 +1,12 @@
+
+x.add('AB','compartment','vol',.0628,'phi',906);
+x.add('LP','compartment','vol',.0628,'phi',906);
+x.add('PY','compartment','vol',.0628,'phi',906);
+
+% set up conductances...
+
+% set up synapses 
+x.connect('AB','LP','Chol','gbar',30);
+x.connect('LP','AB','Glut','gbar',30);
+
+[V,~,~,I_syn] = x.integrate;

### 2018-06-11 paper/paper.tex
added network snippet
went over introduction

@@
 \def\keyFont{\fontsize{8}{11}\helveticabold }
-\def\firstAuthorLast{Hoyland {et~al.}} 
+\def\firstAuthorLast{Gorur-Shandilya, Hoyland \& Marder} 
 \def\Authors{Srinivas Gorur-Shandilya\,$^{1*\dagger}$, Alec Hoyland\,$^{1\dagger}$, and Eve Marder$^{1}$}
-\def\Address{$^{1}$Volen Center and Biology Department \\ Brandeis University \\ Waltham, MA 02453 \\ USA
+\def\Address{$^{1}$Volen Center for Complex Systems and Biology Department \\ Brandeis University \\ Waltham, MA 02453 \\ USA
 \vspace{0.25cm}
@@
 }
-\def\corrAuthor{Srinivas Gorur-Shandilya \vspace{1mm}}
+\def\corrAuthor{Srinivas Gorur-Shandilya }
 %\def\corrAddress{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
@@
 
-	Information processing by neurons relies on the transmission and interaction of electrical signals that arise from the biophysics of ion channels and synapses.
-Electrophysiological characterization of these mechanisms has allowed for the construction of conductance-based models that can reproduce many features of neuronal and circuit behavior.
-However, working with conductance-based models continues to be a challenge due to their high dimensionality, hindering intuition of their dynamical features.
+	Information processing by neurons relies on the generation and transmission of electrical signals that arise from the biophysics of ion channels, synapses and other voltage-coupled components.
+Electrophysiological characterization of these mechanisms has allowed for the construction of conductance-based models that can reproduce several features of neuronal and circuit behavior.
+However, working with conductance-based models continues to be a challenge due to their high dimensionality and large number of parameters, hindering intuition of their dynamical features.
 Here, we present a neuron and network simulator built on a novel automatic type system that binds class templates written in "C++" to object-oriented code in "MATLAB".
-Our approach combines the speed and powerful features of "C++" code with the ease-of-use and rich libraries of scientific programming languages like "MATLAB".
-In our software, models are hierarchical, and components are named and searchable, permitting high-level programmatic control over all parameters.
-The simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter on model behavior.
+Our approach unites the speed and power of "C++" code with the ease-of-use and rich libraries of scientific programming languages like "MATLAB".
+In our software, models are hierarchical, with components that are named and searchable, permitting high-level programmatic control over all parameters.
+The simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately.
 The simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be easily extended to include arbitrary mechanisms and dynamics.
-Finally, the simulator is written in a modular fashion and has been released under a permissive free software license, enabling it to be integrated easily in third party applications.
+Finally, the simulator is written in a modular fashion and has been released under a permissive free software license, enabling it to be integrated easily into the workflow of many researchers.
 
 	\tiny
-	\keyFont{ \section{Keywords:} simulator, MATLAB, C++, conductance-based, neuron, network, pedagogy}
+	\keyFont{ \section{Keywords:} simulator, code:MATLAB, code:C++, conductance-based, neuron, network, pedagogy}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%
-%		INTRODUCTION & OVERVIEW
-%
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%  								%
+%		INTRODUCTION 			%
+% 		& OVERVIEW   			%
+%								%
+%								%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Nervous systems process and transmit information using electrically excitable membranes.
-Conductance-based models are powerful biophysical simplification of an electrically excitable cell (\cite{hodgkinQuantitativeDescriptionMembrane1952}).
-Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small-circuit networks (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}).
-Additionally, these models provide an approachable framework for understanding salient principles of neuroscience.
-However, several challenges remain in simulating biophysically-detailed conductance-based neuron models.
+Conductance-based models are a powerful biophysical simplification of an excitable compartment in a neuron (\cite{hodgkinQuantitativeDescriptionMembrane1952}).
+Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small circuits (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}).
+These models provide an approachable framework for understanding many salient principles of neuroscience, since they explicitly model cell membranes and ion channels as electrical components in a circuit.
+However, several challenges remain in working with biophysically-detailed conductance-based neuron models.
 First, these models can be high-dimensional with many nonlinear differential equations, with several parameters.
-Second, many or all equations are strongly coupled through the variables like the membrane potential, and in multi-compartment models, all membrane potentials and their equations are coupled to each other.
-Finally, the choice of programming language imposes tradeoffs in neuron and network simulators: simulators written in languages like "C", "C++", or "hoc" can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like "Python", "Julia" or "MATLAB".
+Second, many or all equations in these models can be strongly coupled through variables like the membrane potential.
+In multi-compartment models, membrane potentials in every compartment are coupled to each other.
+Finally, the choice of programming language used to implement the model imposes tradeoffs in designing and using neuron and network simulators: simulators written in languages like "C++", "FORTRAN" or "hoc" can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like "Python", "Julia" or "MATLAB".
 
 Two major approaches have dominated the design of neuron simulators.
-First, some simulators, such as "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}) use a custom language to specify modular network components.
+First, some simulators, such as "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}) are written in a fast compiled language like "C" and allow for the construction and simulation of neuron models using object-oriented paradigms.
 These simulators tend to perform fast computations with little overhead, but suffer from a steep learning curve.
-Wrapping "NEURON" in a more approachable language like "Python" and use of graphical interfaces mitigate these drawbacks, at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
-In contrast, many simulators have been designed to be used entirely in popular scientific programming languages.
+Wrapping these simulators in a more approachable language like "Python" and the use of graphical user interfaces (GUIs) mitigate these drawbacks, at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
+Another approach is to design simulators that can be used entirely in popular scientific programming languages.
 These simulators can be easier to use, and can benefit from inter-compatibility with other tools.
-Programs like "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}) allow the user to specify models using strings of equations that are specified in the scientific programming language, that can be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
-This approach permits considerable flexibility for simulating systems of differential equations, but the syntax tends to be verbose and the hierarchical nature of conductance-based neuron models is not naturally reflected in the constructed model.
-Neither approach permits tools that simultaneously maintain efficiency, ease-of-use, and clarity.
+Programs like "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}) allow the user to specify models using strings of equations that are specified in the scientific programming language, that can then be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
+This approach permits considerable flexibility for simulating systems of differential equations.
+Since models need to be translated between the two languages, the hierarchical nature of neuron models is not naturally encapsualted by these tools, and the syntax can be verbose.
+Neither approach facilitates the creation of tools that simultaneously maintain efficiency, ease-of-use, and clarity.
 
@@
 Models can be easily constructed from network building blocks in a few lines of "MATLAB" code using a hierarchical and intuitive syntax.
-Since model objects in the "MATLAB" workspace are automatically linked to the "C++" specification, configuring these objects in "MATLAB" transparently configures the underlying "C++" objects.
+Since models in the "MATLAB" workspace are automatically linked to models in the "C++" implementation, configuring these objects in "MATLAB" transparently configures the underlying "C++" objects.
 "xolotl" comes packaged with hundreds of conductances and synapses, built-in visualization functions, and a graphical user interface (GUI) for real-time manipulation of model parameters.
-Plotting of voltage, intracellular calcium, conductance gating functions, and time constants is provided by built-in "xolotl" methods.
-The GUI permits real-time tuning of any network parameters using numerical sliders in a graphical interface which displays the resultant membrane potential and intracellular calcium traces.
-The ease-of-use of these tools lends them to pedagogical applications and rapid exploration of toy models.
-This tool aims to simplify the investigation of dynamics of biophysically-realistic network models, facilitate collaborative modeling, and complement other tools being developed in the computational neuroscience community.
+Built-in methods allow the user to easily generate plots of voltage and intracellular calcium time traces, and activation functions of constituent conductances.
+The GUI permits real-time manipulation of any network parameter and an immediate visualization of the effect of that parameter on arbitrary dynamical features of the model.
+The ease-of-use of these tools lends them to pedagogical applications and the rapid prototyping of models.
+This tool aims to simplify the investigation of the dynamics of conductance-based network and neuron models, facilitate collaborative modeling, and complement other tools being developed in the computational neuroscience community.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-"xolotl" is designed to be easy-to-use and richly featured and that is fast enough to use in everyday research.
-Our focus was on designing a simulator of conductance-based neurons and networks of these neurons; and simulating arbitrary dynamical systems is therefore beyond the scope of this software.
-The software is designed to be used from within "MATLAB", a scientific programming language popular amongst neuroscientists and engineers for pedagogy and research.
-Our goal is to make "xolotl" completely usable entirely within MATLAB, and all parameters of models can be be inspected and edited before running the model.
-Several features of the simulator are designed to be easily extensible, and the simulator appears in the "MATLAB" workspace as a first-class native object, and is fully compatible with the large library of toolboxes that a scientific programming environment like "MATLAB" provides.
+"xolotl" is designed to be easy-to-use and richly featured while being fast enough to use in everyday research.
+Our focus was on designing a simulator of conductance-based neurons and networks of these neurons; simulating arbitrary dynamical systems is therefore beyond the scope of this software.
+Specifically, the software was designed to simulate models of the form
+
+\[ C_{i}\frac{dV_{i}}{dt}=-\sum_{j}I_{j} \]
+
+where \( C_{i} \) and  \( V_{i} \) are the capacitance and membrane potential of compartment  \( i \) and \( I_{j} \) is the current due to ion channel population \( j \) given by
+
+\[ I_{j}=\bar{g}_{j}m_{j}^{p}h_{j}^{q}(V-E_{j})A_{i} \]
+
+Here, \( \bar{g}_{j} \) is the maximal conductance, and \( E_{j} \) is the reversal potential of the ion channel population \( j \).
+\( A_{i} \) is the area of compartment \( i\) that contains these ion channels.
+\( m_{j} \) and \( h_{j} \) are activation and inactivation variables that change according to
+
+\begin{eqnarray*}
+\tau_{m}\frac{dm}{dt}=m_{\infty}-m & \mathrm{and} & \tau_{h}\frac{dh}{dt}=h_{\infty}-h
+\end{eqnarray*}
+
+Typically, \( \tau_{m} \), \( \tau_{h} \), \( m_{\infty} \), and \( h_{\infty} \) are functions of the membrane potential \( V_{i} \).
+The software uses integration schemes that have been specifically developed to solve equations of this form (cite D\&A, Hines, others), though other schemes can be used.
+
+This software is designed to be used from within "MATLAB", a scientific programming language common amongst neuroscientists and engineers for pedagogy and research.
+Our goal was to make "xolotl" completely usable entirely from within MATLAB, and fully scriptable so that the software can be used as a component of other packages and tools.
+All parameters of models and activation functions of any channel can be be accessed before running the model.
+Finally, we designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by changing a few lines in a template file.
+Models created using this simulator appear in the "MATLAB" workspace as native objects, and are fully compatible with the large library of toolboxes that "MATLAB" provides.
 
@@
 
-\paragraph{A rich library of network components.} "xolotl" comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and controllers) which are used to construct model neurons and networks.
-Compartments represent cylindrical sections of membrane with a single membrane potential, and can represent either entire cells or parts of cells.
-Synapses connect two compartments together by introducing a current in the post-synaptic neuron that depends on the voltage of the presynaptic neuron.
+\paragraph{A rich library of network components.} "xolotl" comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and controllers) that can be used as building blocks to construct model neurons and networks.
+Compartments represent sections of membrane with a single membrane potential, and can represent either entire cells or parts of cells.
 Conductances represent populations of ion channels in a compartment that produce transmembrane currents.
+Synapses connect two compartments together by introducing a current in the post-synaptic neuron that can depend on the voltage of the presynaptic neuron.
 Controllers represent homeostatic mechanisms that integrate a Calcium-dependent signal from the compartment they are in and  change the properties of a conductance or synapse in response.
@@
 "xolotl" is designed to mirror the nested and hierarchical structure of networks and neurons.
-Biological neuronal networks consist of neurons that are connected to each other with synapses.
-Each of these neurons contains within it a set of conductances and synaptic currents that affect its electrical behavior.
+Biological neuronal networks are made up of neurons that are connected to each other with synapses.
+Each of these neurons contains within it a set of conductances and synaptic currents that contribute to its electrical behavior.
+These conductances can be controlled by a feedback system that can homeostatically regulate their properties.
 Similarly, a "xolotl" model can contain a set of compartments that can represent individual neurons.
-Each compartment can contain a set of conductances.
+Each compartment can contain a arbitrary set of conductances, that may be regulated by a set of controllers.
 All objects in the "xolotl" model tree (compartments, conductances, etc.) are bonafide "MATLAB" objects that have their own type, properties and associated methods.
-This object-oriented programming paradigm naturally represents the hierarchical structure of biological neurons and networks.
+This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain neurons that contain populations of ion channels, and makes constructing and integrating these models easier.
 
 \paragraph{Automatic type system.}
-When a network component is added in "MATLAB", the "xolotl" structure is updated, and "cpplab" automatically constructs and compiles the "C++" from the requisite header files.
-"MATLAB" treats "cpplab" constructions as fully-typed first-class objects allowing for symbolic manipulation using only the high-level programming language and graphical interfaces.
-"xolotl" simulations are run entirely from "C++" executables; the inputs and outputs can be represented entirely in "MATLAB".
+Our approach to circumvent the tradeoff between high-performance but hard to use languages like "C++" and richly featured but slow languages like "MATLAB" was to construct an automatic type system that links object oriented code in "C++" to object oriented code in "MATLAB".
+This architecture lets us construct the core of the simulator in "C++", leveraging features of "C++" like pointers that do not exist in "MATLAB".
+A rudimentary way to make this "C++" code useable in "MATLAB" would be to re-write that code in "MATLAB" so that "MATLAB" objects can be bound their "C++" implementations.
+However, this approach is cumbersome, and prone to errors, since the same code needs to be written twice, in two languages.
+Instead, our automatic type system creates objects in "MATLAB" on-the-fly  from "C++" class specifications, obviating the need to rewrite code in "MATLAB" while preserving a tight coupling between objects in the "MATLAB" workspace and their "C++" implementation.
 
-\paragraph{Automatic compiling.}
-The user experience of "xolotl" can stay entirely in "MATLAB".
-Transpiling and compiling of "C++" is performed automatically.
-Cryptographic hashing of the "xolotl" object confirms that compiling occurs only when necessary, and that the correct binary is used during integration.
+\paragraph{Automatic hashing and compiling.}
+Since every model requires a compiled binary to run, a potential stumbling block is the problem of unambiguously linking a model to a binary executable.
+"xolotl" solves this problem by hashing the "C++" header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier.
+Compiled binaries are named with this hash, ensuring both that the correct binary is run to integrate the model, and that compilation occurs only as needed.
+This powerful feature enables the user experience to remain entirely within the "MATLAB" workspace, with compilation and selection of the correct binary occurring silently in the background.
 
@@
 
-The focus on ease-of-use and fast simulations means some features were intentionally neglected in the design process.
+Our focus on ease-of-use and fast simulations means some features were intentionally neglected in the design process.
 
@@
 
-\paragraph{Limited numerical integration strategies.} While the exponential Euler method performs well in neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}, it may be desirable to use other methods under certain conditions.
-"xolotl" does not currently support other integration schemes for its built-in conductances, nor does the software support error-sensitive variable step-sizes `out-of-the-box.'
+\paragraph{Limited numerical integration strategies.} Most components in the software are integrated using the exponential Euler method, which has been shown to perform well in neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}.
+However, it may be desirable to use other methods under certain conditions.
+It is possible to introduce new network components that implement other integration schemes, but that requires writing new "C++"  code, and is still restricted to integration schemes that assume a fixed step size.
 
-\paragraph{Inefficient tools for handling large networks.} While "xolotl" can integrate large networks ($>1000$ compartments), "xolotl" uses string-based comprehension for labeling compartments which is suited to descriptive naming, but prohibits vector operations over compartments.
+\paragraph{Inefficient tools for handling large networks.} Since every compartment and component in a "xolotl" model is $named$, rather than numbered, it is more suited towards simulating small but complex networks, rather than large but simple networks.
+While the software can integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task.
 
-\paragraph{New mechanisms require new \texttt{C++} code.} Adding new network components also requires writing some "C++" code.
+\paragraph{New mechanisms require new \texttt{C++} code.} Adding new network components requires writing new "C++" code.
 Adding a new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinMeasurementCurrentvoltageRelations1952, dayanTheoreticalNeuroscience2001}), requires creating a new "C++" header file, though this is generally trivial.
-Implementing a new integration scheme or "cpplab" class requires much more in-depth knowledge of "C++" and modification of the core code.
+Implementing a new integration scheme or component type requires much more in-depth knowledge of the underlying "C++" core code.
 
@@
 	\caption{\texttt{xolotl} can quickly set up and simulate conductance-based models.
-Schematic representation of a electrotonically-neuron with three populations of ion channels (colored rectangles).
-In the simulator, the soma is represented using ab object of class $compartment$ and the populations of ion channels are represented by $conductance$ objects contained within the compartment.
+Schematic representation of a electrotonically-compact neuron with three populations of ion channels (colored rectangles).
+In the simulator, the soma is represented using an object of class $compartment$ and the populations of ion channels are represented by $conductance$ objects contained within the compartment object.
 Code snippet used to set up this neuron model, inject current, integrate and plot the voltage, and show activation functions demonstrates how various several steps can be run in a few lines of code.
@@
 
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_clamp}
+	\caption{Simulating a voltage-clamp experiment.
+(A) Cartoon of a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998} with experimentally-fixed voltage.
+(B) Structure of \texttt{xolotl} object in A.
+(C) Code snippet depicting integration under voltage clamp.
+(D-E) Current response to steps in voltage from a holding potential of $V_m = -60$ mV.
+(F) Current-voltage relation of the steady-state current ($t = 400$ ms) indicating a reversal potential of $E = -80$ mV and no inactivation.
+(G) Conductance-voltage relation at steady-state takes the form of a sigmoid.
+(H) Sigmoids $m$ fit to the model as $m^n$ data indicating that $n=4$ is the best fit.
+(I) Goodness of fit vs.
+exponent $n$, suggesting $n=4$ as the best fit to the data.}
+	\label{fig:figureclamp}
+\end{figure}
+
 "xolotl" can recapitulate the results of voltage clamp experiments (\cite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}).
@@
 \subsection{Simulating Network Models}
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
+	\caption{Simulating a network of conductance-based model neurons.
+(A) Diagram of a network model of the pyloric rhythm in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004).
+(B) Each neuron is modeled as a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
+(C) \texttt{xolotl} implements conductances as fields of compartments and synapses as connections between compartments.
+(D-F) Simulated voltage trace of a model network for the three compartments.
+(G) Time series of synaptic currents in the simulated network can be obtained from the integration.}
+	\label{fig:figurenetwork}
+\end{figure}
 
@@
 	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_clamp}
-	\caption{Simulating a voltage-clamp experiment.
-(A) Cartoon of a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998} with experimentally-fixed voltage.
-(B) Structure of \texttt{xolotl} object in A.
-(C) Code snippet depicting integration under voltage clamp.
-(D-E) Current response to steps in voltage from a holding potential of $V_m = -60$ mV.
-(F) Current-voltage relation of the steady-state current ($t = 400$ ms) indicating a reversal potential of $E = -80$ mV and no inactivation.
-(G) Conductance-voltage relation at steady-state takes the form of a sigmoid.
-(H) Sigmoids $m$ fit to the model as $m^n$ data indicating that $n=4$ is the best fit.
-(I) Goodness of fit vs.
-exponent $n$, suggesting $n=4$ as the best fit to the data.}
-	\label{fig:figureclamp}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
-	\caption{Simulating a network of conductance-based model neurons.
-(A) Diagram of a network model of the pyloric rhythm in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004).
-(B) Each neuron is modeled as a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
-(C) \texttt{xolotl} implements conductances as fields of compartments and synapses as connections between compartments.
-(D-F) Simulated voltage trace of a model network for the three compartments.
-(G) Time series of synaptic currents in the simulated network can be obtained from the integration.}
-	\label{fig:figurenetwork}
-\end{figure}
-
-\begin{figure}
-	\centering
 	\includegraphics[width=1.0\linewidth]{gfx/figure_integral_control}

### 2018-06-11 figures/figure_network.m

@@ -107,13 +107,13 @@ xlim(ax(7), [0 max(time)]);
 
 %% Post-Processing
 
-prettyFig('fs', 16, 'plw', 1.5,'lw',1.5)
+prettyFig('fs', 14, 'plw', 1.5,'lw',1.5)
 
 % set the positions of the axes
 pos = [ ...
      0.13       0.7127       0.1237       0.2123;
      0.13       0.4131       0.1237       0.2123;
-     0.03       0.2          0.26         0.7;
+     0.03       0.05          0.26         0.92;
    0.3812       0.7932       0.4736       0.1577;
    0.3812       0.5741       0.4736       0.1577;
    0.3812        0.355       0.4736       0.1577;
@@ -128,9 +128,8 @@ for ii = 1:length(ax)
 end
 
 % label the subplots
-% labelFigure('capitalise', true)
+ labelFigure('capitalise', true,'ignore_these',ax(1:3),'x_offset',-.03,'y_offset',-.03)
 
 
-return
 % split the axes for aesthetics
 deintersectAxes(ax(4:7))

### 2018-06-13 paper/paper.tex
Add voltage clamp citations

@@
 "xolotl" can recapitulate the results of voltage clamp experiments (\cite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}).
+Voltage clamp represents a condition in which the membrane potential does not change from a value set by the experimenter, permitting the voltage-dependence of ion channels and their currents to be explicitly studied.
+Channels flux ions until the electrochemical gradient across the membrane is zero.
+Voltage current permits current traces to be gathered at various holding potentials, to characterize the current in isolation, or in combination with other currents (\cite{coleIonsPotentialsNerve1955, coleIonicCurrentMeasurements1960, hodgkinEffectSodiumIons1949, hodgkinMeasurementCurrentvoltageRelations1952a}).
+
 Figure \ref{fig:figureclamp} displays steps in the procedure to clamp the membrane potential of a cell with delayed rectifier potassium conductance.

### 2018-06-25 paper/paper.tex
updated text

--- 2018-06-11 paper/paper.tex
@@
 
+\newcommand{\dd}[1]{\mathrm{d}#1}
+
 \begin{document}
@@
 
-	Information processing by neurons relies on the generation and transmission of electrical signals that arise from the biophysics of ion channels, synapses and other voltage-coupled components.
-Electrophysiological characterization of these mechanisms has allowed for the construction of conductance-based models that can reproduce several features of neuronal and circuit behavior.
-However, working with conductance-based models continues to be a challenge due to their high dimensionality and large number of parameters, hindering intuition of their dynamical features.
+	Conductance-based models of neuron dynamics based on Hodgkin-Huxley formalism have been used extensively in computational neuroscience.
+However, working with these models continues to be a challenge due to their high dimensionality and large number of parameters, hindering intuition of their dynamical features.
 Here, we present a neuron and network simulator built on a novel automatic type system that binds class templates written in "C++" to object-oriented code in "MATLAB".
-Our approach unites the speed and power of "C++" code with the ease-of-use and rich libraries of scientific programming languages like "MATLAB".
-In our software, models are hierarchical, with components that are named and searchable, permitting high-level programmatic control over all parameters.
-The simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately.
+Our approach builds on the tradition of uniting the speed and power of "C++" code with the ease-of-use and rich libraries of scientific programming languages like "MATLAB".
+In our software, models are hierarchical, with components that are named and searchable, permitting intuitive high-level programmatic control over all parameters.
+Our simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately.
 The simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be easily extended to include arbitrary mechanisms and dynamics.
@@
 Nervous systems process and transmit information using electrically excitable membranes.
-Conductance-based models are a powerful biophysical simplification of an excitable compartment in a neuron (\cite{hodgkinQuantitativeDescriptionMembrane1952}).
+Conductance-based models are a powerful biophysical simplification of an electrically excitable compartment in a neuron (\cite{hodgkinQuantitativeDescriptionMembrane1952}).
 Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small circuits (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}).
@@
 Since models in the "MATLAB" workspace are automatically linked to models in the "C++" implementation, configuring these objects in "MATLAB" transparently configures the underlying "C++" objects.
-"xolotl" comes packaged with hundreds of conductances and synapses, built-in visualization functions, and a graphical user interface (GUI) for real-time manipulation of model parameters.
-Built-in methods allow the user to easily generate plots of voltage and intracellular calcium time traces, and activation functions of constituent conductances.
+"xolotl" comes packaged with hundreds of components that can be used to assemble cells and networks;  has built-in visualization functions to inspect voltage time traces and activation functions; and a graphical user interface (GUI) for real-time manipulation of model parameters.
 The GUI permits real-time manipulation of any network parameter and an immediate visualization of the effect of that parameter on arbitrary dynamical features of the model.
 The ease-of-use of these tools lends them to pedagogical applications and the rapid prototyping of models.
-This tool aims to simplify the investigation of the dynamics of conductance-based network and neuron models, facilitate collaborative modeling, and complement other tools being developed in the computational neuroscience community.
+This tool aims to simplify the investigation of the dynamics of conductance-based network and neuron models, facilitate collaborative modeling, and is intended to complement other tools being developed in the computational neuroscience community.
 
@@
 "xolotl" is designed to be easy-to-use and richly featured while being fast enough to use in everyday research.
-Our focus was on designing a simulator of conductance-based neurons and networks of these neurons; simulating arbitrary dynamical systems is therefore beyond the scope of this software.
+Our focus was on designing an approchable simulator of conductance-based neurons and networks of these neurons; simulating arbitrary dynamical systems is therefore beyond the scope of this software.
 Specifically, the software was designed to simulate models of the form
 
-\[ C_{i}\frac{dV_{i}}{dt}=-\sum_{j}I_{j} \]
+\begin{equation}
+C_{i}\frac{dV_{i}}{dt}=-\sum_{j}I_{j} \label{eq:1}
+\end{equation}
 
@@
 
-\[ I_{j}=\bar{g}_{j}m_{j}^{p}h_{j}^{q}(V-E_{j})A_{i} \]
+\begin{equation}
+I_{j}=\bar{g}_{j}m_{j}^{p}h_{j}^{q}(V-E_{j})A_{i}  \label{eq:2}
+\end{equation}
 
@@
 Typically, \( \tau_{m} \), \( \tau_{h} \), \( m_{\infty} \), and \( h_{\infty} \) are functions of the membrane potential \( V_{i} \).
-The software uses integration schemes that have been specifically developed to solve equations of this form (cite D\&A, Hines, others), though other schemes can be used.
+The software uses integration schemes that have been specifically developed to solve equations of this form (cite D\&A, Hines, others), though other schemes can be used if desired.
 
 This software is designed to be used from within "MATLAB", a scientific programming language common amongst neuroscientists and engineers for pedagogy and research.
-Our goal was to make "xolotl" completely usable entirely from within MATLAB, and fully scriptable so that the software can be used as a component of other packages and tools.
-All parameters of models and activation functions of any channel can be be accessed before running the model.
-Finally, we designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by changing a few lines in a template file.
-Models created using this simulator appear in the "MATLAB" workspace as native objects, and are fully compatible with the large library of toolboxes that "MATLAB" provides.
+Our goal was to make "xolotl" completely usable entirely from within MATLAB, and fully scriptable.
+Models created using this simulator appear in the "MATLAB" workspace as native objects, and are thus fully compatible with the large library of toolboxes that "MATLAB" provides, allowing so the software to be used as a component of other packages and tools.
+All parameters of models and activation functions of any channel can be be inspected before running the model.
+We designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by changing a few lines in a template file.
+Finally, "xolotl" is designed to be fully auditable, and comes with several tools to verify model and parameter integrity and aid reproducibility.
 
@@
 \paragraph{A rich library of network components.} "xolotl" comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and controllers) that can be used as building blocks to construct model neurons and networks.
-Compartments represent sections of membrane with a single membrane potential, and can represent either entire cells or parts of cells.
+Compartments represent sections of membrane with a single membrane potential, Calcium level, and set of constituent components; and can represent either entire cells or parts of cells.
 Conductances represent populations of ion channels in a compartment that produce transmembrane currents.
 Synapses connect two compartments together by introducing a current in the post-synaptic neuron that can depend on the voltage of the presynaptic neuron.
-Controllers represent homeostatic mechanisms that integrate a Calcium-dependent signal from the compartment they are in and  change the properties of a conductance or synapse in response.
+Controllers represent intracellular mechanisms that integrate a Calcium-dependent signal from the compartment they are in and  change the properties of a conductance or synapse in response.
 
@@
 Each of these neurons contains within it a set of conductances and synaptic currents that contribute to its electrical behavior.
-These conductances can be controlled by a feedback system that can homeostatically regulate their properties.
+These conductances can be controlled by intracellular regualtory system that can change their properties.
 Similarly, a "xolotl" model can contain a set of compartments that can represent individual neurons.
@@
 Our approach to circumvent the tradeoff between high-performance but hard to use languages like "C++" and richly featured but slow languages like "MATLAB" was to construct an automatic type system that links object oriented code in "C++" to object oriented code in "MATLAB".
-This architecture lets us construct the core of the simulator in "C++", leveraging features of "C++" like pointers that do not exist in "MATLAB".
+This architecture lets us construct the core of the simulator in "C++", leveraging features of "C++" like pointers that are not commonly used in "MATLAB".
 A rudimentary way to make this "C++" code useable in "MATLAB" would be to re-write that code in "MATLAB" so that "MATLAB" objects can be bound their "C++" implementations.
@@
 Instead, our automatic type system creates objects in "MATLAB" on-the-fly  from "C++" class specifications, obviating the need to rewrite code in "MATLAB" while preserving a tight coupling between objects in the "MATLAB" workspace and their "C++" implementation.
+The key advantages of this approach are that it makes developing new code much easier, and simplifies the task of constructing complex frameworks that span these two languages.
 
@@
 Since every model requires a compiled binary to run, a potential stumbling block is the problem of unambiguously linking a model to a binary executable.
-"xolotl" solves this problem by hashing the "C++" header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier.
-Compiled binaries are named with this hash, ensuring both that the correct binary is run to integrate the model, and that compilation occurs only as needed.
+"xolotl" solves this problem by hashing the "C++" header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier (its "hash").
+Compiled binaries are named using this hash, ensuring both that the correct binary is run to integrate the model, and that compilation occurs only as needed.
 This powerful feature enables the user experience to remain entirely within the "MATLAB" workspace, with compilation and selection of the correct binary occurring silently in the background.
@@
 
-Our focus on ease-of-use and fast simulations means some features were intentionally neglected in the design process.
+Our focus on ease-of-use and speed means some features were intentionally neglected in the design process.
 
@@
 However, it may be desirable to use other methods under certain conditions.
-It is possible to introduce new network components that implement other integration schemes, but that requires writing new "C++"  code, and is still restricted to integration schemes that assume a fixed step size.
+It is possible to introduce new network components that implement other integration schemes, or to modify the integration schemes of existing components, but that requires writing new "C++"  code, and is still restricted to integration schemes that assume a fixed step size.
 
-\paragraph{Inefficient tools for handling large networks.} Since every compartment and component in a "xolotl" model is $named$, rather than numbered, it is more suited towards simulating small but complex networks, rather than large but simple networks.
-While the software can integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task.
+\paragraph{Inefficient tools for handling large networks.} "xolotl" was designed to work with small but complex networks and models, where every compartment and component is $named$, rather than numbered.
+It is more therefore suited towards simulating small and heterogeneous networks rather than large but homogenous networks.
+While the software $can$ integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task.
 
@@
 Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}).
-The Hodgkin-Huxley model generates action potentials, and, as one of the simplest models of excitable neural mebranes, often serves as a the first model introduced in pedagogical literature (cite some textbooks).
+As one of the simplest models of excitable neural mebranes, the Hodgkin-Huxley model often serves as a the first model introduced in pedagogical literature (cite some textbooks).
 
@@
 The \texttt{plot} function generates a voltage trace, which is colored by the dominant current (Figure \ref{fig:figurehh}A).
-If the membrane potential is increasing, the strongest instantaneous inward current colors the trace.
-Conversely, if the membrane potential is decreasing, the strongest outward current colors the trace instead.
-These built-in features reveal that the dominant current during the upswing of every action potential is the sodium current, and the dominant current immediately after the peak of the action potential is the potassium current.
+Colors in the voltage trace indicate the strongest instantaneous inward current when the voltage is increasing, and strongest instantaneous outward current when the voltage is decreasing.
+These built-in features reveal that the dominant current during the upswing of every action potential is the sodium current, and the dominant current immediately after the peak of the action potential is the potassium current, but that the leak current contributes to depolarization following an action potential.
 This feature could be useful in quickly understanding the contributions of a number of ion channel types in a complex voltage trace from a more complicated neuron model.
+
+Integrating the model returns the voltage time series for every compartment:
+
+\begin{equation}
+{\tt V = x.integrate;}
+\end{equation}
 
@@
 Figure \ref{fig:figurehh}C shows the f-I curve of this model, obtained by repeated integration of the model.
-Finally, the built-in \texttt{show} function can plot activation (m) and inactivation (h) curves and voltage-dependent timescales of any channel type in the simulator (Figure \ref{fig:figurehh}D-G).
-These plots revleal that activation kinetics of the "NaV" channels are much faster than that of the "Kd" channels (Figure \ref{fig:figurehh}F), which facilitates the transient depolarization in an action potential.
+Finally, the built-in \texttt{show} function can plot activation ($m$) and inactivation ($h$) curves and voltage-dependent timescales of any channel type in the simulator (Figure \ref{fig:figurehh}D-G).
+These plots reveal that activation kinetics of the "NaV" channels are much faster than that of the "Kd" channels (Figure \ref{fig:figurehh}F), which facilitates the transient depolarization in an action potential.
 In summary, the simulator allows the user to construct and integrate this model in a few lines of code, and provides a rich visualization of the dynamics of the model.
@@
 	\caption{Simulating a voltage-clamp experiment.
-(A) Cartoon of a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998} with experimentally-fixed voltage.
-(B) Structure of \texttt{xolotl} object in A.
-(C) Code snippet depicting integration under voltage clamp.
-(D-E) Current response to steps in voltage from a holding potential of $V_m = -60$ mV.
-(F) Current-voltage relation of the steady-state current ($t = 400$ ms) indicating a reversal potential of $E = -80$ mV and no inactivation.
-(G) Conductance-voltage relation at steady-state takes the form of a sigmoid.
-(H) Sigmoids $m$ fit to the model as $m^n$ data indicating that $n=4$ is the best fit.
-(I) Goodness of fit vs.
-exponent $n$, suggesting $n=4$ as the best fit to the data.}
+The cartoon shows a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998}) that is being recorded from in two-electrode voltage clamp.
+The code snippet sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it.
+(A) Voltage steps that the cell is clamped to.
+(B) Current injected into the cell to clamp it to the voltages in (A).
+(C) Asymptotic clamped current vs.
+clamped voltage for this cell.
+(D) Accounting for the reversal potential of Potassium ions yeilds the conductance-voltage curve of this channel type.
+(E) Normalized conductance-voltage curves, with sigmoid fits with various exponents.
+(F) An exponent of $n=4$ yields the best fit, allowing for the characterization of the activation function of this channel type.
+}
 	\label{fig:figureclamp}
@@
 
-"xolotl" can recapitulate the results of voltage clamp experiments (\cite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}).
-Figure \ref{fig:figureclamp} displays steps in the procedure to clamp the membrane potential of a cell with delayed rectifier potassium conductance.
-During an \textit{in-vitro} experiment, confounding currents would be pharmacologically-blocked and two-electrode voltage clamp used to record tail currents at fixed membrane potential (\cite{connorInwardDelayedOutward1971, connorVoltageClampStudies1971}).
+Voltage clamping is a experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (cite D\&A).
+Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp.
+By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized.
+(insert all those citations as needed).
 
-A single-compartment model with a delayed-rectifier conductance (\cite{liuModelNeuronActivitydependent1998}) is simulated at stepped membrane potentials.
-The "integrate" function outputs the membrane potential and current as time series.
-Currents under voltage clamp approach the steady-state holding current (Figure \ref{fig:figureclamp}D-E).
-The current-voltage relation is the steady-state current over the clamped voltage, and the effective conductance is the derivative of that relation (Figure \ref{fig:figureclamp}F-G).
-Since the effective conductance is the product of the maximal conductance and the gating variables (\cite{dayanTheoreticalNeuroscience2001, turrigianoSelectiveRegulationCurrent1995}) and the tail current is monotonically-increasing with time under voltage clamp, the current can be represented as non-inactivating.
-Fitting a sigmoid to various powers yields a model for the current dynamics (Figure \ref{fig:figureclamp}H-I).
-These figures describe graphically the theoretical underpinnings of current analysis through voltage clamp and can serve as an effective pedagogical tool for computational and quantitative neuroscience.
+"xolotl" can reproduce such a voltage clamp experiment $in silico$.
+Figure \ref{fig:figureclamp} illustrates how a simple model with a single compartment and a single ion channel type can be set up and clamped to a desired voltage.
+Integrating the model yields the current required to clamp the cell at that voltage.
+Here, we use a delayed-rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998}) and simulate a voltage-clamp experiment whose goal is to infer the activation function of this channel.
+First, the cell is clamped to a number of different voltages (Fig.
+\ref{fig:figureclamp}A) and the resultant clamping currents are measured by integrating the model (Fig.
+\ref{fig:figureclamp}B).
+The asymptotic clamping currents depend on the clamping voltage in a nonlinear manner (Fig.
+\ref{fig:figureclamp}C), since the open probabilities of the channel are functions of the membrane voltage.
+Assuming the reversal potential is known, Eq.
+\eqref{eq:2} can be used to solve for the total conductance of the channel as a function of the clamping voltage (Fig.
+\ref{fig:figureclamp}D).
+Finally, a sigmoid can be fit to the normalized conductance-voltage curves to obtain the activiation function of the ion channel population (Fig.
+\ref{fig:figureclamp}E-F).
+"xolotl" can therefore be used to describe graphically the theoretical underpinnings of ion channel characterization through voltage clamp and can serve as an effective pedagogical tool in computational neuroscience.
 

### 2018-06-26 paper/paper.tex
more work on benchmark fig

@@
 
-%In "MATLAB", users create a "xolotl" object and populate it with "cpplab"-generated objects which describe compartments, conductances, synapses, and controllers.
-The model is integrated with the "integrate" function where the membrane potential, intracellular calcium concentration, controller states, intrinsic currents, and synaptic currents can be outputs.
-%
-%
-%
-%\subsection{Constructing Networks}
-%
-%The "add" function creates a compartment, conductance, or controller and affixes it as a field in the "xolotl" network structure.
-This function generates a "MATLAB struct" that faithfully represents the underlying "C++" code.
-Compartments add to the "xolotl" object and conductances add to compartments.
-Specific properties can be specified using key-value pair arguments (\eg Figure \ref{fig:figurehh}A).
-%
-%\paragraph{Compartments.}
-%A model neuron consists of one or more compartments, each representing a section of membrane with capacitance and surface area.
-Isopotential models require one compartment, whereas models with multiple neurons, units, or non-trivial morphology require multiple compartments.
-%
-%%\begin{table}
-%%	\begin{center}
-%%		\begin{tabular}{ l | l | c }
-%%			\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
-%%			Membrane potential & "V" & $mV$ \\
-%%			Intracellular calcium concentration & "Ca" & $\mu M$ \\
-%%			Membrane capacitance & "Cm" & $nF/mm^2$ \\
-%%			Volume & "vol" & $mm^3$ \\
-%%			Surface area & "A" & $mm^2$ \\
-%%			Radius of compartment & "radius" & $mm$ \\
-%%			Length of compartment & "len" & $mm$ \\
-%%			Extracellular calcium steady-state & "Ca_out" & $\mu M$ \\
-%%			Intracellular calcium steady-state & "Ca_in" & $\mu M$ \\
-%%			Calcium buffering shell thickness & "shell_thickness" & $\mu M$ \\
-%%			Calcium current to concentration factor & "phi" & ??? \\
-%%			Calcium buffering time constant & "tau_Ca" & $ms$ \\
-%%			Average calcium concentration & "Ca_average" & $\mu M$ \\
-%%			Target calcium concentration & "Ca_target" & $\mu M$ \\
-%%		\end{tabular}
-%%	\end{center}
-%%	\label{tab:compartmentprops}
-%%\end{table}
-%
-%\paragraph{Synapses.}
-%"xolotl" provides some features for generating complex models.
-Synapses can be added with the \texttt{connect} function.
-Synapses possess identifiers to presynaptic and postsynaptic compartments and default to electrical synapses.
-To create axons or transport chains, the \texttt{slice} function splits a compartment into $n$ discrete segments and adds these compartments to the network connected by electrical synapses.
-%
-%%\begin{center}
-%%	\begin{tabular}{ l | l | c }
-%%		\label{tab:synapseprops}
-%%		\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
-%%		Maximal conductance & "gbar" & $\mu S /mm^2$ \\
-%%		Reversal potential & "E" & $mV$ \\
-%%		Synaptic gating variable & "s" &  \\
-%%	\end{tabular}
-%%\end{center}
-%
-%\paragraph{Conductances and controllers.}
-%
-%All conductances contain fields for maximal conductance and reversal potential.
-Conductances with activation and inactivation variables include them as $m$ and $h$ respectively.
-Gating functions and their respective time constants are contained within the conductance header file.
-"xolotl" comes packaged with conductances from several dozen papers (\cite{dethierPositiveFeedbackCellular2015}.
-%
-%%\begin{center}
-%%	\begin{tabular}{ l | l | c }
-%%		\textbf{Name} & \textbf{\texttt{PropertyName}} & Units \\ \hline
-%%		Maximal conductance & "gbar" & $\mu S /mm^2$ \\
-%%		Reversal potential & "E" & $mV$ \\
-%%		Activation gating variable & "m" &  \\
-%%		Inactivation gating variable & "h" &  \\
-%%	\end{tabular}
-%%\end{center}
-%
-%\paragraph{Creating custom \texttt{cpplab} objects.}
-%
-%"xolotl" contains template header files for producing custom conductances. The template contains instructions on how to design novel conductances with arbitrary specifications.
-%
-%\subsection{Simulation}
-%
-%Models are simulated in "xolotl" with the \texttt{integrate} function which outputs as time series the membrane potentials, intracellular calcium concentrations, controller states, intrinsic currents, and synaptic currents.
-The \texttt{integrate} function also accepts an argument which specifies injected current or clamped voltage.
-%
-%\paragraph{Numerical integration.}
-%"xolotl" uses the exponential Euler method for single compartment models, forward Euler for gating variables, and a Crank-Nicholson regime for electrically-coupled compartments (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001, butcherNumericalDifferentialEquation2016}).
-These defaults provide a mix of speed, accuracy, and stability, and are built into the "cpplab" header files.
-Custom "cpplab" header files can be customized with any iterative integration method.
-The simulation time-resolution can be specified to target arbitrary precision, and an output time step can be selected to support automatic down-sampling for memory considerations.
-%
-%\paragraph{`Closed-loop' vs. `open-loop.'}
-%Simulations can be run in `closed-loop' mode where each simulation begins by resetting all dynamical variables to their initial conditions at instantiation, or `open-loop' mode which begins simulation with the current network state.
-%
-%\subsection{Additional Features}
-%
-%\paragraph{Using the graphical interface to manipulate parameters.}
-%
-%"xolotl" comes packaged with a graphical user interface for visualizing parameter changes in real-time.
-The \texttt{manipulate} function opens the GUI, which displays a figure plotting the membrane potential and intracellular calcium concentration of all compartments as time series, and a dialog box with customizable sliders for all parameters of the model, much like the "Manipulate" function in Wolfram "Mathematica".
-Moving the sliders integrates the model in `open-loop` mode with the new parameters.
-Closing the GUI saves the network state of the model to the "xolotl" object.
-This is particularly helpful for rapid prototyping of models.
-%
-%\paragraph{Optimizing parameters.}
-%"xolotl" can use the Global Optimization toolbox for "MATLAB" to optimize any accessible "xolotl" parameters.
-The toolbox is algorithm-agnostic and accepts any function in "MATLAB" with a scalar first output as the objective function.
-Simulations run on multi-core processors or high-performance computing clusters using the Parallel Computing toolbox.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsection{Software use and reproducibility of research}
+\subsection{Design: components vs. equations in simulators }
+
+- components from library vs. equations
+- equations are transparent -- you know what you're solving. but clunky -- string parsing headaches, etc. 
+- components are easy to assemble, but it;s hard to know where they are, how they can be changed, etc.
+- xolotl goes the component route, but makes it very easy to find components (you can click your way into the C++ files).
+components are stored as source files, so you can modify it and immediately see it reflected in your model.
+we didn't invent another synatax for these components -- these are just C++ files.
+
+\subsection{Comparison with other simulators}
+
+- compare with BRIAN, GENESIS, XPPAUT, NEURON, DynaSim
+- selling point is ease of use and integration into MATLAB
+
+\subsection{Auditability and Reproducability}
 
 - reproducibility crisis in computational science 
+- unlike experiemntal research, where reproducability can arise due to deep reasons, reasons for computational research irreproducability are mostly trivial: 
 - multiple sources: software being used is closed-source, researcher's code is not made available; obscure architecture makes improper use possible. 

### 2018-06-26 dynasim/STG.m

@@ -1,5 +1,4 @@
-% simulate DynaSim Hodgkin-Huxley model
-% with same specifications as figure_benchmark_2
+% simulate DynaSim STG like model
 
 % set up DynaSim equation block
 equations = {...

### 2018-06-26 figures/benchmarks/figure_benchmark_2.m

@@ -117,21 +117,6 @@ plot(ax(5), nComps, DynaSim.S, 'b-o')
 ;;    ;; ;;;;;;;;  ;;;;;;;  ;;     ;;  ;;;;;;;  ;;    ;;
 
 
-% now we set up a STG-like neuron
-x = xolotl;
-x.add('compartment','AB','A',0.0628,'phi',90,'vol',.0628);
-
-x.AB.add('liu/NaV','gbar',@() 115/x.AB.A,'E',30);
-x.AB.add('liu/CaT','gbar',@() 1.44/x.AB.A,'E',30);
-x.AB.add('liu/CaS','gbar',@() 1.7/x.AB.A,'E',30);
-x.AB.add('liu/ACurrent','gbar',@() 15.45/x.AB.A,'E',-80);
-x.AB.add('liu/KCa','gbar',@() 61.54/x.AB.A,'E',-80);
-x.AB.add('liu/Kd','gbar',@() 38.31/x.AB.A,'E',-80);
-x.AB.add('liu/HCurrent','gbar',@() .6343/x.AB.A,'E',-20);
-x.AB.add('Leak','gbar',@() 0.0622/x.AB.A,'E',-50);
-x.t_end = 1e4;
-x.integrate;
-x.snapshot('zero');
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % STOMATOGASTRIC MODEL: TIME-STEP

### 2018-06-26 figures/benchmarks/testDynaSimHH.m

@@ -42,8 +42,7 @@ time        = all_dt(end) * (1:(t_end / max(all_dt)));
 
 all_V = NaN(ceil(t_end/max(all_dt)),length(all_dt));
 
-% h = GetMD5(which(mfilename),'File');
-h = ['DS_' GetMD5(all_dt)];
+h = ['DS_' GetMD5(which(mfilename),'File')];
 
 if isempty(cache(h))
 

### 2018-06-26 figures/benchmarks/testXolotlSTG.m

@@ -2,20 +2,24 @@
 % xolotl, and runs the benchmarks on it
 
 
-function testXolotlHH(ax)
+function testXolotlSTG(ax)
 
 
-% set up xolotl object
+% now we set up a STG-like neuron
 x = xolotl;
-x.add('compartment', 'HH', 'Cm', 10, 'A', 0.01);
-x.HH.add('liu/NaV', 'gbar', 1000, 'E', 50);
-x.HH.add('liu/Kd', 'gbar', 300, 'E', -80);
-x.HH.add('Leak', 'gbar', 1, 'E', -50);
-x.I_ext = .2;
+x.add('compartment','AB','A',0.0628,'phi',90,'vol',.0628);
 
+x.AB.add('liu/NaV','gbar',@() 115/x.AB.A,'E',30);
+x.AB.add('liu/CaT','gbar',@() 1.44/x.AB.A,'E',30);
+x.AB.add('liu/CaS','gbar',@() 1.7/x.AB.A,'E',30);
+x.AB.add('liu/ACurrent','gbar',@() 15.45/x.AB.A,'E',-80);
+x.AB.add('liu/KCa','gbar',@() 61.54/x.AB.A,'E',-80);
+x.AB.add('liu/Kd','gbar',@() 38.31/x.AB.A,'E',-80);
+x.AB.add('liu/HCurrent','gbar',@() .6343/x.AB.A,'E',-20);
+x.AB.add('Leak','gbar',@() 0.0622/x.AB.A,'E',-50);
+x.t_end = 1e4;
 x.integrate;
 x.snapshot('zero');
-x.closed_loop = false;
 
 
 x.t_end = 30e3;
@@ -165,86 +169,3 @@ xlabel(ax(4),'t_{end} (ms)')
 ylabel(ax(4),'Speed (X realtime)')
 
 
- ;;;;;;  ;;    ;;  ;;;;;;  ;;;;;;;; ;;;;;;;; ;;     ;; 
-;;    ;;  ;;  ;;  ;;    ;;    ;;    ;;       ;;;   ;;; 
-;;         ;;;;   ;;          ;;    ;;       ;;;; ;;;; 
- ;;;;;;     ;;     ;;;;;;     ;;    ;;;;;;   ;; ;;; ;; 
-      ;;    ;;          ;;    ;;    ;;       ;;     ;; 
-;;    ;;    ;;    ;;    ;;    ;;    ;;       ;;     ;; 
[…]

### 2018-06-26 src/qqq.m

@@ -1,3 +1,0 @@
-close all
-clear
-clc

### 2018-06-26 src/xolotl_comps.m

@@ -1,45 +1,0 @@
-% make figure of slice time over number of slices
-nSlices = [1 linspace(10,100,10)];
-time    = zeros(length(nSlices),1);     % s
-I_ext   = 0.1 * ones(max(nSlices),1);   % nA
-
-for ii = 1:length(nSlices)
-  textbar(ii,length(nSlices))
-  % create xolotl object with two conductances
-  x = xolotl;
-  x.cleanup;
-  x.add('HH','compartment','V', -65, 'Ca', 0.02, 'Cm', 10, ...
-    'radius', 0.025, 'len', 0.050);
-  x.HH.add('liu-approx/NaV', 'gbar', 1000, 'E', 50);
-  x.HH.add('liu-approx/Kd', 'gbar', 300, 'E', -80);
-  x.HH.add('Leak', 'gbar', 1, 'E', -50);
-  if nSlices(ii) ~= 1
-    x.slice('HH',nSlices(ii),100);
-  end
-  x.sha1hash;
-  % set up simulation
-  x.t_end = 100e3;  % ms
-  x.dt = 0.05;      % ms
-  V = zeros(x.t_end/x.dt,nSlices(ii));
-  % test simulation speed
-  tic;
-  V = x.integrate(I_ext(1:nSlices(ii)));
-  time(ii) = toc;
-end
-
-figure;
-
-subplot(1,2,1)
-plot(nSlices,time)
-xlabel('# of slices')
-ylabel('time (s)')
-title('run-time vs. slicing')
-
-subplot(1,2,2)
-hold on
-plot(nSlices,x.t_end/1000/time)
-xlabel('# of slices')
-ylabel('speed factor')
-title('sim-time/run-time vs. slicing')
-
-prettyFig();

### 2018-06-27 paper/paper.tex
minor changes

@@
 Here, we present a neuron and network simulator built on a novel automatic type system that binds class templates written in "C++" to object-oriented code in "MATLAB".
-Our approach builds on the tradition of uniting the speed and power of "C++" code with the ease-of-use and rich libraries of scientific programming languages like "MATLAB".
+Our approach builds on the tradition of uniting the speed and power of languages like "C++" with the ease-of-use and rich libraries of scientific programming languages like "MATLAB".
 In our software, models are hierarchical, with components that are named and searchable, permitting intuitive high-level programmatic control over all parameters.
 Our simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately.
-The simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be easily extended to include arbitrary mechanisms and dynamics.
+Our simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be extended to include arbitrary mechanisms and dynamics.
 Finally, the simulator is written in a modular fashion and has been released under a permissive free software license, enabling it to be integrated easily into the workflow of many researchers.
@@
 Second, many or all equations in these models can be strongly coupled through variables like the membrane potential.
-In multi-compartment models, membrane potentials in every compartment are coupled to each other.
+In multi-compartment models of spatially extended neurons, membrane potentials in every compartment are coupled to each other.
 Finally, the choice of programming language used to implement the model imposes tradeoffs in designing and using neuron and network simulators: simulators written in languages like "C++", "FORTRAN" or "hoc" can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like "Python", "Julia" or "MATLAB".
@@
 These simulators can be easier to use, and can benefit from inter-compatibility with other tools.
-Programs like "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}) allow the user to specify models using strings of equations that are specified in the scientific programming language, that can then be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
+Programs like "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}) allow the user to specify models using strings of equations or using components that are specified in the scientific programming language, that can then be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
 This approach permits considerable flexibility for simulating systems of differential equations.
-Since models need to be translated between the two languages, the hierarchical nature of neuron models is not naturally encapsualted by these tools, and the syntax can be verbose.
+Since models need to be translated between the two languages, the hierarchical nature of neuron models is not naturally encapsulated by these tools, and the syntax can be verbose.
 Neither approach facilitates the creation of tools that simultaneously maintain efficiency, ease-of-use, and clarity.
@@
 "xolotl" is designed to be easy-to-use and richly featured while being fast enough to use in everyday research.
-Our focus was on designing an approchable simulator of conductance-based neurons and networks of these neurons; simulating arbitrary dynamical systems is therefore beyond the scope of this software.
+Our focus was on designing an approachable simulator of conductance-based neurons and networks of these neurons; simulating arbitrary dynamical systems is therefore beyond the scope of this software.
 Specifically, the software was designed to simulate models of the form

### 2018-07-09 paper/paper.tex
added figure caption for snapshot figure

@@
 %
+%		MECHANISMS 
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
+%		SNAPSHOTS
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Using snapshots to explore model dynamics}
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_snapshot}
+	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them when needed.
+The initial state (orange disc) of the single compartment model in the previous example is saved using the snapshot method.
+This saves all parameters and dynamic variables of the model in a named state.
+The first column shows the profile of conductances and the voltage dynamics of the model at this point in time.
+First, the maximal conductances of the Calcium channels are set to zero (purple), changing the voltage dynamics of the neuron.
+After evolving the model for some time (green), the conductance profile and voltage dynamics returns to a state similar to the initial state.
+This configuration is now saved in a state called final and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
+This configuration is now saved in a state called final and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
+In summary, any state can be bookmarked using a descriptive name using the \texttt{snapshot} method, and can be returned to using the \texttt{reset} method.}
+	\label{fig:figuresnapshot}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+%
 %		NETWORK MODELS

### 2018-07-10 paper/paper.tex
added mechanism figure

@@
 
+\subsection{Intracellular mechanisms}
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_mechanism}
+	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them when needed.
+The initial state (orange disc) of the single compartment model in the previous example is saved using the snapshot method.
+This saves all parameters and dynamic variables of the model in a named state.
+The first column shows the profile of conductances and the voltage dynamics of the model at this point in time.
+First, the maximal conductances of the Calcium channels are set to zero (purple), changing the voltage dynamics of the neuron.
+After evolving the model for some time (green), the conductance profile and voltage dynamics returns to a state similar to the initial state.
+This configuration is now saved in a state called final and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
+This configuration is now saved in a state called final and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
+In summary, any state can be bookmarked using a descriptive name using the \texttt{snapshot} method, and can be returned to using the \texttt{reset} method.}
+	\label{fig:figuremechanism}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_integral_control}
-	\caption{Simulating neurons under homeostatic regulation.
-(A) Cartoon of a model neuron (\cite{liuModelNeuronActivitydependent1998} with integral control (\cite{olearyCorrelationsIonChannel2013}.
-(B) Hierarchical structure of a neuronal network considers controllers as components of compartments which act on conductances.
-(C) \texttt{xolotl} implements controllers as properties of conductances and synapses.
-(D) Calcium sensors change maximal conductances to move a neuron from quiescence to a bursting state.
-(E) Voltage trace shows regular bursting activity after integral control.}
-	\label{fig:figureintegralcontrol}
-\end{figure}
-
-\begin{figure}
-	\centering
 	\includegraphics[width=1.0\linewidth]{gfx/puppeteer_screenshot}

### 2018-07-10 paper/paper.tex
wrote caption for mechanism figure

@@
 Our approach builds on the tradition of uniting the speed and power of languages like "C++" with the ease-of-use and rich libraries of scientific programming languages like "MATLAB".
-In our software, models are hierarchical, with components that are named and searchable, permitting intuitive high-level programmatic control over all parameters.
+Our software allows for the creation and manipulation of hierarchical models with components that are named and searchable, permitting intuitive high-level programmatic control over all parameters.
 Our simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately.
@@
 In the simulator, the soma is represented using an object of class $compartment$ and the populations of ion channels are represented by $conductance$ objects contained within the compartment object.
-Code snippet used to set up this neuron model, inject current, integrate and plot the voltage, and show activation functions demonstrates how various several steps can be run in a few lines of code.
+Code snippet used to set up this neuron model, inject current, integrate and plot the voltage, and show activation functions demonstrates how several steps can be run in a few lines of code.
 (A) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current.
@@
 	\caption{Simulating a voltage-clamp experiment.
-The cartoon shows a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998}) that is being recorded from in two-electrode voltage clamp.
-The code snippet sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it.
+The diagram shows a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998}) that is being recorded from in two-electrode voltage clamp.
+The code snippet shown here sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it.
 (A) Voltage steps that the cell is clamped to.
@@
 clamped voltage for this cell.
-(D) Accounting for the reversal potential of Potassium ions yeilds the conductance-voltage curve of this channel type.
+(D) Accounting for the reversal potential of Potassium ions yields the conductance-voltage curve of this channel type.
 (E) Normalized conductance-voltage curves, with sigmoid fits with various exponents.
@@
 	\includegraphics[width=1.0\linewidth]{gfx/figure_mechanism}
-	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them when needed.
-The initial state (orange disc) of the single compartment model in the previous example is saved using the snapshot method.
-This saves all parameters and dynamic variables of the model in a named state.
-The first column shows the profile of conductances and the voltage dynamics of the model at this point in time.
-First, the maximal conductances of the Calcium channels are set to zero (purple), changing the voltage dynamics of the neuron.
-After evolving the model for some time (green), the conductance profile and voltage dynamics returns to a state similar to the initial state.
-This configuration is now saved in a state called final and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
-This configuration is now saved in a state called final and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
-In summary, any state can be bookmarked using a descriptive name using the \texttt{snapshot} method, and can be returned to using the \texttt{reset} method.}
+	\caption{Adding arbitrary intracellular mechanisms.
+(A) A single-compartment neuron model with 8 channel types.
+With these parameters, this model neuron is a tonic spiker.
+Since there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant (top panel).
+(B)  An intracellular buffering of Calcium and the effect of Calcium influx through voltage gated Calcium channels (VGCCs) is modeled using a simple differential equation.
+Adding this mechanism causes the cell to burst periodically, with synchronized oscillations in intracellular Calcium.
+(C) Schematic of Calcium-dependent intregral control homeostastis.
+In this feedback system, the rates of production of mRNA depends on the Calcium levels in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which is a function of the mRNA abundance.
+These integral controllers are implemented as a \texttt{mechanism} object, and can be added to conductances.
+On integrating the model, intracellular calcium levels rise and approach the target (red dashed line).
+This is accompanied by an increase in the conductance densities of all channels being controlled by this homeostatic mechanism.
+The voltage behavior of the cell changes from silence to bursting with truncated spikes to regular bursting.}
 	\label{fig:figuremechanism}
 \end{figure}
+
+"xolotl" can implement homeostatic tuning rules as integral control.
+The controller computes an error signal (typically a function of intracellular calcium concentration), and adjusts the conductance or synapse it controls accordingly (\cite{olearyCorrelationsIonChannel2013}).
+In "xolotl", integral controllers are "cpplab" objects added to the conductance or synapse they regulate.
+
+In a demonstration adapted from \cite{olearyCorrelationsIonChannel2013}, integral control changes maximal conductances to bring a neuron from quiescence into a bursting regime.
+Calcium sensors supervene on maximal conductance density (Figure \ref{fig:figureintegralcontrol}) to change neuronal activity.
+Each conductance in the "xolotl" structure contains a calcium-sensitive controller (Figure \ref{fig:figureintegralcontrol}B-C).
+Maximal conductances increase from random initial conditions to a set which elicits the desired network output by minimizing the error signal (Figure \ref{fig:figureintegralcontrol}D-F).
 
@@
 Synapses are stored in a vector array as a field of the "xolotl" object in "MATLAB" (Figure \ref{fig:figurenetwork}C).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%
-%		INTEGRAL CONTROL
-%
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Simulating Homeostasis}
-
-"xolotl" can implement homeostatic tuning rules as integral control.
-The controller computes an error signal (typically a function of intracellular calcium concentration), and adjusts the conductance or synapse it controls accordingly (\cite{olearyCorrelationsIonChannel2013}).
-In "xolotl", integral controllers are "cpplab" objects added to the conductance or synapse they regulate.
-
-In a demonstration adapted from \cite{olearyCorrelationsIonChannel2013}, integral control changes maximal conductances to bring a neuron from quiescence into a bursting regime.
-Calcium sensors supervene on maximal conductance density (Figure \ref{fig:figureintegralcontrol}) to change neuronal activity.
-Each conductance in the "xolotl" structure contains a calcium-sensitive controller (Figure \ref{fig:figureintegralcontrol}B-C).
-Maximal conductances increase from random initial conditions to a set which elicits the desired network output by minimizing the error signal (Figure \ref{fig:figureintegralcontrol}D-F).
 

### 2018-07-10 paper/paper.tex
figures1-5 done

@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_HH}
-	\caption{\texttt{xolotl} can quickly set up and simulate conductance-based models.
-Schematic representation of a electrotonically-compact neuron with three populations of ion channels (colored rectangles).
-In the simulator, the soma is represented using an object of class $compartment$ and the populations of ion channels are represented by $conductance$ objects contained within the compartment object.
-Code snippet used to set up this neuron model, inject current, integrate and plot the voltage, and show activation functions demonstrates how several steps can be run in a few lines of code.
-(A) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current.
-Colors indicate the dominant current (gold is fast sodium (NaV), blue is delayed rectifier (Kd), red is Leak).
-(C) f-I curve of this neuron.
-(D-E) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
-(F-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables}
-	\label{fig:figurehh}
-\end{figure}
-
 In this section, we demonstrate how to simulate a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model, and gain insight into the underlying biophysical mechanism.
@@
 
-\begin{figure}[!htb]
-	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_clamp}
-	\caption{Simulating a voltage-clamp experiment.
-The diagram shows a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998}) that is being recorded from in two-electrode voltage clamp.
-The code snippet shown here sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it.
-(A) Voltage steps that the cell is clamped to.
-(B) Current injected into the cell to clamp it to the voltages in (A).
-(C) Asymptotic clamped current vs.
-clamped voltage for this cell.
-(D) Accounting for the reversal potential of Potassium ions yields the conductance-voltage curve of this channel type.
-(E) Normalized conductance-voltage curves, with sigmoid fits with various exponents.
-(F) An exponent of $n=4$ yields the best fit, allowing for the characterization of the activation function of this channel type.
-}
-	\label{fig:figureclamp}
-\end{figure}
-
 Voltage clamping is a experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (cite D\&A).
@@
 
-\begin{figure}[!htb]
-	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_mechanism}
-	\caption{Adding arbitrary intracellular mechanisms.
-(A) A single-compartment neuron model with 8 channel types.
-With these parameters, this model neuron is a tonic spiker.
-Since there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant (top panel).
-(B)  An intracellular buffering of Calcium and the effect of Calcium influx through voltage gated Calcium channels (VGCCs) is modeled using a simple differential equation.
-Adding this mechanism causes the cell to burst periodically, with synchronized oscillations in intracellular Calcium.
-(C) Schematic of Calcium-dependent intregral control homeostastis.
-In this feedback system, the rates of production of mRNA depends on the Calcium levels in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which is a function of the mRNA abundance.
-These integral controllers are implemented as a \texttt{mechanism} object, and can be added to conductances.
-On integrating the model, intracellular calcium levels rise and approach the target (red dashed line).
-This is accompanied by an increase in the conductance densities of all channels being controlled by this homeostatic mechanism.
-The voltage behavior of the cell changes from silence to bursting with truncated spikes to regular bursting.}
-	\label{fig:figuremechanism}
-\end{figure}
-
 "xolotl" can implement homeostatic tuning rules as integral control.
@@
 
-\begin{figure}[!htb]
-	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_snapshot}
-	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them when needed.
-The initial state (orange disc) of the single compartment model in the previous example is saved using the snapshot method.
-This saves all parameters and dynamic variables of the model in a named state.
-The first column shows the profile of conductances and the voltage dynamics of the model at this point in time.
-First, the maximal conductances of the Calcium channels are set to zero (purple), changing the voltage dynamics of the neuron.
-After evolving the model for some time (green), the conductance profile and voltage dynamics returns to a state similar to the initial state.
-This configuration is now saved in a state called final and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
-This configuration is now saved in a state called final and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
-In summary, any state can be bookmarked using a descriptive name using the \texttt{snapshot} method, and can be returned to using the \texttt{reset} method.}
-	\label{fig:figuresnapshot}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Simulating Network Models}
-
-\begin{figure}[!htb]
-	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
-	\caption{Simulating a network of conductance-based model neurons.
-(A) Diagram of a network model of the pyloric rhythm in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004).
-(B) Each neuron is modeled as a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
-(C) \texttt{xolotl} implements conductances as fields of compartments and synapses as connections between compartments.
-(D-F) Simulated voltage trace of a model network for the three compartments.
-(G) Time series of synaptic currents in the simulated network can be obtained from the integration.}
-	\label{fig:figurenetwork}
-\end{figure}
 
@@
 
+\newpage
+
+\begin{figure}
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_HH}
+	\caption{\texttt{xolotl} makes it easy tp set up and simulate conductance-based models.
+Schematic representation of a single-compartment neuron with three populations of ion channels (colored rectangles).
+In the simulator, the soma is represented using an object of class $compartment$ and the populations of ion channels are represented by $conductance$ objects contained within the compartment object.
+The code snippet shown sets up this neuron model, injects current, integrates and plots the voltage, and shows activation functions, all in a few lines of code.
+(A) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current.
+Colors indicate the dominant current (gold is fast sodium (NaV), blue is delayed rectifier (Kd), red is Leak).
+(C) f-I curve of this neuron.
+(D-E) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
+(F-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables}
+	\label{fig:figurehh}
+\end{figure}
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_clamp}
+	\caption{Simulating a voltage-clamp experiment.
+The diagram shows a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998}) that is being recorded from in two-electrode voltage clamp.
+The code snippet shown here sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it.
+(A) Voltage steps that the cell is clamped to.
+(B) Current injected into the cell required to clamp it to the voltages in (A).
+(C) Asymptotic clamped current $vs.$ clamped voltage for this cell.
+(D) Accounting for the reversal potential of Potassium ions yields the conductance-voltage curve of this channel type.
+(E) Normalized conductance-voltage curves, with sigmoid fits with various exponents.
+(F) An exponent of $n=4$ yields the best fit, allowing for the characterization of the activation function of this channel type.
+}
+	\label{fig:figureclamp}
+\end{figure}
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_mechanism}
+	\caption{Modeling intracellular mechanisms.
+(box A) A single-compartment neuron model with 8 channel types.
+With these parameters, this model neuron is a tonic spiker.
+Since there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant (top panel).
+(box B)  An intracellular buffering of Calcium and the effect of Calcium influx through voltage gated Calcium channels (VGCCs) is modeled using a simple differential equation.
+Code snippet shows how this mechanism can be added to the neuron model.
+The cell now bursts periodically, with synchronized oscillations in intracellular Calcium.
+(box C) Schematic of Calcium-dependent integral control homeostasis.
+In this feedback system, the rates of production of mRNA depends on the Calcium levels in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which, through translation, is a function of the mRNA abundance.
+These integral controllers are implemented as \texttt{mechanism} objects, and can be added to conductances.
+On integrating the model, intracellular calcium levels rise and approach the target (red dashed line).
+This is accompanied by an increase in the conductance densities of all channels being controlled by this homeostatic mechanism.
+The voltage behavior of the cell changes from silence to bursting with truncated spikes to regular bursting.}
+	\label{fig:figuremechanism}
+\end{figure}
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_snapshot}
+	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them $ad  arbitrium$.
+The initial state (orange disc) of the single compartment model in the previous example is saved using the \texttt{snapshot} method.
+This method saves all parameters and dynamic variables of the model in a named state.
+The first column shows the profile of conductances and the voltage dynamics of the model at this point in time.
+The maximal conductances of the Calcium channels are set to zero (purple), changing the voltage dynamics of the neuron (purple plots).
+After evolving the model for some time (green), the conductance profile and voltage dynamics returns to a state similar to the initial state (green plots).
+This configuration is now saved in a state called $final$ and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
+Another parameter is changed (the Calcium target), and the model is integrated to reach a new state (blue disc) where the voltage dynamics are different from the $initial$ state.
+In summary, any state can be bookmarked using a descriptive name using the \texttt{snapshot} method, and can be returned to using the \texttt{reset} method.}
+	\label{fig:figuresnapshot}
+\end{figure}
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
+	\caption{Simulating a network of conductance-based model neurons.
+The diagram shows a three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004).
+Each neuron is modeled as a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
+Synapses are of two types (Cholinergic (dashed) or Glutamatergic (solid lines)) and have different kinetics.
+The code snippet shows how neurons can be created, wired together using synapses, and the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse.
+(A-C) Simulated voltage trace of a model network for the three compartments obtained from this simulation.
+(D) Time series activation variable of the Glutamatergic synapse between PY and LP (red connection in diagram) shows how the synapse becomes active every time the PY neuron spikes.}
+	\label{fig:figurenetwork}
+\end{figure}
+
 \begin{figure}

### 2018-07-10 figures/figure_mechanism.m

@@ -45,8 +45,8 @@ time = (1:length(V))*x.dt;
 plot(ax.Ca(1),time,Ca,'k')
 
 plot(ax.V(1),time,V,'k')
-set(ax.Ca(1),'XLim',[0 503],'XTick',[0 250 500],'YLim',[-2 15],'YScale','linear')
-set(ax.V(1),'XLim',[0 503],'XTick',[0 250 500],'YLim',[-80 40],'YTick',-80:40:40)
+set(ax.Ca(1),'XLim',[0 202],'XTick',[0 100 200],'YLim',[-2 15],'YScale','linear')
+set(ax.V(1),'XLim',[0 202],'XTick',[0 100 200],'YLim',[-80 40],'YTick',-80:40:40)
 
 
 % now add the calcium mech
@@ -106,7 +106,6 @@ C(:,7) = [];
 
 g = C(:,2:2:end);
 
-%% Make Figure
 
 time = (1:length(Ca))*100*1e-3;
 colours = brighten(othercolor('Mrainbow',8),.1);
@@ -124,7 +123,7 @@ ylabel(ax.Ca(3),'<[Ca^2^+]> (\muM)')
 for i = 1:size(g,2)
 	plot(ax.g,time,g(:,i),'Color',colours(i,:))
 end
-set(ax.g,'XScale','log','YScale','log')
+set(ax.g,'XScale','log','YScale','log','YLim',[1e-2 1e4])
 ylabel(ax.g,'g (\muS/mm^2)')
 xlabel(ax.g,'Time (s)')
 
@@ -149,7 +148,7 @@ plot(ax.V(5),time,V,'k')
 set(ax.V(5),'YLim',[-80 40])
 
 
-axis(ax.V(3:5),'off')
+axis(ax.V(4:5),'off')
 ax.Ca(3).XTickLabel = '';
 prettyFig('plw',1.5,'lw',1.5,'fs',15);
 
@@ -174,10 +173,16 @@ movePlot(ax.V(1:2),'up',.06)
 
 for i = 3:5
 	ax.V(i).Position(3) = .1;
+	ax.V(i).Position(4) = .08;
+	ax.V(i).Position(2) = .08;
 end
+ax.V(3).XTick = [0 .4];
+ax.V(3).XLim = [0 .4];
+ax.V(3).YTick = [-80 0 40];
 ax.V(5).Position(1) = .8;
 ax.V(4).Position(1) = .6;
 ax.V(3).Position(1) = .4;
[…]

### 2018-07-10 paper/paper.tex
figures 1-6 done.

@@
 
+\clearpage
+
 \begin{figure}
 	\centering
-	\includegraphics[width=1.0\linewidth]{gfx/puppeteer_screenshot}
-	\caption{Using the GUI to manipulate neuron parameters.
-(A) Real-time output of the \texttt{plot} function displaying voltage (colored) and intracellular calcium (black) traces of a bursting neuron model (\cite{prinzAlternativeHandtuningConductancebased2003, prinzSimilarNetworkActivity2004}.
-Colors indicate the dominant current.
-(B) Sliders control the maximal conductances, which updates on the figure.}
-	\label{fig:puppeteerscreenshot}
+	\includegraphics[width=1.0\linewidth]{gfx/figure_manipulate}
+	\caption{Manipulating neuron parameters in real time.
+Any set of parameter in the model can be manipulated; here, the maximal conductance of every conductance type in the model from Figure 1 is being manipulated using the code snippet shown here.
+A GUI with sliders for each parameter being manipualted is created, and linked to an atbitrary number of visualization functions.
+In this example, the visualization functions are the built in \texttt{plot} method (A) and a custom function that computes the firing-rate-vs.-injected current curve for this neuron (B).
+Both plots refresh themselves with every movement of any slider, allowing the user to build intuition about how that parameter controls the dynamical behavior of the model.
+A screen recording of this model being manipulated in real time is included in Supplementary Material XX.
+}
+	\label{fig:figuremanipulate}
 \end{figure}
+
+\clearpage
 

### 2018-07-10 paper/paper.tex
ready to make pdf

@@
 The cell now bursts periodically, with synchronized oscillations in intracellular Calcium.
-(box C) Schematic of Calcium-dependent integral control homeostasis.
-In this feedback system, the rates of production of mRNA depends on the Calcium levels in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which, through translation, is a function of the mRNA abundance.
+(box C) Schematic of Calcium-dependent integral control homeostasis (cite O'Leary et al.).
+In this feedback system, the rates of mRNA synthesis depend on the Calcium level in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which, through translation, is a function of the mRNA abundance.
 These integral controllers are implemented as \texttt{mechanism} objects, and can be added to conductances.
@@
 	\includegraphics[width=1.0\linewidth]{gfx/figure_snapshot}
-	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them $ad  arbitrium$.
+	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them $ad$   $arbitrium$.
 The initial state (orange disc) of the single compartment model in the previous example is saved using the \texttt{snapshot} method.
@@
 The diagram shows a three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004).
-Each neuron is modeled as a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
-Synapses are of two types (Cholinergic (dashed) or Glutamatergic (solid lines)) and have different kinetics.
-The code snippet shows how neurons can be created, wired together using synapses, and the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse.
+Each neuron is modeled with a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
+Synapses can be one of two types (Cholinergic (dashed lines) or Glutamatergic (solid lines)) and have different kinetics.
+The code snippet shows how neurons can be created, wired together using synapses, and how the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse.
 (A-C) Simulated voltage trace of a model network for the three compartments obtained from this simulation.
@@
 Any set of parameter in the model can be manipulated; here, the maximal conductance of every conductance type in the model from Figure 1 is being manipulated using the code snippet shown here.
-A GUI with sliders for each parameter being manipualted is created, and linked to an atbitrary number of visualization functions.
-In this example, the visualization functions are the built in \texttt{plot} method (A) and a custom function that computes the firing-rate-vs.-injected current curve for this neuron (B).
-Both plots refresh themselves with every movement of any slider, allowing the user to build intuition about how that parameter controls the dynamical behavior of the model.
+The screenshot shows a GUI with sliders for every parameter of interest that is created by the \texttt{manipulate} method, and linked to an arbitrary number of visualization functions.
+In this example, two visualization functions are used: the built in \texttt{plot} method (A) and a custom function that computes the firing-rate-$vs$.-injected current curve for this neuron (B).
+Both plots refresh themselves with every movement of any slider, allowing the user to build intuition about how every parameter controls the dynamical behavior of the model.
 A screen recording of this model being manipulated in real time is included in Supplementary Material XX.

### 2018-07-10 paper/paper.tex
made fonts on first two figs bigger

@@
 	\includegraphics[width=1.0\linewidth]{gfx/figure_HH}
-	\caption{\texttt{xolotl} makes it easy tp set up and simulate conductance-based models.
-Schematic representation of a single-compartment neuron with three populations of ion channels (colored rectangles).
-In the simulator, the soma is represented using an object of class $compartment$ and the populations of ion channels are represented by $conductance$ objects contained within the compartment object.
-The code snippet shown sets up this neuron model, injects current, integrates and plots the voltage, and shows activation functions, all in a few lines of code.
-(A) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current.
+	\caption{Setting up a single-compartment model and injecting current into it.
+(A) Schematic representation of a single-compartment neuron with three populations of ion channels (colored rectangles).
+(B) In the simulator, the soma is represented using an object of class $compartment$ and the populations of ion channels are represented by $conductance$ objects contained within the compartment object.
+(C) The code snippet shown sets up this neuron model, injects current, integrates and plots the voltage, and displays activation functions, all in a few lines of code.
+(D) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current.
 Colors indicate the dominant current (gold is fast sodium (NaV), blue is delayed rectifier (Kd), red is Leak).
-(C) f-I curve of this neuron.
-(D-E) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
-(F-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables}
+(E) firing rate $vs$.
+current (f-I) curve of this neuron.
+(F-G) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
+(H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables}
 	\label{fig:figurehh}
@@
 	\caption{Simulating a voltage-clamp experiment.
-The diagram shows a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998}) that is being recorded from in two-electrode voltage clamp.
-The code snippet shown here sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it.
-(A) Voltage steps that the cell is clamped to.
-(B) Current injected into the cell required to clamp it to the voltages in (A).
-(C) Asymptotic clamped current $vs.$ clamped voltage for this cell.
-(D) Accounting for the reversal potential of Potassium ions yields the conductance-voltage curve of this channel type.
-(E) Normalized conductance-voltage curves, with sigmoid fits with various exponents.
-(F) An exponent of $n=4$ yields the best fit, allowing for the characterization of the activation function of this channel type.
+(A) The diagram shows a cell with delayed rectifier potassium conductance (from \cite{liuModelNeuronActivitydependent1998}) that is being recorded from in two-electrode voltage clamp.
+(B) The code snippet shown here sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it.
+(C) Voltage steps that the cell is clamped to.
+(D) Current injected into the cell required to clamp it to the voltages in (C).
+(E) Asymptotic clamped current $vs.$ clamped voltage for this cell.
+(F) Accounting for the reversal potential of Potassium ions yields the conductance-voltage curve of this channel type.
+(G) Normalized conductance-voltage curves, with sigmoid fits with various exponents.
+(H) An exponent of $n=4$ yields the best fit, allowing for the characterization of the activation function of this channel type.
 }

### 2018-07-10 figures/figure_HH.m

@@ -31,10 +31,7 @@ for ii = 1:4
 end
 
 %% Make Cartoon Cell
-
-image(ax(3), imread('fig_HH.png'))
-axis(ax(3), 'off');
-ax(3).Tag = 'cartoon';
+showImageInAxes(ax(3),imread('fig_HH.png'))
 
 
 % integrate and obtain the current traces
@@ -122,10 +119,12 @@ for ii = 1:length(conductance)
   end
 
   % plot onto the correct axes
-  plot(ax(7),   V,  minf);
-  plot(ax(8),   V,  hinf);
-  plot(ax(9),   V,  taum);
-  plot(ax(10),  V,  tauh);
+  plot(ax(7),   V,  minf,'Color',c(ii,:));
+  plot(ax(9),   V,  taum,'Color',c(ii,:));
+  if ii == 3
+    plot(ax(8),   V,  hinf,'Color',c(ii,:));
+    plot(ax(10),  V,  tauh,'Color',c(ii,:));
+  end
 end
 
 

### 2018-07-10 figures/figure_clamp.m

@@ -1,7 +1,7 @@
 %% Figure 5: Voltage Clamp
 
 % set up xolotl object
-clearvars
+close all
 x = xolotl;
 x.add('compartment','AB','A',.06)
 x.AB.add('Kd','gbar', 300);
@@ -31,12 +31,6 @@ ax.m_inf =  subplot(3,3,8); hold on
 ax.r2 =  subplot(3,3,9); hold on
 
 
-
-
-% %% Make Xolotl Readout from MATLAB
-
-image(ax.cartoon, imread('clamp_cartoon.png'))
-axis(ax.cartoon, 'off')
 
 %% Set Up Voltage Clamp
 N = floor(x.t_end/x.sim_dt);
@@ -106,6 +100,7 @@ for ii = 1:length(all_V_step)
 end
 % plot the steady state at various powers
 
+clear hplot
 for ii = 1:length(all_n)
 	plot(ax.m_inf, all_V_step, minf.^all_n(ii), 'Color', c(ii, :))
 	hplot(ii) = plot(ax.m_inf, NaN, NaN, 'o', 'MarkerFaceColor', c(ii, :), 'MarkerEdgeColor', c(ii, :), 'MarkerSize', 8);
@@ -138,12 +133,38 @@ set(ax.r2, 'XLim', [0.5 6.5], 'XTick', 1:6)
 
 %% Post-Processing
 
-prettyFig('fs', 18, 'lw', 1)
-return
+prettyFig('fs', 18, 'lw', 1.5,'plw',2)
 
+
+
+showImageInAxes(ax.cartoon,imread('clamp_cartoon.png'))
+
+showImageInAxes(ax.code,imread('clamp_code.png'))
+
+ax.code.Position = [ 0.5500    0.600    0.3346    0.3474];
+ax.cartoon.Position = [ 0.1000    0.6000    0.3346    0.3474];
+
+
+movePlot(ax.V,'left',.03)
+movePlot(ax.g,'left',.03)
+
+
+movePlot(ax.max_I,'right',.03)
+movePlot(ax.r2,'right',.03)
 
 % label the subplots
-labelFigure('capitalise', true,'ignore_these',ax(1:3),'column_first',true,'y_offset',-.035,'x_offset',-.04)
+labelAxes(ax.cartoon,'A','x_offset',-.01,'y_offset',-.13,'font_size',24);
[…]

### 2018-07-11 paper/paper.tex
ready to make network figure (again)

@@
 	\caption{Modeling intracellular mechanisms.
-(box A) A single-compartment neuron model with 8 channel types.
-With these parameters, this model neuron is a tonic spiker.
-Since there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant (top panel).
-(box B)  An intracellular buffering of Calcium and the effect of Calcium influx through voltage gated Calcium channels (VGCCs) is modeled using a simple differential equation.
-Code snippet shows how this mechanism can be added to the neuron model.
-The cell now bursts periodically, with synchronized oscillations in intracellular Calcium.
-(box C) Schematic of Calcium-dependent integral control homeostasis (cite O'Leary et al.).
+(A) A single-compartment neuron model with 8 channel types.
+Since there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant (B), and the cell tonically spikes (C).
+Intracellular Calcium buffering and influx through voltage gated Calcium channels (VGCCs) can be modeled using a simple differential equation (D).
+Code snippet shows how this mechanism can be added to the neuron model (E).
+The cell now bursts periodically, with synchronized oscillations in intracellular Calcium (F-G).
+(H) Schematic of Calcium-dependent integral control homeostasis (cite O'Leary et al.).
 In this feedback system, the rates of mRNA synthesis depend on the Calcium level in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which, through translation, is a function of the mRNA abundance.
-These integral controllers are implemented as \texttt{mechanism} objects, and can be added to conductances.
-On integrating the model, intracellular calcium levels rise and approach the target (red dashed line).
-This is accompanied by an increase in the conductance densities of all channels being controlled by this homeostatic mechanism.
+(I) The code snippet shows how these integral controllers are implemented as \texttt{mechanism} objects, and can be added to conductances.
+(J) On integrating the model, intracellular calcium levels rise and approach the target (red dashed line).
+This is accompanied by an increase in the conductance densities of all channels being controlled by this homeostatic mechanism (K).
 The voltage behavior of the cell changes from silence to bursting with truncated spikes to regular bursting.}

### 2018-07-11 figures/figure_network.m

@@ -3,7 +3,7 @@
 % set up xolotl object
 % conversion from Prinz to phi
 
-clear x
+clear x g
 
 
 x = xolotl;
@@ -14,9 +14,13 @@ g(:,3) = [1000 24  20 500  0   1250 .5];
 g(:,2) = [1000 0   40 200  0   250  .5];
 g(:,1) = [1000 25  60 500  50  1000 .1];
 
-x.add('AB','compartment','vol',.0628,'phi',906);
-x.add('LP','compartment','vol',.0628,'phi',906);
-x.add('PY','compartment','vol',.0628,'phi',906);
+x.add('compartment','AB');
+x.add('compartment','LP');
+x.add('compartment','PY');
+
+x.AB.add('CalciumMech1');
+x.LP.add('CalciumMech1');
+x.PY.add('CalciumMech1');
 
 compartments = x.find('compartment');
 for j = 1:length(compartments)
@@ -99,7 +103,7 @@ synaptic_currents = synaptic_currents(:,2:2:end);
 c = lines(10);
 plot(ax(7), time, synaptic_states(:,6),'r');
 
-xlabel(ax(7), 'time (s)')
+xlabel(ax(7), 'Time (s)')
 ylabel(ax(7), 's_{PY\rightarrowLP} ')
 set(ax(7), 'YScale', 'linear','YLim',[0 1])
 xlim(ax(7), [0 max(time)]);
@@ -109,18 +113,6 @@ xlim(ax(7), [0 max(time)]);
 
 prettyFig('fs', 14, 'plw', 1.5,'lw',1.5)
 
-% set the positions of the axes
-pos = [ ...
-     0.13       0.7127       0.1237       0.2123;
-     0.13       0.4131       0.1237       0.2123;
-     0.03       0.05          0.26         0.92;
-   0.3812       0.7932       0.4736       0.1577;
-   0.3812       0.5741       0.4736       0.1577;
-   0.3812        0.355       0.4736       0.1577;
-   0.3812       0.1359       0.4736       0.1577];
-for ii = 1:length(ax)
-	ax(ii).Position = pos(ii, :);
-end
 
 % remove boxes
 for ii = 1:length(ax)

### 2018-07-11 figures/network_snippet.m

@@ -1,12 +1,11 @@
 
-x.add('AB','compartment','vol'...
-	,.0628,'phi',906);
-x.add('LP','compartment','vol'...
-	,.0628,'phi',906);
-x.add('PY','compartment','vol'...
-	,.0628,'phi',906);
 
-% set up conductances (not shown)
+x.add('compartment','AB');
+x.add('compartment','LP');
+x.add('compartment','PY');
+
+% set up conductances 
+% and calcium mechanisms (not shown)
 
 % set up synapses 
 x.connect('AB','LP','Chol','gbar',30);

### 2018-07-11 paper/paper.tex
figs 1-6 modified; all fonts bigger

@@
 	\caption{Simulating a network of conductance-based model neurons.
-The diagram shows a three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (Prinz \textit{et al.} 2004).
+(A) A three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (cite Prinz \textit{et al.} 2004).
 Each neuron is modeled with a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
 Synapses can be one of two types (Cholinergic (dashed lines) or Glutamatergic (solid lines)) and have different kinetics.
-The code snippet shows how neurons can be created, wired together using synapses, and how the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse.
-(A-C) Simulated voltage trace of a model network for the three compartments obtained from this simulation.
-(D) Time series activation variable of the Glutamatergic synapse between PY and LP (red connection in diagram) shows how the synapse becomes active every time the PY neuron spikes.}
+(B) The code snippet shows how neurons can be created, wired together using synapses, and how the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse.
+(C-E) Simulated voltage trace of a model network for the three compartments obtained from this simulation.
+(F) Time series activation variable of the Glutamatergic synapse between PY and LP (red connection in diagram) shows how the synapse becomes active every time the PY neuron spikes.}
 	\label{fig:figurenetwork}

### 2018-07-11 figures/figure_manipulate.m

@@ -89,11 +89,11 @@ plot(ax.fI,all_I_ext,all_f,'r')
 xlabel(ax.fI,'I_{ext} (nA)')
 ylabel(ax.fI,'Firing rate (Hz)')
 
-ax.V.Position = [.33 .2 .4 .75];
+ax.V.Position = [.4 .2 .34 .75];
 ax.fI.Position = [.8 .2 .16 .75];
 
 
-prettyFig();
+prettyFig('fs',19);
 
 
 

### 2018-07-11 figures/figure_network.m

@@ -54,34 +54,22 @@ c = lines(100);
 
 clear ax
 
-% cartoon cell
-ax(1) = subplot(3,5,1);
-ax(1).Visible = 'off';
-% xolotl structure
-ax(2) = subplot(3,5,6);
-ax(2).Visible = 'off';
-% xolotl printout
-ax(3) = subplot(3,5,11);
-% voltage trace
-ax(4) = subplot(4,5,2:5); hold on;
-ax(5) = subplot(4,5,7:10); hold on;
-ax(6) = subplot(4,5,12:15); hold on;
+ax.network = subplot(4,5,1); hold on
+
+ax.code = subplot(4,5,6);
+
+% voltage traces
+ax.AB = subplot(4,5,2:5); hold on;
+ax.LP = subplot(4,5,7:10); hold on;
+ax.PY = subplot(4,5,12:15); hold on;
+
 % synaptic currents
-ax(7) = subplot(4,5,17:20); hold on;
-
-%% Make Cartoon Cell
-
-% %% Make Xolotl Readout from MATLAB
-
-image(ax(3), imread('network.png'))
-axis(ax(3), 'off')
+ax.S = subplot(4,5,17:20); hold on;
 
 
 %% Make Voltage Trace
+comp_names   = x.find('compartment');
 
-c           = lines(100);
-nameComps   = x.find('compartment');
-nComps      = length(nameComps);
 
 % integrate and obtain the current traces
 x.closed_loop = true;
@@ -91,37 +79,52 @@ x.integrate;
 time        = 1e-3 * x.dt * (1:length(V));
 
 % plot the voltage
-for ii = 1:nComps
-  plot(ax(ii+3), time, V(:,ii), 'k')
-	set(ax(ii+3), 'XTickLabel', [],'XLim',[0 max(time)],'YLim',[-80 50])
-	ylabel(ax(ii+3), ['V_{' nameComps{ii} '} (mV)'])
+for i = 1:3
+  plot(ax.(comp_names{i}), time, V(:,i), 'k')
+	set(ax.(comp_names{i}), 'XTickLabel', [],'XLim',[0 max(time)],'YLim',[-80 50])
+	ylabel(ax.(comp_names{i}), ['V_{' comp_names{i} '} (mV)'])
 end
 
 % plot the synaptic currents
 synaptic_states = synaptic_currents(:,1:2:end);
[…]

### 2018-07-11 paper/paper.tex
wrote new results section for mechanisms

@@
 In this section, we illustrate how "xolotl" can be used to generate a variety of models and simulate them.
-These examples have been chosen to illustrate various features of "xolotl", and are intended to serve as a template that researchers and educators can use for their own needs.
+These examples have been chosen to illustrate various features of "xolotl", and are intended to serve as templates that researchers and educators can build upon.
 
@@
 
-In this section, we demonstrate how to simulate a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model, and gain insight into the underlying biophysical mechanism.
-The schematic in Figure \ref{fig:figurehh} shows how the hierarchical structure of the ion channel populations within the soma is represented in the simulator, and how to create a compartment (here, labelled "HH") and add three populations of ion channels to it.
+In this example, we demonstrate how to simulate the spiking activity of a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model, and gain insight into the underlying biophysical mechanism.
+This simple model consists of a single-compartment neuron with three types of conductances  (Figure \ref{fig:figurehh}A).
+This heirarchical organization of the neuron is mirrored in the structure of the model: an object of type "compartment" is used to represent the cell body, and three objects of type "conductance" are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B).
 These models of ion channels were obtained from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator.
+The syntax of constructing the neuron model (using "add" statements) preserves this herirarchical organization, and the code snippet in (Figure \ref{fig:figurehh}C) shows how this model can be created using just a few lines of code.
+
 Adding an injected current and calling the built-in "plot" function plots the time series of membrane voltage.
@@
 When 0.2 nA is injected, the model tonically spikes (Figure \ref{fig:figurehh}A-B).
-The \texttt{plot} function generates a voltage trace, which is colored by the dominant current (Figure \ref{fig:figurehh}A).
+The \texttt{plot} function displays a voltage trace colored by the dominant current (Figure \ref{fig:figurehh}A).
 Colors in the voltage trace indicate the strongest instantaneous inward current when the voltage is increasing, and strongest instantaneous outward current when the voltage is decreasing.
@@
 The model can be integrated for various amplitudes of injected current to determine its f-I curve, or the relationship between injected current and its firing rate (cite).
-Figure \ref{fig:figurehh}C shows the f-I curve of this model, obtained by repeated integration of the model.
+Figure \ref{fig:figurehh}E shows the f-I curve of this model, obtained by repeated integration of the model.
 Finally, the built-in \texttt{show} function can plot activation ($m$) and inactivation ($h$) curves and voltage-dependent timescales of any channel type in the simulator (Figure \ref{fig:figurehh}D-G).
@@
 Voltage clamping is a experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (cite D\&A).
-Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp.
+Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp (Figure \ref{fig:figureclamp}A).
 By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized.
-(insert all those citations as needed).
+(alec:insert all those citations as needed).
 
 "xolotl" can reproduce such a voltage clamp experiment $in silico$.
-Figure \ref{fig:figureclamp} illustrates how a simple model with a single compartment and a single ion channel type can be set up and clamped to a desired voltage.
+Figure \ref{fig:figureclamp}B illustrates how a simple model with a single compartment and a single ion channel type can be set up and clamped to a desired voltage.
 Integrating the model yields the current required to clamp the cell at that voltage.
@@
 First, the cell is clamped to a number of different voltages (Fig.
-\ref{fig:figureclamp}A) and the resultant clamping currents are measured by integrating the model (Fig.
-\ref{fig:figureclamp}B).
+\ref{fig:figureclamp}C) and the resultant clamping currents are measured by integrating the model (Fig.
+\ref{fig:figureclamp}D).
 The asymptotic clamping currents depend on the clamping voltage in a nonlinear manner (Fig.
-\ref{fig:figureclamp}C), since the open probabilities of the channel are functions of the membrane voltage.
+\ref{fig:figureclamp}E), since the open probabilities of the channel are functions of the membrane voltage.
 Assuming the reversal potential is known, Eq.
 \eqref{eq:2} can be used to solve for the total conductance of the channel as a function of the clamping voltage (Fig.
-\ref{fig:figureclamp}D).
+\ref{fig:figureclamp}F).
 Finally, a sigmoid can be fit to the normalized conductance-voltage curves to obtain the activiation function of the ion channel population (Fig.
-\ref{fig:figureclamp}E-F).
+\ref{fig:figureclamp}G-H).
 "xolotl" can therefore be used to describe graphically the theoretical underpinnings of ion channel characterization through voltage clamp and can serve as an effective pedagogical tool in computational neuroscience.
@@
 
-"xolotl" can implement homeostatic tuning rules as integral control.
-The controller computes an error signal (typically a function of intracellular calcium concentration), and adjusts the conductance or synapse it controls accordingly (\cite{olearyCorrelationsIonChannel2013}).
-In "xolotl", integral controllers are "cpplab" objects added to the conductance or synapse they regulate.
+So far, the models we described only considered the voltage dynamics of a cell (the solution to Eq.
+\ref{eq:1}).
+However, real neurons possess a several dynamical features, stemming from a number of intracellular mechanisms.
+"xolotl" makes it possible to model and include arbitrary intracellular mechanisms.
+In our simulator, these intracellular mechanisms are represented by the "mechanism" object, and can be bound to "compartments", "conductances", and other object types.
 
-In a demonstration adapted from \cite{olearyCorrelationsIonChannel2013}, integral control changes maximal conductances to bring a neuron from quiescence into a bursting regime.
-Calcium sensors supervene on maximal conductance density (Figure \ref{fig:figureintegralcontrol}) to change neuronal activity.
-Each conductance in the "xolotl" structure contains a calcium-sensitive controller (Figure \ref{fig:figureintegralcontrol}B-C).
-Maximal conductances increase from random initial conditions to a set which elicits the desired network output by minimizing the error signal (Figure \ref{fig:figureintegralcontrol}D-F).
+One of the most important intracellular mechanisms is the cytosolic buffering of Calcium and its influx through voltage-gated Calcium channels (VGCCs).
+Figure \ref{fig:figuremechanism}A shows a model of a single-compartment model with 8 populations of ion channels (alec: cite Liu).
+Without any mechanism for Calcium influx or buffering, the intracellular Calcium levels in this model do not change (Figure \ref{fig:figuremechanism}B) and the model tonically spikes (Figure \ref{fig:figuremechanism}C).
+A common method of modelling Calcium buffering and influx is to use an differential equation that increases intracellular Calcium with the current through Calcium channels, and relaxes back to a baseline value (alec: cite Liu, Prinz, DayanAbbott) (Figure \ref{fig:figuremechanism}D).
+This mechanism exists in the "xolotl" library as "CalciumMech1" and can be added to the model using a simple "add" statement (Figure \ref{fig:figuremechanism}E).
+The intracellular Calcium in the model now oscillates periodically (Figure \ref{fig:figuremechanism}F), synchronised to bursts in action potentials in this cell (Figure \ref{fig:figuremechanism}G).
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Another intracellular mechanism of interest is a neuron's ability to regulate its electrical activity by controlling the spectrum of ion channels its expresses.
+Here, we will show "xolotl" can be used to model a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration and some target Calcium (alec: cite Tim) (Figure \ref{fig:figuremechanism}H).
+Since this mechanism affects each ion channel population individually, object corresponding to this mechanism is added to each "condutance" object in the neuron (Figure \ref{fig:figuremechanism}I).
+Setting all maximal conductance densities to some low value and integrating the model shows that the intracellular Calcium levels rise over time and approach the target (Figure \ref{fig:figuremechanism}J), while all conductance densities increase and then remain bounded (Figure \ref{fig:figuremechanism}K).
+Examining the voltage dynamics of the cell reveal that it transitions from silence to truncated bursts of action potentials to periodic bursting as this mechanism regualtes the neuron's ion channel spectrum.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
@@
 %
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2018-07-11 paper/paper.tex
wrote snapshot section

@@
 
-\subsection{Using snapshots to explore model dynamics}
+\subsection{Using snapshots to explore model dynamics and parameters}
+
+Switching back and forth points in parameter space and state space of a neuron model is a common occurrence in working with neuron models.
+"xolotl" makes it easy to remember bookmark configurations of a model and return to them at will.
+Internally, "xolotl" does not distinguish between parameters and dynamic variables of a model, and uses the "serialize" method to gather all parameters and dynamic variables into a vector of values, that is passed to the underlying "C++" implementation.
+A paired "deserialize" method is used to update all parameters and variables in the object tree from this vector.
+This architecture provides a natural framework for representing the state of any model, no matter how complex, using a vector of numbers.
+The "snapshot" method built into "xolotl" leverages this schema to save the entire state of the model in a named variable, that can be accessed using another built in method called "reset".
+
+(Figure \ref{fig:figuresnapshot}) illustrates how these features can be used to explore the dynamics of the model presented in the previous section.
+First, the current state of the model is saved using the "snapshot" method into a namespace called "initial".
+In this state, the model exhibits periodic bursting activity due to a particular configuration of maximal conductance densities (Figure \ref{fig:figuremechanism}, orange).
+On setting the maximal conductances of the Calcium-permissive channels to 0, the model switches to a tonic spiking activity (Figure \ref{fig:figuremechanism}, purple).
+Integrating the model for a longer duration allows the homeostatic control mechanism in the cell to restore the conductance profile and bursting activity to a state close to the original state (Figure \ref{fig:figuremechanism}, green).
+This state is saved using the name "final".
+
+The initial state can be returned to using the "reset" method, and a new manipulation to the model can be explored.
+Here, the intracellular Calcium target is modified, and the model is re-integrated, to yield a different voltage activity (Figure \ref{fig:figuremechanism}, blue).
+At the end of this numerical exploration, any of the saved states can be quickly returned to using the "reset" method, making the process of re-initializing models to desired states and parameters both error-free and painless.
 
@@
 
-\subsection{Auditability and Reproducability}
+\subsection{Auditability and Reproducibility}
 

### 2018-07-11 paper/supplementary/supplementary.tex
Add supplementary information .tex file and auxiliaries

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when producing your supplementary material for your Frontiers article.                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 2.5 Generated 2018/06/15 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
%%%  NB logo1.jpg is required in the path in order to correctly compile front page header %%%

\documentclass[utf8]{frontiers_suppmat} % for all articles
\usepackage{url,hyperref,lineno,microtype}
\usepackage[onehalfspacing]{setspace}

% bibliography
\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage[
backend=biber, 
bibencoding=utf8, 
style=authoryear, 
date=year, 
maxbibnames=6, 
minbibnames=6, 
url=false, 
isbn=false, 
eprint=false,
uniquename = false, 
autocite=inline]{biblatex}
\addbibresource{conductances.bib}

% make pretty MATLAB code
\usepackage{listings}
\usepackage{matlab-prettifier}
\lstMakeShortInline[style=Matlab-editor]"

% Leave a blank line between paragraphs instead of using \\

\begin{document}
\onecolumn
\firstpage{1}

\title[Supplementary Material]{{\helveticaitalic{Supplementary Material}}:
\\ \helvetica{Xolotl: An Intuitive and Approachable Neuron \& Network Simulator in MATLAB}} %Please insert the title of your article here

\maketitle

\section{Database of Conductances}

Dynamics for model conductances in the following papers have been implemented in "xolotl".

\begin{enumerate}
	\item \cite{bronkRegulationEagCa22018}
	\item \cite{brookingsAutomaticParameterEstimation2014}
	\item \cite{caplanManyParameterSets2014a}
	\item \cite{dethierPositiveFeedbackCellular2015a}
	\item \cite{drionDopaminePacemakerNeuron2017}
	\item \cite{goldmanGlobalStructureRobustness2001a}
	\item \cite{herasModulationVoltagedependentConductances2018}
	\item \cite{kisperskyIncreaseSodiumConductance2012a}
	\item \cite{linActivityDependentAlternativeSplicing2012}
	\item \cite{liuModelNeuronActivityDependent1998}
	\item \cite{nadimFrequencyRegulationSlow1998}
	\item \cite{nadimSynapticDepressionCreates1999}
	\item \cite{odowdVoltageclampAnalysisSodium1988}
	\item \cite{prinzAlternativeHandtuningConductancebased2003a}
	\item \cite{rodriguezConvergentRhythmGeneration2013}
	\item \cite{sharpDynamicClampComputergenerated1993a}
	\item \cite{soto-trevinoActivitydependentModificationInhibitory2001}
	\item \cite{soto-trevinoComputationalModelElectrically2005a}
	\item \cite{swensenModulatorsConvergentCellular2001a}
	\item \cite{turrigianoSelectiveRegulationCurrent1995a}
	\item \cite{wicherNonsynapticIonChannels2001}
\end{enumerate}

\printbibliography

\end{document}

### 2018-07-11 paper/supplementary/supplementary.tex

@@
 \begin{enumerate}
-	\item \cite{bronkRegulationEagCa22018}
-	\item \cite{brookingsAutomaticParameterEstimation2014}
-	\item \cite{caplanManyParameterSets2014a}
-	\item \cite{dethierPositiveFeedbackCellular2015a}
-	\item \cite{drionDopaminePacemakerNeuron2017}
-	\item \cite{goldmanGlobalStructureRobustness2001a}
-	\item \cite{herasModulationVoltagedependentConductances2018}
-	\item \cite{kisperskyIncreaseSodiumConductance2012a}
-	\item \cite{linActivityDependentAlternativeSplicing2012}
-	\item \cite{liuModelNeuronActivityDependent1998}
-	\item \cite{nadimFrequencyRegulationSlow1998}
-	\item \cite{nadimSynapticDepressionCreates1999}
-	\item \cite{odowdVoltageclampAnalysisSodium1988}
-	\item \cite{prinzAlternativeHandtuningConductancebased2003a}
-	\item \cite{rodriguezConvergentRhythmGeneration2013}
-	\item \cite{sharpDynamicClampComputergenerated1993a}
-	\item \cite{soto-trevinoActivitydependentModificationInhibitory2001}
-	\item \cite{soto-trevinoComputationalModelElectrically2005a}
-	\item \cite{swensenModulatorsConvergentCellular2001a}
-	\item \cite{turrigianoSelectiveRegulationCurrent1995a}
-	\item \cite{wicherNonsynapticIonChannels2001}
+	\item EAGes, EAGmut, EAGwt (\cite{bronkRegulationEagCa22018})
+	\item A-type (\cite{brookingsAutomaticParameterEstimation2014})
+	\item MICurrent (\cite{caplanManyParameterSets2014a})
+	\item CaT, H, Kd, NaV (\cite{dethierPositiveFeedbackCellular2015a})
+	\item CaL, CaPump, KCa, Kd, NaV (\cite{drionDopaminePacemakerNeuron2017})
+	\item Ap, At, Cal, KCa, KCaT, Proc (\cite{goldmanGlobalStructureRobustness2001a})
+	\item Shaker (\cite{hardieNovelPotassiumChannels1991})
+	\item Novel, Shab, Shaker (\cite{herasModulationVoltagedependentConductances2018})
+	\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{kisperskyIncreaseSodiumConductance2012a})
+	\item NaP, NaT (\cite{linActivityDependentAlternativeSplicing2012})
+	\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{liuModelNeuronActivityDependent1998})
+	\item \cite{liuModelNeuronActivityDependent1998} conductances with cached gating functions
+	\item \cite{liuModelNeuronActivityDependent1998} conductances with forward Euler integration
+	\item \cite{liuModelNeuronActivityDependent1998} conductances with temperature dependence
+	\item Kd and NaV conductances from Int1, LG, and MCN1 cells (\cite{nadimFrequencyRegulationSlow1998})
+	\item Cal (\cite{nadimSynapticDepressionCreates1999})
+	\item Drosophila NaV (\cite{odowdVoltageclampAnalysisSodium1988})
+	\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{prinzAlternativeHandtuningConductancebased2003a})
+	\item \cite{prinzAlternativeHandtuningConductancebased2003a} conductances with cached gating functions
+	\item \cite{prinzAlternativeHandtuningConductancebased2003a} conductances with temperature dependence
+	\item Kd and NaV from Int1 cells and modulatory input and transient LTS conductances from LG cells (\cite{rodriguezConvergentRhythmGeneration2013})
+	\item proctolin modulatory input (\cite{sharpDynamicClampComputergenerated1993a})
+	\item Cal and Kd from AB-PD, LP, and PY cells (\cite{soto-trevinoActivitydependentModificationInhibitory2001})
+	\item A-type, CaT, and KCa from AB and PD cells, generic H, Kd, NaP, NaV, and proctolin conductances (\cite{soto-trevinoComputationalModelElectrically2005a})
+	\item proctolin modulatory input (\cite{swensenModulatorsConvergentCellular2001a})
+	\item A-type, CaS, CaT, H, KCa, Kd, NaP, NaV (\cite{turrigianoSelectiveRegulationCurrent1995a})
+	\item Drosophila NaV (\cite{wicherNonsynapticIonChannels2001})
 \end{enumerate}

### 2018-07-27 paper/paper.tex
Add citations and begin editing benchmarks section

@@
 These simulators tend to perform fast computations with little overhead, but suffer from a steep learning curve.
-Wrapping these simulators in a more approachable language like "Python" and the use of graphical user interfaces (GUIs) mitigate these drawbacks, at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
+Wrapping these simulators in a more approachable language like "Python" and the use of graphical user interfaces (GUIs) mitigate these drawbacks (\cite{hinesNEURONPython2009}), at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
 Another approach is to design simulators that can be used entirely in popular scientific programming languages.
 These simulators can be easier to use, and can benefit from inter-compatibility with other tools.
-Programs like "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}) allow the user to specify models using strings of equations or using components that are specified in the scientific programming language, that can then be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
+Programs like "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}), "morphforge" (\cite{hullMorphforgeToolboxSimulating2014}), "PyNN" (\cite{davisonPyNNCommonInterface2009}) allow the user to specify models using strings of equations or using components that are specified in the scientific programming language, that can then be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
 This approach permits considerable flexibility for simulating systems of differential equations.
@@
 Here, \( \bar{g}_{j} \) is the maximal conductance, and \( E_{j} \) is the reversal potential of the ion channel population \( j \).
-\( A_{i} \) is the area of compartment \( i\) that contains these ion channels.
+\( A_{i} \) is the surface area of compartment \( i\) that contains these ion channels.
 \( m_{j} \) and \( h_{j} \) are activation and inactivation variables that change according to
@@
 Typically, \( \tau_{m} \), \( \tau_{h} \), \( m_{\infty} \), and \( h_{\infty} \) are functions of the membrane potential \( V_{i} \).
-The software uses integration schemes that have been specifically developed to solve equations of this form (cite D\&A, Hines, others), though other schemes can be used if desired.
+The software uses integration schemes that have been specifically developed to solve equations of this form (\cite{dayanTheoreticalNeuroscience2001, hinesEfficientComputationBranched1984, ohErrorAnalysisSpecialized2006}), though other schemes can be used if desired.
 
@@
 \paragraph{New mechanisms require new \texttt{C++} code.} Adding new network components requires writing new "C++" code.
-Adding a new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinMeasurementCurrentvoltageRelations1952, dayanTheoreticalNeuroscience2001}), requires creating a new "C++" header file, though this is generally trivial.
+Adding a new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, dayanTheoreticalNeuroscience2001}), requires creating a new "C++" header file, though this is generally trivial.
 Implementing a new integration scheme or component type requires much more in-depth knowledge of the underlying "C++" core code.
@@
 Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}).
-As one of the simplest models of excitable neural mebranes, the Hodgkin-Huxley model often serves as a the first model introduced in pedagogical literature (cite some textbooks).
+As one of the simplest models of excitable neural mebranes, the Hodgkin-Huxley model often serves as a the first model introduced in pedagogical literature (\cite{dayanTheoreticalNeuroscience2001, sterrattPrinciplesComputationalModelling2011, trappenbergFundamentalsComputationalNeuroscience2010}).
 
@@
 The "manipulate" function opens the GUI which permits visualization of changing parameters in real-time.
-Moving sliders representing the values of network parameters updates a plot (Figure \ref{fig:puppeteerscreenshot}B).
-By default, the function opens a figure displaying the results of the "plot" function, which shows the voltage and intracellular calcium traces for each compartment (Figure \ref{fig:puppeteerscreenshot}A).
+Moving sliders representing the values of network parameters updates a plot (Figure \ref{fig:figuremanipulate}B).
+By default, the function opens a figure displaying the results of the "plot" function, which shows the voltage and intracellular calcium traces for each compartment (Figure \ref{fig:figuremanipulate}A).
 "manipulate" grants slider control over all "xolotl" parameters by default, but specific ones can be selected by passing them as arguments.
@@
 	\caption{\texttt{xolotl} benchmarked against \texttt{DynaSim} and \texttt{NEURON}.
-(A) Ratio of 'simulated' time to runtime (speed factor) and accuracy, measured by spike train coincidence plotted against decreasing time-resolution.
-(B) Speed factor for models at increasing simulation times.
-(C) Speed factor over number of compartments.}
+The top row benchmarks using a tonically-firing Hodgkin-Huxley model with 3 conductances with constant injected current.
+The bottom row repeats the same benchmarks using a bursting stomatogastric ganglion model with 8 conductances.
+(A,E) Ratio of run-time to simulation time (speed factor) over increasing integration time step.
+(B,F) Error over increasing integration time step as computed by the LeMasson matrix where the baseline condition is the voltage trace at $\Delta t = 0.001$ ms.
+(C,G) Simulations over increasing simulation time where $\Delta t = 0.1$ ms.
+(D,H) Simulations over increasing number of unconnected compartments.}
 	\label{fig:figurebenchmark}

### 2018-07-27 figures/benchmarks/model_stability.m
Add second panel to stability analysis

@@ -85,11 +85,33 @@ end
 % generate a figure
 figure('outerposition',[100 100 1550 666],'PaperUnits','points','PaperSize',[1000 1000]); hold on
 % create subplots
-ax(1) = subplot(1,2,1);
+ax(1) = subplot(1,2,1); hold on; ax(2) = subplot(1,2,2); hold on;
+
 % plot the error over time-step
 plot(ax(1), all_dt, Q, '-o')
 set(ax(1), 'XScale','log','YScale','log')
 xlabel(ax(1), '\Deltat (ms)')
 ylabel(ax(1), 'Simulation error (\epsilon_{HH})')
 
+% simulate the first and last traces for the first set of parameters
+model = 1;
+x.t_end = 10e3;
+x.set('*gbar', params(1:8, model));
+x.AB.phi = params(9, model);
+x.sim_dt = all_dt(1);
+first = x.integrate;
+x.sim_dt = all_dt(end);
+last = x.integrate;
+time = x.dt / 1000 * (1:length(last));
+% plot the first and last traces for the first set of parameters
+plot(ax(2), time, first, 'k');
+plot(ax(2), time, last, 'r');
+xlabel(ax(2), 'time (ms)')
+ylabel(ax(2), 'V_m (mV)')
+xlim(ax(2), [2.9 3.0])
+
 prettyFig()
+
+for ii = 1:length(ax)
+  box(ax(ii), 'off');
+end

### 2018-07-27 paper/paper.tex
Add many citations

@@
 Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}).
-As one of the simplest models of excitable neural mebranes, the Hodgkin-Huxley model often serves as a the first model introduced in pedagogical literature (\cite{dayanTheoreticalNeuroscience2001, sterrattPrinciplesComputationalModelling2011, trappenbergFundamentalsComputationalNeuroscience2010}).
+As one of the simplest models of excitable neural membranes, the Hodgkin-Huxley model often serves as a the first model introduced in pedagogical literature (\cite{dayanTheoreticalNeuroscience2001, sterrattPrinciplesComputationalModelling2011, trappenbergFundamentalsComputationalNeuroscience2010}).
 
@@
 This heirarchical organization of the neuron is mirrored in the structure of the model: an object of type "compartment" is used to represent the cell body, and three objects of type "conductance" are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B).
-These models of ion channels were obtained from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator.
+These models of ion channels were obtained from \cite{liuModelNeuronActivityDependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator.
 The syntax of constructing the neuron model (using "add" statements) preserves this herirarchical organization, and the code snippet in (Figure \ref{fig:figurehh}C) shows how this model can be created using just a few lines of code.
@@
 Integrating the model yields the current required to clamp the cell at that voltage.
-Here, we use a delayed-rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998}) and simulate a voltage-clamp experiment whose goal is to infer the activation function of this channel.
+Here, we use a delayed-rectifier potassium conductance (\cite{liuModelNeuronActivityDependent1998}) and simulate a voltage-clamp experiment whose goal is to infer the activation function of this channel.
 First, the cell is clamped to a number of different voltages (Fig.
@@
 One of the most important intracellular mechanisms is the cytosolic buffering of Calcium and its influx through voltage-gated Calcium channels (VGCCs).
-Figure \ref{fig:figuremechanism}A shows a model of a single-compartment model with 8 populations of ion channels (alec: cite Liu).
+Figure \ref{fig:figuremechanism}A shows a model of a single-compartment model with 8 populations of ion channels (\cite{liuModelNeuronActivityDependent1998}).
 Without any mechanism for Calcium influx or buffering, the intracellular Calcium levels in this model do not change (Figure \ref{fig:figuremechanism}B) and the model tonically spikes (Figure \ref{fig:figuremechanism}C).
-A common method of modelling Calcium buffering and influx is to use an differential equation that increases intracellular Calcium with the current through Calcium channels, and relaxes back to a baseline value (alec: cite Liu, Prinz, DayanAbbott) (Figure \ref{fig:figuremechanism}D).
+A common method of modelling Calcium buffering and influx is to use an differential equation that increases intracellular Calcium with the current through Calcium channels, and relaxes back to a baseline value (\cite{liuModelNeuronActivityDependent1998, prinzAlternativeHandtuningConductancebased2003, dayanTheoreticalNeuroscience2001}) (Figure \ref{fig:figuremechanism}D).
 This mechanism exists in the "xolotl" library as "CalciumMech1" and can be added to the model using a simple "add" statement (Figure \ref{fig:figuremechanism}E).
@@
 Another intracellular mechanism of interest is a neuron's ability to regulate its electrical activity by controlling the spectrum of ion channels its expresses.
-Here, we will show "xolotl" can be used to model a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration and some target Calcium (alec: cite Tim) (Figure \ref{fig:figuremechanism}H).
+Here, we will show "xolotl" can be used to model a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration and some target Calcium (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H).
 Since this mechanism affects each ion channel population individually, object corresponding to this mechanism is added to each "condutance" object in the neuron (Figure \ref{fig:figuremechanism}I).
@@
 
-Single-compartment Hodgkin-Huxley-like models were generated using conductance dynamics from \cite{liuModelNeuronActivitydependent1998}) in the simulation environments.
+Single-compartment Hodgkin-Huxley-like models with three conductances and stomatogastric ganglion neuron models with eight conductances were generated using conductance dynamics from \cite{liuModelNeuronActivityDependent1998}) in the simulation environments.
 "xolotl" uses the exponential Euler method for integrating membrane potential (\cite{dayanTheoreticalNeuroscience2001}).
@@
 
-To compare the integration methods, models were simulated for 5 s at varying time-resolution (Figure \ref{fig:figurebenchmark}A).
+To compare the integration methods, models were simulated for 30 s at varying time-resolution (Figure \ref{fig:figurebenchmark}A,E).
 The speed factor was defined as the ratio between time represented in the simulation and actual runtime (simulation-time).
 Therefore, the speed factor represents how many times faster the simulation is than a real-time observation.
-The coincidence factor determines the correlative overlap between two spike trains (\cite{jolivetBenchmarkTestQuantitative2008}).
-In this context, it is a measure of accuracy in the numerical simulation.
-Unity indicates perfect correlation.
-To assess accuracy over decreasing time-resolution for the three simulation environments, spike trains at each resolution were compared to a `canonical' spike train (exponential Euler at a time-step of $dt = 0.001$ ms).
+The error resulting from coarser time-resolution was computed as the distance between two LeMasson matrices (\cite{lemassonIntroductionEquationSolving2000}), one representing the canonical voltage trace with time step $\Delta t = 0.001$ ms and the other representing the voltage trace simulated at time steps up to $1$ ms (Figure \ref{fig:figurebenchmark}B,F).
 
-To assess the performance of the simulators in absence of set-up overhead, models were simulated with a time-resolution of $0.1$ ms over increasing simulation time (Figure \ref{fig:figurebenchmark}B).
+To assess the performance of the simulators in absence of set-up overhead, models were simulated with a time-resolution of $0.1$ ms over increasing simulation time (Figure \ref{fig:figurebenchmark}C,G).
 
 Many simulators perform well in simulations of many compartments (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}).
-To assess how "xolotl" under these conditions, networks of up to 1,000 Hodgkin-Huxley cells were simulated for 5 s at a time-resolution of 0.1 ms (Figure \ref{fig:figurebenchmark}C).
+To assess how "xolotl" performs under these conditions, networks of up to 1,000 neuron models were simulated for 30 s at a time-resolution of 0.1 ms (Figure \ref{fig:figurebenchmark}D,H).
 

### 2018-07-27 paper/paper.tex
More citations

@@
 Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp (Figure \ref{fig:figureclamp}A).
-By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized.
-(alec:insert all those citations as needed).
+By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized (\cite{coleIonicCurrentMeasurements1960, coleIonsPotentialsNerve1955, hodgkinEffectSodiumIons1949, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, turrigianoSelectiveRegulationCurrent1995}).
 

### 2018-07-30 paper/paper.tex
Add to comparison section

@@
 
-- compare with BRIAN, GENESIS, XPPAUT, NEURON, DynaSim
+- compare with BRIAN, GENESIS, XPPAUT, NEURON, DynaSim \\
 - selling point is ease of use and integration into MATLAB
+
+\begin{quote}
+	\textit{``About half the time spent on a typical simulation project involves creating and tuning the model.
+Thus, a good user interface may contribute more to the overall efficiency of a project than pure computation speed.''}
+	\begin{flushright}
+		\cite{deschutterConsumerGuideNeuronal1992}
+	\end{flushright}
+\end{quote}
+
+"xolotl" is designed with ease-of-use in mind.
+This includes the time it takes to install, setup, and learn the software, the time to write and debug scripts, and the time to perform the simulations.
+An easy-to-use simulation environment must minimize time spent in all these domains, especially during human engagement with the software.
+Complicated software remains broadly inaccessible and time-consuming even to perform single-compartment simulations, though the actual simulation time may be very small.
+
+The front-end interface to "xolotl" is built in "MATLAB" to facilitate interoperability with existing tools.
+"MATLAB" provides a comprehensive distribution from a single vendor with straightforward and powerful toolboxes for time series analysis, optimization, and parallel computing.
+For these reasons, "MATLAB" has remained a popular tool for data analysis and visualization in neuroscience.
+Since "xolotl" can be run directly from the "C++" executable, it would be possible to have implemented the interface in other popular scientific computing languages (\viz "Python" or "Julia").
+Ultimately due to the popularity and ease of "MATLAB" for pedagogy and research, and its extensive support for vector and matrix mathematics, "xolotl" supports a "MATLAB" interface.
+
+In contrast to "BRIAN" and "DynaSim" and in concordance with "NEURON", "xolotl" represents models in a compartment-oriented manner.
+Compartment-oriented simulators consider membrane compartments first, representing the neuron model hierarchically.
+Compartments contain conductances, synaptic connections, and mechanisms (injected current, \etc).
+The dynamical variables hide within the hierarchical structure.
+Generally, the user specifies the features of and in the compartment to define the model; the dynamical variables are hidden within the structure.
+
+"xolotl" allows the user to explore the structure in the "MATLAB" console and see all properties and defining equations at each level of the model.
+Numerical properties can be changed in the console and are reflected in future simulations.
+The equations, structure, parameters, and methods by which the simulation proceeds are transparently available to the user.
+
+In equation-oriented specifications such as "BRIAN" and "DynaSim", all equations which define the model must be explicitly written.
+The equations maintain primacy in the model structure and are parsed and interpreted by the numerical integration algorithm.
+This has the added feature of total customization, as the equations can be edited in-place, but long strings of characters can become unwieldy and are prone to syntax mistakes.
+In addition, editing a feature of the model or simulation requires directly editing the mechanism or equation script and cannot be done in the console.
+
+"xolotl" cannot match the total flexibility of equation-oriented specifications, which can integrate arbitrary dynamical systems, but avoids some of the most common problems associated with equation-oriented specification.
+While the equations "xolotl" solves cannot be changed without adding a mechanism or editing the specifying text file, numerical parameters and network components can be added at any time.
+The user can inspect the "xolotl" network object in the console and represent its structure.
+Changes can be made on-the-fly.
+In addition, "xolotl" components come specified by a "C++" header file, greatly reducing typing errors during model specification.
 

### 2018-07-31 paper/paper.tex
Fix broken citations and minor fixes

@@
 \def\Authors{Srinivas Gorur-Shandilya\,$^{1*\dagger}$, Alec Hoyland\,$^{1\dagger}$, and Eve Marder$^{1}$}
-\def\Address{$^{1}$Volen Center for Complex Systems and Biology Department \\ Brandeis University \\ Waltham, MA 02453 \\ USA
+\def\Address{$^{1}$Volen National Center for Complex Systems and Biology Department \\ Brandeis University \\ Waltham, MA 02453 \\ USA
 \vspace{0.25cm}
@@
 %\def\corrAddress{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
-\def\corrAddress{Volen Center for Complex Systems, Brandeis University, Waltham MA 02453}
+\def\corrAddress{Volen National Center for Complex Systems, Brandeis University, Waltham MA 02453}
 \def\corrEmail{\texttt{srinivasgs@brandeis.edu}}
@@
 Nervous systems process and transmit information using electrically excitable membranes.
-Conductance-based models are a powerful biophysical simplification of an electrically excitable compartment in a neuron (\cite{hodgkinQuantitativeDescriptionMembrane1952}).
+Conductance-based models are a powerful biophysical simplification of an electrically excitable compartment in a neuron (\cite{hodgkinQuantitativeDescriptionMembrane1952a}).
 Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small circuits (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}).
@@
 \paragraph{New mechanisms require new \texttt{C++} code.} Adding new network components requires writing new "C++" code.
-Adding a new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, dayanTheoreticalNeuroscience2001}), requires creating a new "C++" header file, though this is generally trivial.
+Adding a new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952a, hodgkinQuantitativeDescriptionMembrane1952a, dayanTheoreticalNeuroscience2001}), requires creating a new "C++" header file, though this is generally trivial.
 Implementing a new integration scheme or component type requires much more in-depth knowledge of the underlying "C++" core code.
@@
 The axon of the giant squid contains contains a fast inactivating sodium conductance ("NaV"), a slower non-inactivating potassium conductance ("Kd"), and a passive leak current.
-Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}).
+Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952a}).
 As one of the simplest models of excitable neural membranes, the Hodgkin-Huxley model often serves as a the first model introduced in pedagogical literature (\cite{dayanTheoreticalNeuroscience2001, sterrattPrinciplesComputationalModelling2011, trappenbergFundamentalsComputationalNeuroscience2010}).
@@
 Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp (Figure \ref{fig:figureclamp}A).
-By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized (\cite{coleIonicCurrentMeasurements1960, coleIonsPotentialsNerve1955, hodgkinEffectSodiumIons1949, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, turrigianoSelectiveRegulationCurrent1995}).
+By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized (\cite{coleIonicCurrentMeasurements1960, coleIonsPotentialsNerve1955, hodgkinEffectSodiumIons1949, hodgkinMeasurementCurrentvoltageRelations1952a, hodgkinQuantitativeDescriptionMembrane1952a, turrigianoSelectiveRegulationCurrent1995}).
 
@@
 
-- reproducibility crisis in computational science 
-- unlike experiemntal research, where reproducability can arise due to deep reasons, reasons for computational research irreproducability are mostly trivial: 
-- multiple sources: software being used is closed-source, researcher's code is not made available; obscure architecture makes improper use possible. 
-- how xolotl conributes to reoridubile science: (1) freely available, code for this paper available too for others to inspect and replicate (2) intutitive syntax, solid documentation.
+- reproducibility crisis in computational science \\
+- unlike experimental research, where reproducibility can arise due to deep reasons \\
+- reasons for computational research irreproducibility are mostly trivial:  \\
+- multiple sources: software being used is closed-source, researcher's code is not made available; obscure architecture makes improper use possible.  \\
+- how xolotl contributes to reproducible science: (1) freely available, code for this paper available too for others to inspect and replicate (2) intuitive syntax, solid documentation.
 (3) architectural choices that ensure accountability: object hashing, parameter hashing possible.

### 2018-07-31 paper/paper.tex
Add to supplementary information, fix some spelling

@@
 This simple model consists of a single-compartment neuron with three types of conductances  (Figure \ref{fig:figurehh}A).
-This heirarchical organization of the neuron is mirrored in the structure of the model: an object of type "compartment" is used to represent the cell body, and three objects of type "conductance" are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B).
+This hierarchical organization of the neuron is mirrored in the structure of the model: an object of type "compartment" is used to represent the cell body, and three objects of type "conductance" are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B).
 These models of ion channels were obtained from \cite{liuModelNeuronActivityDependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator.
-The syntax of constructing the neuron model (using "add" statements) preserves this herirarchical organization, and the code snippet in (Figure \ref{fig:figurehh}C) shows how this model can be created using just a few lines of code.
+The syntax of constructing the neuron model (using "add" statements) preserves this hierarchical organization, and the code snippet in (Figure \ref{fig:figurehh}C) shows how this model can be created using just a few lines of code.
 
@@
 \ref{fig:figureclamp}F).
-Finally, a sigmoid can be fit to the normalized conductance-voltage curves to obtain the activiation function of the ion channel population (Fig.
+Finally, a sigmoid can be fit to the normalized conductance-voltage curves to obtain the activation function of the ion channel population (Fig.
 \ref{fig:figureclamp}G-H).
@@
 Here, we will show "xolotl" can be used to model a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration and some target Calcium (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H).
-Since this mechanism affects each ion channel population individually, object corresponding to this mechanism is added to each "condutance" object in the neuron (Figure \ref{fig:figuremechanism}I).
+Since this mechanism affects each ion channel population individually, object corresponding to this mechanism is added to each "conductance" object in the neuron (Figure \ref{fig:figuremechanism}I).
 Setting all maximal conductance densities to some low value and integrating the model shows that the intracellular Calcium levels rise over time and approach the target (Figure \ref{fig:figuremechanism}J), while all conductance densities increase and then remain bounded (Figure \ref{fig:figuremechanism}K).
@@
 
-"C" is statically-typed, with procedural syntax that provides low-level access to memory (\cite{kernighanProgrammingLanguage1978}, providing significant advantages for time-intensive computations.
+"C" is statically-typed, with procedural syntax that provides low-level access to memory (\cite{kernighanProgrammingLanguage1978}), providing significant advantages for time-intensive computations.
 Unfortunately, automatic code-generation is limited by the supervening language.
@@
 
-AH received funding from National Institute on Drug Abuse (NIDA) through the undergraduate training grant in computational neuroscience (1R90DA033463-01).
+AH received funding from the National Institute on Drug Abuse (NIDA) through the undergraduate training grant in computational neuroscience (1R90DA033463-01).
 
@@
 	\caption{Simulating a network of conductance-based model neurons.
-(A) A three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (cite Prinz \textit{et al.} 2004).
+(A) A three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (\cite{prinzSimilarNetworkActivity2004}).
 Each neuron is modeled with a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.

### 2018-07-31 paper/supplementary/supplementary.tex
Add to supplementary information, fix some spelling

@@
 		\item A-type (\cite{brookingsAutomaticParameterEstimation2014})
-	\item MICurrent (\cite{caplanManyParameterSets2014a})
-	\item CaT, H, Kd, NaV (\cite{dethierPositiveFeedbackCellular2015a})
+		\item Proctolin modulatory input current (\cite{caplanManyParameterSets2014})
+		\item CaT, H, Kd, NaV (\cite{dethierPositiveFeedbackCellular2015})
 		\item CaL, CaPump, KCa, Kd, NaV (\cite{drionDopaminePacemakerNeuron2017})
-	\item Ap, At, Cal, KCa, KCaT, Proc (\cite{goldmanGlobalStructureRobustness2001a})
+		\item Ap, At, Cal, KCa, KCaT, Proc (\cite{goldmanGlobalStructureRobustness2001})
 		\item Shaker (\cite{hardieNovelPotassiumChannels1991})
 		\item Novel, Shab, Shaker (\cite{herasModulationVoltagedependentConductances2018})
-	\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{kisperskyIncreaseSodiumConductance2012a})
+		\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{kisperskyIncreaseSodiumConductance2012})
 		\item NaP, NaT (\cite{linActivityDependentAlternativeSplicing2012})
@@
 		\item Kd and NaV conductances from Int1, LG, and MCN1 cells (\cite{nadimFrequencyRegulationSlow1998})
-	\item Cal (\cite{nadimSynapticDepressionCreates1999})
+		\item Calcium current (\cite{nadimSynapticDepressionCreates1999})
 		\item Drosophila NaV (\cite{odowdVoltageclampAnalysisSodium1988})
-	\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{prinzAlternativeHandtuningConductancebased2003a})
-	\item \cite{prinzAlternativeHandtuningConductancebased2003a} conductances with cached gating functions
-	\item \cite{prinzAlternativeHandtuningConductancebased2003a} conductances with temperature dependence
+		\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{prinzAlternativeHandtuningConductancebased2003})
+		\item \cite{prinzAlternativeHandtuningConductancebased2003} conductances with cached gating functions
+		\item \cite{prinzAlternativeHandtuningConductancebased2003} conductances with temperature dependence
 		\item Kd and NaV from Int1 cells and modulatory input and transient LTS conductances from LG cells (\cite{rodriguezConvergentRhythmGeneration2013})
-	\item proctolin modulatory input (\cite{sharpDynamicClampComputergenerated1993a})
+		\item proctolin modulatory input (\cite{sharpDynamicClampComputergenerated1993})
 		\item Cal and Kd from AB-PD, LP, and PY cells (\cite{soto-trevinoActivitydependentModificationInhibitory2001})
-	\item A-type, CaT, and KCa from AB and PD cells, generic H, Kd, NaP, NaV, and proctolin conductances (\cite{soto-trevinoComputationalModelElectrically2005a})
-	\item proctolin modulatory input (\cite{swensenModulatorsConvergentCellular2001a})
-	\item A-type, CaS, CaT, H, KCa, Kd, NaP, NaV (\cite{turrigianoSelectiveRegulationCurrent1995a})
+		\item A-type, CaT, and KCa from AB and PD cells, generic H, Kd, NaP, NaV, and proctolin conductances (\cite{soto-trevinoComputationalModelElectrically2005})
+		\item Proctolin modulatory input (\cite{swensenModulatorsConvergentCellular2001})
+		\item A-type, CaS, CaT, H, KCa, Kd, NaP, NaV (\cite{turrigianoSelectiveRegulationCurrent1995})
 		\item Drosophila NaV (\cite{wicherNonsynapticIonChannels2001})
 	\end{enumerate}
+
+\section{Parameters for Simulations}
+
+	Two single-compartment models were simulated in this paper.
+The first, a Hodgkin-Huxley model with three conductances ("NaV", "Kd", and "Leak") and injected current was simulated for Figure 1 and 7.
+The second is a stomatogastric neuron model with eight conductances ("NaV", "CaT", "CaS", "ACurrent", "KCa", "Kd", "HCurrent", "Leak") and was simulated for Figure 3 and 7.
+Both models have dynamics as described in \cite{liuModelNeuronActivityDependent1998}.
+
+	\subsection{Parameters for Hodgkin-Huxley model}
+
+		\begin{center}
+			\begin{tabular}{|l|l|}
+				\hline 
+				\textbf{Parameter Name} & \textbf{Value} \\ 
+				\hline 
+				Membrane capacitance ("HH.Cm") & 10 ${nF}/{mm^2}$ \\ 
+				\hline 
+				Surface area ("HH.A") & 0.01 $mm^2$ \\ 
+				\hline 
+				Fast sodium maximal conductance ("HH.NaV.gbar") & 1000 $\mu S/mm^2$ \\ 
+				\hline 
+				Delayed rectifier maximal conductance ("HH.Kd.gbar") & 300 $\mu S/mm^2$ \\ 
+				\hline 
+				Passive leak maximal conductance ("HH.Leak.gbar") & 1 $\mu S/mm^2$ \\ 
+				\hline 
+				Sodium reversal potential ("HH.NaV.E") & 50 $mV$ \\ 
+				\hline 
+				Potassium reversal potential ("HH.Kd.E") & -80 $mV$ \\ 
+				\hline
+				Leak reversal potential ("HH.Leak.E") & -50 $mV$ \\
+				\hline
+				Injected current ("I_ext") & 0.2 $nA$ \\
+				\hline 
+			\end{tabular} 
+		\end{center}
+
+	\subsection{Parameters for Stomatogastric Model}
+
+		\begin{center}
+			\begin{tabular}{|l|l|}
+				\hline 
+				\textbf{Parameter Name} & \textbf{Value} \\ 
+				\hline 
+				Membrane capacitance ("AB.Cm") & 10 ${nF}/{mm^2}$ \\ 
+				\hline 
+				Surface area ("AB.A") & 0.0628 $mm^2$ \\
+				\hline
+				Calcium buffering $\phi$ ("phi") & 90 $\mu M / nA$ \\
+				\hline
+				Calcium buffering shell volume ("vol") & 0.0628 $mm^3$ \\
+				\hline 
+				Fast sodium maximal conductance ("AB.NaV.gbar") & 00 $\mu S/mm^2$ \\ 
+				\hline
+				Transient calcium maximal conductance ("AB.CaT.gbar") & 00 $\mu S/mm^2$ \\ 
+				\hline 
+				Slow calcium maximal conductance ("AB.CaS.gbar") & 00 $\mu S/mm^2$ \\ 
+				\hline 
+				Transient potassium maximal conductance ("AB.ACurrent.gbar") & 00 $\mu S/mm^2$ \\ 
+				\hline 
+				Calcium-gated potassium maximal conductance ("AB.KCa.gbar") & 00 $\mu S/mm^2$ \\ 
+				\hline 
+				Delayed rectifier maximal conductance ("AB.Kd.gbar") & 00 $\mu S/mm^2$ \\ 
+				\hline
+				Hyperpolarization-activated maximal conductance ("AB.HCurrent.gbar") & 00 $\mu S/mm^2$ \\ 
+				\hline 
+				Passive leak maximal conductance ("AB.Leak.gbar") & 00 $\mu S/mm^2$ \\ 
+				\hline 
+				Sodium reversal potential ("AB.NaV.E") & 30 $mV$ \\ 
+				\hline 
+				Potassium reversal potential ("AB.Kd.E") & -80 $mV$ \\ 
+				\hline
+				Hyperpolarization-activated reversal potential ("AB.HCurrent.E") & -20 $mV$ \\
+				\hline
+				Leak reversal potential ("AB.Leak.E") & -50 $mV$ \\
+				\hline
+			\end{tabular} 
+		\end{center}
+
+	\subsection{Parameters for Network Model}
+		The network model displayed in Figure 4 is described in \cite{prinzAlternativeHandtuningConductancebased2003, prinzSimilarNetworkActivity2004}.
+It is comprised of three compartments "ABPD", "LP", and "PY" with the same dynamics but differing parameters and synaptic inputs.
+
+		\begin{center}
+			\begin{tabular}{|l|c|c|c|c|}
+				\hline 
+				\textbf{Parameter Name} & "ABPD" & "LP" & "PY" & \textbf{Units} \\ 
+				\hline 
+				Fast sodium maximal conductance & 1000 & 1000 & 1000 & $\mu S/mm^2$ \\ 
+				\hline 
+				Transient calcium maximal conductance & 25 & 0 & 24 & $\mu S/mm^2$ \\  
+				\hline 
+				Slow calcium maximal conductance & 60 & 40 & 20 & $\mu S/mm^2$ \\  
+				\hline 
+				Transient calcium maximal conductance & 500 & 200 & 500 & $\mu S/mm^2$ \\  
+				\hline 
+				Calcium-gated potassium maximal conductance & 50 & 0 & 0 & $\mu S/mm^2$ \\  
+				\hline 
+				Delayed rectifier maximal conductance & 1000 & 250 & 1250 & $\mu S/mm^2$ \\  
+				\hline 
+				Hyperpolarization-activated maximal conductance & 0.1 & 0.5 & 0.5 & $\mu S/mm^2$ \\  
+				\hline 
+				Leak maximal conductance & 0 & 0.3 & 0.1 & $\mu S/mm^2$ \\  
+				\hline 
+			\end{tabular}
+		\end{center}
+
+		There are seven synapses in the model of two types: glutamatergic ("Glut") and cholinergic ("Chol") with differing dynamics.
+The synapse produces a current in the postsynaptic compartment based on the membrane potential in the presynaptic compartment.
+
+		\begin{center}
+			\begin{tabular}{c|c|c|c|}
+				\hline
+				\textbf{Presynaptic} & \textbf{Postsynaptic} & \textbf{Type} & \textbf{Maximal Conductance} \\
+				\hline
+
+			\end{tabular}
+		\end{center}
 

### 2018-08-02 paper/paper.tex
Fix spelling errors

@@
 However, several challenges remain in working with biophysically-detailed conductance-based neuron models.
-First, these models can be high-dimensional with many nonlinear differential equations, with several parameters.
+First, these models can be high-dimensional with many nonlinear differential equations, each with several parameters.
 Second, many or all equations in these models can be strongly coupled through variables like the membrane potential.
@@
 Since models in the "MATLAB" workspace are automatically linked to models in the "C++" implementation, configuring these objects in "MATLAB" transparently configures the underlying "C++" objects.
-"xolotl" comes packaged with hundreds of components that can be used to assemble cells and networks;  has built-in visualization functions to inspect voltage time traces and activation functions; and a graphical user interface (GUI) for real-time manipulation of model parameters.
+"xolotl" comes packaged with hundreds of components that can be used to assemble cells and networks;  has built-in visualization functions to inspect voltage time traces and activation functions; and a GUI for real-time manipulation of model parameters.
 The GUI permits real-time manipulation of any network parameter and an immediate visualization of the effect of that parameter on arbitrary dynamical features of the model.

### 2018-08-06 paper/paper.tex
Add citations

@@
 This architecture automatically creates objects in "MATLAB" that represent the underlying object-oriented "C++" code, without having to re-specify these objects in "MATLAB".
-In this paper, we introduce "xolotl", an implementation of the "cpplab" system specialized in integrating conductance-based neuron and network models.
+In this paper, we introduce "xolotl", an implementationclose of the "cpplab" system specialized in integrating conductance-based neuron and network models.
 Models can be easily constructed from network building blocks in a few lines of "MATLAB" code using a hierarchical and intuitive syntax.
@@
 
-The model can be integrated for various amplitudes of injected current to determine its f-I curve, or the relationship between injected current and its firing rate (cite).
+The model can be integrated for various amplitudes of injected current to determine its f-I curve, or the relationship between injected current and its firing rate (\cite{kisperskyIncreaseSodiumConductance2012}).
 Figure \ref{fig:figurehh}E shows the f-I curve of this model, obtained by repeated integration of the model.
@@
 
-Voltage clamping is a experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (cite D\&A).
+Voltage clamping is a experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (\cite{dayanTheoreticalNeuroscience2001}).
 Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp (Figure \ref{fig:figureclamp}A).
@@
 	\caption{Simulating a voltage-clamp experiment.
-(A) The diagram shows a cell with delayed rectifier potassium conductance (from \cite{liuModelNeuronActivitydependent1998}) that is being recorded from in two-electrode voltage clamp.
+(A) The diagram shows a cell with delayed rectifier potassium conductance (from \cite{liuModelNeuronActivityDependent1998}) that is being recorded from in two-electrode voltage clamp.
 (B) The code snippet shown here sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it.
@@
 The cell now bursts periodically, with synchronized oscillations in intracellular Calcium (F-G).
-(H) Schematic of Calcium-dependent integral control homeostasis (cite O'Leary et al.).
+(H) Schematic of Calcium-dependent integral control homeostasis (\cite{olearyCellTypesNetwork2014}).
 In this feedback system, the rates of mRNA synthesis depend on the Calcium level in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which, through translation, is a function of the mRNA abundance.

### 2018-08-06 paper/paper.tex
minor fixes

@@
 
-- failure of averaging
-- complex interrelationships b/w parameters and dynamics
-- not clear what parameter does what
-- traditional approach: vary parameters, but for more than 2 D, how do you even visualize this? also slow. expensive to set up. 
-- need something where parameter variation can be linked to dynamics, easy to set up, easy to use.
+- failure of averaging \\
+- complex interrelationships b/w parameters and dynamics\\
+- not clear what parameter does what\\
+- traditional approach: vary parameters, but for more than 2 D, how do you even visualize this? also slow. expensive to set up. \\
+- need something where parameter variation can be linked to dynamics, easy to set up, easy to use. \\
 
@@
 
-- components from library vs. equations
-- equations are transparent -- you know what you're solving. but clunky -- string parsing headaches, etc. 
-- components are easy to assemble, but it;s hard to know where they are, how they can be changed, etc.
+- components from library vs. equations\\
+- equations are transparent -- you know what you're solving. but clunky -- string parsing headaches, etc. \\
+- components are easy to assemble, but it;s hard to know where they are, how they can be changed, etc.\\
 - xolotl goes the component route, but makes it very easy to find components (you can click your way into the C++ files).
 components are stored as source files, so you can modify it and immediately see it reflected in your model.
-we didn't invent another synatax for these components -- these are just C++ files.
+we didn't invent another syntax for these components -- these are just C++ files.
+\\
 

### 2018-08-07 paper/paper.tex
wrote accountability section

--- 2018-08-02 paper/paper.tex
@@
 
-- reproducibility crisis in computational science \\
-- unlike experimental research, where reproducibility can arise due to deep reasons \\
-- reasons for computational research irreproducibility are mostly trivial:  \\
-- multiple sources: software being used is closed-source, researcher's code is not made available; obscure architecture makes improper use possible.  \\
-- how xolotl contributes to reproducible science: (1) freely available, code for this paper available too for others to inspect and replicate (2) intuitive syntax, solid documentation.
-(3) architectural choices that ensure accountability: object hashing, parameter hashing possible.
+The use of computational tools is increasingly central to the scientific method; yet, the lack of auditability and accountability in their use has led to a crisis of credibility affecting many scientific fields (cite Victoria Stodden).
+Unlike in experimental research, where a lack of reproducibility can arise due to deep reasons like unknown differences in experimental protocols or intrinsic variability, the reasons for irreproducibility in computational research are largely trivial and include: a) typographical errors from transcribing out model parameters and equations, b) obscure software design that leads to users not knowing precisely what equations the software is solving, or the parameters it using, c) incompatibility with version control systems and rolling software development which leads to ambiguity in which version of the software was used to generate a particular result, and d) convoluted architectures that make it too complex for non-experts to understand the inner workings of the software.
 
-"xolotl" fosters reproducibility in science.
-While the availability of hosting sites with version control (\viz GitHub (\url{https://github.com}), GitLab (\url{https://gitlab.com/}), and Open Science Framework (\url{https://osf.io/})) and the push for reproducibility in computational science (\cite{eklundClusterFailureWhy2016, stoddenEnhancingReproducibilityComputational2016, bakerWhyScientistsMust2016}) has resulted in the availability of source code, much of this code base is bespoke and difficult to implement (\cite{sedanoCodeReadabilityTesting2016, xuMeasurementSourceCode2017}).
+Our software was designed to tackle all these threats, and has features that allow the user to answer the following questions in the affirmative: ``Am I doing what I think I am doing?'' and ``Is it possible for others to reproduce exactly what I have just done?''.
+Our goal was to design software that would allow the user to verify for herself that the software was running as she intended, and to be able to reproduce results from others quickly and unambiguously.
+The primary design choice of our software that enables auditability and reproducibility is that every simulation is tied to a alphanumeric checksum, or hash, using the md5 algorithm (cite Rivest 1992).
+The hash is computed from every C++ file that is part of the model, and any changes in the C++ file will trigger a recompilation of that model.
+Thus, the hash guarantees that a given model is derived from a set of source files, obviating any ambiguity about the code used to generate a model.
+In addition, every parameter and state variable in the model can also be hashed together with underlying code, allowing the user to generate a short checksum that guarantees with high probability that every aspect of the model -- code, parameters and initial conditions -- are exactly as they should be.
 
-To this end, "xolotl" provides an environment with readability and reproducibility in mind.
-Each network is hashed to provide a unique alphanumeric identifier.
-Conductance header files are easily viewed in the "xolotl" source files; conductances in "MATLAB" contain links to the full path of the generating file.
-
-\subsection{Circumventing Language Tradeoffs}
-
-Executing "C/C++" code in higher-level languages such as "MATLAB" or "Python" often provides speed improvements for iterative code in algorithms.
-
-"C" is statically-typed, with procedural syntax that provides low-level access to memory (\cite{kernighanProgrammingLanguage1978}), providing significant advantages for time-intensive computations.
-Unfortunately, automatic code-generation is limited by the supervening language.
-"MATLAB", for instance, cannot use pointers or pass by reference, which limits the efficiency of "C" code automatically generated from "MATLAB".
-Conversely, custom "C/C++" code provides significant increases in performance and memory conservation, but lacks the ease-of-use and flexibility of scripting languages.
-
-"xolotl" handles this problem through symbolic manipulation of "C++" objects in "MATLAB".
-Built from the ground up in "C++", "xolotl" maintains all the advantages of custom compiled code, but can run in "MATLAB" without the user having to touch the "C++" code.
-"xolotl" represents compartment, conductance, synapse, and controllers as "cpplab" objects, which template from underlying "C++" header files.
-In this way, properties of the "xolotl" network can be examined and changed using object-oriented paradigms.
-The object specifies the "integrate" function, not the other way around.
-
-\subsection{Applications of the \texttt{cpplab} architecture}
-
-NEED TO WRITE THIS
+Typographical errors from transcribing model parameters and equations can also be detected using hashes, and the component-oriented architecture of our software makes it easy to debug code and spot errors.
+Our software has been designed so that it is possible to explore the model interactively in the command line, and it is possible to ``click through'' from the highest level of the model in the command line all the way down to the underlying code of any component in the model.
+This design allows the user to know precisely the equations being solved, and view the code that numerically integrates them.
+Finally, the core of our software is written in a few hundred lines of code and contains just four classes: compartments, conductances, synapses and mechanisms.
+This architectural simplicity permits a motivated user to understand the entire organization of our code in short period of time.
 
@@
 
-SG-S \& AH designed and wrote the software, and wrote the manuscript. AH created the online user documentation. EM supervised the project. All authors reviewed the paper.
+SG-S \& AH designed and wrote the software. AH created the online user documentation. EM supervised the project. All authors wrote and reviewed the paper.
 

### 2018-08-07 paper/supplementary/supplementary.tex
Add more to tables

@@
 		\item Kd and NaV from Int1 cells and modulatory input and transient LTS conductances from LG cells (\cite{rodriguezConvergentRhythmGeneration2013})
-		\item proctolin modulatory input (\cite{sharpDynamicClampComputergenerated1993})
+		\item Proctolin modulatory input (\cite{sharpDynamicClampComputergenerated1993})
 		\item Cal and Kd from AB-PD, LP, and PY cells (\cite{soto-trevinoActivitydependentModificationInhibitory2001})
@@
 		\begin{center}
-			\begin{tabular}{|l|l|}
+			\begin{tabular}{|l|r|l|}
 				\hline 
-				\textbf{Parameter Name} & \textbf{Value} \\ 
+				\textbf{Parameter Name} & \textbf{Value} & \textbf{Units} \\ 
 				\hline 
-				Membrane capacitance ("HH.Cm") & 10 ${nF}/{mm^2}$ \\ 
+				Membrane capacitance ("HH.Cm") & 10 & ${nF}/{mm^2}$ \\ 
 				\hline 
-				Surface area ("HH.A") & 0.01 $mm^2$ \\ 
+				Surface area ("HH.A") & 0.01 & $mm^2$ \\ 
 				\hline 
-				Fast sodium maximal conductance ("HH.NaV.gbar") & 1000 $\mu S/mm^2$ \\ 
+				Fast sodium maximal conductance ("HH.NaV.gbar") & 1000 & $\mu S/mm^2$ \\ 
 				\hline 
-				Delayed rectifier maximal conductance ("HH.Kd.gbar") & 300 $\mu S/mm^2$ \\ 
+				Delayed rectifier maximal conductance ("HH.Kd.gbar") & 300 & $\mu S/mm^2$ \\ 
 				\hline 
-				Passive leak maximal conductance ("HH.Leak.gbar") & 1 $\mu S/mm^2$ \\ 
+				Passive leak maximal conductance ("HH.Leak.gbar") & 1 & $\mu S/mm^2$ \\ 
 				\hline 
-				Sodium reversal potential ("HH.NaV.E") & 50 $mV$ \\ 
+				Sodium reversal potential ("HH.NaV.E") & 50 & $mV$ \\ 
 				\hline 
-				Potassium reversal potential ("HH.Kd.E") & -80 $mV$ \\ 
+				Potassium reversal potential ("HH.Kd.E") & -80 & $mV$ \\ 
 				\hline
-				Leak reversal potential ("HH.Leak.E") & -50 $mV$ \\
+				Leak reversal potential ("HH.Leak.E") & -50 & $mV$ \\
 				\hline
-				Injected current ("I_ext") & 0.2 $nA$ \\
+				Injected current ("I_ext") & 0.2 & $nA$ \\
 				\hline 
@@
 		\begin{center}
-			\begin{tabular}{|l|l|}
+			\begin{tabular}{|l|r|l|}
 				\hline 
-				\textbf{Parameter Name} & \textbf{Value} \\ 
+				\textbf{Parameter Name} & \textbf{Value} & \textbf{Units} \\ 
 				\hline 
-				Membrane capacitance ("AB.Cm") & 10 ${nF}/{mm^2}$ \\ 
+				Membrane capacitance ("AB.Cm") & 10 & ${nF}/{mm^2}$ \\ 
 				\hline 
-				Surface area ("AB.A") & 0.0628 $mm^2$ \\
+				Surface area ("AB.A") & 0.0628 & $mm^2$ \\
 				\hline
-				Calcium buffering $\phi$ ("phi") & 90 $\mu M / nA$ \\
+				Calcium buffering $\phi$ ("AB.phi") & 90 & $\mu M / nA$ \\
 				\hline
-				Calcium buffering shell volume ("vol") & 0.0628 $mm^3$ \\
+				Calcium buffering shell volume ("AB.vol") & 0.0628 & $mm^3$ \\
 				\hline 
-				Fast sodium maximal conductance ("AB.NaV.gbar") & 00 $\mu S/mm^2$ \\ 
+				Fast sodium maximal conductance ("AB.NaV.gbar") & 00 & $\mu S/mm^2$ \\ 
 				\hline
-				Transient calcium maximal conductance ("AB.CaT.gbar") & 00 $\mu S/mm^2$ \\ 
+				Transient calcium maximal conductance ("AB.CaT.gbar") & 00 & $\mu S/mm^2$ \\ 
 				\hline 
-				Slow calcium maximal conductance ("AB.CaS.gbar") & 00 $\mu S/mm^2$ \\ 
+				Slow calcium maximal conductance ("AB.CaS.gbar") & 00 & $\mu S/mm^2$ \\ 
 				\hline 
-				Transient potassium maximal conductance ("AB.ACurrent.gbar") & 00 $\mu S/mm^2$ \\ 
+				Transient potassium maximal conductance ("AB.ACurrent.gbar") & 00 & $\mu S/mm^2$ \\ 
 				\hline 
-				Calcium-gated potassium maximal conductance ("AB.KCa.gbar") & 00 $\mu S/mm^2$ \\ 
+				Calcium-gated potassium maximal conductance ("AB.KCa.gbar") & 00 & $\mu S/mm^2$ \\ 
 				\hline 
-				Delayed rectifier maximal conductance ("AB.Kd.gbar") & 00 $\mu S/mm^2$ \\ 
+				Delayed rectifier maximal conductance ("AB.Kd.gbar") & 00 & $\mu S/mm^2$ \\ 
 				\hline
-				Hyperpolarization-activated maximal conductance ("AB.HCurrent.gbar") & 00 $\mu S/mm^2$ \\ 
+				Hyperpolarization-activated maximal conductance ("AB.HCurrent.gbar") & 00 & $\mu S/mm^2$ \\ 
 				\hline 
-				Passive leak maximal conductance ("AB.Leak.gbar") & 00 $\mu S/mm^2$ \\ 
+				Passive leak maximal conductance ("AB.Leak.gbar") & 00 & $\mu S/mm^2$ \\ 
 				\hline 
-				Sodium reversal potential ("AB.NaV.E") & 30 $mV$ \\ 
+				Sodium reversal potential ("AB.NaV.E") & 30 & $mV$ \\ 
 				\hline 
-				Potassium reversal potential ("AB.Kd.E") & -80 $mV$ \\ 
+				Potassium reversal potential ("AB.Kd.E") & -80 & $mV$ \\ 
 				\hline
-				Hyperpolarization-activated reversal potential ("AB.HCurrent.E") & -20 $mV$ \\
+				Hyperpolarization-activated reversal potential ("AB.HCurrent.E") & -20 & $mV$ \\
 				\hline
-				Leak reversal potential ("AB.Leak.E") & -50 $mV$ \\
+				Leak reversal potential ("AB.Leak.E") & -50 & $mV$ \\
 				\hline
@@
 		The network model displayed in Figure 4 is described in \cite{prinzAlternativeHandtuningConductancebased2003, prinzSimilarNetworkActivity2004}.
-It is comprised of three compartments "ABPD", "LP", and "PY" with the same dynamics but differing parameters and synaptic inputs.
+It is comprised of three compartments "AB", "LP", and "PY" with the same dynamics but differing parameters and synaptic inputs.
 
@@
 				\hline 
-				\textbf{Parameter Name} & "ABPD" & "LP" & "PY" & \textbf{Units} \\ 
+				\textbf{Parameter Name} & \textbf{AB} & \textbf{LP} & \textbf{PY} & \textbf{Units} \\ 
 				\hline 
@@
 				\hline 
+				Sodium reversal potential & 50 & 50 & 50 & $mV$ \\ 
+				\hline 
+				Potassium reversal potential & -80 & -80 & -80 & $mV$ \\ 
+				\hline
+				Hyperpolarization-activated reversal potential & -20 & -20 & -20 & $mV$ \\
+				\hline
+				Leak reversal potential & -50 & -50 & -50 & $mV$ \\
+				\hline
 			\end{tabular}
@@
 		\begin{center}
-			\begin{tabular}{c|c|c|c|}
+			\begin{tabular}{|c|c|c|c|}
 				\hline
-				\textbf{Presynaptic} & \textbf{Postsynaptic} & \textbf{Type} & \textbf{Maximal Conductance} \\
+				\textbf{Presynaptic} & \textbf{Postsynaptic} & \textbf{Type} & \textbf{Maximal Conductance} ($\mu S/mm^2$) \\
 				\hline
-
+				AB & LP & Chol & 30 \\ \hline
+				AB & PY & Chol & 3 \\ \hline
+				AB & LP & Glut & 30 \\ \hline
+				AB & PY & Glut & 10 \\ \hline
+				LP & PY & Glut & 1 \\ \hline
+				PY & LP & Glut & 30 \\ \hline
+				LP & AB & Glut & 30 \\ \hline
 			\end{tabular}

### 2018-08-07 paper/paper.tex
wrote more text

@@
 
-	Conductance-based models of neuron dynamics based on Hodgkin-Huxley formalism have been used extensively in computational neuroscience.
-However, working with these models continues to be a challenge due to their high dimensionality and large number of parameters, hindering intuition of their dynamical features.
+	Conductance-based models of neurons have been used extensively in computational neuroscience.
+However, working with these models continues to be a challenge due to their high dimensionality and large number of parameters, that hinders intuition of their dynamical features.
 Here, we present a neuron and network simulator built on a novel automatic type system that binds class templates written in "C++" to object-oriented code in "MATLAB".
-Our approach builds on the tradition of uniting the speed and power of languages like "C++" with the ease-of-use and rich libraries of scientific programming languages like "MATLAB".
+Our approach builds on the tradition of uniting the speed and power of languages like "C++" with the ease-of-use and feature-set of scientific programming languages like "MATLAB".
 Our software allows for the creation and manipulation of hierarchical models with components that are named and searchable, permitting intuitive high-level programmatic control over all parameters.
 Our simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately.
-Our simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be extended to include arbitrary mechanisms and dynamics.
-Finally, the simulator is written in a modular fashion and has been released under a permissive free software license, enabling it to be integrated easily into the workflow of many researchers.
+Our simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be extended to include arbitrary conductances, synapses, and mechanisms.
+Finally, our simulator is written in a modular fashion and has been released under a permissive free software license, enabling seamless integration into the workflow of many researchers.
 
@@
 First, these models can be high-dimensional with many nonlinear differential equations, each with several parameters.
-Second, many or all equations in these models can be strongly coupled through variables like the membrane potential.
-In multi-compartment models of spatially extended neurons, membrane potentials in every compartment are coupled to each other.
-Finally, the choice of programming language used to implement the model imposes tradeoffs in designing and using neuron and network simulators: simulators written in languages like "C++", "FORTRAN" or "hoc" can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like "Python", "Julia" or "MATLAB".
+Second, many or all equations in these models can be strongly coupled through dynamic variables like the membrane potential.
+In multi-compartment models of spatially extended neurons, membrane potentials in every compartment can be different, and are coupled to each other.
+Finally, the choice of programming language used to implement the model imposes tradeoffs in designing and using neuron and network simulators: simulators written in languages like "C++" or "FORTRAN" can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like "Python", "Julia" or "MATLAB".
 
@@
 These simulators can be easier to use, and can benefit from inter-compatibility with other tools.
-Programs like "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}), "morphforge" (\cite{hullMorphforgeToolboxSimulating2014}), "PyNN" (\cite{davisonPyNNCommonInterface2009}) allow the user to specify models using strings of equations or using components that are specified in the scientific programming language, that can then be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
+Programs like "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}), "morphforge" (\cite{hullMorphforgeToolboxSimulating2014}), "PyNN" (\cite{davisonPyNNCommonInterface2009}) allow the user to specify models with strings of equations or with components that are constructed using a special syntax, that can then be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
 This approach permits considerable flexibility for simulating systems of differential equations.
@@
 "xolotl" comes packaged with hundreds of components that can be used to assemble cells and networks;  has built-in visualization functions to inspect voltage time traces and activation functions; and a GUI for real-time manipulation of model parameters.
-The GUI permits real-time manipulation of any network parameter and an immediate visualization of the effect of that parameter on arbitrary dynamical features of the model.
-The ease-of-use of these tools lends them to pedagogical applications and the rapid prototyping of models.
-This tool aims to simplify the investigation of the dynamics of conductance-based network and neuron models, facilitate collaborative modeling, and is intended to complement other tools being developed in the computational neuroscience community.
+The ease-of-use of the tools built into our software lends them to pedagogical applications and the rapid prototyping of models.
+Our software aims to simplify the investigation of the dynamics of conductance-based network and neuron models, facilitate collaborative modeling, and is intended to complement other tools being developed in the computational neuroscience community.
 
@@
 All parameters of models and activation functions of any channel can be be inspected before running the model.
-We designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by changing a few lines in a template file.
+We designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by calling functions that generate new "C++" files on the fly.
 Finally, "xolotl" is designed to be fully auditable, and comes with several tools to verify model and parameter integrity and aid reproducibility.
@@
 
-\paragraph{A rich library of network components.} "xolotl" comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and controllers) that can be used as building blocks to construct model neurons and networks.
+\paragraph{A rich library of network components.} "xolotl" comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and mechanisms) that can be used as building blocks to construct model neurons and networks.
 Compartments represent sections of membrane with a single membrane potential, Calcium level, and set of constituent components; and can represent either entire cells or parts of cells.
-Conductances represent populations of ion channels in a compartment that produce transmembrane currents.
-Synapses connect two compartments together by introducing a current in the post-synaptic neuron that can depend on the voltage of the presynaptic neuron.
-Controllers represent intracellular mechanisms that integrate a Calcium-dependent signal from the compartment they are in and  change the properties of a conductance or synapse in response.
+Objects of type "conductance" represent populations of ion channels in a compartment that produce transmembrane currents.
+Objects of type "synapse" connect two compartments together by introducing a current in the post-synaptic neuron that can depend on the voltage of the presynaptic neuron.
+Objects of type "mechanism" can represent any intracellular mechanism and can read and modify any other component in the cell, and can run arbitrary dynamical systems within them.
 
@@
 Each of these neurons contains within it a set of conductances and synaptic currents that contribute to its electrical behavior.
-These conductances can be controlled by intracellular regualtory system that can change their properties.
+Intracellular mechanisms can act within the cell to modify and regulate conductances, synapses, or the dynamic properties of the cell.
 Similarly, a "xolotl" model can contain a set of compartments that can represent individual neurons.
-Each compartment can contain a arbitrary set of conductances, that may be regulated by a set of controllers.
+Each compartment can contain a arbitrary set of conductances, and compartments, conductances and synapses can contain mechanisms that affect anything in the model.
 All objects in the "xolotl" model tree (compartments, conductances, etc.) are bonafide "MATLAB" objects that have their own type, properties and associated methods.
-This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain neurons that contain populations of ion channels, and makes constructing and integrating these models easier.
+This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain neurons that contain populations of ion channels, and makes constructing, integrating and thinking about these models easier.
 
 \paragraph{Automatic type system.}
-Our approach to circumvent the tradeoff between high-performance but hard to use languages like "C++" and richly featured but slow languages like "MATLAB" was to construct an automatic type system that links object oriented code in "C++" to object oriented code in "MATLAB".
-This architecture lets us construct the core of the simulator in "C++", leveraging features of "C++" like pointers that are not commonly used in "MATLAB".
+Our approach to circumvent the tradeoff between high-performance but hard to use languages like "C++" and richly featured but potentially slow languages like "MATLAB" was to construct an automatic type system that links object oriented code in "C++" to object oriented code in "MATLAB".
+This architecture lets us construct the core of the simulator in "C++", leveraging features of "C++" like pointers that are not readily available in "MATLAB".
 A rudimentary way to make this "C++" code useable in "MATLAB" would be to re-write that code in "MATLAB" so that "MATLAB" objects can be bound their "C++" implementations.
@@
 It is more therefore suited towards simulating small and heterogeneous networks rather than large but homogenous networks.
-While the software $can$ integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task.
+While the software $can$ integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task, offer more natural frameworks for dealing with a large number of identical units.
 
@@
 
-The axon of the giant squid contains contains a fast inactivating sodium conductance ("NaV"), a slower non-inactivating potassium conductance ("Kd"), and a passive leak current.
+The axon of the giant squid contains contains a fast inactivating sodium conductance ("NaV"), a slower non-inactivating potassium conductance ("Kd"), and a passive leak current ("Leak").
 Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952a}).
@@
 
-One of the most important intracellular mechanisms is the cytosolic buffering of Calcium and its influx through voltage-gated Calcium channels (VGCCs).
+A key intracellular mechanisms is the cytosolic buffering of Calcium and its influx through voltage-gated Calcium channels (VGCCs).
 Figure \ref{fig:figuremechanism}A shows a model of a single-compartment model with 8 populations of ion channels (\cite{liuModelNeuronActivityDependent1998}).
 Without any mechanism for Calcium influx or buffering, the intracellular Calcium levels in this model do not change (Figure \ref{fig:figuremechanism}B) and the model tonically spikes (Figure \ref{fig:figuremechanism}C).
-A common method of modelling Calcium buffering and influx is to use an differential equation that increases intracellular Calcium with the current through Calcium channels, and relaxes back to a baseline value (\cite{liuModelNeuronActivityDependent1998, prinzAlternativeHandtuningConductancebased2003, dayanTheoreticalNeuroscience2001}) (Figure \ref{fig:figuremechanism}D).
+A common method of modeling Calcium buffering and influx is by an differential equation that increases intracellular Calcium with the current through Calcium channels, and relaxes back to a baseline value (\cite{liuModelNeuronActivityDependent1998, prinzAlternativeHandtuningConductancebased2003, dayanTheoreticalNeuroscience2001}) (Figure \ref{fig:figuremechanism}D).
 This mechanism exists in the "xolotl" library as "CalciumMech1" and can be added to the model using a simple "add" statement (Figure \ref{fig:figuremechanism}E).
-The intracellular Calcium in the model now oscillates periodically (Figure \ref{fig:figuremechanism}F), synchronised to bursts in action potentials in this cell (Figure \ref{fig:figuremechanism}G).
+The intracellular Calcium in the model now oscillates periodically (Figure \ref{fig:figuremechanism}F), synchronized to bursts in action potentials in this cell (Figure \ref{fig:figuremechanism}G).
 
-Another intracellular mechanism of interest is a neuron's ability to regulate its electrical activity by controlling the spectrum of ion channels its expresses.
-Here, we will show "xolotl" can be used to model a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration and some target Calcium (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H).
-Since this mechanism affects each ion channel population individually, object corresponding to this mechanism is added to each "conductance" object in the neuron (Figure \ref{fig:figuremechanism}I).
+Another intracellular mechanism is a neuron's ability to regulate its electrical activity by controlling the spectrum of ion channels its expresses.
+Here, we will show how "xolotl" can be used to model a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration and some target Calcium (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H).
+Since this mechanism affects each ion channel population individually, an object corresponding to this mechanism is added to each "conductance" object in the neuron (Figure \ref{fig:figuremechanism}I).
 Setting all maximal conductance densities to some low value and integrating the model shows that the intracellular Calcium levels rise over time and approach the target (Figure \ref{fig:figuremechanism}J), while all conductance densities increase and then remain bounded (Figure \ref{fig:figuremechanism}K).
-Examining the voltage dynamics of the cell reveal that it transitions from silence to truncated bursts of action potentials to periodic bursting as this mechanism regualtes the neuron's ion channel spectrum.
+Examining the voltage dynamics of the cell reveal that it transitions from silence to truncated bursts of action potentials to periodic bursting as this mechanism regulates the neuron's ion channel spectrum.
 
@@
 
-Switching back and forth points in parameter space and state space of a neuron model is a common occurrence in working with neuron models.
+Switching back and forth points in parameter space and state space of a neuron model is a common occurrence in working with neuron models (alec:cite DeSchutter 1992).
 "xolotl" makes it easy to remember bookmark configurations of a model and return to them at will.
@@
 (Figure \ref{fig:figuresnapshot}) illustrates how these features can be used to explore the dynamics of the model presented in the previous section.
-First, the current state of the model is saved using the "snapshot" method into a namespace called "initial".
+First, the current state of the model is saved using the "snapshot" method into a namespace called ``initial''.
 In this state, the model exhibits periodic bursting activity due to a particular configuration of maximal conductance densities (Figure \ref{fig:figuremechanism}, orange).
@@
 Integrating the model for a longer duration allows the homeostatic control mechanism in the cell to restore the conductance profile and bursting activity to a state close to the original state (Figure \ref{fig:figuremechanism}, green).
-This state is saved using the name "final".
+This state is saved using the name ``final''.
 
@@
 \subsection{Simulating Network Models}
+
+- neurons communicate via synapses \\
+- electrical and chemical synapses \\
+- new behaviour in neywork that emerges from interactions
+
+- how synapses are implemnted \\
+- the STG model \\
+- explain figure \\
 

### 2018-08-07 paper/supplementary/supplementary.tex
Add maximal conductances for Liu models

@@
 				\hline 
-				Fast sodium maximal conductance ("AB.NaV.gbar") & 00 & $\mu S/mm^2$ \\ 
+				Fast sodium maximal conductance ("AB.NaV.gbar") & 1831.2 & $\mu S/mm^2$ \\ 
 				\hline
-				Transient calcium maximal conductance ("AB.CaT.gbar") & 00 & $\mu S/mm^2$ \\ 
+				Transient calcium maximal conductance ("AB.CaT.gbar") & 22.93 & $\mu S/mm^2$ \\ 
 				\hline 
-				Slow calcium maximal conductance ("AB.CaS.gbar") & 00 & $\mu S/mm^2$ \\ 
+				Slow calcium maximal conductance ("AB.CaS.gbar") & 27.07 & $\mu S/mm^2$ \\ 
 				\hline 
-				Transient potassium maximal conductance ("AB.ACurrent.gbar") & 00 & $\mu S/mm^2$ \\ 
+				Transient potassium maximal conductance ("AB.ACurrent.gbar") & 246.02 & $\mu S/mm^2$ \\ 
 				\hline 
-				Calcium-gated potassium maximal conductance ("AB.KCa.gbar") & 00 & $\mu S/mm^2$ \\ 
+				Calcium-gated potassium maximal conductance ("AB.KCa.gbar") & 979.94 & $\mu S/mm^2$ \\ 
 				\hline 
-				Delayed rectifier maximal conductance ("AB.Kd.gbar") & 00 & $\mu S/mm^2$ \\ 
+				Delayed rectifier maximal conductance ("AB.Kd.gbar") & 610.03 & $\mu S/mm^2$ \\ 
 				\hline
-				Hyperpolarization-activated maximal conductance ("AB.HCurrent.gbar") & 00 & $\mu S/mm^2$ \\ 
+				Hyperpolarization-activated maximal conductance ("AB.HCurrent.gbar") & 10.1 & $\mu S/mm^2$ \\ 
 				\hline 
-				Passive leak maximal conductance ("AB.Leak.gbar") & 00 & $\mu S/mm^2$ \\ 
+				Passive leak maximal conductance ("AB.Leak.gbar") & 0.99045 & $\mu S/mm^2$ \\ 
 				\hline 
@@
 		\begin{center}
-			\begin{tabular}{|c|c|c|c|}
+			\begin{tabular}{|c|c|c|r|}
 				\hline
-				\textbf{Presynaptic} & \textbf{Postsynaptic} & \textbf{Type} & \textbf{Maximal Conductance} ($\mu S/mm^2$) \\
+				\textbf{Presynaptic} & \textbf{Postsynaptic} & \textbf{Type} & $\bar{g}$ ($\mu S/mm^2$) \\
 				\hline
@@
 
+\section{Stability of Exponential Euler Method}
+
+PUT FIGURE HERE
+
 \printbibliography

### 2018-08-08 paper/paper.tex

--- 2018-08-06 paper/paper.tex
@@
 "xolotl" uses the exponential Euler method for integrating membrane potential (\cite{dayanTheoreticalNeuroscience2001}).
-"DynaSim" was implemented with a 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance in the documentation.
+"DynaSim" with "C"-compilation was implemented with a 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance in the documentation.
 "NEURON" used an implicit Euler regime (\cite{hinesNEURONSimulationEnvironment1997}).
@@
 
-DESCRIBE BENCHMARK RESULTS
+"xolotl" performed well in the benchmark simulations.
+Over increasing time step, "xolotl" was faster than "DynaSim" and "NEURON" with smaller error.
+During longer simulations "xolotl" and "DynaSim" increased in speed.
+This is likely due to the decreased impact of set-up overhead from "MATLAB", an interpreted language.
+For most simulation times, "xolotl" is roughly ten times faster than "DynaSim" and about twice as fast in the long-simulation time limit.
+"NEURON" performed consistently at high speed, likely due to a lack of programmatic overhead, though "DynaSim" and "xolotl" were significantly faster at large time steps and during long simulations.
 
@@
 (A) Schematic representation of a single-compartment neuron with three populations of ion channels (colored rectangles).
-(B) In the simulator, the soma is represented using an object of class $compartment$ and the populations of ion channels are represented by $conductance$ objects contained within the compartment object.
+(B) In the simulator, the soma is represented using an object of class \texttt{compartment} and the populations of ion channels are represented by \texttt{conductance} objects contained within the compartment object.
 (C) The code snippet shown sets up this neuron model, injects current, integrates and plots the voltage, and displays activation functions, all in a few lines of code.
@@
 Colors indicate the dominant current (gold is fast sodium (NaV), blue is delayed rectifier (Kd), red is Leak).
-(E) firing rate $vs$.
+(E) firing rate \vs.
 current (f-I) curve of this neuron.
@@
 (D) Current injected into the cell required to clamp it to the voltages in (C).
-(E) Asymptotic clamped current $vs.$ clamped voltage for this cell.
+(E) Asymptotic clamped current \vs clamped voltage for this cell.
 (F) Accounting for the reversal potential of Potassium ions yields the conductance-voltage curve of this channel type.
@@
 	\includegraphics[width=1.0\linewidth]{gfx/figure_snapshot}
-	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them $ad$   $arbitrium$.
+	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them \textit{ad arbitrium}.
 The initial state (orange disc) of the single compartment model in the previous example is saved using the \texttt{snapshot} method.
@@
 After evolving the model for some time (green), the conductance profile and voltage dynamics returns to a state similar to the initial state (green plots).
-This configuration is now saved in a state called $final$ and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
-Another parameter is changed (the Calcium target), and the model is integrated to reach a new state (blue disc) where the voltage dynamics are different from the $initial$ state.
+This configuration is now saved in a state called \textit{final} and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
+Another parameter is changed (the Calcium target), and the model is integrated to reach a new state (blue disc) where the voltage dynamics are different from the \textit{initial} state.
 In summary, any state can be bookmarked using a descriptive name using the \texttt{snapshot} method, and can be returned to using the \texttt{reset} method.}
@@
 The screenshot shows a GUI with sliders for every parameter of interest that is created by the \texttt{manipulate} method, and linked to an arbitrary number of visualization functions.
-In this example, two visualization functions are used: the built in \texttt{plot} method (A) and a custom function that computes the firing-rate-$vs$.-injected current curve for this neuron (B).
+In this example, two visualization functions are used: the built in \texttt{plot} method (A) and a custom function that computes the firing-rate-\vs-injected current curve for this neuron (B).
 Both plots refresh themselves with every movement of any slider, allowing the user to build intuition about how every parameter controls the dynamical behavior of the model.
-A screen recording of this model being manipulated in real time is included in Supplementary Material XX.
+A screen recording of this model being manipulated in real time is included in Supplementary Material.
 }
@@
 (C,G) Simulations over increasing simulation time where $\Delta t = 0.1$ ms.
-(D,H) Simulations over increasing number of unconnected compartments.}
+(D,H) Simulations over increasing number of unconnected compartments in the same network.}
 	\label{fig:figurebenchmark}

### 2018-08-08 paper/supplementary/supplementary.tex

@@
 
-PUT FIGURE HERE
+Bursting models from the Prinz stomatogastric ganglion model database (\cite{prinzAlternativeHandtuningConductancebased2003}) were simulated for 15 seconds at varying time resolution in order to characterize the stability of the exponential Euler numerical integration method.
+The exponential Euler method is generally accurate up to $\approx 0.1$ ms for most models tested.
+A time step of $\approx 0.05$ ms gives a good compromise between accuracy and speed (\cite{prinzAlternativeHandtuningConductancebased2003, prinzSimilarNetworkActivity2004}).
+
+\begin{figure}
+\centering
+\includegraphics[width=1.0\linewidth]{gfx/figure_sup_stability}
+\caption{Stability of exponential Euler method using models randomly chosen from the Prinz database.
+(A, C, E) Three metrics of bursting neurons: burst frequency, average spikes per burst, and duty cycle were computed over increasing time step.
+(B, D, F) The metrics were normalized to the first value in for each model.
+A line of zero slope would indicate no significant change in integration error over increasing time step.
+Each line indicates a model with different maximal conductances.}
+\label{fig:figuresupstability}
+\end{figure}
+
+\cleardoublepage
 

### 2018-08-09 paper/supplementary/supplementary.tex
Fix citations

@@
 		\item Shaker (\cite{hardieNovelPotassiumChannels1991})
-		\item Novel, Shab, Shaker (\cite{herasModulationVoltagedependentConductances2018})
+		\item A novel potassium conductance, Shab, Shaker (\cite{herasModulationVoltagedependentConductances2018})
 		\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{kisperskyIncreaseSodiumConductance2012})
@@
 		\item \cite{liuModelNeuronActivityDependent1998} conductances with forward Euler integration
-		\item \cite{liuModelNeuronActivityDependent1998} conductances with temperature dependence
+		\item \cite{liuModelNeuronActivityDependent1998} conductances with temperature dependence (\textit{cf.} \cite{caplanManyParameterSets2014})
 		\item Kd and NaV conductances from Int1, LG, and MCN1 cells (\cite{nadimFrequencyRegulationSlow1998})
@@
 		\item \cite{prinzAlternativeHandtuningConductancebased2003} conductances with cached gating functions
-		\item \cite{prinzAlternativeHandtuningConductancebased2003} conductances with temperature dependence
+		\item \cite{prinzAlternativeHandtuningConductancebased2003} conductances with temperature dependence (\textit{cf.} \cite{caplanManyParameterSets2014})
 		\item Kd and NaV from Int1 cells and modulatory input and transient LTS conductances from LG cells (\cite{rodriguezConvergentRhythmGeneration2013})

### 2018-08-09 paper/paper.tex

@@
 Nervous systems process and transmit information using electrically excitable membranes.
-Conductance-based models are a powerful biophysical simplification of an electrically excitable compartment in a neuron (\cite{hodgkinQuantitativeDescriptionMembrane1952a}).
+Conductance-based models are a powerful biophysical simplification of an electrically excitable compartment in a neuron (\cite{hodgkinQuantitativeDescriptionMembrane1952}).
 Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small circuits (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}).
@@
 \paragraph{New mechanisms require new \texttt{C++} code.} Adding new network components requires writing new "C++" code.
-Adding a new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952a, hodgkinQuantitativeDescriptionMembrane1952a, dayanTheoreticalNeuroscience2001}), requires creating a new "C++" header file, though this is generally trivial.
+Adding a new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, dayanTheoreticalNeuroscience2001}), requires creating a new "C++" header file, though this is generally trivial.
 Implementing a new integration scheme or component type requires much more in-depth knowledge of the underlying "C++" core code.
@@
 The axon of the giant squid contains contains a fast inactivating sodium conductance ("NaV"), a slower non-inactivating potassium conductance ("Kd"), and a passive leak current.
-Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952a}).
+Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}).
 As one of the simplest models of excitable neural membranes, the Hodgkin-Huxley model often serves as a the first model introduced in pedagogical literature (\cite{dayanTheoreticalNeuroscience2001, sterrattPrinciplesComputationalModelling2011, trappenbergFundamentalsComputationalNeuroscience2010}).
@@
 Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp (Figure \ref{fig:figureclamp}A).
-By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized (\cite{coleIonicCurrentMeasurements1960, coleIonsPotentialsNerve1955, hodgkinEffectSodiumIons1949, hodgkinMeasurementCurrentvoltageRelations1952a, hodgkinQuantitativeDescriptionMembrane1952a, turrigianoSelectiveRegulationCurrent1995}).
+By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized (\cite{coleIonicCurrentMeasurements1960, coleIonsPotentialsNerve1955, hodgkinEffectSodiumIons1949, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, turrigianoSelectiveRegulationCurrent1995}).
 
@@
 We set out to design a neuron and network simulator that could be useful in the classroom setting, especially for students of computational students, while also being powerful, fast and extensible enough to be used for research.
-By using a novel architecture that .., we showed how to create and simulate (1) single-compartment models of Hogkin-Huxley like neurons (Figure \ref{fig:figurehh}), (2) voltage clamp experiments and recover activation functions of single channel types (Figure 2), (3) a network of neurons with multiple synapse types (Figure 3), (4) a multi-compartment neuron model where the voltage of every compartment is coupled, and (5) a neuron model where homeostatic mechanisms control the maximal conductance of multiple channel types.
+By using a novel architecture that .., we showed how to create and simulate (1) single-compartment models of Hogkin-Huxley like neurons (Figure \ref{fig:figurehh}), (2) voltage clamp experiments and recover activation functions of single channel types (Figure \ref{fig:figureclamp}), (3) a network of neurons with multiple synapse types (Figure \ref{fig:figurenetwork}), (4) a multi-compartment neuron model where the voltage of every compartment is coupled, and (5) a neuron model where homeostatic mechanisms control the maximal conductance of multiple channel types (Figure \ref{fig:figuremechanism}).
 

### 2018-08-09 paper/paper.tex
Manual merging for srinivas-text

--- 2018-08-07 paper/paper.tex
@@
 The axon of the giant squid contains contains a fast inactivating sodium conductance ("NaV"), a slower non-inactivating potassium conductance ("Kd"), and a passive leak current ("Leak").
-Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952a}).
+Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}).
 As one of the simplest models of excitable neural membranes, the Hodgkin-Huxley model often serves as a the first model introduced in pedagogical literature (\cite{dayanTheoreticalNeuroscience2001, sterrattPrinciplesComputationalModelling2011, trappenbergFundamentalsComputationalNeuroscience2010}).
@@
 
-Voltage clamping is a experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (cite D\&A).
+Voltage clamping is a experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (\cite{dayanTheoreticalNeuroscience2001}).
 Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp (Figure \ref{fig:figureclamp}A).
-By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized (\cite{coleIonicCurrentMeasurements1960, coleIonsPotentialsNerve1955, hodgkinEffectSodiumIons1949, hodgkinMeasurementCurrentvoltageRelations1952a, hodgkinQuantitativeDescriptionMembrane1952a, turrigianoSelectiveRegulationCurrent1995}).
+By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized (\cite{coleIonicCurrentMeasurements1960, coleIonsPotentialsNerve1955, hodgkinEffectSodiumIons1949, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, turrigianoSelectiveRegulationCurrent1995}).
 
@@
 
-Switching back and forth points in parameter space and state space of a neuron model is a common occurrence in working with neuron models (alec:cite DeSchutter 1992).
+Switching back and forth points in parameter space and state space of a neuron model is a common occurrence in working with neuron models (\cite{deschutterConsumerGuideNeuronal1992}).
 "xolotl" makes it easy to remember bookmark configurations of a model and return to them at will.
@@
 
-- failure of averaging
-- complex interrelationships b/w parameters and dynamics
-- not clear what parameter does what
-- traditional approach: vary parameters, but for more than 2 D, how do you even visualize this? also slow. expensive to set up. 
-- need something where parameter variation can be linked to dynamics, easy to set up, easy to use.
-
 The "manipulate" function opens the GUI which permits visualization of changing parameters in real-time.
@@
 
-The use of computational tools is increasingly central to the scientific method; yet, the lack of auditability and accountability in their use has led to a crisis of credibility affecting many scientific fields (cite Victoria Stodden).
-Unlike in experimental research, where a lack of reproducibility can arise due to deep reasons like unknown differences in experimental protocols or intrinsic variability, the reasons for irreproducibility in computational research are largely trivial and include: a) typographical errors from transcribing out model parameters and equations, b) obscure software design that leads to users not knowing precisely what equations the software is solving, or the parameters it using, c) incompatibility with version control systems and rolling software development which leads to ambiguity in which version of the software was used to generate a particular result, and d) convoluted architectures that make it too complex for non-experts to understand the inner workings of the software.
+The use of computational tools is increasingly central to the scientific method; yet, the lack of auditability and accountability in their use has led to a crisis of credibility affecting many scientific fields (\cite{stoddenEnhancingReproducibilityComputational2016, bakerWhyScientistsMust2016}).
+Unlike in experimental research, where a lack of reproducibility can arise due to deep reasons like unknown differences in experimental protocols or intrinsic variability, the reasons for irreproducibility in computational research are largely trivial and include: a) typographical errors from transcribing out model parameters and equations, b) obscure software design that leads to users not knowing precisely what equations the software is solving, or the parameters it using, c) incompatibility with version control systems and rolling software development which leads to ambiguity in which version of the software was used to generate a particular result, and d) convoluted architectures that make it too complex for non-experts to understand the inner workings of the software (\cite{xuMeasurementSourceCode2017, sedanoCodeReadabilityTesting2016, vikstromStudyAutomaticTranslation2009}).
 
@@
 Our goal was to design software that would allow the user to verify for herself that the software was running as she intended, and to be able to reproduce results from others quickly and unambiguously.
-The primary design choice of our software that enables auditability and reproducibility is that every simulation is tied to a alphanumeric checksum, or hash, using the md5 algorithm (cite Rivest 1992).
+The primary design choice of our software that enables auditability and reproducibility is that every simulation is tied to a alphanumeric checksum, or hash, using the md5 algorithm (\cite{rivestMD5MessageDigestAlgorithm1992}).
 The hash is computed from every C++ file that is part of the model, and any changes in the C++ file will trigger a recompilation of that model.
@@
 The cell now bursts periodically, with synchronized oscillations in intracellular Calcium (F-G).
-(H) Schematic of Calcium-dependent integral control homeostasis (cite O'Leary et al.).
+(H) Schematic of Calcium-dependent integral control homeostasis (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}).
 In this feedback system, the rates of mRNA synthesis depend on the Calcium level in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which, through translation, is a function of the mRNA abundance.

### 2018-08-09 neuron/HH_benchmark2.py
Add neuron benchmarking script

@@ -74,3 +74,5 @@ for ii in range(0,len(t_end)):
 
 # save the results
 np.savetxt("neuron_HH_benchmark2.csv", S, delimiter=",")
+
+exit()

### 2018-08-09 neuron/STG_benchmark2.py

@@ -92,3 +92,5 @@ for ii in range(0,len(t_end)):
 
 # save the results
 np.savetxt("neuron_STG_benchmark2.csv", S, delimiter=",")
+
+exit()

### 2018-08-09 paper/paper.tex
Add neuron benchmarking script

@@
 First, these models can be high-dimensional with many nonlinear differential equations, each with several parameters.
-Second, many or all equations in these models can be strongly coupled through dynamic variables like the membrane potential.
+Second, many or all equations in these models can be strongly coupled through dynamical variables like the membrane potential.
 In multi-compartment models of spatially extended neurons, membrane potentials in every compartment can be different, and are coupled to each other.
-Finally, the choice of programming language used to implement the model imposes tradeoffs in designing and using neuron and network simulators: simulators written in languages like "C++" or "FORTRAN" can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like "Python", "Julia" or "MATLAB".
+Finally, the choice of programming language used to implement the model imposes tradeoffs in designing and using neuron and network simulators: simulators written in languages like "C++" or "FORTRAN" can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like "Python", "Julia", or "MATLAB".
 
@@
 These simulators tend to perform fast computations with little overhead, but suffer from a steep learning curve.
-Wrapping these simulators in a more approachable language like "Python" and the use of graphical user interfaces (GUIs) mitigate these drawbacks (\cite{hinesNEURONPython2009}), at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
-Another approach is to design simulators that can be used entirely in popular scientific programming languages.
-These simulators can be easier to use, and can benefit from inter-compatibility with other tools.
-Programs like "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}), "morphforge" (\cite{hullMorphforgeToolboxSimulating2014}), "PyNN" (\cite{davisonPyNNCommonInterface2009}) allow the user to specify models with strings of equations or with components that are constructed using a special syntax, that can then be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
+Wrapping these simulators in a more approachable language like "Python"  or using graphical user interfaces (GUIs) mitigates these drawbacks (\cite{hinesNEURONPython2009}) at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
+In contrast, simulators designed from the ground up in popular scientific computing languages can be easier to use and benefit from interoperability with other commonly-used tools.
+Simulators like "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}), "morphforge" (\cite{hullMorphforgeToolboxSimulating2014}), and "PyNN" (\cite{davisonPyNNCommonInterface2009}) allow the user to specify models with strings of equations or components that are constructed using a special syntax, that can then be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
 This approach permits considerable flexibility for simulating systems of differential equations.
@@
 To overcome these design limitations, we have developed a novel automatic type system, "cpplab", which binds "MATLAB" code to "C++" header files.
-This architecture automatically creates objects in "MATLAB" that represent the underlying object-oriented "C++" code, without having to re-specify these objects in "MATLAB".
+This architecture automatically creates objects in "MATLAB" that represent the underlying object-oriented "C++" code, allowing the symbolic manipulation of "C++" objects in the "MATLAB" interface.
 In this paper, we introduce "xolotl", an implementation of the "cpplab" system specialized in integrating conductance-based neuron and network models.
@@
 This software is designed to be used from within "MATLAB", a scientific programming language common amongst neuroscientists and engineers for pedagogy and research.
-Our goal was to make "xolotl" completely usable entirely from within MATLAB, and fully scriptable.
-Models created using this simulator appear in the "MATLAB" workspace as native objects, and are thus fully compatible with the large library of toolboxes that "MATLAB" provides, allowing so the software to be used as a component of other packages and tools.
-All parameters of models and activation functions of any channel can be be inspected before running the model.
-We designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by calling functions that generate new "C++" files on the fly.
-Finally, "xolotl" is designed to be fully auditable, and comes with several tools to verify model and parameter integrity and aid reproducibility.
+We aimed to make "xolotl" completely usable entirely from within "MATLAB", and fully scriptable.
+Models created using this simulator appear in the "MATLAB" workspace as native objects, and are thus fully compatible with the large library of toolboxes that "MATLAB" provides, allowing the software to be used as a component of other packages and tools.
+All parameters of models and activation functions of any channel can be inspected before running the model.
+We designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by calling functions that generate new "C++" files on-the-fly.
+Finally, "xolotl" fully auditable by design, with several tools to verify model and parameter integrity and aid in reproducibility.
 
@@
 \paragraph{A rich library of network components.} "xolotl" comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and mechanisms) that can be used as building blocks to construct model neurons and networks.
-Compartments represent sections of membrane with a single membrane potential, Calcium level, and set of constituent components; and can represent either entire cells or parts of cells.
+Compartments represent sections of membrane with a single membrane potential, Calcium level, and set of constituent components, representing either entire cells or parts of cells.
 Objects of type "conductance" represent populations of ion channels in a compartment that produce transmembrane currents.
-Objects of type "synapse" connect two compartments together by introducing a current in the post-synaptic neuron that can depend on the voltage of the presynaptic neuron.
+"synapse" objects connect two compartments together by introducing a current in the postsynaptic compartment that depends on the presynaptic compartment's membrane potential.
 Objects of type "mechanism" can represent any intracellular mechanism and can read and modify any other component in the cell, and can run arbitrary dynamical systems within them.
@@
 Similarly, a "xolotl" model can contain a set of compartments that can represent individual neurons.
-Each compartment can contain a arbitrary set of conductances, and compartments, conductances and synapses can contain mechanisms that affect anything in the model.
-All objects in the "xolotl" model tree (compartments, conductances, etc.) are bonafide "MATLAB" objects that have their own type, properties and associated methods.
-This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain neurons that contain populations of ion channels, and makes constructing, integrating and thinking about these models easier.
+Each compartment can contain an arbitrary set of conductances, and compartments, conductances and synapses can contain mechanisms that affect anything in the model.
+All objects in the "xolotl" model tree (compartments, conductances, etc.) are genuine "MATLAB" objects with their own type, properties, and associated methods.
+This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain neurons that contain populations of ion channels, and makes constructing, integrating, and thinking about these models easier.
 
 \paragraph{Automatic type system.}
-Our approach to circumvent the tradeoff between high-performance but hard to use languages like "C++" and richly featured but potentially slow languages like "MATLAB" was to construct an automatic type system that links object oriented code in "C++" to object oriented code in "MATLAB".
+To circumvent the tradeoff between high-performance but hard to use languages like "C++" and richly-featured but potentially slow languages like "MATLAB", we constructed an automatic type system that links object oriented code in "C++" to object oriented code in "MATLAB".
 This architecture lets us construct the core of the simulator in "C++", leveraging features of "C++" like pointers that are not readily available in "MATLAB".
 A rudimentary way to make this "C++" code useable in "MATLAB" would be to re-write that code in "MATLAB" so that "MATLAB" objects can be bound their "C++" implementations.
-However, this approach is cumbersome, and prone to errors, since the same code needs to be written twice, in two languages.
+However, this approach is cumbersome and inefficient since only features common between "MATLAB" and "C++" can be used.
 Instead, our automatic type system creates objects in "MATLAB" on-the-fly from "C++" class specifications, obviating the need to rewrite code in "MATLAB" while preserving a tight coupling between objects in the "MATLAB" workspace and their "C++" implementation.
-The key advantages of this approach are that it makes developing new code much easier, and simplifies the task of constructing complex frameworks that span these two languages.
+Crucially, this method makes developing new code much easier and simplifies the task of constructing complex frameworks that span these two languages.
 
@@
 
-\paragraph{Limited numerical integration strategies.} Most components in the software are integrated using the exponential Euler method, which has been shown to perform well in neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}.
+\paragraph{Limited numerical integration strategies.} Most components in the software are integrated using the exponential Euler method, which has been shown to perform well in neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}).
 However, it may be desirable to use other methods under certain conditions.
-It is possible to introduce new network components that implement other integration schemes, or to modify the integration schemes of existing components, but that requires writing new "C++"  code, and is still restricted to integration schemes that assume a fixed step size.
+It is possible to introduce new network components that implement other integration schemes, or to modify the integration schemes of existing components, but that requires writing new "C++"  code.
+Currently, "xolotl" can only implement integration schemes with fixed step size.
 
-\paragraph{Inefficient tools for handling large networks.} "xolotl" was designed to work with small but complex networks and models, where every compartment and component is $named$, rather than numbered.
+\paragraph{Inefficient tools for handling large networks.} "xolotl" was designed to work with small but complex networks and models, where every compartment and component is named, rather than numbered.
 It is more therefore suited towards simulating small and heterogeneous networks rather than large but homogenous networks.
-While the software $can$ integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task, offer more natural frameworks for dealing with a large number of identical units.
+While the software can integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task, offering more natural frameworks for dealing with a large number of identical units.
 
 \paragraph{New mechanisms require new \texttt{C++} code.} Adding new network components requires writing new "C++" code.
-Adding a new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, dayanTheoreticalNeuroscience2001}), requires creating a new "C++" header file, though this is generally trivial.
+A new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, dayanTheoreticalNeuroscience2001}) requires creating a new "C++" header file, though this is generally trivial.
 Implementing a new integration scheme or component type requires much more in-depth knowledge of the underlying "C++" core code.
@@
 In this section, we illustrate how "xolotl" can be used to generate a variety of models and simulate them.
-These examples have been chosen to illustrate various features of "xolotl", and are intended to serve as templates that researchers and educators can build upon.
+These examples have been chosen to illustrate various features of "xolotl", and are intended to serve as templates upon which researchers and educators can build.
 
@@
 
-In this example, we demonstrate how to simulate the spiking activity of a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model, and gain insight into the underlying biophysical mechanism.
+In this example, we demonstrate how to simulate the spiking activity of a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model and gain insight into the underlying biophysical mechanisms.
 This simple model consists of a single-compartment neuron with three types of conductances (Figure \ref{fig:figurehh}A).
@@
 These plots reveal that activation kinetics of the "NaV" channels are much faster than that of the "Kd" channels (Figure \ref{fig:figurehh}F), which facilitates the transient depolarization in an action potential.
-In summary, the simulator allows the user to construct and integrate this model in a few lines of code, and provides a rich visualization of the dynamics of the model.
+In summary, the simulator allows the user to construct and integrate this model in a few lines of code, and provides rich visualization of the dynamics of the model.
 
@@
 
-Voltage clamping is a experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (\cite{dayanTheoreticalNeuroscience2001}).
+Voltage clamping is an experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (\cite{dayanTheoreticalNeuroscience2001}).
 Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp (Figure \ref{fig:figureclamp}A).
@@
 
-"xolotl" can reproduce such a voltage clamp experiment $in silico$.
+"xolotl" can reproduce such a voltage clamp experiment \textit{in-silico}.
 Figure \ref{fig:figureclamp}B illustrates how a simple model with a single compartment and a single ion channel type can be set up and clamped to a desired voltage.
@@
 "xolotl" makes it possible to model and include arbitrary intracellular mechanisms.
-In our simulator, these intracellular mechanisms are represented by the "mechanism" object, and can be bound to "compartments", "conductances", and other object types.
+In our simulator, these intracellular mechanisms are represented by the "mechanism" object, and can be bound to compartments, conductances, and other object types.
 
-A key intracellular mechanisms is the cytosolic buffering of Calcium and its influx through voltage-gated Calcium channels (VGCCs).
+A key intracellular mechanisms is the cytosolic buffering of Calcium and its influx through voltage-gated Calcium channels.
 Figure \ref{fig:figuremechanism}A shows a model of a single-compartment model with 8 populations of ion channels (\cite{liuModelNeuronActivityDependent1998}).
 Without any mechanism for Calcium influx or buffering, the intracellular Calcium levels in this model do not change (Figure \ref{fig:figuremechanism}B) and the model tonically spikes (Figure \ref{fig:figuremechanism}C).
-A common method of modeling Calcium buffering and influx is by an differential equation that increases intracellular Calcium with the current through Calcium channels, and relaxes back to a baseline value (\cite{liuModelNeuronActivityDependent1998, prinzAlternativeHandtuningConductancebased2003, dayanTheoreticalNeuroscience2001}) (Figure \ref{fig:figuremechanism}D).
+Calcium buffering and influx can be modeled by a differential equation that increases intracellular Calcium with the current through Calcium channels and relaxes back to a baseline value (\cite{liuModelNeuronActivityDependent1998, prinzAlternativeHandtuningConductancebased2003, dayanTheoreticalNeuroscience2001}) (Figure \ref{fig:figuremechanism}D).
 This mechanism exists in the "xolotl" library as "CalciumMech1" and can be added to the model using a simple "add" statement (Figure \ref{fig:figuremechanism}E).
@@
 
-Another intracellular mechanism is a neuron's ability to regulate its electrical activity by controlling the spectrum of ion channels its expresses.
-Here, we will show how "xolotl" can be used to model a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration and some target Calcium (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H).
+Neurons can regulate their electrical activity by controlling the spectrum of ion channels they express (\cite{macleanActivityIndependentHomeostasisRhythmically2003, turrigianoSelectiveRegulationCurrent1995, schulzVariableChannelExpression2006}).
+Here, we will show how "xolotl" can be used to represent a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H).
 Since this mechanism affects each ion channel population individually, an object corresponding to this mechanism is added to each "conductance" object in the neuron (Figure \ref{fig:figuremechanism}I).
-Setting all maximal conductance densities to some low value and integrating the model shows that the intracellular Calcium levels rise over time and approach the target (Figure \ref{fig:figuremechanism}J), while all conductance densities increase and then remain bounded (Figure \ref{fig:figuremechanism}K).
-Examining the voltage dynamics of the cell reveal that it transitions from silence to truncated bursts of action potentials to periodic bursting as this mechanism regulates the neuron's ion channel spectrum.
+Setting all maximal conductance densities to some low value and integrating the model shows that the intracellular Calcium levels rise over time and approach the target Calcium concentration (Figure \ref{fig:figuremechanism}J), while all conductance densities increase and then remain bounded (Figure \ref{fig:figuremechanism}K).
+Examining the voltage dynamics of the cell reveal that it transitions from quiescence to truncated bursts of action potentials to periodic bursting as this mechanism regulates the neuron's ion channel spectrum.
 
@@
 Switching back and forth points in parameter space and state space of a neuron model is a common occurrence in working with neuron models (\cite{deschutterConsumerGuideNeuronal1992}).
-"xolotl" makes it easy to remember bookmark configurations of a model and return to them at will.
-Internally, "xolotl" does not distinguish between parameters and dynamic variables of a model, and uses the "serialize" method to gather all parameters and dynamic variables into a vector of values, that is passed to the underlying "C++" implementation.
+"xolotl" makes it easy to bookmark configurations of a model and return to them at-will.
+Internally, "xolotl" does not distinguish between parameters and dynamical variables of a model, and uses the "serialize" method to gather all parameters and dynamic variables into a vector of values that is passed to the underlying "C++" implementation.
 A paired "deserialize" method is used to update all parameters and variables in the object tree from this vector.
@@
 
-- neurons communicate via synapses \\
-- electrical and chemical synapses \\
-- new behaviour in neywork that emerges from interactions
+Neurons communicate via synapses, electrochemical junctions between cells (\cite{gjorgjievaComputationalImplicationsBiophysical2016, huaNeuralActivityDynamics2004}).
+In chemical synapses, the presynaptic neuron releases neurotransmitter across the synaptic cleft, which activate receptors on the postsynaptic neuron.
+In electrical synapses, no chemical intermediary is involved.
+Ions flux directly from the presynaptic neuron to the postsynaptic one.
+New patterns of activity emerge from the characteristics of the connecting synapses (\cite{liComodulationSynapsesVoltagegated2018, nadimSynapticDepressionCreates1999, gutierrezRectifyingElectricalSynapses2013, gutierrezMultipleMechanismsSwitch2013}).
 
-- how synapses are implemnted \\
-- the STG model \\
-- explain figure \\
+Network models in "xolotl" consist of "compartment" objects connected by "synapse" objects.
+Synapses are stored in a vector array as a field of the "xolotl" object in "MATLAB".
+Presynaptic and postsynaptic labels indicate the connectivity of the synapse.
+Like conductances, the dynamics of the synapse is specified by the type of "cpplab" object added to the network structure.
+For example, the "Electrical" synapse type describes equipotent and rectifying electrical synapses.
 
-Network models in "xolotl" consist of compartment objects connected by synapses.
-Presynaptic and postsynaptic labels indicate the connectivity of the synapse.
+Postsynaptic currents are a function of the membrane potential of both the presynaptic and postsynaptic neurons. "xolotl" handles synaptic current equations in the form
+
+\begin{equation}
+	I_{syn} = \bar{g}_{syn} s(V_{pre}) (V_{post} - E_{syn})
+\end{equation}
+
+Here $I_{syn}$ is the postsynaptic current, $\bar{g}_{syn}$ is the maximal conductance of the synaptic connection, $s$ is a dynamical variable dependent on the presynaptic membrane potential.
+The electrochemical driving force is comprised of the postsynaptic membrane potential and the reversal potential of the synapse.
+
 Figure \ref{fig:figurenetwork} implements a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans (\cite{prinzSimilarNetworkActivity2004}).
-The pyloric model contains three compartments and seven synapses (Figure \ref{fig:figurenetwork}A).
+The pyloric model contains three compartments ("AB", "LP", and "PY") and seven synapses (Figure \ref{fig:figurenetwork}A).
 This structure is recapitulated in the hierarchy of the "xolotl" object, where conductances are contained within compartments (Figure \ref{fig:figurenetwork}B).
-Synapses are stored in a vector array as a field of the "xolotl" object in "MATLAB" (Figure \ref{fig:figurenetwork}C).
+The membrane potentials show triphasic rhythmicity in the three compartments (Figure \ref{fig:figurenetwork}C-E).
+Synaptic inhibition generates this phase-locked rhythm.
+When the "PY" is depolarized, the dynamical variable mediating the  glutamatergic ("Glut") synapse between "PY" and "LP" is close to 1 and "LP" is inhibited (Figure \ref{fig:figurenetwork}F).
+Inversely, when "PY" is hyperpolarized, the dynamical variable is close to 0.
+When "PY" spikes, IPSPs can be seen in the "LP" trace.
 
@@
 
-We set out to design a neuron and network simulator that could be useful in the classroom setting, especially for students of computational students, while also being powerful, fast and extensible enough to be used for research.
-By using a novel architecture that .., we showed how to create and simulate (1) single-compartment models of Hogkin-Huxley like neurons (Figure \ref{fig:figurehh}), (2) voltage clamp experiments and recover activation functions of single channel types (Figure 2), (3) a network of neurons with multiple synapse types (Figure 3), (4) a multi-compartment neuron model where the voltage of every compartment is coupled, and (5) a neuron model where homeostatic mechanisms control the maximal conductance of multiple channel types.
+We set out to design a neuron and network simulator that could be useful in the classroom setting, especially for students of computational science, while also being powerful, fast, and extensible enough to be used for research.
+By using a novel architecture that permits the symbolic manipulation of "C++" objects in a comprehensible "MATLAB" interface, we showed how to create and simulate single-compartment models of Hogkin-Huxley like neurons (Figure \ref{fig:figurehh}), voltage clamp experiments and recover activation functions of single channel types (Figure \ref{fig:figureclamp}), a network of neurons with multiple synapse types (Figure \ref{fig:figurenetwork}), a multi-compartment neuron model where the voltage of every compartment is coupled, and a neuron model where homeostatic mechanisms control the maximal conductance of multiple channel types (Figure (\ref{fig:figuremechanism}).
 
@@
 Our goal was to design software that would allow the user to verify for herself that the software was running as she intended, and to be able to reproduce results from others quickly and unambiguously.
-The primary design choice of our software that enables auditability and reproducibility is that every simulation is tied to a alphanumeric checksum, or hash, using the md5 algorithm (\cite{rivestMD5MessageDigestAlgorithm1992}).
-The hash is computed from every C++ file that is part of the model, and any changes in the C++ file will trigger a recompilation of that model.
+The primary design choice of our software that enables auditability and reproducibility is that every simulation is tied to an alphanumeric checksum, or hash, using the "MD5" algorithm (\cite{rivestMD5MessageDigestAlgorithm1992}).
+The hash is computed from every "C++" file that is part of the model, and any changes in the "C++" file will trigger a recompilation of that model.
 Thus, the hash guarantees that a given model is derived from a set of source files, obviating any ambiguity about the code used to generate a model.
-In addition, every parameter and state variable in the model can also be hashed together with underlying code, allowing the user to generate a short checksum that guarantees with high probability that every aspect of the model -- code, parameters and initial conditions -- are exactly as they should be.
+In addition, every parameter and state variable in the model can also be hashed together with underlying code, allowing the user to generate a short checksum that guarantees with high probability that every aspect of the model -- code, parameters, and initial conditions -- are exactly as they should be.
 
@@
 This design allows the user to know precisely the equations being solved, and view the code that numerically integrates them.
-Finally, the core of our software is written in a few hundred lines of code and contains just four classes: compartments, conductances, synapses and mechanisms.
+Finally, the core of our software is written in a few hundred lines of code and contains just four classes: compartments, conductances, synapses, and mechanisms.
 This architectural simplicity permits a motivated user to understand the entire organization of our code in short period of time.

### 2018-08-09 figures/benchmarks/figure_sup_stability.m
minor fixes

@@ -186,25 +186,37 @@ for model = 1:length(sol)
   canonical_duty_cycle(model) = burst_metrics(9);
   canonical_n_spikes_b(model) = burst_metrics(2);
 end
-canonical_burst_freq(canonical_burst_freq<0) = NaN;
-canonical_duty_cycle(canonical_duty_cycle<0) = NaN;
-canonical_n_spikes_b(canonical_n_spikes_b<3) = NaN;
+
+% manual override
+passingModels = canonical_burst_freq >= 0.5 & canonical_burst_freq <= 2.0 & canonical_n_spikes_b >= 3 & canonical_n_spikes_b <= 10;
+params = params(:, passingModels);
+nV = nV(:, passingModels);
+nCa = nCa(:, passingModels);
 
 % simulate xolotl traces at low time-resolution
-disp('simulating xolotl traces...')
-xV   = NaN(20e3, length(sol));
-xCa  = NaN(20e3, length(sol));
+xV   = NaN(20e3, size(params, 2));
+xCa  = NaN(20e3, size(params, 2));
 x.dt = 1;
 x.sim_dt = 0.05;
-for model = 1:length(sol)
-  % set up the xolotl object with the new conductances
-  for qq = 1:length(conds)
-    x.AB.(conds{qq}).gbar = params(qq, model);
+
+h = GetMD5([GetMD5(params) x.hash]);
+if isempty(cache(h))
+  disp('simulating xolotl traces...')
+  for model = 1:size(params, 2)
+    textbar(model, size(params, 2))
+    % set up the xolotl object with the new conductances
+    for qq = 1:length(conds)
+      x.AB.(conds{qq}).gbar = params(qq, model);
+    end
+    % simulate and store
+    x.reset;
+    [xV(:, model), Ca] = x.integrate;
+    xCa(:, model) = Ca(:, 1);
   end
-  % simulate and store
-  x.reset;
-  [xV(:, model), Ca] = x.integrate;
-  xCa(:, model) = Ca(:, 1);
+  cache(h, xV, xCa);
+else
+  disp('loading xolotl traces...')
+  [xCa, xV] = cache(h);
 end
 
 % remove transient
@@ -227,14 +239,13 @@ end
 
[…]

### 2018-08-09 paper/paper.tex
minor fixes

@@
 
-\begin{equation}
-{\tt V = x.integrate;}
-\end{equation}
+\begin{lstlisting}[style=Matlab-editor]
+	V = x.integrate;
+\end{lstlisting}
 
@@
 
-- compare with BRIAN, GENESIS, XPPAUT, NEURON, DynaSim \\
-- selling point is ease of use and integration into MATLAB
-
 \begin{quote}

### 2018-08-13 paper/paper.tex
cleanup, trying to fix the NEURON mess

@@
 	Conductance-based models of neurons have been used extensively in computational neuroscience.
-However, working with these models continues to be a challenge due to their high dimensionality and large number of parameters, that hinders intuition of their dynamical features.
-Here, we present a neuron and network simulator built on a novel automatic type system that binds class templates written in "C++" to object-oriented code in "MATLAB".
+However, working with these models continues to be a challenge due to their high dimensionality and large number of parameters.
+Here, we present a neuron and network simulator built on a novel automatic type system that binds object-oriented code written in "C++" to objects in "MATLAB".
 Our approach builds on the tradition of uniting the speed and power of languages like "C++" with the ease-of-use and feature-set of scientific programming languages like "MATLAB".
@@
 Our simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be extended to include arbitrary conductances, synapses, and mechanisms.
-Finally, our simulator is written in a modular fashion and has been released under a permissive free software license, enabling seamless integration into the workflow of many researchers.
+Finally, our simulator is written in a modular fashion and has been released under a permissive free software license, enabling seamless integration into the workflow of other researchers.
 
@@
 Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small circuits (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}).
-These models provide an approachable framework for understanding many salient principles of neuroscience, since they explicitly model cell membranes and ion channels as electrical components in a circuit.
+These models provide an approachable framework for understanding many salient principles of electrophysiology, since they explicitly model cell membranes and ion channels as electrical components in a circuit.
 However, several challenges remain in working with biophysically-detailed conductance-based neuron models.
@@
 
-To overcome these design limitations, we have developed a novel automatic type system, "cpplab", which binds "MATLAB" code to "C++" header files.
+To overcome these design limitations, we have developed a novel automatic type system, that we call "cpplab", which binds "MATLAB" code to classes specified in "C++" header files.
 This architecture automatically creates objects in "MATLAB" that represent the underlying object-oriented "C++" code, allowing the symbolic manipulation of "C++" objects in the "MATLAB" interface.
@@
 
-where \( C_{i} \) and  \( V_{i} \) are the capacitance and membrane potential of compartment  \( i \) and \( I_{j} \) is the current due to ion channel population \( j \) given by
+where \( C_{i} \) and  \( V_{i} \) are the capacitance and membrane potential of compartment  \( i \).
+Compartments can represent whole neurons or parts of neurons.
+\( I_{j} \) is the current due to ion channel population \( j \) and is given by
 
@@
 This software is designed to be used from within "MATLAB", a scientific programming language common amongst neuroscientists and engineers for pedagogy and research.
-We aimed to make "xolotl" completely usable entirely from within "MATLAB", and fully scriptable.
-Models created using this simulator appear in the "MATLAB" workspace as native objects, and are thus fully compatible with the large library of toolboxes that "MATLAB" provides, allowing the software to be used as a component of other packages and tools.
-All parameters of models and activation functions of any channel can be inspected before running the model.
+Our goal was to make "xolotl" completely usable entirely from within "MATLAB".
+Models created using this simulator appear in the "MATLAB" workspace as native objects, are are thus fully scriptable, and are fully compatible with the large library of toolboxes that "MATLAB" provides, allowing the software to be used as a component of other packages and tools.
+All parameters of models and activation functions of any channel can be inspected at any point.
 We designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by calling functions that generate new "C++" files on-the-fly.
@@
 Each of these neurons contains within it a set of conductances and synaptic currents that contribute to its electrical behavior.
-Intracellular mechanisms can act within the cell to modify and regulate conductances, synapses, or the dynamic properties of the cell.
+Intracellular mechanisms can act within the cell, or parts of the cell, to modify and regulate conductances, synapses, or other dynamic properties of the cell.
 Similarly, a "xolotl" model can contain a set of compartments that can represent individual neurons.
-Each compartment can contain an arbitrary set of conductances, and compartments, conductances and synapses can contain mechanisms that affect anything in the model.
-All objects in the "xolotl" model tree (compartments, conductances, etc.) are genuine "MATLAB" objects with their own type, properties, and associated methods.
+Each compartment can contain an arbitrary set of conductances.
+Compartments, conductances and synapses can contain mechanisms that can affect anything in the model.
+All objects in the "xolotl" model tree (compartments, conductances, etc.) are bonafide "MATLAB" objects with their own type and properties.
 This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain neurons that contain populations of ion channels, and makes constructing, integrating, and thinking about these models easier.
@@
 This architecture lets us construct the core of the simulator in "C++", leveraging features of "C++" like pointers that are not readily available in "MATLAB".
-A rudimentary way to make this "C++" code useable in "MATLAB" would be to re-write that code in "MATLAB" so that "MATLAB" objects can be bound their "C++" implementations.
-However, this approach is cumbersome and inefficient since only features common between "MATLAB" and "C++" can be used.
+A rudimentary way to make this "C++" code useable in "MATLAB" would be to re-write that code in "MATLAB" so that "MATLAB" objects can be bound to their "C++" implementations.
+However, this approach is cumbersome and inefficient, and can introduce errors.
 Instead, our automatic type system creates objects in "MATLAB" on-the-fly from "C++" class specifications, obviating the need to rewrite code in "MATLAB" while preserving a tight coupling between objects in the "MATLAB" workspace and their "C++" implementation.
@@
 Since every model requires a compiled binary to run, a potential stumbling block is the problem of unambiguously linking a model to a binary executable.
-"xolotl" solves this problem by hashing the "C++" header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier (its "hash").
+"xolotl" solves this problem by hashing (\cite{rivestMD5MessageDigestAlgorithm1992}) the "C++" header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier (its ``hash'').
 Compiled binaries are named using this hash, ensuring both that the correct binary is run to integrate the model, and that compilation occurs only as needed.
@@
 While the software can integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task, offering more natural frameworks for dealing with a large number of identical units.
+Similarly, "xolotl" is not optimized to solve coupled ODEs on complex branched morphologies, and other simulators like (\cite{hinesNEURONSimulationEnvironment1997}) are more suited for this task.
 
@@
 
-In this section, we illustrate how "xolotl" can be used to generate a variety of models and simulate them.
+In this section, we illustrate how "xolotl" can be used to generate, inspect, and simulate a variety of models.
 These examples have been chosen to illustrate various features of "xolotl", and are intended to serve as templates upon which researchers and educators can build.
@@
 The axon of the giant squid contains contains a fast inactivating sodium conductance ("NaV"), a slower non-inactivating potassium conductance ("Kd"), and a passive leak current ("Leak").
-Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}).
+Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}).
 As one of the simplest models of excitable neural membranes, the Hodgkin-Huxley model often serves as a the first model introduced in pedagogical literature (\cite{dayanTheoreticalNeuroscience2001, sterrattPrinciplesComputationalModelling2011, trappenbergFundamentalsComputationalNeuroscience2010}).
@@
 In this example, we demonstrate how to simulate the spiking activity of a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model and gain insight into the underlying biophysical mechanisms.
-This simple model consists of a single-compartment neuron with three types of conductances (Figure \ref{fig:figurehh}A).
-This hierarchical organization of the neuron is mirrored in the structure of the model: an object of type "compartment" is used to represent the cell body, and three objects of type "conductance" are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B).
+This simple model consists of a single electrical compartment with three types of conductances (Figure \ref{fig:figurehh}A).
+This hierarchical organization of the neuron is mirrored in the structure of the model in our simulator: an object of type "compartment" is used to represent the cell body, and three objects of type "conductance" are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B).
 These models of ion channels were obtained from \cite{liuModelNeuronActivityDependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator.
-The syntax of constructing the neuron model (using "add" statements) preserves this hierarchical organization, and the code snippet in (Figure \ref{fig:figurehh}C) shows how this model can be created using just a few lines of code.
+The code to set up this model is terse, idiomatic, relies on no special markup, and preserves the heirarchical nature of the model (Figure \ref{fig:figurehh}C).
 
@@
 Colors in the voltage trace indicate the strongest instantaneous inward current when the voltage is increasing, and strongest instantaneous outward current when the voltage is decreasing.
-These built-in features reveal that the dominant current during the upswing of every action potential is the sodium current, and the dominant current immediately after the peak of the action potential is the potassium current, but that the leak current contributes to depolarization following an action potential.
+This built-in feature reveals that the dominant current during the upswing of every action potential is the sodium current, and the dominant current immediately after the peak of the action potential is the potassium current, but that the leak current contributes to depolarization following an action potential.
 This feature could be useful in quickly understanding the contributions of a number of ion channel types in a complex voltage trace from a more complicated neuron model.
@@
 \ref{fig:figureclamp}D).
-The asymptotic clamping currents depend on the clamping voltage in a nonlinear manner (Fig.
+Since the compartment is being voltage clamped, integrating the model returns the clamping current:
+
+\begin{lstlisting}[style=Matlab-editor]
+I_clamp = x.integrate;
+\end{lstlisting}
+
+By repeating the integration at a number of clamping voltages, we observe that the asymptotic clamping currents depend on the clamping voltage in a nonlinear manner (Fig.
 \ref{fig:figureclamp}E), since the open probabilities of the channel are functions of the membrane voltage.
@@
 \ref{eq:1}).
-However, real neurons possess a several dynamical features, stemming from a number of intracellular mechanisms.
+However, real neurons possess several dynamical features, arising from a variety of intracellular mechanisms.
 "xolotl" makes it possible to model and include arbitrary intracellular mechanisms.
@@
 Figure \ref{fig:figuremechanism}A shows a model of a single-compartment model with 8 populations of ion channels (\cite{liuModelNeuronActivityDependent1998}).
-Without any mechanism for Calcium influx or buffering, the intracellular Calcium levels in this model do not change (Figure \ref{fig:figuremechanism}B) and the model tonically spikes (Figure \ref{fig:figuremechanism}C).
-Calcium buffering and influx can be modeled by a differential equation that increases intracellular Calcium with the current through Calcium channels and relaxes back to a baseline value (\cite{liuModelNeuronActivityDependent1998, prinzAlternativeHandtuningConductancebased2003, dayanTheoreticalNeuroscience2001}) (Figure \ref{fig:figuremechanism}D).
+Without any explicit mechanism for Calcium influx or buffering, the intracellular Calcium levels in this model do not change (Figure \ref{fig:figuremechanism}B) and the model tonically spikes (Figure \ref{fig:figuremechanism}C).
+Calcium buffering and influx can be modelled by a differential equation that increases intracellular Calcium with the current through Calcium channels and relaxes back to a baseline value (\cite{liuModelNeuronActivityDependent1998, prinzAlternativeHandtuningConductancebased2003, dayanTheoreticalNeuroscience2001}) (Figure \ref{fig:figuremechanism}D).
 This mechanism exists in the "xolotl" library as "CalciumMech1" and can be added to the model using a simple "add" statement (Figure \ref{fig:figuremechanism}E).
@@
 Examining the voltage dynamics of the cell reveal that it transitions from quiescence to truncated bursts of action potentials to periodic bursting as this mechanism regulates the neuron's ion channel spectrum.
+In summary, our simulator can be used to construct neuron models with intracellular mechanisms such as Calcium influx and buffering, and homeostatic regulation.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Switching back and forth points in parameter space and state space of a neuron model is a common occurrence in working with neuron models (\cite{deschutterConsumerGuideNeuronal1992}).
-"xolotl" makes it easy to bookmark configurations of a model and return to them at-will.
-Internally, "xolotl" does not distinguish between parameters and dynamical variables of a model, and uses the "serialize" method to gather all parameters and dynamic variables into a vector of values that is passed to the underlying "C++" implementation.
+Switching back and forth points in parameter space and state space of a neuron model is a common occurrence in working with neuron models, and a significant fraction of a modeller's time is spent in a feedback loop of running simulations, viewing the output, changing parameters, and re-running simulations.
+(\cite{deschutterConsumerGuideNeuronal1992}).
+"xolotl" makes it easy to bookmark configurations of a model and return to them at will.
+Internally, "xolotl" uses the "serialize" method to gather all parameters and dynamic variables into a vector of values that is passed to the underlying "C++" implementation.
 A paired "deserialize" method is used to update all parameters and variables in the object tree from this vector.
 This architecture provides a natural framework for representing the state of any model, no matter how complex, using a vector of numbers.
-The "snapshot" method built into "xolotl" leverages this schema to save the entire state of the model in a named variable, that can be accessed using another built in method called "reset".
+The "snapshot" method built into "xolotl" leverages this schema to save the entire state of the model in a named variable, that can be accessed using another built-in method called "reset".
 
 (Figure \ref{fig:figuresnapshot}) illustrates how these features can be used to explore the dynamics of the model presented in the previous section.
-First, the current state of the model is saved using the "snapshot" method into a namespace called ``initial''.
+First, the current state of the model is saved using the "snapshot" method into a state called ``initial''.
 In this state, the model exhibits periodic bursting activity due to a particular configuration of maximal conductance densities (Figure \ref{fig:figuremechanism}, orange).
@@
 Here, the intracellular Calcium target is modified, and the model is re-integrated, to yield a different voltage activity (Figure \ref{fig:figuremechanism}, blue).
-At the end of this numerical exploration, any of the saved states can be quickly returned to using the "reset" method, making the process of re-initializing models to desired states and parameters both error-free and painless.
+At the end of this numerical exploration, any of the saved states can be quickly returned to using the "reset" method, making the process of re-initializing models to desired states and parameters both error-free and efficient.
 
@@
 
-Neurons communicate via synapses, electrochemical junctions between cells (\cite{gjorgjievaComputationalImplicationsBiophysical2016, huaNeuralActivityDynamics2004}).
-In chemical synapses, the presynaptic neuron releases neurotransmitter across the synaptic cleft, which activate receptors on the postsynaptic neuron.
+Neurons communicate and interact using synapses, electrochemical junctions between cells (\cite{gjorgjievaComputationalImplicationsBiophysical2016, huaNeuralActivityDynamics2004}).
+In chemical synapses, the presynaptic neuron releases packets of neurotransmitter across the synaptic cleft, which activate receptors on the postsynaptic neuron.
 In electrical synapses, no chemical intermediary is involved.
-Ions flux directly from the presynaptic neuron to the postsynaptic one.
-New patterns of activity emerge from the characteristics of the connecting synapses (\cite{liComodulationSynapsesVoltagegated2018, nadimSynapticDepressionCreates1999, gutierrezRectifyingElectricalSynapses2013, gutierrezMultipleMechanismsSwitch2013}).
+Ions flux directly from the presynaptic neuron to the postsynaptic one, setting up a current in the post-synaptic cell.
+New patterns of activity can emerge from the characteristics of the connecting synapses (\cite{liComodulationSynapsesVoltagegated2018, nadimSynapticDepressionCreates1999, gutierrezRectifyingElectricalSynapses2013, gutierrezMultipleMechanismsSwitch2013}).
 
-Network models in "xolotl" consist of "compartment" objects connected by "synapse" objects.
-Synapses are stored in a vector array as a field of the "xolotl" object in "MATLAB".
-Presynaptic and postsynaptic labels indicate the connectivity of the synapse.
-Like conductances, the dynamics of the synapse is specified by the type of "cpplab" object added to the network structure.
-For example, the "Electrical" synapse type describes equipotent and rectifying electrical synapses.
+Network models in "xolotl" consist of "compartment" objects that can be connected by "synapse" objects.
+Two compartments representing different neurons can be connected using synapses using the built-in "connect" method.
+For example, to connect two single-compartment neurons called "LP" and "PY" using a chemical synapse of type "Cholinergic" with strength 30 nS, we can use
 
-Postsynaptic currents are a function of the membrane potential of both the presynaptic and postsynaptic neurons. "xolotl" handles synaptic current equations in the form
+\begin{lstlisting}[style=Matlab-editor]
+x.connect('LP','PY','Cholinergic','gbar',30); 
+\end{lstlisting}
 
-\begin{equation}
-	I_{syn} = \bar{g}_{syn} s(V_{pre}) (V_{post} - E_{syn})
-\end{equation}
-
-Here $I_{syn}$ is the postsynaptic current, $\bar{g}_{syn}$ is the maximal conductance of the synaptic connection, $s$ is a dynamical variable dependent on the presynaptic membrane potential.
-The electrochemical driving force is comprised of the postsynaptic membrane potential and the reversal potential of the synapse.
-
-Figure \ref{fig:figurenetwork} implements a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans (\cite{prinzSimilarNetworkActivity2004}).
+Our simulator has several types of synapses like Cholinergic, Glutamatergic, AMPA and NMDA built-in, and other synapse classes can be easily added using templates.
+Figure \ref{fig:figurenetwork} demonstrates the implementation of a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans (\cite{prinzSimilarNetworkActivity2004}).
 The pyloric model contains three compartments ("AB", "LP", and "PY") and seven synapses (Figure \ref{fig:figurenetwork}A).
@@
 The membrane potentials show triphasic rhythmicity in the three compartments (Figure \ref{fig:figurenetwork}C-E).
-Synaptic inhibition generates this phase-locked rhythm.
-When the "PY" is depolarized, the dynamical variable mediating the  glutamatergic ("Glut") synapse between "PY" and "LP" is close to 1 and "LP" is inhibited (Figure \ref{fig:figurenetwork}F).
-Inversely, when "PY" is hyperpolarized, the dynamical variable is close to 0.
-When "PY" spikes, IPSPs can be seen in the "LP" trace.
+When the "PY" is depolarized, the dynamical variable mediating the  glutamatergic ("Glut") synapse model between "PY" and "LP" is close to 1 and "LP" is inhibited (Figure \ref{fig:figurenetwork}F).
+Conversely, when "PY" is hyperpolarized, the dynamical variable is close to 0.
+When "PY" spikes, IPSPs can be seen in the "LP" voltage trace (Figure \ref{fig:figurenetwork}D-E).
 
@@
 
-To assess speed and accuracy, "xolotl", "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), and "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}) were compared in simulations over varied time resolution,  simulation time, and number of compartments (Figure \ref{fig:figurebenchmark}).
+Our goal in designing "xolotl" was to create an easy-to-use neuron and network simulator that was fast enough and accurate enough for routine use by computational neuroscientists.
+In this section we compare the speed of "xolotl" in simulating single and large numbers of Hodgkin-Huxley-like neuron models (as in Figure \ref{fig:figurehh}) and bursting neuron models based on the bursting neurons in the lobster STG (\cite{prinzSimilarNetworkActivity2004}).
+For each type of neuron model, we compared our software to  "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}), a high-performance and powerful neuron simulator and "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), a general-purpose simulator that can solve coupled differential equations numerically.
+All simulators were run on the same hardware using fixed time step solvers: "xolotl" used the Exponential Euler method  (\cite{dayanTheoreticalNeuroscience2001}), "NEURON" used the ALEC_FIX_ME and "DynaSim" used the ALEC_FIX_ME solvers.
+We measured the speed of each simulator by dividing the time simulated for by the time it took for the simulator to complete.
+For example, if a simulator could simulate 10 seconds of neuron data in 1 second, its speed would be 10X.
+We measured the speed of every simulator as a function of simulation time step, total length of simulation, and system size.
+
+We also measured the
+
+To assess speed and accuracy, "xolotl",  were compared in simulations over varied time resolution,  simulation time, and number of compartments (Figure \ref{fig:figurebenchmark}).
 
@@
 \section*{Acknowledgments}
-The authors would like to thank Mara CP Rue and Hillary Rodgers for beta-testing the "xolotl" software. Janis Li helped to prepare many conductance header files.
+The authors would like to thank Timothy O'Leary and Leandro Alonso for useful discussions. Janis Li helped to prepare many conductance header files.
 

### 2018-08-13 figures/benchmarks/figure_benchmark.m

@@ -2,109 +2,86 @@
 % and other sim. software using a HH model and a bursting
 % STG model
 
-figure('outerposition',[100 100 1550 666],'PaperUnits','points','PaperSize',[1550 666]); hold on
-for i = 10:-1:1
-	ax(i) = subplot(2,5,i); hold on
+figure('outerposition',[100 100 1550 700],'PaperUnits','points','PaperSize',[1550 700]); hold on
+for i = 8:-1:1
+	ax(i) = subplot(2,4,i); hold on
+    set(ax(i),'XScale','log','YScale','log')
 end
+
+for i = [1 2 5 6]
+    xlabel(ax(i),'\Deltat (ms)')
+end
+
+for i = [1 3 5 7]
+    ylabel(ax(i),'Speed (X realtime)')
+end
+
+
+for i = [2 6]
+    ylabel(ax(i),'Simulation error (a.u.)')
+end
+
+
+for i = [3 7]
+    xlabel(ax(i),'t_{end} (ms)')
+end
+
+for i = [4 8]
+    ylabel(ax(i),'Speed*N (X realtime)')
+    xlabel(ax(i),'N')
+end
+
+
+
+
 
 % perform benchmarking and plot data
 disp('Begin xolotl HH')
 testXolotlHH(ax); drawnow
-disp('Begin DynaSim HH')
-testDynaSimHH(ax); drawnow
+
+
 disp('Begin NEURON HH')
 testNeuronHH(ax); drawnow
 
+
+
+
 disp('Begin xolotl STG')
-testXolotlSTG(ax); drawnow
+testXolotlSTG(ax(5:end)); drawnow
+
+
+prettyFig('lw',.5,'plw',1.5,'tick_length',.03);
+return
+
+
+
+disp('Begin DynaSim HH')
+testDynaSimHH(ax); drawnow
+
+
+
+
+
+
+
+
 disp('Begin DynaSim STG')
 testDynaSimSTG(ax); drawnow
+
+
+
+
+
+
+return
+
+
+
+
+
+
+
+
+
 disp('Begin NEURON STG')
 testNeuronSTG(ax); drawnow
-
-% add axis labels and scaling
-ax(1).Visible = 'off';
-
-set(ax(2),'XScale','log','YScale','log')
-xlabel(ax(2),'\Deltat (ms)')
-ylabel(ax(2),'Speed (X realtime)')
-
-set(ax(3),'XScale','log','YScale','log')
[…]

### 2018-08-13 figures/benchmarks/loadNeuronData.m

@@ -1,54 +1,65 @@
-function [Q,S, all_dt] = loadNeuronData(filename)
+function [Q,S, dt] = loadNeuronData(filename)
 
-  % make a vector of dt to vary
-  max_dt = 1e3;
-  K = 1:max_dt;
-  all_dt = K(rem(max_dt,K) == 0);
-  all_dt = all_dt/1e3;
+S = csvread([filename '.csv'])
 
-  % simulation time
-  t_end = 30000;
 
-  % set up a hash
-  % checks the filename and the contents of this function
-  h = [GetMD5(filename) GetMD5(which(mfilename),'File')];
+% make a vector of dt to vary
+max_dt = 1e3;
+K = 1:max_dt;
+all_dt = K(rem(max_dt,K) == 0);
+all_dt = all_dt/1e3;
 
-  % if there is no cached data, perform computation
-  if isempty(cache(h))
+% simulation time
+t_end = 30000;
 
-    % load the NEURON data
-    disp('loading NEURON data...')
-    all_V = csvread([filename '_raw.csv']); % this is 3.3 GB
-    disp('NEURON data loaded...')
-    for i = length(all_dt):-1:1
-      textbar(length(all_dt)-i, length(all_dt))
-      V = nonnans(all_V(:,i));
-    	all_f(i) = xolotl.findNSpikes(V,-20);
-    	all_f(i) = all_f(i)/(t_end*1e-3);
-    end
 
-    % measure the baseline error using the LeMasson matrix
-    V0 = nonnans(all_V(:,1));
-    [M0, V_lim, dV_lim] = xolotl.V2matrix(V0);
+% load the NEURON data
+disp('loading NEURON data...')
 
-    % compute the matrix error relative to the baseline error
-    for i = length(all_dt):-1:2
-      V = nonnans(all_V(:,i));
-    	M = xolotl.V2matrix(V, V_lim, dV_lim);
-    	matrix_error(i) = xolotl.matrixCost(M0,M);
-    end
 
-    % store the matrix error
-    Q = matrix_error;
+allfiles = dir([pwd '/neuron/neuron_HH_raw*.npy']);
+all_V = {};
+all_t = {};
+for i = 1:length(allfiles)
+	a = strfind(allfiles(i).name,'raw')+3;
+	z = strfind(allfiles(i).name,'.npy')-1;
 
-    % store the speed
-    S = csvread([filename '.csv'])
[…]

### 2018-08-13 figures/benchmarks/neuron/HH_benchmark1.py

@@ -1,14 +1,16 @@
 # simulate a hodgkin-huxley model in NEURON
-# measure the speed and accuracy with increasing simulation time
+# measure the speed and accuracy with increasing time-step
 
-# import the graphical interface
+
 from neuron import h, gui
 import numpy as np
 from matplotlib import pyplot
 import time
 from pathlib import Path
 
-benchmarks_file = "neuron_HH_benchmark2.csv"
+
+benchmarks_file = "neuron_HH_benchmark1.csv"
+
 
 my_file1 = Path(benchmarks_file)
 if my_file1.is_file():
@@ -17,10 +19,10 @@ if my_file1.is_file():
 
 
 # create the neuron
-soma = h.Section(name='soma');
+soma        = h.Section(name='soma');
 
 # set the size of the soma
-soma.L = 28.209; # microns
+soma.L      = 28.209; # microns
 soma.diam   = 28.209; # microns
 
 # set up the capacitance
@@ -40,46 +42,61 @@ soma(0.5).kd.gbar  = 300/10000
 h.psection(sec=soma)
 
 # set up injected current
-stim = h.IClamp(soma(0.5))
+stim        = h.IClamp(soma(0.5))
 stim.amp    = 0.2 # nA
-stim.dur    = 10000 # ms
-
-# set up recording variables
-v_vec       = h.Vector()
-t_vec       = h.Vector()
-v_vec.record(soma(0.5)._ref_v)
-t_vec.record(h._ref_t)
+stim.dur    = 30000 # ms
 
 # set up simulation
 h.dt        = 0.1 # ms
-h.tstop     = 10000 # ms
+h.tstop     = 30000 # ms
 
 tic         = time.perf_counter() # s
 h.run()
 toc         = time.perf_counter() # s
 
 # set up vectors to hold outputs
-t_end = np.array([1, 2, 4, 9, 18, 38, 78, 162, 336, 695, 1438, 2976,
-6158, 12743, 26367, 54556, 112884, 233572, 483293, 1000000]) # ms
+dt = [0.0010, 0.0020, 0.0040, 0.0050, 0.0080, 0.0100, 0.0200, 0.0250, .05, .1] # ms
 
-sim_time        = np.zeros((len(t_end),1))
-S               = np.zeros((len(t_end),1))
 
+
+
+S  = np.zeros((len(dt), 1))
 
 # perform the simulation
-for ii in range(0,len(t_end)):
-    percent = 100*ii/len(t_end)
+for ii in range(0, len(dt)):
[…]

### 2018-08-13 figures/benchmarks/neuron/HH_benchmark2.py

@@ -1,29 +1,26 @@
 # simulate a hodgkin-huxley model in NEURON
-# measure the speed and accuracy with increasing time-step
+# measure the speed and accuracy with increasing simulation time
 
-
+# import the graphical interface
 from neuron import h, gui
 import numpy as np
 from matplotlib import pyplot
 import time
 from pathlib import Path
 
-raw_file = "neuron_HH_benchmark1_raw.csv"
-benchmarks_file = "neuron_HH_benchmark1.csv"
-
+benchmarks_file = "neuron_HH_benchmark2.csv"
 
 my_file1 = Path(benchmarks_file)
-my_file2 = Path(raw_file)
-if my_file1.is_file() and my_file2.is_file():
+if my_file1.is_file():
     print("It looks like the benchmark is already done, so aborting...")
     exit()
 
 
 # create the neuron
-soma        = h.Section(name='soma');
+soma = h.Section(name='soma');
 
 # set the size of the soma
-soma.L      = 28.209; # microns
+soma.L = 28.209; # microns
 soma.diam   = 28.209; # microns
 
 # set up the capacitance
@@ -43,9 +40,9 @@ soma(0.5).kd.gbar  = 300/10000
 h.psection(sec=soma)
 
 # set up injected current
-stim        = h.IClamp(soma(0.5))
+stim = h.IClamp(soma(0.5))
 stim.amp    = 0.2 # nA
-stim.dur    = 30000 # ms
+stim.dur    = 10000 # ms
 
 # set up recording variables
 v_vec       = h.Vector()
@@ -55,51 +52,36 @@ t_vec.record(h._ref_t)
 
 # set up simulation
 h.dt        = 0.1 # ms
-h.tstop     = 30000 # ms
+h.tstop     = 10000 # ms
 
 tic         = time.perf_counter() # s
 h.run()
 toc         = time.perf_counter() # s
 
 # set up vectors to hold outputs
-dt          = np.array([0.0010, 0.0020, 0.0040, 0.0050, 0.0080, 0.0100, 0.0200, 0.0250, 0.0400, 0.0500, 0.1000, 0.1250, 0.2000, 0.2500, 0.5000, 1.0000]) # ms
-Vtrace      = np.empty((int(np.round(h.tstop/dt[0]+1)),len(dt)))
-Vtrace[:,:] = np.nan
+t_end = np.array([1, 2, 4, 9, 18, 38, 78, 162, 336, 695, 1438, 2976,
[…]

### 2018-08-13 figures/benchmarks/neuron/STG_benchmark1.py

@@ -7,6 +7,16 @@ from neuron import h, gui
 import numpy as np
 from matplotlib import pyplot
 import time
+from pathlib import Path
+
+benchmarks_file = "neuron_STG_benchmark1.csv"
+
+my_file1 = Path(benchmarks_file)
+if my_file1.is_file():
+    print("It looks like the benchmark is already done, so aborting...")
+    exit()
+
+
 
 # create the neuron
 soma        = h.Section(name='soma');
@@ -73,12 +83,10 @@ h.run()
 toc         = time.perf_counter() # s
 
 # set up vectors to hold outputs
-dt          = np.array([0.0010, 0.0020, 0.0040, 0.0050, 0.0080, 0.0100, 0.0200, 0.0250, 0.0400, 0.0500, 0.1000, 0.1250, 0.2000, 0.2500, 0.5000, 1.0000]) # ms
-Vtrace      = np.empty((int(np.round(h.tstop/dt[0]+1)),len(dt)))
-Vtrace[:,:] = np.nan
+dt = [0.0010, 0.0020, 0.0040, 0.0050, 0.0080, 0.0100, 0.0200, 0.0250, .05, .1] # ms
 
-sim_time    = np.zeros((len(dt), 1))
-S           = np.zeros((len(dt), 1))
+
+S  = np.zeros((len(dt), 1))
 
 # perform the simulation
 for ii in range(0, len(dt)):
@@ -88,23 +96,30 @@ for ii in range(0, len(dt)):
     # set up recording variable
     V = h.Vector()
     V.record(soma(0.5)._ref_v)
+    T = h.Vector()
+    T.record(h._ref_t)
 
     # set up independent parameter
     h.dt = dt[ii]
+    h.tstop  = 30000 # ms
+
 
     # perform simulation & capture time
     tic = time.perf_counter() # s
     h.run()
     toc = time.perf_counter() # s
 
-    # save voltage trace
-    for qq in range(0,len(V)-1):
-        Vtrace[qq,ii] = V[qq]
+
+    np.save('neuron_STG_raw' + str(ii+1),V)
+    np.save('neuron_STG_raw_time' + str(ii+1),T)
 
     # process simulation time
-    sim_time[ii] = (toc - tic) * 1000; # ms
-    S[ii]        = h.tstop / sim_time[ii] # unitless
+    sim_time = (toc - tic) * 1000; # ms
[…]

### 2018-08-13 figures/benchmarks/neuron/benchmark-neuron.sh

@@ -19,3 +19,7 @@ python HH_benchmark2.py
 
 # HH mdoel -- varying n_comp
 python HH_benchmark3.py
+
+
+# run STG test 
+python STG_benchmark1.py

### 2018-08-13 figures/benchmarks/testDynaSimHH.m

@@ -43,7 +43,7 @@ time        = all_dt(end) * (1:(t_end / max(all_dt)));
 
 all_V = NaN(ceil(t_end/max(all_dt)),length(all_dt));
 
-h = ['DS_HH' GetMD5(which(mfilename),'File') GetMD5(all_dt)];
+h = GetMD5([GetMD5([equations{:}]) GetMD5(all_dt)]);
 
 if isempty(cache(h))
 
@@ -93,9 +93,9 @@ S = t_end ./ all_sim_time;
 S = S * 1e-3;
 
 
-plot(ax(2),all_dt,S,'r-o')
+plot(ax(1),all_dt,S,'r-o')
 
-plot(ax(3),all_dt,matrix_error,'r-o')
+plot(ax(2),all_dt,matrix_error,'r-o')
 
 
 %% Increasing Simulation Time
@@ -105,7 +105,7 @@ dt          = 0.1;
 all_t_end   = unique(round(logspace(0,6,20)));
 all_sim_time = NaN*all_t_end;
 
-h = ['DS_HH' GetMD5(which(mfilename),'File') GetMD5(all_t_end)];
+h = ['DS_HH' GetMD5(all_t_end)];
 
 if isempty(cache(h))
 
@@ -130,7 +130,7 @@ else
 	S = cache(h);
 end
 
-plot(ax(4),all_t_end,S,'r-o')
+plot(ax(3),all_t_end,S,'r-o')
 
 
 %% Increasing Number of Compartments
@@ -174,4 +174,6 @@ else
 	S = cache(h);
 end
 
-plot(ax(5),nComps,S,'r-o')
+
+
+plot(ax(4),nComps,S,'r-o')

### 2018-08-13 figures/benchmarks/testNeuronHH.m

@@ -21,13 +21,13 @@ function testNeuronHH(ax)
 ;;;;;;;;  ;;;;;;;; ;;;;;;;;    ;;    ;;     ;;    ;;
 
 % time-consuming step, requires analyzing ~ 3.3 GB of data
-[Q, S, all_dt] = loadNeuronData('../../neuron/neuron_HH_benchmark1');
+[Q, S, all_dt] = loadNeuronData('./neuron/neuron_HH_benchmark1');
 
 % plot simulation speed vs. time step on axes #2
-plot(ax(2), all_dt(:), S, 'b-o')
+plot(ax(1), all_dt(:), S, 'b-o')
 
 % plot simulation error vs time step on axes #3
-plot(ax(3),all_dt, Q, 'b-o')
+plot(ax(2),all_dt, Q, 'b-o')
 
 
 
@@ -50,10 +50,10 @@ plot(ax(3),all_dt, Q, 'b-o')
 
 
 all_t_end   = unique(round(logspace(0,6,20)));
-S           = csvread('../../neuron/neuron_HH_benchmark2.csv');
+S           = csvread('./neuron/neuron_HH_benchmark2.csv');
 
 % plot simulation speed vs. simulation time on axes #4
-plot(ax(4),all_t_end, S, 'b-o')
+plot(ax(3),all_t_end, S, 'b-o')
 
 
  ;;;;;;  ;;    ;;  ;;;;;;  ;;;;;;;; ;;;;;;;; ;;     ;;
@@ -74,7 +74,10 @@ plot(ax(4),all_t_end, S, 'b-o')
 
 
 nComps      = [1, 2, 4, 8, 16, 32, 64, 128, 250, 500, 1000];
-S           = csvread('../../neuron/neuron_HH_benchmark3.csv');
+S           = csvread('./neuron/neuron_HH_benchmark3.csv');
+
+S = S(:).*nComps(:);
+
 
 % plot simulation speed vs. number of compartments on axes #5
-plot(ax(5),nComps, S, 'b-o')
+plot(ax(4),nComps, S, 'b-o')

### 2018-08-13 figures/benchmarks/testXolotlHH.m

@@ -68,12 +68,6 @@ else
 end
 
 
-
-for i = length(all_dt):-1:1
-	all_f(i) = xolotl.findNSpikes(all_V(:,i),-20);
-	all_f(i) = all_f(i)/(x.t_end*1e-3);
-end
-
 % now measure the errors using the LeMasson matrix
 [M0, V_lim, dV_lim] = xolotl.V2matrix(all_V(:,1));
 
@@ -85,7 +79,7 @@ end
 
 % delete the last one because the first sim is slow for
 % trivial reasons involving matlab compiling
-all_f(end) = [];
+
 matrix_error(end) = [];
 all_sim_time(end) = [];
 all_dt(end) = [];
@@ -96,10 +90,10 @@ S = x.t_end./all_sim_time;
 S = S*1e-3;
 
 % plot simulation speed vs. time step on axes #2
-plot(ax(2),all_dt,S,'k-o')
+plot(ax(1),all_dt,S,'k-o')
 
 % plot simulation error vs time step on axes #3
-plot(ax(3),all_dt,Q,'k-o')
+plot(ax(2),all_dt,Q,'k-o')
 
 
 
@@ -156,7 +150,7 @@ S = all_t_end./all_sim_time;
 S = S*1e-3;
 
 % plot simulation speed vs. simulation time on axes #4
-plot(ax(4),all_t_end,S,'k-o')
+plot(ax(3),all_t_end,S,'k-o')
 
  ;;;;;;  ;;    ;;  ;;;;;;  ;;;;;;;; ;;;;;;;; ;;     ;;
 ;;    ;;  ;;  ;;  ;;    ;;    ;;    ;;       ;;;   ;;;
@@ -237,5 +231,7 @@ end
 
 S = t_end./(all_sim_time*1e3);
 
+S = S.*nComps;
+
 % plot simulation speed vs. number of compartments on axes #5
-plot(ax(5),nComps,S,'k-o')
+plot(ax(4),nComps,S,'k-o')

### 2018-08-13 figures/benchmarks/testXolotlSTG.m

@@ -99,9 +99,9 @@ Q = matrix_error;
 S = x.t_end./all_sim_time;
 S = S*1e-3;
 
-plot(ax(2+5),all_dt,S,'k-o')
+plot(ax(1),all_dt,S,'k-o')
 
-plot(ax(3+5),all_dt,Q,'k-o')
+plot(ax(2),all_dt,Q,'k-o')
 
 
 
@@ -160,7 +160,7 @@ end
 S = all_t_end./all_sim_time;
 S = S*1e-3;
 
-plot(ax(4+5),all_t_end,S,'k-o')
+plot(ax(3),all_t_end,S,'k-o')
 
 
  ;;;;;;  ;;    ;;  ;;;;;;  ;;;;;;;; ;;;;;;;; ;;     ;;
@@ -245,5 +245,7 @@ end
 
 S = x.t_end./(all_sim_time*1e3);
 
+S = S.*nComps;
+
 % plot simulation speed vs. number of compartments on axes #5
-plot(ax(5+5),nComps,S,'k-o')
+plot(ax(4),nComps,S,'k-o')

### 2018-08-13 figures/integral_control_snippet.m

@@ -1,22 +1,0 @@
-
-
-% add controllers to conductances
-x.AB.NaV.add('IntegralController');
-x.AB.CaT.add('IntegralController');
-% and so on...
-
-% configure controller parameters
-x.set('*tau_m',1e4./[1 .22 .18 .08 8 5 15])
-x.set('*Controller.m',1e-1+1e-2*rand(7,1))
-
-x.t_end = .5e3;
-[~,Ca0,C0] = x.integrate;
-x.snapshot('before');
-
-x.t_end = 9.5e3;
-[~,Ca1,C1] = x.integrate;
-x.snapshot('during');
-
-x.t_end = 990e3;
-[~,Ca2,C2] = x.integrate;
-x.snapshot('after');

### 2018-08-13 figures/makeFigures.m

@@ -1,22 +1,0 @@
-%% Make All Figures
-
-clear
-figure_HH
-
-clear
-figure_network
-
-clear
-figure_clamp
-
-clear
-figure_mechanism
-
-clear
-figure_manipulate
-
-clear
-figure_snapshot
-
-clear
-figure_puppeteer

### 2018-08-13 neuron/x86_64/acurrent.c

@@ -1,520 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* VECTORIZED */
-#define NRN_VECTORIZED 1
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__acurrent
-#define _nrn_initial _nrn_initial__acurrent
-#define nrn_cur _nrn_cur__acurrent
-#define _nrn_current _nrn_current__acurrent
-#define nrn_jacob _nrn_jacob__acurrent
-#define nrn_state _nrn_state__acurrent
-#define _net_receive _net_receive__acurrent 
-#define rates rates__acurrent 
-#define states states__acurrent 
- 
-#define _threadargscomma_ _p, _ppvar, _thread, _nt,
-#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
-#define _threadargs_ _p, _ppvar, _thread, _nt
-#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- /* Thread safe. No static _p or _ppvar. */
- 
-#define t _nt->_t
-#define dt _nt->_dt
-#define gbar _p[0]
-#define Erev _p[1]
-#define i _p[2]
-#define m _p[3]
-#define h _p[4]
-#define g _p[5]
-#define m_inf _p[6]
-#define tau_m _p[7]
-#define h_inf _p[8]
-#define tau_h _p[9]
-#define Dm _p[10]
-#define Dh _p[11]
-#define v _p[12]
-#define _g _p[13]
- 
-#if MAC
-#if !defined(v)
-#define v _mlhv
-#endif
-#if !defined(h)
-#define h _mlhh
-#endif
-#endif
- 
-#if defined(__cplusplus)
-extern "C" {
-#endif
[…]

### 2018-08-13 neuron/x86_64/cad.c

@@ -1,418 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* NOT VECTORIZED */
-#define NRN_VECTORIZED 0
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__cad
-#define _nrn_initial _nrn_initial__cad
-#define nrn_cur _nrn_cur__cad
-#define _nrn_current _nrn_current__cad
-#define nrn_jacob _nrn_jacob__cad
-#define nrn_state _nrn_state__cad
-#define _net_receive _net_receive__cad 
-#define state state__cad 
- 
-#define _threadargscomma_ /**/
-#define _threadargsprotocomma_ /**/
-#define _threadargs_ /**/
-#define _threadargsproto_ /**/
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- static double *_p; static Datum *_ppvar;
- 
-#define t nrn_threads->_t
-#define dt nrn_threads->_dt
-#define ica _p[0]
-#define Dcai _p[1]
-#define Ca_inf _p[2]
-#define _g _p[3]
-#define _ion_ica	*_ppvar[0]._pval
-#define _ion_cai	*_ppvar[1]._pval
-#define _style_ca	*((int*)_ppvar[2]._pvoid)
- 
-#if MAC
-#if !defined(v)
-#define v _mlhv
-#endif
-#if !defined(h)
-#define h _mlhh
-#endif
-#endif
- 
-#if defined(__cplusplus)
-extern "C" {
-#endif
- static int hoc_nrnpointerindex =  -1;
- /* external NEURON variables */
- /* declaration of user functions */
- static int _mechtype;
-extern void _nrn_cacheloop_reg(int, int);
-extern void hoc_register_prop_size(int, int, int);
[…]

### 2018-08-13 neuron/x86_64/cadecay.c

@@ -1,440 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* NOT VECTORIZED */
-#define NRN_VECTORIZED 0
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__cad
-#define _nrn_initial _nrn_initial__cad
-#define nrn_cur _nrn_cur__cad
-#define _nrn_current _nrn_current__cad
-#define nrn_jacob _nrn_jacob__cad
-#define nrn_state _nrn_state__cad
-#define _net_receive _net_receive__cad 
-#define state state__cad 
- 
-#define _threadargscomma_ /**/
-#define _threadargsprotocomma_ /**/
-#define _threadargs_ /**/
-#define _threadargsproto_ /**/
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- static double *_p; static Datum *_ppvar;
- 
-#define t nrn_threads->_t
-#define dt nrn_threads->_dt
-#define depth _p[0]
-#define ica _p[1]
-#define channel_flow _p[2]
-#define B _p[3]
-#define Dcai _p[4]
-#define _g _p[5]
-#define _ion_ica	*_ppvar[0]._pval
-#define _ion_cai	*_ppvar[1]._pval
-#define _style_ca	*((int*)_ppvar[2]._pvoid)
- 
-#if MAC
-#if !defined(v)
-#define v _mlhv
-#endif
-#if !defined(h)
-#define h _mlhh
-#endif
-#endif
- 
-#if defined(__cplusplus)
-extern "C" {
-#endif
- static int hoc_nrnpointerindex =  -1;
- /* external NEURON variables */
- /* declaration of user functions */
- static int _mechtype;
-extern void _nrn_cacheloop_reg(int, int);
[…]

### 2018-08-13 neuron/x86_64/cas.c

@@ -1,482 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* VECTORIZED */
-#define NRN_VECTORIZED 1
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__cas
-#define _nrn_initial _nrn_initial__cas
-#define nrn_cur _nrn_cur__cas
-#define _nrn_current _nrn_current__cas
-#define nrn_jacob _nrn_jacob__cas
-#define nrn_state _nrn_state__cas
-#define _net_receive _net_receive__cas 
-#define castate castate__cas 
-#define evaluate_fct evaluate_fct__cas 
- 
-#define _threadargscomma_ _p, _ppvar, _thread, _nt,
-#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
-#define _threadargs_ _p, _ppvar, _thread, _nt
-#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- /* Thread safe. No static _p or _ppvar. */
- 
-#define t _nt->_t
-#define dt _nt->_dt
-#define gbar _p[0]
-#define i _p[1]
-#define carev _p[2]
-#define m_inf _p[3]
-#define tau_m _p[4]
-#define h_inf _p[5]
-#define tau_h _p[6]
-#define m _p[7]
-#define h _p[8]
-#define cai _p[9]
-#define Dm _p[10]
-#define Dh _p[11]
-#define ica _p[12]
-#define v _p[13]
-#define _g _p[14]
-#define _ion_cai	*_ppvar[0]._pval
-#define _ion_ica	*_ppvar[1]._pval
-#define _ion_dicadv	*_ppvar[2]._pval
- 
-#if MAC
[…]

### 2018-08-13 neuron/x86_64/cat.c

@@ -1,482 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* VECTORIZED */
-#define NRN_VECTORIZED 1
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__cat
-#define _nrn_initial _nrn_initial__cat
-#define nrn_cur _nrn_cur__cat
-#define _nrn_current _nrn_current__cat
-#define nrn_jacob _nrn_jacob__cat
-#define nrn_state _nrn_state__cat
-#define _net_receive _net_receive__cat 
-#define castate castate__cat 
-#define evaluate_fct evaluate_fct__cat 
- 
-#define _threadargscomma_ _p, _ppvar, _thread, _nt,
-#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
-#define _threadargs_ _p, _ppvar, _thread, _nt
-#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- /* Thread safe. No static _p or _ppvar. */
- 
-#define t _nt->_t
-#define dt _nt->_dt
-#define gbar _p[0]
-#define i _p[1]
-#define carev _p[2]
-#define m_inf _p[3]
-#define tau_m _p[4]
-#define h_inf _p[5]
-#define tau_h _p[6]
-#define m _p[7]
-#define h _p[8]
-#define cai _p[9]
-#define Dm _p[10]
-#define Dh _p[11]
-#define ica _p[12]
-#define v _p[13]
-#define _g _p[14]
-#define _ion_cai	*_ppvar[0]._pval
-#define _ion_ica	*_ppvar[1]._pval
-#define _ion_dicadv	*_ppvar[2]._pval
- 
-#if MAC
[…]

### 2018-08-13 neuron/x86_64/hcurrent.c

@@ -1,443 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* VECTORIZED */
-#define NRN_VECTORIZED 1
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__hcurrent
-#define _nrn_initial _nrn_initial__hcurrent
-#define nrn_cur _nrn_cur__hcurrent
-#define _nrn_current _nrn_current__hcurrent
-#define nrn_jacob _nrn_jacob__hcurrent
-#define nrn_state _nrn_state__hcurrent
-#define _net_receive _net_receive__hcurrent 
-#define rates rates__hcurrent 
-#define states states__hcurrent 
- 
-#define _threadargscomma_ _p, _ppvar, _thread, _nt,
-#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
-#define _threadargs_ _p, _ppvar, _thread, _nt
-#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- /* Thread safe. No static _p or _ppvar. */
- 
-#define t _nt->_t
-#define dt _nt->_dt
-#define gbar _p[0]
-#define Erev _p[1]
-#define i _p[2]
-#define m _p[3]
-#define g _p[4]
-#define minf _p[5]
-#define tau_m _p[6]
-#define Dm _p[7]
-#define v _p[8]
-#define _g _p[9]
- 
-#if MAC
-#if !defined(v)
-#define v _mlhv
-#endif
-#if !defined(h)
-#define h _mlhh
-#endif
-#endif
- 
-#if defined(__cplusplus)
-extern "C" {
-#endif
- static int hoc_nrnpointerindex =  -1;
[…]

### 2018-08-13 neuron/x86_64/kca.c

@@ -1,484 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* VECTORIZED */
-#define NRN_VECTORIZED 1
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__kca
-#define _nrn_initial _nrn_initial__kca
-#define nrn_cur _nrn_cur__kca
-#define _nrn_current _nrn_current__kca
-#define nrn_jacob _nrn_jacob__kca
-#define nrn_state _nrn_state__kca
-#define _net_receive _net_receive__kca 
-#define rates rates__kca 
-#define states states__kca 
- 
-#define _threadargscomma_ _p, _ppvar, _thread, _nt,
-#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
-#define _threadargs_ _p, _ppvar, _thread, _nt
-#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- /* Thread safe. No static _p or _ppvar. */
- 
-#define t _nt->_t
-#define dt _nt->_dt
-#define gbar _p[0]
-#define Erev _p[1]
-#define i _p[2]
-#define m _p[3]
-#define cai _p[4]
-#define g _p[5]
-#define m_inf _p[6]
-#define tau_m _p[7]
-#define Dm _p[8]
-#define v _p[9]
-#define _g _p[10]
-#define _ion_cai	*_ppvar[0]._pval
- 
-#if MAC
-#if !defined(v)
-#define v _mlhv
-#endif
-#if !defined(h)
-#define h _mlhh
-#endif
-#endif
- 
-#if defined(__cplusplus)
-extern "C" {
-#endif
- static int hoc_nrnpointerindex =  -1;
- static Datum* _extcall_thread;
[…]

### 2018-08-13 neuron/x86_64/kd.c

@@ -1,443 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* VECTORIZED */
-#define NRN_VECTORIZED 1
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__kd
-#define _nrn_initial _nrn_initial__kd
-#define nrn_cur _nrn_cur__kd
-#define _nrn_current _nrn_current__kd
-#define nrn_jacob _nrn_jacob__kd
-#define nrn_state _nrn_state__kd
-#define _net_receive _net_receive__kd 
-#define rates rates__kd 
-#define states states__kd 
- 
-#define _threadargscomma_ _p, _ppvar, _thread, _nt,
-#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
-#define _threadargs_ _p, _ppvar, _thread, _nt
-#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- /* Thread safe. No static _p or _ppvar. */
- 
-#define t _nt->_t
-#define dt _nt->_dt
-#define gbar _p[0]
-#define Erev _p[1]
-#define i _p[2]
-#define m _p[3]
-#define g _p[4]
-#define minf _p[5]
-#define tau_m _p[6]
-#define Dm _p[7]
-#define v _p[8]
-#define _g _p[9]
- 
-#if MAC
-#if !defined(v)
-#define v _mlhv
-#endif
-#if !defined(h)
-#define h _mlhh
-#endif
-#endif
- 
-#if defined(__cplusplus)
-extern "C" {
-#endif
- static int hoc_nrnpointerindex =  -1;
- static Datum* _extcall_thread;
- static Prop* _extcall_prop;
[…]

### 2018-08-13 neuron/x86_64/na.c

@@ -1,478 +1,0 @@
-/* Created by Language version: 6.2.0 */
-/* VECTORIZED */
-#define NRN_VECTORIZED 1
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include "scoplib_ansi.h"
-#undef PI
-#define nil 0
-#include "md1redef.h"
-#include "section.h"
-#include "nrniv_mf.h"
-#include "md2redef.h"
- 
-#if METHOD3
-extern int _method3;
-#endif
-
-#if !NRNGPU
-#undef exp
-#define exp hoc_Exp
-extern double hoc_Exp(double);
-#endif
- 
-#define nrn_init _nrn_init__na
-#define _nrn_initial _nrn_initial__na
-#define nrn_cur _nrn_cur__na
-#define _nrn_current _nrn_current__na
-#define nrn_jacob _nrn_jacob__na
-#define nrn_state _nrn_state__na
-#define _net_receive _net_receive__na 
-#define rates rates__na 
-#define states states__na 
- 
-#define _threadargscomma_ _p, _ppvar, _thread, _nt,
-#define _threadargsprotocomma_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt,
-#define _threadargs_ _p, _ppvar, _thread, _nt
-#define _threadargsproto_ double* _p, Datum* _ppvar, Datum* _thread, _NrnThread* _nt
- 	/*SUPPRESS 761*/
-	/*SUPPRESS 762*/
-	/*SUPPRESS 763*/
-	/*SUPPRESS 765*/
-	 extern double *getarg();
- /* Thread safe. No static _p or _ppvar. */
- 
-#define t _nt->_t
-#define dt _nt->_dt
-#define gbar _p[0]
-#define Erev _p[1]
-#define i _p[2]
-#define m _p[3]
-#define h _p[4]
-#define g _p[5]
-#define minf _p[6]
-#define hinf _p[7]
-#define tau_h _p[8]
-#define tau_m _p[9]
-#define Dm _p[10]
-#define Dh _p[11]
-#define v _p[12]
-#define _g _p[13]
- 
-#if MAC
-#if !defined(v)
-#define v _mlhv
-#endif
-#if !defined(h)
-#define h _mlhh
-#endif
-#endif
- 
-#if defined(__cplusplus)
-extern "C" {
-#endif
[…]

### 2018-08-13 src/getFigPosition.m

@@ -1,10 +1,0 @@
-function pos = getFigPosition(ax)
-  % creates a matrix of positions for a vector of axes
-
-  pos = zeros(length(ax), 4);
-
-  for ii = 1:length(ax)
-    pos(ii,:) = ax(ii).Position;
-  end
-
-end

### 2018-08-13 src/setFigPosition.m

@@ -1,8 +1,0 @@
-function pos = setFigPosition(ax, pos)
-  % sets axis positions to a matrix of positions
-
-  for ii = 1:length(ax)
-    ax(ii).Position = pos(ii, :);
-  end
-
-end

### 2018-08-14 paper/paper.tex
inserted benchmark figure

@@
 Our goal in designing "xolotl" was to create an easy-to-use neuron and network simulator that was fast enough and accurate enough for routine use by computational neuroscientists.
-In this section we compare the speed of "xolotl" in simulating single and large numbers of Hodgkin-Huxley-like neuron models (as in Figure \ref{fig:figurehh}) and bursting neuron models based on the bursting neurons in the lobster STG (\cite{prinzSimilarNetworkActivity2004}).
-For each type of neuron model, we compared our software to  "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}), a high-performance and powerful neuron simulator and "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), a general-purpose simulator that can solve coupled differential equations numerically.
-All simulators were run on the same hardware using fixed time step solvers: "xolotl" used the Exponential Euler method  (\cite{dayanTheoreticalNeuroscience2001}), "NEURON" used the ALEC_FIX_ME and "DynaSim" used the ALEC_FIX_ME solvers.
+In this section, we compare the speed of "xolotl" in simulating single and large numbers of Hodgkin-Huxley-like neuron models (as in Figure \ref{fig:figurehh}) and bursting neuron models based on the bursting neurons in the lobster STG (\cite{prinzSimilarNetworkActivity2004}).
+For each type of neuron model, we compared our software to  "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}), a high-performance and powerful neuron simulator specialised in simulating neurons with complex morphologies and "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), a general-purpose simulator that can solve coupled differential equations numerically.
+All simulators were run on the same hardware using fixed time step solvers: "xolotl" used the Exponential Euler method  (\cite{dayanTheoreticalNeuroscience2001}), "NEURON" used the  implicit Euler solver (\cite{hinesNEURONSimulationEnvironment1997}) and "DynaSim" used "C"-compiled 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance in the documentation.
 We measured the speed of each simulator by dividing the time simulated for by the time it took for the simulator to complete.
@@
 
-We also measured the
+All three simulators were faster with larger time steps, since fewer iterations were needed (Figure \ref{fig:figurebenchmark}A-B), and were approximately linear in the region tested.
+Our simulator compared favourably to "NEURON" and "DynaSim" in this task.
+We also measured the quality of the simulated output by comparing it to the simulated output at the smallest time step.
+Simulation error was measured using the LeMasson cost (\cite{lemassonIntroductionEquationSolving2000}), and was comparable amongst the three simulators (Figure \ref{fig:figurebenchmark}C-D).
+Our simulator sets up and runs the simulation in "C++", and needs to transfer parameters and data to and from the underlying implementation.
+To measure the performance cost of this overhead, we repeated these simulations on all three simulators at a afixed time step of $0.1$ ms and varied the length of time simulated for.
+Speed increased with simulation duration up to a limit, and then saturated, indicating a fixed performance cost of the overhead (Figure \ref{fig:figurebenchmark}E-F, black lines).
+Simualtions using "DynaSim", which also used a similar architecture, showed a similar increase in speed with simulation length (Figure \ref{fig:figurebenchmark}E-F, red lines).
+In contrast, simulations using "NEURON" ran at a constant speed irrespective of simulation length, presumably due to differences in the underlying implementation (Figure \ref{fig:figurebenchmark}E-F, blue lines).
 
-To assess speed and accuracy, "xolotl",  were compared in simulations over varied time resolution,  simulation time, and number of compartments (Figure \ref{fig:figurebenchmark}).
+Many simulators have been designed to simualte large numbers of compartments, either as networks with many identical neurons or in a large mulit-compartmentel neuron model (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}).
+While our software is not designed for this task {\em per se}, we measured its performance as a function of the number of compartments simulated.
+"xolotl" can quickly create a number of identical copies of a compartment using the "replicate" method:
 
-Single-compartment Hodgkin-Huxley-like models with three conductances and stomatogastric ganglion neuron models with eight conductances were generated using conductance dynamics from \cite{liuModelNeuronActivityDependent1998}) in the simulation environments.
-"xolotl" uses the exponential Euler method for integrating membrane potential (\cite{dayanTheoreticalNeuroscience2001}).
-"DynaSim" with "C"-compilation was implemented with a 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance in the documentation.
-"NEURON" used an implicit Euler regime (\cite{hinesNEURONSimulationEnvironment1997}).
+\begin{lstlisting}[style=Matlab-editor]
+x.replicate('compartment_to_replicate',n_copies); 
+\end{lstlisting}
 
-To compare the integration methods, models were simulated for 30 s at varying time-resolution (Figure \ref{fig:figurebenchmark}A,E).
-The speed factor was defined as the ratio between time represented in the simulation and actual runtime (simulation-time).
-Therefore, the speed factor represents how many times faster the simulation is than a real-time observation.
-The error resulting from coarser time-resolution was computed as the distance between two LeMasson matrices (\cite{lemassonIntroductionEquationSolving2000}), one representing the canonical voltage trace with time step $\Delta t = 0.001$ ms and the other representing the voltage trace simulated at time steps up to $1$ ms (Figure \ref{fig:figurebenchmark}B,F).
-
-To assess the performance of the simulators in absence of set-up overhead, models were simulated with a time-resolution of $0.1$ ms over increasing simulation time (Figure \ref{fig:figurebenchmark}C,G).
-
-Many simulators perform well in simulations of many compartments (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}).
-To assess how "xolotl" performs under these conditions, networks of up to 1,000 neuron models were simulated for 30 s at a time-resolution of 0.1 ms (Figure \ref{fig:figurebenchmark}D,H).
-
-"xolotl" performed well in the benchmark simulations.
-Over increasing time step, "xolotl" was faster than "DynaSim" and "NEURON" with smaller error.
-During longer simulations "xolotl" and "DynaSim" increased in speed.
-This is likely due to the decreased impact of set-up overhead from "MATLAB", an interpreted language.
-For most simulation times, "xolotl" is roughly ten times faster than "DynaSim" and about twice as fast in the long-simulation time limit.
-"NEURON" performed consistently at high speed, likely due to a lack of programmatic overhead, though "DynaSim" and "xolotl" were significantly faster at large time steps and during long simulations.
+We used the "replicate" method to create simulations with varying numbers of neurons (either Hodgkin-Huxley-like or bursting neurons) and measured the speed of all three simulators as a function of system size.
+Plotting the speed normalised by the system size, we observed that the speed of integration of "xolotl" is linear with system size (Figure \ref{fig:figurebenchmark}G-H, black lines), for up to 1000 single-compartment neurons (up to 13,000 ODEs).
+Its performance compares favourably with that of "NEURON" and "DynaSim" as a function of system size (Figure \ref{fig:figurebenchmark}G-H).
 
@@
 We set out to design a neuron and network simulator that could be useful in the classroom setting, especially for students of computational science, while also being powerful, fast, and extensible enough to be used for research.
-By using a novel architecture that permits the symbolic manipulation of "C++" objects in a comprehensible "MATLAB" interface, we showed how to create and simulate single-compartment models of Hogkin-Huxley like neurons (Figure \ref{fig:figurehh}), voltage clamp experiments and recover activation functions of single channel types (Figure \ref{fig:figureclamp}), a network of neurons with multiple synapse types (Figure \ref{fig:figurenetwork}), a multi-compartment neuron model where the voltage of every compartment is coupled, and a neuron model where homeostatic mechanisms control the maximal conductance of multiple channel types (Figure (\ref{fig:figuremechanism}).
+By using a novel architecture that permits the symbolic manipulation of "C++" objects in a intuitive "MATLAB" interface, we demonstrated some of the features of our simulator using simulations of single-compartment models of Hogkin-Huxley like neurons (Fig.
+\ref{fig:figurehh}); voltage clamp experiments to recover activation functions of single channel types (Fig.
+\ref{fig:figureclamp}); a neuron model where intracellular mechanisms can control the dynamics of Calcium and can regulate the maximal conductance of ion channel types in an activity-dependent manner (Fig.
+(\ref{fig:figuremechanism}); and a network of neurons with multiple synapse types (Fig.
+\ref{fig:figurenetwork}).
+We also illustrated how built-in features of our simulators make it easy to bookmark and jump between model configurations (Fig.
+\ref{fig:figuremechanism}, purple), and how parameters of the model can be changed using sliders and their effect can be viewed in real time (Fig.
+\ref{fig:figuremanipulate}).
 
-\subsection{Design: components vs. equations in simulators }
+\subsection{A focus on usability}
 
-- components from library vs. equations
-- equations are transparent -- you know what you're solving. but clunky -- string parsing headaches, etc.
-- components are easy to assemble, but it;s hard to know where they are, how they can be changed, etc.
-- xolotl goes the component route, but makes it very easy to find components (you can click your way into the C++ files).
-components are stored as source files, so you can modify it and immediately see it reflected in your model.
-we didn't invent another synatax for these components -- these are just C++ files.
-
-\subsection{Comparison with other simulators}
-
-\begin{quote}
-	\textit{``About half the time spent on a typical simulation project involves creating and tuning the model.
-Thus, a good user interface may contribute more to the overall efficiency of a project than pure computation speed.''}
-	\begin{flushright}
-		\cite{deschutterConsumerGuideNeuronal1992}
-	\end{flushright}
-\end{quote}
-
-"xolotl" is designed with ease-of-use in mind.
+``About half the time spent on a typical simulation project involves creating and tuning the model.
+Thus, a good user interface may contribute more to the overall efficiency of a project than pure computation speed.'' \cite{deschutterConsumerGuideNeuronal1992} "xolotl" is designed primarily with ease-of-use in mind.
 This includes the time it takes to install, setup, and learn the software, the time to write and debug scripts, and the time to perform the simulations.
@@
 
-The front-end interface to "xolotl" is built in "MATLAB" to facilitate interoperability with existing tools.
-"MATLAB" provides a comprehensive distribution from a single vendor with straightforward and powerful toolboxes for time series analysis, optimization, and parallel computing.
-For these reasons, "MATLAB" has remained a popular tool for data analysis and visualization in neuroscience.
-Since "xolotl" can be run directly from the "C++" executable, it would be possible to have implemented the interface in other popular scientific computing languages (\viz "Python" or "Julia").
-Ultimately due to the popularity and ease of "MATLAB" for pedagogy and research, and its extensive support for vector and matrix mathematics, "xolotl" supports a "MATLAB" interface.
+We have focussed on making our software as easy to use as possible, without sacrificing performance or extensibility.
+For example, the software and all dependencies can be installed using a single-line installer script from within the "MATLAB" command line.
+The installation includes worked example scripts that demonstrate various features of the simulator, that can be run without any configuration.
+We decided to built the the front-end interface to "xolotl" in "MATLAB" to facilitate interoperability with existing tools for time series analysis, optimization, and parallel computing.
+This allowed us to build rich tools for visualisation and to interact with the simulation.
+
+\subsection{Comparison with other simulators}
+
+Over the years, several simulators have been developed to integrate systems of coupled differential equations,  those that model the spiking activity of neurons (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003, hinesNEURONSimulationEnvironment1997}).
+A critical design choice
 
@@
 In addition, editing a feature of the model or simulation requires directly editing the mechanism or equation script and cannot be done in the console.
+
+- components from library vs. equations
+- equations are transparent -- you know what you're solving. but clunky -- string parsing headaches, etc.
+- components are easy to assemble, but it;s hard to know where they are, how they can be changed, etc.
+- xolotl goes the component route, but makes it very easy to find components (you can click your way into the C++ files).
+components are stored as source files, so you can modify it and immediately see it reflected in your model.
+we didn't invent another synatax for these components -- these are just C++ files.
+- scaffolding
 

### 2018-08-14 figures/benchmarks/figure_benchmark.m

@@ -13,7 +13,7 @@ for i = [1 2 5 6]
 end
 
 for i = [1 3 5 7]
-    ylabel(ax(i),'Speed (X realtime)')
+    ylabel(ax(i),'Relative Speed')
 end
 
 
@@ -27,7 +27,7 @@ for i = [3 7]
 end
 
 for i = [4 8]
-    ylabel(ax(i),'Speed*N (X realtime)')
+    ylabel(ax(i),'Relative Speed x N')
     xlabel(ax(i),'N')
 end
 
@@ -50,38 +50,40 @@ disp('Begin xolotl STG')
 testXolotlSTG(ax(5:end)); drawnow
 
 
-prettyFig('lw',.5,'plw',1.5,'tick_length',.03);
-return
-
-
 
 disp('Begin DynaSim HH')
 testDynaSimHH(ax); drawnow
 
 
-
-
-
-
-
-
 disp('Begin DynaSim STG')
-testDynaSimSTG(ax); drawnow
-
-
-
-
-
-
-return
-
-
-
-
-
-
+testDynaSimSTG(ax(5:end)); drawnow
 
 
 
 disp('Begin NEURON STG')
-testNeuronSTG(ax); drawnow
+testNeuronSTG(ax(5:end)); drawnow
+
+
+prettyFig('lw',.5,'plw',1.5,'tick_length',.03);
+
+
+movePlot(ax([1 5]),'left',.05)
+movePlot(ax([2 6]),'left',.04)
+movePlot(ax([3 7]),'left',.03)
+movePlot(ax([4 8]),'left',.02)
+
+for i = 1:8
+	axis(ax(i),'square')
+end
+
+% nice legends
+clear l L 
+
+l(1) = plot(ax(4),NaN,NaN,'k.','MarkerSize',34);
+l(3) = plot(ax(4),NaN,NaN,'b.','MarkerSize',34);
+l(2) = plot(ax(4),NaN,NaN,'r.','MarkerSize',34);
+
+lh = legend(l,{'xolotl','DynaSim','NEURON'});
+lh.Position = [0.9071    0.7416    0.0690    0.1037];
+
+labelFigure('capitalise',true,'column_first',true,'x_offset',-.02,'y_offset',-.02,'font_size',28)

### 2018-08-14 figures/benchmarks/loadNeuronData.m

@@ -17,7 +17,12 @@ t_end = 30000;
 disp('loading NEURON data...')
 
 
-allfiles = dir([pwd '/neuron/neuron_HH_raw*.npy']);
+if any(strfind(filename,'HH'))
+	allfiles = dir([pwd '/neuron/neuron_HH_raw*.npy']);
+else
+	allfiles = dir([pwd '/neuron/neuron_STG_raw*.npy']);
+end
+
 all_V = {};
 all_t = {};
 for i = 1:length(allfiles)

### 2018-08-14 figures/benchmarks/neuron/HH_benchmark3.py

@@ -17,15 +17,18 @@ if my_file1.is_file():
 # set up vectors to hold outputs
 nComps = np.array([1, 2, 4, 8, 16, 32, 64, 128, 250, 500, 1000])
 
-sim_time        = np.zeros((len(nComps),1))
 S               = np.zeros((len(nComps),1))
-sec             = []
-stim            = []
+
 
 # perform the simulation
 for ii in range(0,len(nComps)):
     percent = 100*ii/len(nComps)
     print('percent complete:  ' + repr(percent) + '%')
+
+    sec   = []
+    stim  = []
+
+
     for qq in range(0,nComps[ii]-1):
         # create the neuron
         compName = 'comp #'+repr(qq)
@@ -60,8 +63,10 @@ for ii in range(0,len(nComps)):
     tic = time.perf_counter() # s
     h.run()
     toc = time.perf_counter() # s
-    sim_time[ii] = (toc - tic) * 1000; # ms
-    S[ii] = h.tstop / sim_time[ii] # unitless
+
+
+    sim_time = (toc - tic) * 1000; # ms
+    S[ii] = h.tstop / sim_time # unitless
 
 
 print("All done, saving results...")

### 2018-08-14 figures/benchmarks/neuron/STG_benchmark1.py

@@ -9,6 +9,7 @@ from matplotlib import pyplot
 import time
 from pathlib import Path
 
+
 benchmarks_file = "neuron_STG_benchmark1.csv"
 
 my_file1 = Path(benchmarks_file)

### 2018-08-14 figures/benchmarks/neuron/STG_benchmark2.py

@@ -7,6 +7,15 @@ from neuron import h, gui
 import numpy as np
 from matplotlib import pyplot
 import time
+from pathlib import Path
+
+benchmarks_file = "neuron_STG_benchmark2.csv"
+
+my_file1 = Path(benchmarks_file)
+if my_file1.is_file():
+    print("It looks like the benchmark is already done, so aborting...")
+    exit()
+
 
 # create the neuron
 soma        = h.Section(name='soma');
@@ -75,20 +84,26 @@ toc         = time.perf_counter() # s
 t_end = np.array([1, 2, 4, 9, 18, 38, 78, 162, 336, 695, 1438, 2976,
 6158, 12743, 26367, 54556, 112884, 233572, 483293, 1000000]) # ms
 
-sim_time        = np.zeros((len(t_end),1))
 S               = np.zeros((len(t_end),1))
 
 # perform the simulation
 for ii in range(0,len(t_end)):
     percent = 100*ii/len(t_end)
     print('percent complete:  ' + repr(percent) + '%')
-    h.tstop         = t_end[ii]
-    stim.dur        = t_end[ii]
+    h.tstop  = t_end[ii]
+    stim.dur = t_end[ii]
+
+
     tic             = time.perf_counter() # s
     h.run()
     toc             = time.perf_counter() # s
-    sim_time[ii]    = (toc-tic) * 1000; # ms
-    S[ii]           = t_end[ii] / sim_time[ii] # unitless
+
+
+    sim_time    = (toc-tic) * 1000; # ms
+    S[ii]  = t_end[ii] / sim_time # unitless
 
 # save the results
-np.savetxt("neuron_STG_benchmark2.csv", S, delimiter=",")
+np.savetxt(benchmarks_file, S, delimiter=",")
+
+
+exit()

### 2018-08-14 figures/benchmarks/neuron/STG_benchmark3.py

@@ -4,17 +4,26 @@
 import numpy as np
 from neuron import h, gui
 import time
+from pathlib import Path
+
+benchmarks_file = "neuron_STG_benchmark3.csv"
+
+my_file1 = Path(benchmarks_file)
+if my_file1.is_file():
+    print("It looks like the benchmark is already done, so aborting...")
+    exit()
 
 # set up vectors to hold outputs
 nComps = np.array([1, 2, 4, 8, 16, 32, 64, 128, 250, 500, 1000])
 
-sim_time        = np.zeros((len(nComps),1))
-S               = np.zeros((len(nComps),1))
-sec             = []
-stim            = []
+S    = np.zeros((len(nComps),1))
 
 # perform the simulation
 for ii in range(0,len(nComps)):
+
+    sec  = []
+    stim = []
+
     percent = 100*ii/len(nComps)
     print('percent complete:  ' + repr(percent) + '%')
     for qq in range(0,nComps[ii]-1):
@@ -73,8 +82,8 @@ for ii in range(0,len(nComps)):
     tic = time.perf_counter() # s
     h.run()
     toc = time.perf_counter() # s
-    sim_time[ii] = (toc - tic) * 1000; # ms
-    S[ii] = h.tstop / sim_time[ii] # unitless
+    sim_time = (toc - tic) * 1000; # ms
+    S[ii] = h.tstop / sim_time # unitless
 
 # save the results
-np.savetxt("neuron_STG_benchmark3.csv", S, delimiter=",")
+np.savetxt(benchmarks_file, S, delimiter=",")

### 2018-08-14 figures/benchmarks/neuron/benchmark-neuron.sh

@@ -6,20 +6,38 @@
 # import neuron 
 
 # compile the .mod files
+clear
 nrnivmodl na.mod cat.mod cas.mod acurrent.mod kca.mod kd.mod hcurrent.mod cad.mod
+echo "Compiled NEURON files"
 
 # basic check to make sure things are working
+clear
+echo "Starting a basic check..."
 python test_STG.py
 
-# HH model -- varying dt
+
+clear
+echo "HH model -- varying dt."
 python HH_benchmark1.py
 
-# HH mdoel -- varying t_end
+
+clear
+echo "HH model -- varying t_end"
 python HH_benchmark2.py
 
-# HH mdoel -- varying n_comp
+clear
+echo "HH model -- varying n_comp"
 python HH_benchmark3.py
 
 
-# run STG test 
+clear
+echo "STG model -- varying dt"
 python STG_benchmark1.py
+
+clear
+echo "STG model -- varying t_end"
+python STG_benchmark2.py
+
+clear
+echo "STG model -- varying n_comp"
+python STG_benchmark3.py

### 2018-08-14 figures/benchmarks/testDynaSimHH.m

@@ -174,6 +174,7 @@ else
 	S = cache(h);
 end
 
+S = S.*nComps;
 
 
 plot(ax(4),nComps,S,'r-o')

### 2018-08-14 figures/benchmarks/testDynaSimSTG.m

@@ -146,10 +146,10 @@ S = t_end ./ all_sim_time;
 S = S * 1e-3;
 
 % plot speed over increasing time step
-plot(ax(2+5),all_dt,S,'r-o')
+plot(ax(1),all_dt,S,'r-o')
 
 % plot error over increasing time-step
-plot(ax(3+5),all_dt,matrix_error,'r-o')
+plot(ax(2),all_dt,matrix_error,'r-o')
 
 
 %% Increasing Simulation Time
@@ -184,7 +184,7 @@ end
 S = all_t_end./all_sim_time;
 S = S*1e-3;
 
-plot(ax(4+5),all_t_end,S,'r-o')
+plot(ax(3),all_t_end,S,'r-o')
 
 
 %% Increasing Number of Compartments
@@ -230,4 +230,7 @@ else
 	S = cache(h);
 end
 
-plot(ax(5+5),nComps,S,'r-o')
+
+S = S(:).*nComps(:);
+
+plot(ax(4),nComps,S,'r-o')

### 2018-08-14 figures/benchmarks/testNeuronSTG.m

@@ -20,14 +20,14 @@ function testNeuronSTG(ax)
 ;;     ;; ;;       ;;          ;;    ;;     ;;    ;;
 ;;;;;;;;  ;;;;;;;; ;;;;;;;;    ;;    ;;     ;;    ;;
 
-% time-consuming step, requires analyzing ~ 3.3 GB of data
-[Q, S, all_dt] = loadNeuronData('../../neuron/neuron_STG_benchmark1');
+
+[Q, S, all_dt] = loadNeuronData('./neuron/neuron_STG_benchmark1');
 
 % plot simulation speed vs. time step on axes #2
-plot(ax(2+5), all_dt, S, 'b-o')
+plot(ax(1), all_dt, S, 'b-o')
 
 % plot simulation error vs time step on axes #3
-plot(ax(3+5),all_dt, Q, 'b-o')
+plot(ax(2),all_dt, Q, 'b-o')
 
 
 
@@ -50,10 +50,10 @@ plot(ax(3+5),all_dt, Q, 'b-o')
 
 
 all_t_end   = unique(round(logspace(0,6,20)));
-S           = csvread('../../neuron/neuron_STG_benchmark2.csv');
+S           = csvread('./neuron/neuron_STG_benchmark2.csv');
 
 % plot simulation speed vs. simulation time on axes #4
-plot(ax(4+5),all_t_end, S, 'b-o')
+plot(ax(3),all_t_end, S, 'b-o')
 
  ;;;;;;  ;;    ;;  ;;;;;;  ;;;;;;;; ;;;;;;;; ;;     ;;
 ;;    ;;  ;;  ;;  ;;    ;;    ;;    ;;       ;;;   ;;;
@@ -74,7 +74,9 @@ plot(ax(4+5),all_t_end, S, 'b-o')
 
 % nComps      = unique(round(logspace(0,3,21)));
 nComps      = [1, 2, 4, 8, 16, 32, 64, 128, 250, 500, 1000];
-S           = csvread('../../neuron/neuron_STG_benchmark3.csv');
+S           = csvread('./neuron/neuron_STG_benchmark3.csv');
+
+S = S(:).*nComps(:);
 
 % plot simulation speed vs. number of compartments on axes #5
-plot(ax(5+5),nComps, S, 'b-o')
+plot(ax(4),nComps, S, 'b-o')

### 2018-08-14 paper/paper.tex
getting close to the final stretch

@@
 
-\title[xolotl: neuron \& network simulator]{Xolotl: An Intuitive and Approachable Neuron \& Network Simulator in MATLAB}
+\title[xolotl: neuron \& network simulator]{xolotl: An Intuitive and Approachable Neuron \& Network Simulator}
 \author[\firstAuthorLast ]{\Authors}
@@
 
-	Conductance-based models of neurons have been used extensively in computational neuroscience.
-However, working with these models continues to be a challenge due to their high dimensionality and large number of parameters.
-Here, we present a neuron and network simulator built on a novel automatic type system that binds object-oriented code written in "C++" to objects in "MATLAB".
-Our approach builds on the tradition of uniting the speed and power of languages like "C++" with the ease-of-use and feature-set of scientific programming languages like "MATLAB".
+	Conductance-based models of neurons are used extensively in computational neuroscience, and working with these models continues to be a challenge due to their high dimensionality and large number of parameters.
+Here, we present a neuron and network simulator built on a novel automatic type system that binds object-oriented code written in C++ to objects in MATLAB.
+Our approach builds on the tradition of uniting the speed and power of languages like C++ with the ease-of-use and feature-set of scientific programming languages like MATLAB.
 Our software allows for the creation and manipulation of hierarchical models with components that are named and searchable, permitting intuitive high-level programmatic control over all parameters.
@@
 Our simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be extended to include arbitrary conductances, synapses, and mechanisms.
+Several core features like bookmarking of parameters and automatic hashing of source code promote reproducible and auditable research.
 Finally, our simulator is written in a modular fashion and has been released under a permissive free software license, enabling seamless integration into the workflow of other researchers.
@@
 	\tiny
-	\keyFont{ \section{Keywords:} simulator, code:MATLAB, code:C++, conductance-based, neuron, network, pedagogy}
+	\keyFont{ \section{Keywords:} simulator, code:MATLAB, code:C++, conductance-based, neuron, network, neuron models}
 
@@
 In multi-compartment models of spatially extended neurons, membrane potentials in every compartment can be different, and are coupled to each other.
-Finally, the choice of programming language used to implement the model imposes tradeoffs in designing and using neuron and network simulators: simulators written in languages like "C++" or "FORTRAN" can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like "Python", "Julia", or "MATLAB".
+Finally, the choice of programming language used to implement the model imposes tradeoffs in designing and using neuron and network simulators: simulators written in languages like C++ or FORTRAN can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like Python, Julia, or MATLAB (Mathworks).
 
 Two major approaches have dominated the design of neuron simulators.
-First, some simulators, such as "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}) are written in a fast compiled language like "C" and allow for the construction and simulation of neuron models using object-oriented paradigms.
-These simulators tend to perform fast computations with little overhead, but suffer from a steep learning curve.
+One approach is to write the simulator in a fast compiled language like C and allow for the construction and simulation of neuron models using object-oriented paradigms.
+This approach has been implemented in NEURON (\cite{hinesNEURONSimulationEnvironment1997}).
+Simulators designed in this way tend to perform fast computations with little overhead, but suffer from a steep learning curve.
 Wrapping these simulators in a more approachable language like "Python"  or using graphical user interfaces (GUIs) mitigates these drawbacks (\cite{hinesNEURONPython2009}) at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
 In contrast, simulators designed from the ground up in popular scientific computing languages can be easier to use and benefit from interoperability with other commonly-used tools.
-Simulators like "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}), "morphforge" (\cite{hullMorphforgeToolboxSimulating2014}), and "PyNN" (\cite{davisonPyNNCommonInterface2009}) allow the user to specify models with strings of equations or components that are constructed using a special syntax, that can then be translated into a faster implementation language such as "C" or "C++" (\cite{stimbergEquationorientedSpecificationNeural2014}).
+Simulators like DynaSim (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}), "morphforge" (\cite{hullMorphforgeToolboxSimulating2014}), and "PyNN" (\cite{davisonPyNNCommonInterface2009}) allow the user to specify models with strings of equations or components that are constructed using a special syntax, that can then be translated into a faster implementation language such as "C" or C++ (\cite{stimbergEquationorientedSpecificationNeural2014}).
 This approach permits considerable flexibility for simulating systems of differential equations.
@@
 
-To overcome these design limitations, we have developed a novel automatic type system, that we call "cpplab", which binds "MATLAB" code to classes specified in "C++" header files.
-This architecture automatically creates objects in "MATLAB" that represent the underlying object-oriented "C++" code, allowing the symbolic manipulation of "C++" objects in the "MATLAB" interface.
-In this paper, we introduce "xolotl", an implementation of the "cpplab" system specialized in integrating conductance-based neuron and network models.
-Models can be easily constructed from network building blocks in a few lines of "MATLAB" code using a hierarchical and intuitive syntax.
-Since models in the "MATLAB" workspace are automatically linked to models in the "C++" implementation, configuring these objects in "MATLAB" transparently configures the underlying "C++" objects.
-"xolotl" comes packaged with hundreds of components that can be used to assemble cells and networks;  has built-in visualization functions to inspect voltage time traces and activation functions; and a GUI for real-time manipulation of model parameters.
+To overcome these design limitations, we have developed a novel automatic type system, that we call "cpplab", which binds MATLAB code to classes specified in C++ header files.
+This architecture automatically creates objects in MATLAB that represent the underlying object-oriented C++ code, allowing the symbolic manipulation of C++ objects in the MATLAB interface.
+In this paper, we introduce xolotl, an implementation of the "cpplab" system specialized in integrating conductance-based neuron and network models.
+Models can be easily constructed from network building blocks in a few lines of MATLAB code using a hierarchical and intuitive syntax.
+Since models in the MATLAB workspace are automatically linked to models in the C++ implementation, configuring these objects in MATLAB transparently configures the underlying C++ objects.
+xolotl comes packaged with hundreds of components that can be used to assemble cells and networks;  has built-in visualization functions to inspect voltage time traces and activation functions; and a GUI for real-time manipulation of model parameters.
 The ease-of-use of the tools built into our software lends them to pedagogical applications and the rapid prototyping of models.
@@
 
-"xolotl" is designed to be easy-to-use and richly featured while being fast enough to use in everyday research.
+xolotl is designed to be easy-to-use and richly featured while being fast enough to use in everyday research.
 Our focus was on designing an approachable simulator of conductance-based neurons and networks of these neurons; simulating arbitrary dynamical systems is therefore beyond the scope of this software.
@@
 
-This software is designed to be used from within "MATLAB", a scientific programming language common amongst neuroscientists and engineers for pedagogy and research.
-Our goal was to make "xolotl" completely usable entirely from within "MATLAB".
-Models created using this simulator appear in the "MATLAB" workspace as native objects, are are thus fully scriptable, and are fully compatible with the large library of toolboxes that "MATLAB" provides, allowing the software to be used as a component of other packages and tools.
+This software is designed to be used from within MATLAB, a scientific programming language common amongst neuroscientists and engineers for pedagogy and research.
+Our goal was to make xolotl completely usable entirely from within MATLAB.
+Models created using this simulator appear in the MATLAB workspace as native objects, are are thus fully scriptable, and are fully compatible with the large library of toolboxes that MATLAB provides, allowing the software to be used as a component of other packages and tools.
 All parameters of models and activation functions of any channel can be inspected at any point.
-We designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by calling functions that generate new "C++" files on-the-fly.
-Finally, "xolotl" fully auditable by design, with several tools to verify model and parameter integrity and aid in reproducibility.
+We designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by calling functions that generate new C++ files on-the-fly.
+Finally, xolotl fully auditable by design, with several tools to verify model and parameter integrity and aid in reproducibility.
 
@@
 
-\paragraph{A rich library of network components.} "xolotl" comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and mechanisms) that can be used as building blocks to construct model neurons and networks.
+\paragraph{A rich library of network components.} xolotl comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and mechanisms) that can be used as building blocks to construct model neurons and networks.
 Compartments represent sections of membrane with a single membrane potential, Calcium level, and set of constituent components, representing either entire cells or parts of cells.
@@
 \paragraph{Object-oriented.}
-"xolotl" is designed to mirror the nested and hierarchical structure of networks and neurons.
+xolotl is designed to mirror the nested and hierarchical structure of networks and neurons.
 Biological neuronal networks are made up of neurons that are connected to each other with synapses.
@@
 Intracellular mechanisms can act within the cell, or parts of the cell, to modify and regulate conductances, synapses, or other dynamic properties of the cell.
-Similarly, a "xolotl" model can contain a set of compartments that can represent individual neurons.
+Similarly, a xolotl model can contain a set of compartments that can represent individual neurons.
 Each compartment can contain an arbitrary set of conductances.
 Compartments, conductances and synapses can contain mechanisms that can affect anything in the model.
-All objects in the "xolotl" model tree (compartments, conductances, etc.) are bonafide "MATLAB" objects with their own type and properties.
+All objects in the xolotl model tree (compartments, conductances, etc.) are bonafide MATLAB objects with their own type and properties.
 This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain neurons that contain populations of ion channels, and makes constructing, integrating, and thinking about these models easier.
@@
 \paragraph{Automatic type system.}
-To circumvent the tradeoff between high-performance but hard to use languages like "C++" and richly-featured but potentially slow languages like "MATLAB", we constructed an automatic type system that links object oriented code in "C++" to object oriented code in "MATLAB".
-This architecture lets us construct the core of the simulator in "C++", leveraging features of "C++" like pointers that are not readily available in "MATLAB".
-A rudimentary way to make this "C++" code useable in "MATLAB" would be to re-write that code in "MATLAB" so that "MATLAB" objects can be bound to their "C++" implementations.
+To circumvent the tradeoff between high-performance but hard to use languages like C++ and richly-featured but potentially slow languages like MATLAB, we constructed an automatic type system that links object oriented code in C++ to object oriented code in MATLAB.
+This architecture lets us construct the core of the simulator in C++, leveraging features of C++ like pointers that are not readily available in MATLAB.
+A rudimentary way to make this C++ code useable in MATLAB would be to re-write that code in MATLAB so that MATLAB objects can be bound to their C++ implementations.
 However, this approach is cumbersome and inefficient, and can introduce errors.
-Instead, our automatic type system creates objects in "MATLAB" on-the-fly from "C++" class specifications, obviating the need to rewrite code in "MATLAB" while preserving a tight coupling between objects in the "MATLAB" workspace and their "C++" implementation.
+Instead, our automatic type system creates objects in MATLAB on-the-fly from C++ class specifications, obviating the need to rewrite code in MATLAB while preserving a tight coupling between objects in the MATLAB workspace and their C++ implementation.
 Crucially, this method makes developing new code much easier and simplifies the task of constructing complex frameworks that span these two languages.
@@
 Since every model requires a compiled binary to run, a potential stumbling block is the problem of unambiguously linking a model to a binary executable.
-"xolotl" solves this problem by hashing (\cite{rivestMD5MessageDigestAlgorithm1992}) the "C++" header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier (its ``hash'').
+xolotl solves this problem by hashing (\cite{rivestMD5MessageDigestAlgorithm1992}) the C++ header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier (its ``hash'').
 Compiled binaries are named using this hash, ensuring both that the correct binary is run to integrate the model, and that compilation occurs only as needed.
-This powerful feature enables the user experience to remain entirely within the "MATLAB" workspace, with compilation and selection of the correct binary occurring silently in the background.
+This powerful feature enables the user experience to remain entirely within the MATLAB workspace, with compilation and selection of the correct binary occurring silently in the background.
 
@@
 
-Our focus on ease-of-use and speed means some features were intentionally neglected in the design process.
+Our focus on our simulator's ease-of-use and speed imposed some limitations on its feature set.
 
-\paragraph{Limited to conductance-based models.} "xolotl" has been developed specifically for conductance-based models.
+\paragraph{Limited to conductance-based models.} xolotl has been developed specifically for conductance-based models.
 It does not currently support rate- or current-based models, or arbitrary dynamical systems.
@@
 However, it may be desirable to use other methods under certain conditions.
-It is possible to introduce new network components that implement other integration schemes, or to modify the integration schemes of existing components, but that requires writing new "C++"  code.
-Currently, "xolotl" can only implement integration schemes with fixed step size.
+It is possible to introduce new network components that implement other integration schemes, or to modify the integration schemes of existing components, but that requires writing new C++  code.
+Currently, xolotl can only implement integration schemes with fixed step size.
 
-\paragraph{Inefficient tools for handling large networks.} "xolotl" was designed to work with small but complex networks and models, where every compartment and component is named, rather than numbered.
+\paragraph{Inefficient tools for handling large networks.} xolotl was designed to work with small but complex networks and models, where every compartment and component is named, rather than numbered.
 It is more therefore suited towards simulating small and heterogeneous networks rather than large but homogenous networks.
 While the software can integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task, offering more natural frameworks for dealing with a large number of identical units.
-Similarly, "xolotl" is not optimized to solve coupled ODEs on complex branched morphologies, and other simulators like (\cite{hinesNEURONSimulationEnvironment1997}) are more suited for this task.
+Similarly, xolotl is not optimized to solve coupled ODEs on complex branched morphologies, and other simulators like (\cite{hinesNEURONSimulationEnvironment1997}) are more suited for this task.
 
-\paragraph{New mechanisms require new \texttt{C++} code.} Adding new network components requires writing new "C++" code.
-A new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, dayanTheoreticalNeuroscience2001}) requires creating a new "C++" header file, though this is generally trivial.
-Implementing a new integration scheme or component type requires much more in-depth knowledge of the underlying "C++" core code.
+\paragraph{New mechanisms require new \texttt{C++} code.} Adding new network components requires writing new C++ code.
+A new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, dayanTheoreticalNeuroscience2001}) requires creating a new C++ header file, though this is generally trivial.
+Implementing a new integration scheme or component type requires much more in-depth knowledge of the underlying C++ core code.
 
@@
 
-In this section, we illustrate how "xolotl" can be used to generate, inspect, and simulate a variety of models.
-These examples have been chosen to illustrate various features of "xolotl", and are intended to serve as templates upon which researchers and educators can build.
+In this section, we illustrate how xolotl can be used to generate, inspect, and simulate a variety of models.
+These examples have been chosen to illustrate various features of xolotl, and are intended to serve as templates upon which researchers and educators can build.
 
@@
 
-"xolotl" can reproduce such a voltage clamp experiment \textit{in-silico}.
+xolotl can reproduce such a voltage clamp experiment \textit{in-silico}.
 Figure \ref{fig:figureclamp}B illustrates how a simple model with a single compartment and a single ion channel type can be set up and clamped to a desired voltage.
@@
 \ref{fig:figureclamp}G-H).
-"xolotl" can therefore be used to describe graphically the theoretical underpinnings of ion channel characterization through voltage clamp and can serve as an effective pedagogical tool in computational neuroscience.
+xolotl can therefore be used to describe graphically the theoretical underpinnings of ion channel characterization through voltage clamp and can serve as an effective pedagogical tool in computational neuroscience.
 
@@
 However, real neurons possess several dynamical features, arising from a variety of intracellular mechanisms.
-"xolotl" makes it possible to model and include arbitrary intracellular mechanisms.
+xolotl makes it possible to model and include arbitrary intracellular mechanisms.
 In our simulator, these intracellular mechanisms are represented by the "mechanism" object, and can be bound to compartments, conductances, and other object types.
@@
 Calcium buffering and influx can be modelled by a differential equation that increases intracellular Calcium with the current through Calcium channels and relaxes back to a baseline value (\cite{liuModelNeuronActivityDependent1998, prinzAlternativeHandtuningConductancebased2003, dayanTheoreticalNeuroscience2001}) (Figure \ref{fig:figuremechanism}D).
-This mechanism exists in the "xolotl" library as "CalciumMech1" and can be added to the model using a simple "add" statement (Figure \ref{fig:figuremechanism}E).
+This mechanism exists in the xolotl library as "CalciumMech1" and can be added to the model using a simple "add" statement (Figure \ref{fig:figuremechanism}E).
 The intracellular Calcium in the model now oscillates periodically (Figure \ref{fig:figuremechanism}F), synchronized to bursts in action potentials in this cell (Figure \ref{fig:figuremechanism}G).
@@
 Neurons can regulate their electrical activity by controlling the spectrum of ion channels they express (\cite{macleanActivityIndependentHomeostasisRhythmically2003, turrigianoSelectiveRegulationCurrent1995, schulzVariableChannelExpression2006}).
-Here, we will show how "xolotl" can be used to represent a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H).
+Here, we will show how xolotl can be used to represent a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H).
 Since this mechanism affects each ion channel population individually, an object corresponding to this mechanism is added to each "conductance" object in the neuron (Figure \ref{fig:figuremechanism}I).
@@
 (\cite{deschutterConsumerGuideNeuronal1992}).
-"xolotl" makes it easy to bookmark configurations of a model and return to them at will.
-Internally, "xolotl" uses the "serialize" method to gather all parameters and dynamic variables into a vector of values that is passed to the underlying "C++" implementation.
+xolotl makes it easy to bookmark configurations of a model and return to them at will.
+Internally, xolotl uses the "serialize" method to gather all parameters and dynamic variables into a vector of values that is passed to the underlying C++ implementation.
 A paired "deserialize" method is used to update all parameters and variables in the object tree from this vector.
 This architecture provides a natural framework for representing the state of any model, no matter how complex, using a vector of numbers.
-The "snapshot" method built into "xolotl" leverages this schema to save the entire state of the model in a named variable, that can be accessed using another built-in method called "reset".
+The "snapshot" method built into xolotl leverages this schema to save the entire state of the model in a named variable, that can be accessed using another built-in method called "reset".
 
@@
 
-Network models in "xolotl" consist of "compartment" objects that can be connected by "synapse" objects.
+Network models in xolotl consist of "compartment" objects that can be connected by "synapse" objects.
 Two compartments representing different neurons can be connected using synapses using the built-in "connect" method.
@@
 The pyloric model contains three compartments ("AB", "LP", and "PY") and seven synapses (Figure \ref{fig:figurenetwork}A).
-This structure is recapitulated in the hierarchy of the "xolotl" object, where conductances are contained within compartments (Figure \ref{fig:figurenetwork}B).
+This structure is recapitulated in the hierarchy of the xolotl object, where conductances are contained within compartments (Figure \ref{fig:figurenetwork}B).
 The membrane potentials show triphasic rhythmicity in the three compartments (Figure \ref{fig:figurenetwork}C-E).
@@
 
-The "manipulate" function opens the GUI which permits visualization of changing parameters in real-time.
-Moving sliders representing the values of network parameters updates a plot (Figure \ref{fig:figuremanipulate}B).
-By default, the function opens a figure displaying the results of the "plot" function, which shows the voltage and intracellular calcium traces for each compartment (Figure \ref{fig:figuremanipulate}A).
-"manipulate" grants slider control over all "xolotl" parameters by default, but specific ones can be selected by passing them as arguments.
+Conductance-based neuron models are typically high dimensional and contain several parameters.
+Changing a single parameter monotonically can cause non-monotonic changes in behavior of the model, and certain dynamical features may only emerge when in specific non-convex regions of parameter space (\cite{golowasch2002failure}).
+It is often challenging to build intuition about what effect a parameter has on the model under these conditions.
+Traditionally, the technique used by computational neuroscientists in building intuition about these models is to iteratively run simulations, view outputs and change parameters.
+In practice, this meant writing a script, running it, inspecting the output, changing parameters in the script, and repeating this process.
+It can be cumbersome, and every step in this process involves ``mode'' changes: between a text editor, viewing a graphical output, and the command line that can frustrate the researcher.
+
+Our software is dfesigned to streamline this process and allows for any parameter in any model to be manipulated using graphical sliders, with immediate, real-time feedback of its behavior.
+Any model in our simulator can be manipulated using
+
+\begin{lstlisting}[style=Matlab-editor]
+x.manipulate 
+\end{lstlisting}
+
+This method creates a GUI element with sliders for every parameter, and also creates a set of plots that shows the dynamical behavior of the model (Fig.
+\ref{fig:figuremanipulate}).
+By default, time series of the voltage and the calcium of every compartment are shown, though this can be modified.
+Moving any of the sliders updates the value of that parameter in the model, and also triggers a function call that reintegrates the model and updates the output plots.
+This function call can also update custom plots, like the one shown in (Fig.
+\ref{fig:figuremanipulate}B).
+Any model can be manipulated in this way without writing any additional code.
+
+This feature is only possible due to our architectural decision to split the code base across two programming environments.
+A rich scientific programming environment like MATLAB makes it possible to generate user interface elements and to bind them to data in plots, while the sheer speed of compiled languages like C++ allow for the immediate, real-time feedback and updating of plots.
+By default, any parameter in the model can be manipulated, including parameters in user-defined mechanisms that do not exist in the base simulator.
+The GUI can be constrained to arbitrary subsets of model parameters using wild card matching (as shown in Fig.
+\ref{fig:figuremanipulate}) or by manually specifying the parameters of interest.
 
@@
 
-Our goal in designing "xolotl" was to create an easy-to-use neuron and network simulator that was fast enough and accurate enough for routine use by computational neuroscientists.
-In this section, we compare the speed of "xolotl" in simulating single and large numbers of Hodgkin-Huxley-like neuron models (as in Figure \ref{fig:figurehh}) and bursting neuron models based on the bursting neurons in the lobster STG (\cite{prinzSimilarNetworkActivity2004}).
-For each type of neuron model, we compared our software to  "NEURON" (\cite{hinesNEURONSimulationEnvironment1997}), a high-performance and powerful neuron simulator specialised in simulating neurons with complex morphologies and "DynaSim" (\cite{sherfeyDynaSimMATLABToolbox2018}), a general-purpose simulator that can solve coupled differential equations numerically.
-All simulators were run on the same hardware using fixed time step solvers: "xolotl" used the Exponential Euler method  (\cite{dayanTheoreticalNeuroscience2001}), "NEURON" used the  implicit Euler solver (\cite{hinesNEURONSimulationEnvironment1997}) and "DynaSim" used "C"-compiled 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance in the documentation.
-We measured the speed of each simulator by dividing the time simulated for by the time it took for the simulator to complete.
-For example, if a simulator could simulate 10 seconds of neuron data in 1 second, its speed would be 10X.
+Our goal in designing xolotl was to create an easy-to-use neuron and network simulator that was fast enough and accurate enough for routine use by computational neuroscientists.
+In this section, we benchmarked the speed and accuracy of xolotl in simulating single and large numbers of Hodgkin-Huxley-like neuron models (as in Fig.
+\ref{fig:figurehh}) and bursting neuron models based on the bursting neurons in the lobster stomatogastric ganglion (STG) (\cite{prinzSimilarNetworkActivity2004}).
+For each type of neuron model, we compared our software to  NEURON (\cite{hinesNEURONSimulationEnvironment1997}), a high-performance and powerful neuron simulator specialized in simulating neurons with complex morphologies and DynaSim (\cite{sherfeyDynaSimMATLABToolbox2018}), a general-purpose simulator that can solve coupled differential equations numerically.
+All simulators were run on the same hardware using fixed time-step solvers: xolotl used the Exponential Euler method (\cite{dayanTheoreticalNeuroscience2001}), NEURON used the  implicit Euler solver (\cite{hinesNEURONSimulationEnvironment1997}) and DynaSim used C-compiled 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance (\cite{sherfeyDynaSimMATLABToolbox2018}).
+We measured the speed of each simulator by dividing the time simulated for, by the time it took for the simulator to complete integration.
+For example, if a simulator could simulate 10 seconds of model data in 1 second, its speed would be 10X.
 We measured the speed of every simulator as a function of simulation time step, total length of simulation, and system size.
 
-All three simulators were faster with larger time steps, since fewer iterations were needed (Figure \ref{fig:figurebenchmark}A-B), and were approximately linear in the region tested.
-Our simulator compared favourably to "NEURON" and "DynaSim" in this task.
+All three simulators were faster with larger time steps, since fewer iterations were needed (Fig.
+\ref{fig:figurebenchmark}A-B), and were approximately linear in the region tested.
+Our simulator compared favorably to NEURON and DynaSim in this task.
 We also measured the quality of the simulated output by comparing it to the simulated output at the smallest time step.
-Simulation error was measured using the LeMasson cost (\cite{lemassonIntroductionEquationSolving2000}), and was comparable amongst the three simulators (Figure \ref{fig:figurebenchmark}C-D).
-Our simulator sets up and runs the simulation in "C++", and needs to transfer parameters and data to and from the underlying implementation.
-To measure the performance cost of this overhead, we repeated these simulations on all three simulators at a afixed time step of $0.1$ ms and varied the length of time simulated for.
-Speed increased with simulation duration up to a limit, and then saturated, indicating a fixed performance cost of the overhead (Figure \ref{fig:figurebenchmark}E-F, black lines).
-Simualtions using "DynaSim", which also used a similar architecture, showed a similar increase in speed with simulation length (Figure \ref{fig:figurebenchmark}E-F, red lines).
-In contrast, simulations using "NEURON" ran at a constant speed irrespective of simulation length, presumably due to differences in the underlying implementation (Figure \ref{fig:figurebenchmark}E-F, blue lines).
+Simulation error was measured using the LeMasson cost (\cite{lemassonIntroductionEquationSolving2000}), and was comparable amongst the three simulators (Fig.
+\ref{fig:figurebenchmark}C-D).
+Since our simulator sets up and runs the simulation in C++, it needs to transfer parameters and data to and from the underlying implementation.
+To measure the performance cost of this overhead, we repeated these benchmarks on all three simulators at a fixed time step of $0.1$ ms and varied the length of time simulated for.
+Speed increased with simulation duration up to a point, and then saturated, indicating a fixed performance cost to the overhead (Figure \ref{fig:figurebenchmark}E-F, black lines).
+Simulations using DynaSim, which also used a similar architecture and need to move data between C implementations and the MATLAB workspace, showed a similar increase in speed with simulation length (Figure \ref{fig:figurebenchmark}E-F, red lines).
+However, simulations using NEURON ran at a constant speed irrespective of simulation length, presumably due to differences in the underlying implementation (Figure \ref{fig:figurebenchmark}E-F, blue lines).
 
-Many simulators have been designed to simualte large numbers of compartments, either as networks with many identical neurons or in a large mulit-compartmentel neuron model (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}).
+Many simulators have been designed with a focus on simulate large numbers of compartments, either as networks with many identical neurons or in a large multi-compartment neuron model (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}).
 While our software is not designed for this task {\em per se}, we measured its performance as a function of the number of compartments simulated.
-"xolotl" can quickly create a number of identical copies of a compartment using the "replicate" method:
+xolotl can quickly create a number of identical copies of a compartment using the "replicate" method:
 
@@
 
-We used the "replicate" method to create simulations with varying numbers of neurons (either Hodgkin-Huxley-like or bursting neurons) and measured the speed of all three simulators as a function of system size.
-Plotting the speed normalised by the system size, we observed that the speed of integration of "xolotl" is linear with system size (Figure \ref{fig:figurebenchmark}G-H, black lines), for up to 1000 single-compartment neurons (up to 13,000 ODEs).
-Its performance compares favourably with that of "NEURON" and "DynaSim" as a function of system size (Figure \ref{fig:figurebenchmark}G-H).
+We used the "replicate" method to create and run models with varying numbers of neurons (either Hodgkin-Huxley-like or bursting neurons) and measured the speed of all three simulators as a function of system size.
+Plotting the speed normalized by the system size vs.
+the system size, we observed that the speed of integration of xolotl is linear with system size (Figure \ref{fig:figurebenchmark}G-H, black lines), for up to 1000 single-compartment neurons (up to 13,000 ODEs).
+Its performance compares favorably with that of NEURON and DynaSim as a function of system size (Figure \ref{fig:figurebenchmark}G-H).
 
@@
 We set out to design a neuron and network simulator that could be useful in the classroom setting, especially for students of computational science, while also being powerful, fast, and extensible enough to be used for research.
-By using a novel architecture that permits the symbolic manipulation of "C++" objects in a intuitive "MATLAB" interface, we demonstrated some of the features of our simulator using simulations of single-compartment models of Hogkin-Huxley like neurons (Fig.
+By using a novel architecture that permits the symbolic manipulation of C++ objects in a intuitive MATLAB interface, we demonstrated some of the features of our simulator using simulations of single-compartment models of Hogkin-Huxley like neurons (Fig.
 \ref{fig:figurehh}); voltage clamp experiments to recover activation functions of single channel types (Fig.
@@
 ``About half the time spent on a typical simulation project involves creating and tuning the model.
-Thus, a good user interface may contribute more to the overall efficiency of a project than pure computation speed.'' \cite{deschutterConsumerGuideNeuronal1992} "xolotl" is designed primarily with ease-of-use in mind.
-This includes the time it takes to install, setup, and learn the software, the time to write and debug scripts, and the time to perform the simulations.
+Thus, a good user interface may contribute more to the overall efficiency of a project than pure computation speed.'' (\cite{deschutterConsumerGuideNeuronal1992}).
+xolotl is designed primarily with ease-of-use in mind.
+This includes the time it takes to install, setup, and learn how to use the software, the time to write and debug scripts, and the time to perform the simulations (\cite{rudolph2007much}).
 An easy-to-use simulation environment must minimize time spent in all these domains, especially during human engagement with the software.
@@
 We have focussed on making our software as easy to use as possible, without sacrificing performance or extensibility.
-For example, the software and all dependencies can be installed using a single-line installer script from within the "MATLAB" command line.
+For example, the software and all dependencies can be installed using a single-line installer script from within the MATLAB command line.
 The installation includes worked example scripts that demonstrate various features of the simulator, that can be run without any configuration.
-We decided to built the the front-end interface to "xolotl" in "MATLAB" to facilitate interoperability with existing tools for time series analysis, optimization, and parallel computing.
-This allowed us to build rich tools for visualisation and to interact with the simulation.
+We decided to built the the front-end interface to xolotl in MATLAB to facilitate interoperability with existing tools for time series analysis, optimization, and parallel computing.
+This allowed us to build rich tools for visualization and to interact with the simulation.
 
@@
 Over the years, several simulators have been developed to integrate systems of coupled differential equations, those that model the spiking activity of neurons (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003, hinesNEURONSimulationEnvironment1997}).
-A critical design choice
+A critical architectural choice in designing a simulator is how much ``scaffolding'' a user is provided to construct a model; whether a model is specified by equations or by components, or by some combination of the two.
+In an equation-oriented architecture, the user starts with a blank slate and the primary method of specifying a model is to write out its differential equations.
+In contrast, in a component-oriented architecture, the primary method of specifying a model is to assemble it from pre-existing components, each of which include differential equations, parameters, and solvers.
+Both approaches have advantages and disadvantages that are illustrated below.
 
-In contrast to "BRIAN" and "DynaSim" and in concordance with "NEURON", "xolotl" represents models in a compartment-oriented manner.
-Compartment-oriented simulators consider membrane compartments first, representing the neuron model hierarchically.
-Compartments contain conductances, synaptic connections, and mechanisms (injected current, \etc).
-The dynamical variables hide within the hierarchical structure.
-Generally, the user specifies the features of and in the compartment to define the model; the dynamical variables are hidden within the structure.
+An equation-oriented simulator can be more transparent and allows the user to know exactly what is being solved, but equations can be cumbersome to write out, read, or to debug.
+In most commonly used programming environments, these equations have to be entered as strings, and complex parsing has to be carried out by the simulator to check that these strings constitute valid equations.
+In addition, parameters have to be written explicitly into equations, and it is usually not trivial to change parameters after initial specification.
+In contrast, components are easy to assemble into a model, but it can be hard to know what they contain, where they are physically located on the user's computer, and how they can be changed.
+NEURON is primarily a component-oriented neuronal simulator, and new components are specified in special model files that can be ``inserted'' into a model.
+BRIAN is an equation-oriented neuronal simulator meant to be used from within Python (\cite{goodmanBrianSimulator2009}).
+XPP is a general purpose dynamical system simulator that is equation-oriented (\cite{ermentrout2002simulating}).
+DynaSim is an equation-oriented simulator with some component-oriented capabilities.
+Models can be specified by both strings of equations or components, but since models do not exist as objects in the workspace, parameters and variables have to be reinitialized when changed (\cite{sherfeyDynaSimMATLABToolbox2018}).
 
-"xolotl" allows the user to explore the structure in the "MATLAB" console and see all properties and defining equations at each level of the model.
-Numerical properties can be changed in the console and are reflected in future simulations.
-The equations, structure, parameters, and methods by which the simulation proceeds are transparently available to the user.
+Our simulator is a purely component-based simulator, and all equations need to be included in a C++ header file that specifies a object.
+Since our automatic type system binds MATLAB objects to the underlying C++ header files, objects can be inspected and parameters can be modified in the MATLAB workspace.
+To mitigate some of the drawbacks of the component-oriented paradigm, we have implemented an architecture that allows the user to access the underlying C++ code of any object by simply clicking on the object tree in the MATLAB command line.
+This feature removes the uncertainty inherent in other component-oriented simulators of the equations underlying each component, and allows the user to modify these equations if needed.
 
-In equation-oriented specifications such as "BRIAN" and "DynaSim", all equations which define the model must be explicitly written.
-The equations maintain primacy in the model structure and are parsed and interpreted by the numerical integration algorithm.
-This has the added feature of total customization, as the equations can be edited in-place, but long strings of characters can become unwieldy and are prone to syntax mistakes.
-In addition, editing a feature of the model or simulation requires directly editing the mechanism or equation script and cannot be done in the console.
-
-- components from library vs. equations
-- equations are transparent -- you know what you're solving. but clunky -- string parsing headaches, etc.
-- components are easy to assemble, but it;s hard to know where they are, how they can be changed, etc.
-- xolotl goes the component route, but makes it very easy to find components (you can click your way into the C++ files).
-components are stored as source files, so you can modify it and immediately see it reflected in your model.
-we didn't invent another synatax for these components -- these are just C++ files.
-- scaffolding
-
-"xolotl" cannot match the total flexibility of equation-oriented specifications, which can integrate arbitrary dynamical systems, but avoids some of the most common problems associated with equation-oriented specification.
-While the equations "xolotl" solves cannot be changed without adding a mechanism or editing the specifying text file, numerical parameters and network components can be added at any time.
-The user can inspect the "xolotl" network object in the console and represent its structure.
-Changes can be made on-the-fly.
-In addition, "xolotl" components come specified by a "C++" header file, greatly reducing typing errors during model specification.
+Another architectural choice in simulators is the syntax required for specifying models.
+Equation-oriented simulators specify models as strings of equations, so must invent a new syntax to specify derivatives, variables, and other common elements in coupled differential equations.
+Some component-oriented simulators like NEURON also specify their own syntax, or have invented their own language to specify components and write out equations.
+While this allows for powerful features such as support for units in NEURON, the user is required to learn a novel syntax and vocabulary, hindering ease of use.
+In general, the syntax for model specification in different simulator can be different.
+For example, DynaSim, BRIAN, XPP, and NEURON all use different, incompatible formalisms to represent equations, increasing the cognitive load of users using more than one simulator.
+In our software, we have elected not to specify a domain-specific ``middleware'' layer, and instead specify and implement models and equations in idiomatic C++.
+This greatly decreases the learning curve and allows users with a general familiarity with programming languages to quickly acquaint themselves even with the most technical parts of our simulator.
 
@@
 The use of computational tools is increasingly central to the scientific method; yet, the lack of auditability and accountability in their use has led to a crisis of credibility affecting many scientific fields (\cite{stoddenEnhancingReproducibilityComputational2016, bakerWhyScientistsMust2016}).
-Unlike in experimental research, where a lack of reproducibility can arise due to deep reasons like unknown differences in experimental protocols or intrinsic variability, the reasons for irreproducibility in computational research are largely trivial and include: a) typographical errors from transcribing out model parameters and equations, b) obscure software design that leads to users not knowing precisely what equations the software is solving, or the parameters it using, c) incompatibility with version control systems and rolling software development which leads to ambiguity in which version of the software was used to generate a particular result, and d) convoluted architectures that make it too complex for non-experts to understand the inner workings of the software (\cite{xuMeasurementSourceCode2017, sedanoCodeReadabilityTesting2016, vikstromStudyAutomaticTranslation2009}).
+Unlike in experimental research, where a lack of reproducibility can manifest due to meaningful reasons like uncharted differences in experimental protocols or intrinsic variability, the reasons for irreproducibility in computational research are often trivial and include: a) typographical errors from transcribing model parameters and equations, b) obscure software design that leads to users not knowing precisely what equations the software is solving, or what parameters it is using, c) incompatibility with version control systems and rolling software development that leads to ambiguity in which the version of the software that was used to generate a particular result is unclear, and d) convoluted architectures that make it too complex for non-experts to understand the inner workings of the software (\cite{xuMeasurementSourceCode2017, sedanoCodeReadabilityTesting2016, vikstromStudyAutomaticTranslation2009}).
 
-Our software was designed to tackle all these threats, and has features that allow the user to answer the following questions in the affirmative: ``Am I doing what I think I am doing?'' and ``Is it possible for others to reproduce exactly what I have just done?''.
+Our software was designed with this threat model in mind, and has features that allow the user to answer the following questions in the affirmative: ``Can I be sure that I am doing what I think I am doing?'' and ``Is it possible for others to reproduce {\em exactly} what I have just done?''.
 Our goal was to design software that would allow the user to verify for herself that the software was running as she intended, and to be able to reproduce results from others quickly and unambiguously.
-The primary design choice of our software that enables auditability and reproducibility is that every simulation is tied to an alphanumeric checksum, or hash, using the "MD5" algorithm (\cite{rivestMD5MessageDigestAlgorithm1992}).
-The hash is computed from every "C++" file that is part of the model, and any changes in the "C++" file will trigger a recompilation of that model.
+The primary design choice in our software that enables auditability and reproducibility is that every simulation is tied to an alphanumeric checksum, or hash, using the "MD5" algorithm (\cite{rivestMD5MessageDigestAlgorithm1992}).
+The hash is computed from every C++ file that is part of the model, and any changes in any C++ file included in the model will trigger a recompilation of that model.
 Thus, the hash guarantees that a given model is derived from a set of source files, obviating any ambiguity about the code used to generate a model.
@@
 Finally, the core of our software is written in a few hundred lines of code and contains just four classes: compartments, conductances, synapses, and mechanisms.
-This architectural simplicity permits a motivated user to understand the entire organization of our code in short period of time.
+This architectural simplicity lets a motivated user understand the entirety of our code quickly.
+
+\subsection{Outlook and future directions}
+
+In its current form, xolotl is an efficient and easy to use neuron and network simulator that is actively being used in research. 
+Results from an early version of this simulator guided intuition in the modeling of a recently characterized Calcium-dependent Potassium channel found in {\em Drosophila} neuromuscular presynaptic terminals (\cite{bronk2018regulation}).
+Work on the simulator continues in the open at a publicly accessible repository (https://github.com/sg-s/xolotl/), and the library of conductances, synapses and mechanisms that xolotl ships with grows continuously.
+
+Since xolotl models are bonafide MATLAB objects, they are compatible with most of the powerful tools that exist within MATLAB.
+For example, it is possible to write simple scripts that run xolotl models in parallel using MATLAB's parallel processing toolbox, speeding up large simulations.
+xolotl models are also compatible with the global optimization toolbox in MATLAB, allowing parameters in xolotl models to be optimized, enabling the creation of toolboxes that efficiently tune parameters in neuron models to satisfy arbitrary constraints (\cite{Achard:2006ia, Krichmar:2014jj, Keren:2005ja, VanGeit:2007ea, Druckmann:2008kg}).
+
+Care has been taken to reduce the amount of technical debt (\cite{suryanarayana2014refactoring}) associated with this project, with all parts of the simulator and dependencies written in a modular, objected oriented fashion.
+As a result, many of the key features and architectures of xolotl can be reused by others in their own applications.
+For example, cpplab, the automatic type system that binds C++ code to MATLAB objects, is independent of this simulator, and exists as a distinct repository that has been made freely available (https://github.com/sg-s/cpplab); and the framework for generating a GUI with sliders for each parameter that is hooked up to function callbacks also exists as an independent repository (https://github.com/sg-s/puppeteer) that can be easily integrated into other applications.
 
@@
 
-SG-S \& AH designed and wrote the software. AH created the online user documentation. EM supervised the project. All authors wrote and reviewed the paper.
+SG-S \& AH wrote the software and created the online user documentation. EM supervised the project. All authors wrote and reviewed the paper.
 
@@
 \section*{Acknowledgments}
-The authors would like to thank Timothy O'Leary and Leandro Alonso for useful discussions. Janis Li helped to prepare many conductance header files.
-
-\section*{Supplemental Data}
-Tables including all conductances packaged with "xolotl" should be put in the supplementary material.
+The authors would like to thank Timothy O'Leary and Leandro Alonso for useful discussions, and Janis Li for help transcribing many conductance models.
 
@@
 
-The code to generate all figures is available at (\url{https://github.com/marderlab/xolotl-paper}). "xolotl" is freely available at (\url{https://github.com/marderlab/xolotl}).
+The code to generate all figures is available at (\url{https://github.com/marderlab/xolotl-paper}). xolotl is freely available at (\url{https://github.com/sg-s/xolotl}).
 
@@
 	\includegraphics[width=1.0\linewidth]{gfx/figure_HH}
-	\caption{Setting up a single-compartment model and injecting current into it.
+	\caption{Simulating a single-compartment Hodgkin-Huxley spiking neuron model.
 (A) Schematic representation of a single-compartment neuron with three populations of ion channels (colored rectangles).
-(B) In the simulator, the soma is represented using an object of class $compartment$ and the populations of ion channels are represented by $conductance$ objects contained within the compartment object.
+(B) In our simulator, the soma is represented using an object of class ``compartment'' and populations of ion channels are represented by ``conductance'' objects contained within the compartment object.
 (C) The code snippet shown sets up this neuron model, injects current, integrates and plots the voltage, and displays activation functions, all in a few lines of code.
@@
 	\caption{Simulating a voltage-clamp experiment.
-(A) The diagram shows a cell with delayed rectifier potassium conductance (from \cite{liuModelNeuronActivityDependent1998}) that is being recorded from in two-electrode voltage clamp.
-(B) The code snippet shown here sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it.
-(C) Voltage steps that the cell is clamped to.
-(D) Current injected into the cell required to clamp it to the voltages in (C).
-(E) Asymptotic clamped current $vs.$ clamped voltage for this cell.
-(F) Accounting for the reversal potential of Potassium ions yields the conductance-voltage curve of this channel type.
-(G) Normalized conductance-voltage curves, with sigmoid fits with various exponents.
-(H) An exponent of $n=4$ yields the best fit, allowing for the characterization of the activation function of this channel type.
+The diagram shows a cell with delayed rectifier potassium conductance (from \cite{liuModelNeuronActivityDependent1998}) that is being recorded from in two-electrode voltage clamp (A).
+The code snippet shown here sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it (B).
+Voltage steps that the cell is clamped to (C).
+Clamping currents as a function of time (D).
+Asymptotic clamped current {\em vs.} clamped voltage for this cell (E).
+Accounting for the reversal potential of Potassium ions yields the conductance-voltage curve of this channel type (F).
+Normalized conductance-voltage curves, with sigmoid fits with various exponents (G).
+An exponent of $n=4$ yields the best fit, allowing for the characterization of the activation function of this channel type (H).
 }
@@
 	\caption{Modeling intracellular mechanisms.
-(A) A single-compartment neuron model with 8 channel types.
+A single-compartment neuron model with 8 channel types (A).
 Since there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant (B), and the cell tonically spikes (C).
@@
 The cell now bursts periodically, with synchronized oscillations in intracellular Calcium (F-G).
-(H) Schematic of Calcium-dependent integral control homeostasis (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}).
-In this feedback system, the rates of mRNA synthesis depend on the Calcium level in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which, through translation, is a function of the mRNA abundance.
+Schematic of Calcium-dependent integral control homeostasis (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (H).
+In this feedback system, the rates of mRNA synthesis depend on the Calcium level in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which, through translation, depends on the mRNA abundance.
 (I) The code snippet shows how these integral controllers are implemented as \texttt{mechanism} objects, and can be added to conductances.
@@
 	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them $ad$   $arbitrium$.
-The initial state (orange disc) of the single compartment model in the previous example is saved using the \texttt{snapshot} method.
+The initial state (orange node) of the single compartment model in the previous example is saved using the \texttt{snapshot} method.
 This method saves all parameters and dynamic variables of the model in a named state.
-The first column shows the profile of conductances and the voltage dynamics of the model at this point in time.
-The maximal conductances of the Calcium channels are set to zero (purple), changing the voltage dynamics of the neuron (purple plots).
-After evolving the model for some time (green), the conductance profile and voltage dynamics returns to a state similar to the initial state (green plots).
+The first column (orange plots) shows the profile of conductances and the voltage dynamics of the model at this point at this time.
+The maximal conductances of the Calcium channels are then set to zero (purple node), changing the voltage dynamics of the neuron (purple plots).
+After evolving the model for some time (green node), the conductance profile and voltage dynamics returns to a state similar to the initial state (green plots).
 This configuration is now saved in a state called $final$ and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
-Another parameter is changed (the Calcium target), and the model is integrated to reach a new state (blue disc) where the voltage dynamics are different from the $initial$ state.
+Another parameter is now changed (the Calcium target), and the model is integrated to reach a new state (blue node) where the voltage dynamics are different from the $initial$ state.
 In summary, any state can be bookmarked using a descriptive name using the \texttt{snapshot} method, and can be returned to using the \texttt{reset} method.}
@@
 	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
-	\caption{Simulating a network of conductance-based model neurons.
-(A) A three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (\cite{prinzSimilarNetworkActivity2004}).
+	\caption{Simulating a network of conductance-based model neurons, coupled by voltage-dependent dynamic chemical synapses.
+A three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (\cite{prinzSimilarNetworkActivity2004}) (A).
 Each neuron is modeled with a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
 Synapses can be one of two types (Cholinergic (dashed lines) or Glutamatergic (solid lines)) and have different kinetics.
-(B) The code snippet shows how neurons can be created, wired together using synapses, and how the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse.
-(C-E) Simulated voltage trace of a model network for the three compartments obtained from this simulation.
-(F) Time series activation variable of the Glutamatergic synapse between PY and LP (red connection in diagram) shows how the synapse becomes active every time the PY neuron spikes.}
+The code snippet shows how neurons can be created, wired together using synapses, and how the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse (B).
+Simulated voltage trace of a model network for the three compartments obtained from this simulation (C-E) .
+Time series activation variable of the Glutamatergic synapse between PY and LP (red connection in diagram) shows how the synapse becomes active every time the PY neuron spikes (F).}
 	\label{fig:figurenetwork}
@@
 	\caption{Manipulating neuron parameters in real time.
-Any set of parameter in the model can be manipulated; here, the maximal conductance of every conductance type in the model from Figure 1 is being manipulated using the code snippet shown here.
-The screenshot shows a GUI with sliders for every parameter of interest that is created by the \texttt{manipulate} method, and linked to an arbitrary number of visualization functions.
+Any set of parameter in the model can be manipulated; here, the maximal conductance of every conductance type in the model from Fig.
+\ref{fig:figurehh} is being manipulated using the code snippet shown here.
+The screenshot shows a GUI with sliders for every parameter of interest that is created by the \texttt{manipulate} method.
+These sliders can be linked to an arbitrary number of visualization functions.
 In this example, two visualization functions are used: the built in \texttt{plot} method (A) and a custom function that computes the firing-rate-$vs$.-injected current curve for this neuron (B).
 Both plots refresh themselves with every movement of any slider, allowing the user to build intuition about how every parameter controls the dynamical behavior of the model.
-A screen recording of this model being manipulated in real time is included in Supplementary Material XX.
-}
+A screen recording of this model being manipulated in real time is included in Supplementary Material}
 	\label{fig:figuremanipulate}
@@
 	\includegraphics[width=1.0\linewidth]{gfx/figure_benchmark}
-	\caption{\texttt{xolotl} benchmarked against \texttt{DynaSim} and \texttt{NEURON}.
-The top row benchmarks using a tonically-firing Hodgkin-Huxley model with 3 conductances with constant injected current.
-The bottom row repeats the same benchmarks using a bursting stomatogastric ganglion model with 8 conductances.
-(A,E) Ratio of run-time to simulation time (speed factor) over increasing integration time step.
-(B,F) Error over increasing integration time step as computed by the LeMasson matrix where the baseline condition is the voltage trace at $\Delta t = 0.001$ ms.
-(C,G) Simulations over increasing simulation time where $\Delta t = 0.1$ ms.
-(D,H) Simulations over increasing number of unconnected compartments.}
+	\caption{Comparison of speed and accuracy of xolotl, NEURON and DynaSim.
+The top row shows simulations of a tonically-firing Hodgkin-Huxley model with three conductances with constant injected current (as in Fig.
+\ref{fig:figurehh}).
+The bottom row shows simulations of  a bursting stomatogastric ganglion neuron model with 8 conductances (as in Fig.
+\ref{fig:figurenetwork}).
+Ratio of run-time to simulation time (relative speed) as a function of simulation time step (A, B).
+Simulation error as a function of the step size (C, D).
+Relative speed of integration as a function of the simulation length (E, F).
+Relative speed of integration, normalised by system size, as a function of the number of compartments simulated simultaneously (G, H).
+All benchmarks were performed on the same computer, and all simulators using fixed time-step integration methods.
+NEURON was run using the Python wrapper.
+}
 	\label{fig:figurebenchmark}

### 2018-08-14 figures/HH/figure_HH.m

@@ -2,6 +2,7 @@
 
 % set up xolotl object
 % with Hodgkin-Huxley type dynamics
+clearvars
 x = xolotl;
 
 
@@ -152,7 +153,7 @@ end
 %% Post-Processing
 
 % beautify
-prettyFig('fs', 16, 'plw', 2,'lw',1.5)
+prettyFig('fs', 16, 'plw', 2,'lw',1,'tick_length',.02)
 
 % remove boxes around subplots
 for ii = 3:length(ax)

### 2018-08-14 figures/benchmarks/figure_benchmark.m

@@ -2,7 +2,9 @@
 % and other sim. software using a HH model and a bursting
 % STG model
 
-figure('outerposition',[100 100 1550 700],'PaperUnits','points','PaperSize',[1550 700]); hold on
+
+clearvars
+figure('outerposition',[100 100 1600 850],'PaperUnits','points','PaperSize',[1600 850]); hold on
 for i = 8:-1:1
 	ax(i) = subplot(2,4,i); hold on
     set(ax(i),'XScale','log','YScale','log')
@@ -64,13 +66,15 @@ disp('Begin NEURON STG')
 testNeuronSTG(ax(5:end)); drawnow
 
 
-prettyFig('lw',.5,'plw',1.5,'tick_length',.03);
+prettyFig('lw',.5,'plw',1.5,'tick_length',.03,'fs',24);
 
 
-movePlot(ax([1 5]),'left',.05)
-movePlot(ax([2 6]),'left',.04)
-movePlot(ax([3 7]),'left',.03)
-movePlot(ax([4 8]),'left',.02)
+movePlot(ax([1 5]),'left',.075)
+movePlot(ax([2 6]),'left',.05)
+movePlot(ax([3 7]),'left',.035)
+movePlot(ax([4 8]),'left',.01)
+
+movePlot(ax(1:4),'up',.02)
 
 for i = 1:8
 	axis(ax(i),'square')
@@ -85,5 +89,5 @@ l(2) = plot(ax(4),NaN,NaN,'r.','MarkerSize',34);
 
 lh = legend(l,{'xolotl','DynaSim','NEURON'});
 lh.Position = [0.9071    0.7416    0.0690    0.1037];
-
-labelFigure('capitalise',true,'column_first',true,'x_offset',-.02,'y_offset',-.02,'font_size',28)
+lh.Box = 'off';
+labelFigure('capitalise',true,'column_first',true,'x_offset',-.02,'y_offset',-.02,'font_size',28,'font_weight','bold')

### 2018-08-14 figures/benchmarks/figure_sup_stability.m

@@ -8,7 +8,7 @@ rng(prng);
 
 % use zoidberg to view the Prinz database
 z = zoidberg;
-z.path_to_neuron_model_db = '~/code/prinz-database/neuron-db/neuron properties';
+
 G = z.findNeurons('burster');
 % generate 50 models from the database
 conds = {'NaV', 'CaT', 'CaS', 'ACurrent', 'KCa', 'Kd', 'HCurrent', 'Leak'};

### 2018-08-14 figures/benchmarks/testXolotlSTG.m

@@ -62,8 +62,8 @@ if isempty(cache(h))
 		x.sim_dt = all_dt(i);
 		x.dt = 1;
 
-    % trial run
-    V = x.integrate;
+    	% trial run
+    	V = x.integrate;
 
 		tic
 		all_V(:,i) = x.integrate;
@@ -74,10 +74,6 @@ else
 	[all_V,all_sim_time] = cache(h);
 end
 
-for i = length(all_dt):-1:1
-	all_f(i) = xolotl.findNSpikes(all_V(:,i),-20);
-	all_f(i) = all_f(i)/(x.t_end*1e-3);
-end
 
 % now measure the errors using the LeMasson matrix
 [M0, V_lim, dV_lim] = xolotl.V2matrix(all_V(:,1));
@@ -89,7 +85,6 @@ end
 
 % delete the last one because the first sim is slow for
 % trivial reasons involving matlab compiling
-all_f(end) = [];
 matrix_error(end) = [];
 all_sim_time(end) = [];
 all_dt(end) = [];

### 2018-08-14 figures/snapshots/figure_snapshot.m

@@ -1,7 +1,7 @@
 % tests a neuron that reproduces Fig 3 in Tim's paper
 
 
-
+clearvars
 A = 0.0628; % mm^2
 Ca_target = 7; % used only when we add in homeostatic control 
 
@@ -100,7 +100,7 @@ ylabel(ax.g(1),'g (\muS/mm^2)','interpreter','tex')
 ylabel(ax.V(1),'V_m (mV)','interpreter','tex')
 xlabel(ax.V(1),'Time (ms)','interpreter','tex')
 
-prettyFig('fs',12,'plw',1.5,'lw',1.5);
+prettyFig('fs',18,'plw',1.5,'lw',1);
 
 
 ax.cartoon.Units = 'pixels';
@@ -112,6 +112,14 @@ for i = 1:length(ax.g)
 	ax.g(i).Position(4) = .14;
 	ax.g(i).Position(2) = ax.g(i).Position(2) + .02;
 end
+
+
+movePlot([ax.g(1) ax.V(1)],'left',.03)
+movePlot([ax.g(2) ax.V(2)],'left',.01)
+movePlot([ax.g(3) ax.V(3)],'right',.01)
+movePlot([ax.g(4) ax.V(4)],'right',.03)
+
+movePlot(ax.g,'up',.05)
 
 for i = 1:length(ax.g)
 	deintersectAxes(ax.V(i));

### 2018-08-14 figures/benchmarks/figure_sup_stability.m
working on supp figure comparison to ode23t

@@ -2,6 +2,18 @@
 % creates a supplementary figure that shows the stability of
 % a model with varied maximal conductances over increasing time-step
 
+clearvars
+
+t_end = 30e3; % ms
+t_transient = 10e3; % ms
+dt = 1; % ms
+
+% make a vector of dt to vary
+max_dt = 1e3;
+K = 1:max_dt;
+all_dt = K(rem(max_dt,K) == 0);
+all_dt = all_dt/1e3;
+
 % fix pseudorandom number generation
 prng = 353;
 rng(prng);
@@ -25,70 +37,73 @@ x.AB.add('prinz/KCa', 'gbar', G(1,5), 'E', -80);
 x.AB.add('prinz/Kd', 'gbar', G(1,6), 'E', -80);
 x.AB.add('prinz/HCurrent', 'gbar', G(1,7), 'E', -20);
 x.AB.add('Leak', 'gbar', G(1,8), 'E', -50);
-x.t_end = 20e3;
+x.t_end = t_end;
 x.sim_dt = 0.001;
-x.dt = 1;
+x.dt = dt;
 
 % check to make sure that they are actually bursting
 h = GetMD5([GetMD5(prng) GetMD5(G) x.hash]);
 disp('checking models for bursting...')
 
 if isempty(cache(h))
-  disp('running bursting tests...')
-  passingModels = [];
-  % set up the conductances
-  while length(passingModels) <= 100
-    model = randi(length(G),1);
-    params = G(:, model);
-    for qq = 1:length(conds)
-      x.AB.(conds{qq}).gbar = params(qq);
-    end
-    % simulate at low time resolution
-    x.sim_dt = 0.1;
-    % simulate each model
-    [V, Ca] = x.integrate;
-    V = V(10e3/x.dt:end);
-    Ca = Ca(10e3/x.dt:end,1);
-    burst_metrics = psychopomp.findBurstMetrics(V, Ca);
-    burst_freq = 1 / (burst_metrics(1) * 1e-3);
-    % confirm that burst frequency is in [0.5, 2.0]
-    if burst_freq >= 0.5 & burst_freq <= 2.0 & burst_metrics(10) == 0 & burst_metrics(9) >= 0.2 & burst_metrics(2) >= 3 & burst_metrics(2) <= 10;
-      % simulate at high time resolution
-      disp('simulating at high time-resolution...')
-      x.sim_dt = 0.001;
-      % simulate each model
-      [V, Ca] = x.integrate;
-      V = V(10e3/x.dt:end);
[…]

### 2018-08-14 figures/benchmarks/neuron_standalone.m

@@ -21,7 +21,6 @@ function [xprime]=  neuron_standalone(t, x, g)
 % g = [100 2.5 6 50 5  100 0.01 0.00]; % ABPD#2
 % %g = [100 0 4  20 0 25  0.05 0.03]; % LP#4
 
-input=0;
 
 C = 0.628*10^(-9); % in Farads
 A = 0.628*10^(-7); % in sq. m. (area of cell)
@@ -32,13 +31,18 @@ Cazero = 0.05*10^(-3); %milliMol
 tauCa = 0.200; % in seconds.
 % now the reversal potential for Ca
 ECa = 12.9*(log((extCa)/(x(13)))); % in milliVolts, corrected for divalent Calcium
-% begin reversal potentials.
+
 E = [50 ECa ECa -80 -80 -80 -20 -50]; % in milliVolts. ECa needs to be computed from the Nernst Equation
 
 %% correct units..
- g = g*10; % now all in Siemens/sq.m.
+g = g*10; % now all in Siemens/sq.m.
 %% begin voltage dependency...
 %minf
+minf = zeros(7,1);
+hinf = zeros(4,1);
+taum = zeros(7,1);
+tauh = zeros(4,1);
+
 minf(1) = 1/(1+exp((x(12) + 25.5)/(-5.29)));
 minf(2) = 1/(1+exp((x(12) + 27.1)/(-7.20)));
 minf(3) = 1/(1+exp((x(12) + 33.0)/(-8.10)));
@@ -59,30 +63,24 @@ taum(4) = 23.20 - (20.8/(1 + exp((x(12) + 32.9)/(-15.2))));
 taum(5) = 180.6 - (150.2/(1 + exp((x(12) + 46)/(-22.7))));
 taum(6) = 14.40 - (12.8/(1 + exp((x(12) + 28.3)/(-19.2))));
 taum(7) = 0.000 + (2/(exp((x(12) + 169.7)/(-11.6)) + exp((x(12) - 26.7)/(14.3))));
-taum = taum/1000; % all taums now in seconds
+
 % tauh
 tauh(1) = (1.34/(1+exp((x(12) + 62.9)/(-10))))*((1.5)+(1/(1+(exp((x(12)+34.9)/(3.6)))))); % hopefully, this is correct
 tauh(2) = 210 - (179.6/(1 + exp((x(12) + 55)/(-16.9))));
 tauh(3) = 120 + (300/(exp((x(12) + 55)/(9)) + exp((x(12) + 65)/(-16))));
[…]

### 2018-08-15 paper/paper.tex
working on supp figure comparison to ode23t
minor changes to paper; still working on benchmarks supp

@@
 Here, we present a neuron and network simulator built on a novel automatic type system that binds object-oriented code written in C++ to objects in MATLAB.
-Our approach builds on the tradition of uniting the speed and power of languages like C++ with the ease-of-use and feature-set of scientific programming languages like MATLAB.
+Our approach builds on the tradition of uniting the speed of languages like C++ with the ease-of-use and feature-set of scientific programming languages like MATLAB.
 Our software allows for the creation and manipulation of hierarchical models with components that are named and searchable, permitting intuitive high-level programmatic control over all parameters.
-Our simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately.
+Our simulator's architecture allows for the interactive manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately.
 Our simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be extended to include arbitrary conductances, synapses, and mechanisms.
-Several core features like bookmarking of parameters and automatic hashing of source code promote reproducible and auditable research.
-Finally, our simulator is written in a modular fashion and has been released under a permissive free software license, enabling seamless integration into the workflow of other researchers.
+Several core features like bookmarking of parameters and automatic hashing of source code facilitate reproducible and auditable research.
+Finally, our simulator is written in a modular fashion, includes detailed tutorials and worked examples, and is freely available at https://github.com/sg-s/xolotl, enabling seamless integration into the workflow of other researchers.
 
@@
 Simulators designed in this way tend to perform fast computations with little overhead, but suffer from a steep learning curve.
-Wrapping these simulators in a more approachable language like "Python"  or using graphical user interfaces (GUIs) mitigates these drawbacks (\cite{hinesNEURONPython2009}) at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
+Wrapping these simulators in a more approachable language like Python or using graphical user interfaces (GUIs) mitigates these drawbacks (\cite{hinesNEURONPython2009, gratiy2018bionet}) at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
 In contrast, simulators designed from the ground up in popular scientific computing languages can be easier to use and benefit from interoperability with other commonly-used tools.
-Simulators like DynaSim (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}), "morphforge" (\cite{hullMorphforgeToolboxSimulating2014}), and "PyNN" (\cite{davisonPyNNCommonInterface2009}) allow the user to specify models with strings of equations or components that are constructed using a special syntax, that can then be translated into a faster implementation language such as "C" or C++ (\cite{stimbergEquationorientedSpecificationNeural2014}).
+Simulators like DynaSim (\cite{sherfeyDynaSimMATLABToolbox2018}), ANNarchy (\cite{vitayANNarchyCodeGeneration2015}), BRIAN (\cite{stimbergBrianSecondComing2013}), morphforge (\cite{hullMorphforgeToolboxSimulating2014}), and PyNN (\cite{davisonPyNNCommonInterface2009}) allow the user to specify models with strings of equations or components that are constructed using a special syntax, that can then be translated into a faster implementation language such as C or C++ (\cite{stimbergEquationorientedSpecificationNeural2014}).
 This approach permits considerable flexibility for simulating systems of differential equations.
@@
 
-To overcome these design limitations, we have developed a novel automatic type system, that we call "cpplab", which binds MATLAB code to classes specified in C++ header files.
+To overcome these design limitations, we have developed a novel automatic type system, that we call cpplab, which binds MATLAB code to classes specified in C++ header files.
 This architecture automatically creates objects in MATLAB that represent the underlying object-oriented C++ code, allowing the symbolic manipulation of C++ objects in the MATLAB interface.
-In this paper, we introduce xolotl, an implementation of the "cpplab" system specialized in integrating conductance-based neuron and network models.
+In this paper, we introduce xolotl, an implementation of the cpplab system specialized in integrating conductance-based neuron and network models.
 Models can be easily constructed from network building blocks in a few lines of MATLAB code using a hierarchical and intuitive syntax.
@@
 
-\paragraph{A rich library of network components.} xolotl comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and mechanisms) that can be used as building blocks to construct model neurons and networks.
-Compartments represent sections of membrane with a single membrane potential, Calcium level, and set of constituent components, representing either entire cells or parts of cells.
-Objects of type "conductance" represent populations of ion channels in a compartment that produce transmembrane currents.
-"synapse" objects connect two compartments together by introducing a current in the postsynaptic compartment that depends on the presynaptic compartment's membrane potential.
-Objects of type "mechanism" can represent any intracellular mechanism and can read and modify any other component in the cell, and can run arbitrary dynamical systems within them.
-
 \paragraph{Object-oriented.}
@@
 This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain neurons that contain populations of ion channels, and makes constructing, integrating, and thinking about these models easier.
+
+\paragraph{A rich library of network components.} xolotl comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and mechanisms) that can be used as building blocks to construct model neurons and networks.
+Compartments represent sections of membrane with a single membrane potential, Calcium level, and set of constituent components, representing either entire cells or parts of cells.
+Objects of type "conductance" represent populations of ion channels in a compartment that produce transmembrane currents.
+"synapse" objects connect two compartments together by introducing a current in the postsynaptic compartment that depends on the presynaptic compartment's membrane potential.
+Objects of type "mechanism" can represent any intracellular mechanism and can read and modify any other component in the cell, and can run arbitrary dynamical systems within them.
 
@@
 
-The axon of the giant squid contains contains a fast inactivating sodium conductance ("NaV"), a slower non-inactivating potassium conductance ("Kd"), and a passive leak current ("Leak").
+The axon of the giant squid contains a fast inactivating sodium conductance ("NaV"), a slower non-inactivating potassium conductance ("Kd"), and a passive leak current ("Leak").
 Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}).
@@
 These models of ion channels were obtained from \cite{liuModelNeuronActivityDependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator.
-The code to set up this model is terse, idiomatic, relies on no special markup, and preserves the heirarchical nature of the model (Figure \ref{fig:figurehh}C).
+The code to set up this model is terse, idiomatic, relies on no special markup, and preserves the hierarchical nature of the model (Figure \ref{fig:figurehh}C).
 
@@
 In electrical synapses, no chemical intermediary is involved.
-Ions flux directly from the presynaptic neuron to the postsynaptic one, setting up a current in the post-synaptic cell.
 New patterns of activity can emerge from the characteristics of the connecting synapses (\cite{liComodulationSynapsesVoltagegated2018, nadimSynapticDepressionCreates1999, gutierrezRectifyingElectricalSynapses2013, gutierrezMultipleMechanismsSwitch2013}).
@@
 
-Our software is dfesigned to streamline this process and allows for any parameter in any model to be manipulated using graphical sliders, with immediate, real-time feedback of its behavior.
+Our software is designed to streamline this process and allows for any parameter in any model to be manipulated using graphical sliders, with immediate, real-time feedback of its behavior.
 Any model in our simulator can be manipulated using
@@
 We set out to design a neuron and network simulator that could be useful in the classroom setting, especially for students of computational science, while also being powerful, fast, and extensible enough to be used for research.
-By using a novel architecture that permits the symbolic manipulation of C++ objects in a intuitive MATLAB interface, we demonstrated some of the features of our simulator using simulations of single-compartment models of Hogkin-Huxley like neurons (Fig.
+By using a novel architecture that permits the symbolic manipulation of C++ objects in a intuitive MATLAB interface, we demonstrated some of the features of our simulator using simulations of single-compartment models of Hodgkin-Huxley like neurons (Fig.
 \ref{fig:figurehh}); voltage clamp experiments to recover activation functions of single channel types (Fig.

### 2018-08-15 figures/benchmarks/figure_sup_stability.m

@@ -2,24 +2,28 @@
 % creates a supplementary figure that shows the stability of
 % a model with varied maximal conductances over increasing time-step
 
-clearvars
+clearvars -except z
+
+redo = false;
 
 t_end = 30e3; % ms
 t_transient = 10e3; % ms
-dt = 1; % ms
+dt = .1; % ms
 
 % make a vector of dt to vary
-max_dt = 1e3;
+max_dt = 1e2;
 K = 1:max_dt;
 all_dt = K(rem(max_dt,K) == 0);
 all_dt = all_dt/1e3;
 
 % fix pseudorandom number generation
-prng = 353;
+prng = 1984;
 rng(prng);
 
-% use zoidberg to view the Prinz database
-z = zoidberg;
+if ~exist('z','var')
+	% use zoidberg to view the Prinz database
+	z = zoidberg;
+end
 
 G = z.findNeurons('burster');
 % generate 50 models from the database
@@ -29,23 +33,23 @@ conds = {'NaV', 'CaT', 'CaS', 'ACurrent', 'KCa', 'Kd', 'HCurrent', 'Leak'};
 x = xolotl;
 x.add('compartment', 'AB', 'A', 0.0628,'vol', 0.0628);
 x.AB.add('CalciumMech1');
-x.AB.add('prinz/NaV', 'gbar', G(1,1), 'E', 50);
-x.AB.add('prinz/CaT', 'gbar', G(1,2), 'E', 30);
-x.AB.add('prinz/CaS', 'gbar', G(1,3), 'E', 30);
-x.AB.add('prinz/ACurrent', 'gbar', G(1,4), 'E', -80);
-x.AB.add('prinz/KCa', 'gbar', G(1,5), 'E', -80);
-x.AB.add('prinz/Kd', 'gbar', G(1,6), 'E', -80);
-x.AB.add('prinz/HCurrent', 'gbar', G(1,7), 'E', -20);
-x.AB.add('Leak', 'gbar', G(1,8), 'E', -50);
+x.AB.add('prinz/NaV', 'gbar', 1, 'E', 50);
+x.AB.add('prinz/CaT', 'gbar', 1, 'E', 30);
+x.AB.add('prinz/CaS', 'gbar', 1, 'E', 30);
+x.AB.add('prinz/ACurrent', 'gbar', 1, 'E', -80);
+x.AB.add('prinz/KCa', 'gbar', 1, 'E', -80);
[…]

### 2018-08-15 figures/benchmarks/neuron_standalone.m

@@ -1,103 +1,102 @@
-function [xprime]=  neuron_standalone(t, x, g)
+function xprime = neuron_standalone(t, x, g)
 % Neuron_standalone.m
 % you need to call it using an ode solver, e.g.:
 %[t n] = ode23t(@neuron_standalone, [0, tend], n0);
 % last edit @ 19:00 on Saturday the 27th
 % this is supposed to be called by STGnetwork.m
 % this function works for any neuron : g specifies which one it is
-% 06.10.2009
-% 07.10.2009
-% Dimensional Analysis and Unit checking
+
 % making sense of this:
 % x(1)...x(7) : m(1)...m(7)
 % x(7)...x(11): h(1)...h(4)
-% I(i)s are not treated as a variable
+
 % x(12) : The Voltage V
 % x(13) : The intracellular Calcium concentration [Ca]_i
-% future changes :
-% - fast switching of parameters, neuron-type by function calls.
-%%  begin parameters
+
 
 % g = [100 2.5 6 50 5  100 0.01 0.00]; % ABPD#2
 % %g = [100 0 4  20 0 25  0.05 0.03]; % LP#4
 
+% unpack some stuff
+V = x(12);
+Ca = x(13);
 
+% parameters
 C = 0.628*10^(-9); % in Farads
 A = 0.628*10^(-7); % in sq. m. (area of cell)
 f = 14.96*10^(6); % in milliMol/Ampere;
-p = [3 3 3 3 4 4 1]; % dimensionless
+p = [3 3 3 3 4 4 1]; % dimensionless, exponents
 extCa = 3; % in milliMol
 Cazero = 0.05*10^(-3); %milliMol
 tauCa = 0.200; % in seconds.
 % now the reversal potential for Ca
-ECa = 12.9*(log((extCa)/(x(13)))); % in milliVolts, corrected for divalent Calcium
+ECa = 12.2423*(log((extCa)/(Ca))); % in milliVolts, corrected for divalent Calcium
 
[…]

### 2018-08-15 paper/paper.tex
switching to new V_clamp snippet

@@
 Without any explicit mechanism for Calcium influx or buffering, the intracellular Calcium levels in this model do not change (Figure \ref{fig:figuremechanism}B) and the model tonically spikes (Figure \ref{fig:figuremechanism}C).
-Calcium buffering and influx can be modelled by a differential equation that increases intracellular Calcium with the current through Calcium channels and relaxes back to a baseline value (\cite{liuModelNeuronActivityDependent1998, prinzAlternativeHandtuningConductancebased2003, dayanTheoreticalNeuroscience2001}) (Figure \ref{fig:figuremechanism}D).
+Calcium buffering and influx can be modeled by a differential equation that increases intracellular Calcium with the current through Calcium channels and relaxes back to a baseline value (\cite{liuModelNeuronActivityDependent1998, prinzAlternativeHandtuningConductancebased2003, dayanTheoreticalNeuroscience2001}) (Figure \ref{fig:figuremechanism}D).
 This mechanism exists in the xolotl library as "CalciumMech1" and can be added to the model using a simple "add" statement (Figure \ref{fig:figuremechanism}E).
@@
 
-Switching back and forth points in parameter space and state space of a neuron model is a common occurrence in working with neuron models, and a significant fraction of a modeller's time is spent in a feedback loop of running simulations, viewing the output, changing parameters, and re-running simulations.
-(\cite{deschutterConsumerGuideNeuronal1992}).
+Switching back and forth points in parameter space and state space of a neuron model is a common occurrence in working with neuron models, and a significant fraction of a modeler's time is spent in a feedback loop of running simulations, viewing the output, changing parameters, and re-running simulations (\cite{deschutterConsumerGuideNeuronal1992}).
 xolotl makes it easy to bookmark configurations of a model and return to them at will.

### 2018-08-15 figures/voltage-clamp/clamp_snippet.m

@@ -2,7 +2,7 @@
 x.add('compartment','AB','A',.06)
 x.AB.add('Kd','gbar', 300);
 n = floor(x.t_end/x.sim_dt);
-V_clamp = repmat(20,n,1);
-V_clamp(1:1e3) = -60;
+V_clamp = zeros(n,1);
+V_clamp(1:1e4) = -60;
 x.V_clamp = V_clamp;
-I = x.integrate;
+I_clamp = x.integrate;

### 2018-08-15 paper/paper.tex
General revisions of the manuscript

--- 2018-08-14 paper/paper.tex
@@
 \newcommand{\viz}{\textit{viz.}\xspace}
+\newcommand{\cf}{\textit{cf.}\xspace}
 
@@
 
-	Conductance-based models of neurons are used extensively in computational neuroscience, and working with these models continues to be a challenge due to their high dimensionality and large number of parameters.
-Here, we present a neuron and network simulator built on a novel automatic type system that binds object-oriented code written in C++ to objects in MATLAB.
+	Conductance-based models of neurons are used extensively in computational neuroscience, though working with these models continues to be a challenge due to their high dimensionality and large number of parameters.
+We present a neuron and network simulator built on a novel automatic type system that binds object-oriented code written in C++ to objects in MATLAB.
 Our approach builds on the tradition of uniting the speed and power of languages like C++ with the ease-of-use and feature-set of scientific programming languages like MATLAB.
@@
 Our simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be extended to include arbitrary conductances, synapses, and mechanisms.
-Several core features like bookmarking of parameters and automatic hashing of source code promote reproducible and auditable research.
+Several core features like bookmarking of parameters and automatic hashing of source code facilitate reproducible and auditable research.
 Finally, our simulator is written in a modular fashion and has been released under a permissive free software license, enabling seamless integration into the workflow of other researchers.
@@
 Two major approaches have dominated the design of neuron simulators.
-One approach is to write the simulator in a fast compiled language like C and allow for the construction and simulation of neuron models using object-oriented paradigms.
-This approach has been implemented in NEURON (\cite{hinesNEURONSimulationEnvironment1997}).
+Component-based neuronal simulators, like NEURON (\cite{hinesNEURONSimulationEnvironment1997}), implemented in a fast compiled language like C, allow for the construction and simulation of neuron models using object-oriented paradigms.
 Simulators designed in this way tend to perform fast computations with little overhead, but suffer from a steep learning curve.
@@
 
-This software is designed to be used from within MATLAB, a scientific programming language common amongst neuroscientists and engineers for pedagogy and research.
-Our goal was to make xolotl completely usable entirely from within MATLAB.
-Models created using this simulator appear in the MATLAB workspace as native objects, are are thus fully scriptable, and are fully compatible with the large library of toolboxes that MATLAB provides, allowing the software to be used as a component of other packages and tools.
+xolotl is designed to be completely usable entirely from within MATLAB, a scientific programming language common among neuroscientists and engineers for research and pedagogy.
+Models created using this simulator appear in the MATLAB workspace as native objects, are are thus fully scriptable, and fully compatible with the large library of toolboxes that MATLAB provides, allowing the software to be used as a component of other packages and tools.
 All parameters of models and activation functions of any channel can be inspected at any point.
@@
 \paragraph{A rich library of network components.} xolotl comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and mechanisms) that can be used as building blocks to construct model neurons and networks.
-Compartments represent sections of membrane with a single membrane potential, Calcium level, and set of constituent components, representing either entire cells or parts of cells.
+Compartments represent sections of membrane with a single membrane potential, intracellular Calcium concentration, and set of constituent components, representing either entire cells or parts of cells.
 Objects of type "conductance" represent populations of ion channels in a compartment that produce transmembrane currents.
@@
 \paragraph{Object-oriented.}
-xolotl is designed to mirror the nested and hierarchical structure of networks and neurons.
+xolotl mirrors the nested and hierarchical structure of networks and neurons.
 Biological neuronal networks are made up of neurons that are connected to each other with synapses.
@@
 Compartments, conductances and synapses can contain mechanisms that can affect anything in the model.
-All objects in the xolotl model tree (compartments, conductances, etc.) are bonafide MATLAB objects with their own type and properties.
-This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain neurons that contain populations of ion channels, and makes constructing, integrating, and thinking about these models easier.
+All objects in the xolotl model tree (compartments, conductances, etc.) are genuine MATLAB objects with their own type and properties.
+This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain whole neurons or sections of membrane which contain populations of ion channels, making constructing, integrating, and thinking about these models easier.
 
@@
 
-Our focus on our simulator's ease-of-use and speed imposed some limitations on its feature set.
+Our focus on ease-of-use and speed imposed some limitations on our simulator's feature set.
 
@@
 While the software can integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task, offering more natural frameworks for dealing with a large number of identical units.
-Similarly, xolotl is not optimized to solve coupled ODEs on complex branched morphologies, and other simulators like (\cite{hinesNEURONSimulationEnvironment1997}) are more suited for this task.
+Similarly, xolotl is not optimized to solve coupled ODEs on complex branched morphologies, and other simulators like NEURON (\cite{hinesNEURONSimulationEnvironment1997}) are more suited for this task.
 
@@
 These models of ion channels were obtained from \cite{liuModelNeuronActivityDependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator.
-The code to set up this model is terse, idiomatic, relies on no special markup, and preserves the heirarchical nature of the model (Figure \ref{fig:figurehh}C).
+The code to set up this model is terse, idiomatic, relies on no special markup, and preserves the hierarchical nature of the model (Figure \ref{fig:figurehh}C).
 
-Adding an injected current and calling the built-in "plot" function plots the time series of membrane voltage.
+Calling the built-in "plot" function plots the time series of membrane voltage.
 In the absence of injected current, the model is quiescent.
@@
 
-Voltage clamping is an experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (\cite{dayanTheoreticalNeuroscience2001}).
-Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp (Figure \ref{fig:figureclamp}A).
+Voltage clamping is an experimental technique in which an amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (\cite{dayanTheoreticalNeuroscience2001}).
+Under this paradigm, the membrane voltage is ``clamped'' or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp (Figure \ref{fig:figureclamp}A).
 By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized (\cite{coleIonicCurrentMeasurements1960, coleIonsPotentialsNerve1955, hodgkinEffectSodiumIons1949, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, turrigianoSelectiveRegulationCurrent1995}).
@@
 
-So far, the models we described only considered the voltage dynamics of a cell (the solution to Eq.
-\ref{eq:1}).
+So far, the models we described only considered the voltage dynamics of a cell (the solution to Equation \ref{eq:1}).
 However, real neurons possess several dynamical features, arising from a variety of intracellular mechanisms.
@@
 Neurons can regulate their electrical activity by controlling the spectrum of ion channels they express (\cite{macleanActivityIndependentHomeostasisRhythmically2003, turrigianoSelectiveRegulationCurrent1995, schulzVariableChannelExpression2006}).
-Here, we will show how xolotl can be used to represent a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H).
+Here, we show how xolotl can be used to represent a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H).
 Since this mechanism affects each ion channel population individually, an object corresponding to this mechanism is added to each "conductance" object in the neuron (Figure \ref{fig:figuremechanism}I).
@@
 
-(Figure \ref{fig:figuresnapshot}) illustrates how these features can be used to explore the dynamics of the model presented in the previous section.
+Figure \ref{fig:figuresnapshot} illustrates how these features can be used to explore the dynamics of the model presented in the previous section.
 First, the current state of the model is saved using the "snapshot" method into a state called ``initial''.
 In this state, the model exhibits periodic bursting activity due to a particular configuration of maximal conductance densities (Figure \ref{fig:figuremechanism}, orange).
-On setting the maximal conductances of the Calcium-permissive channels to 0, the model switches to a tonic spiking activity (Figure \ref{fig:figuremechanism}, purple).
+Upon setting the maximal conductances of the Calcium-permissive channels to 0, the model switches to a tonic spiking activity (Figure \ref{fig:figuremechanism}, purple).
 Integrating the model for a longer duration allows the homeostatic control mechanism in the cell to restore the conductance profile and bursting activity to a state close to the original state (Figure \ref{fig:figuremechanism}, green).
@@
 In practice, this meant writing a script, running it, inspecting the output, changing parameters in the script, and repeating this process.
-It can be cumbersome, and every step in this process involves ``mode'' changes: between a text editor, viewing a graphical output, and the command line that can frustrate the researcher.
+It can be cumbersome; every step in this process involves ``mode'' changes: between a text editor, viewing a graphical output, and the command line that can frustrate the researcher.
 
@@
 \ref{fig:figurehh}) and bursting neuron models based on the bursting neurons in the lobster stomatogastric ganglion (STG) (\cite{prinzSimilarNetworkActivity2004}).
-For each type of neuron model, we compared our software to  NEURON (\cite{hinesNEURONSimulationEnvironment1997}), a high-performance and powerful neuron simulator specialized in simulating neurons with complex morphologies and DynaSim (\cite{sherfeyDynaSimMATLABToolbox2018}), a general-purpose simulator that can solve coupled differential equations numerically.
-All simulators were run on the same hardware using fixed time-step solvers: xolotl used the Exponential Euler method (\cite{dayanTheoreticalNeuroscience2001}), NEURON used the  implicit Euler solver (\cite{hinesNEURONSimulationEnvironment1997}) and DynaSim used C-compiled 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance (\cite{sherfeyDynaSimMATLABToolbox2018}).
-We measured the speed of each simulator by dividing the time simulated for, by the time it took for the simulator to complete integration.
+For each type of neuron model, we compared our software to  NEURON (\cite{hinesNEURONSimulationEnvironment1997}), a high-performance and powerful neuron simulator specialized in simulating neurons with complex morphologies, and DynaSim (\cite{sherfeyDynaSimMATLABToolbox2018}), a general-purpose simulator that can solve coupled differential equations numerically.
+All simulators were run on the same hardware using fixed time-step solvers: xolotl used an exponential Euler method (\cite{dayanTheoreticalNeuroscience2001}), NEURON used an implicit Euler solver (\cite{hinesNEURONSimulationEnvironment1997}) and DynaSim used C-compiled 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance (\cite{sherfeyDynaSimMATLABToolbox2018}).
+We measured the speed of each simulator by dividing the time simulated, by the time it took for the simulator to complete integration.
 For example, if a simulator could simulate 10 seconds of model data in 1 second, its speed would be 10X.
@@
 All three simulators were faster with larger time steps, since fewer iterations were needed (Fig.
-\ref{fig:figurebenchmark}A-B), and were approximately linear in the region tested.
+\ref{fig:figurebenchmark}A-B).
+Speeds were approximately linear in the region tested.
 Our simulator compared favorably to NEURON and DynaSim in this task.
@@
 
-Our simulator is a purely component-based simulator, and all equations need to be included in a C++ header file that specifies a object.
+Our simulator is a purely component-based simulator; all equations need to be included in a C++ header file that specifies a object.
 Since our automatic type system binds MATLAB objects to the underlying C++ header files, objects can be inspected and parameters can be modified in the MATLAB workspace.
@@
 
-Another architectural choice in simulators is the syntax required for specifying models.
+The syntax required for specifying models is an architectural choice which varies between simulators.
 Equation-oriented simulators specify models as strings of equations, so must invent a new syntax to specify derivatives, variables, and other common elements in coupled differential equations.
@@
 
-Since xolotl models are bonafide MATLAB objects, they are compatible with most of the powerful tools that exist within MATLAB.
+Since xolotl models are genuine MATLAB objects, they are compatible with most of the powerful tools that exist within MATLAB.
 For example, it is possible to write simple scripts that run xolotl models in parallel using MATLAB's parallel processing toolbox, speeding up large simulations.
@@
 As a result, many of the key features and architectures of xolotl can be reused by others in their own applications.
-For example, cpplab, the automatic type system that binds C++ code to MATLAB objects, is independent of this simulator, and exists as a distinct repository that has been made freely available (https://github.com/sg-s/cpplab); and the framework for generating a GUI with sliders for each parameter that is hooked up to function callbacks also exists as an independent repository (https://github.com/sg-s/puppeteer) that can be easily integrated into other applications.
+For example, "cpplab", the automatic type system that binds C++ code to MATLAB objects, is independent of this simulator, and exists as a distinct repository that has been made freely available (https://github.com/sg-s/cpplab); and the framework for generating a GUI with sliders for each parameter that is hooked up to function callbacks also exists as an independent repository (https://github.com/sg-s/puppeteer) that can be easily integrated into other applications.
 
@@
 Colors indicate the dominant current (gold is fast sodium (NaV), blue is delayed rectifier (Kd), red is Leak).
-(E) firing rate $vs$.
-current (f-I) curve of this neuron.
+(E) firing rate \vs current (f-I) curve of this neuron.
 (F-G) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
@@
 
+\clearpage
+
 \begin{figure}[!htb]
@@
 	\caption{Simulating a voltage-clamp experiment.
-The diagram shows a cell with delayed rectifier potassium conductance (from \cite{liuModelNeuronActivityDependent1998}) that is being recorded from in two-electrode voltage clamp (A).
-The code snippet shown here sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it (B).
-Voltage steps that the cell is clamped to (C).
-Clamping currents as a function of time (D).
-Asymptotic clamped current {\em vs.} clamped voltage for this cell (E).
-Accounting for the reversal potential of Potassium ions yields the conductance-voltage curve of this channel type (F).
-Normalized conductance-voltage curves, with sigmoid fits with various exponents (G).
-An exponent of $n=4$ yields the best fit, allowing for the characterization of the activation function of this channel type (H).
-}
+(A) The diagram shows a cell with delayed rectifier potassium conductance (\cf \cite{liuModelNeuronActivityDependent1998}) that is being recorded from in two-electrode voltage clamp.
+(B) The code snippet sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it.
+(C) Voltage steps that the cell is clamped to.
+(D) Current while in voltage clamp as a function of time.
+(E) Asymptotic clamped current \vs clamped voltage for this cell.
+(F) Accounting for the reversal potential of Potassium ions yields the conductance-voltage curve of this channel type.
+(G) Normalized conductance-voltage curves, with sigmoid fits with various exponents.
+(H) An exponent of $n=4$ yields the best fit, allowing for the characterization of the activation function of this channel type.}
 	\label{fig:figureclamp}
 \end{figure}
+
+\clearpage
 
@@
 	\caption{Modeling intracellular mechanisms.
-A single-compartment neuron model with 8 channel types (A).
-Since there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant (B), and the cell tonically spikes (C).
-Intracellular Calcium buffering and influx through voltage gated Calcium channels (VGCCs) can be modeled using a simple differential equation (D).
-Code snippet shows how this mechanism can be added to the neuron model (E).
-The cell now bursts periodically, with synchronized oscillations in intracellular Calcium (F-G).
-Schematic of Calcium-dependent integral control homeostasis (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (H).
+(A) Cartoon of a  single-compartment neuron model with eight channel types.
+(B-C) Since there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant and the cell tonically spikes.
+(D) Intracellular Calcium buffering and influx through voltage gated Calcium channels can be modeled using a simple differential equation.
+(E) Code snippet shows how this mechanism can be added to the neuron model.
+(F-G) The cell now bursts periodically, with synchronized intracellular Calcium concentration and membrane potential oscillations.
+(H) Schematic of Calcium-dependent integral control homeostasis (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}).
 In this feedback system, the rates of mRNA synthesis depend on the Calcium level in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which, through translation, depends on the mRNA abundance.
-(I) The code snippet shows how these integral controllers are implemented as \texttt{mechanism} objects, and can be added to conductances.
-(J) On integrating the model, intracellular calcium levels rise and approach the target (red dashed line).
-This is accompanied by an increase in the conductance densities of all channels being controlled by this homeostatic mechanism (K).
-The voltage behavior of the cell changes from silence to bursting with truncated spikes to regular bursting.}
+(I) The code snippet shows how these integral controllers are implemented as "mechanism" objects, and can be added to conductances.
+(J) Upon integrating the model, intracellular calcium levels rise and approach the target (red dashed line).
+(K) This is accompanied by an increase in the conductance densities of all channels being controlled by this homeostatic mechanism.
+The voltage behavior of the cell changes from silence to bursting with truncated spikes to regular bursting (insets).}
 	\label{fig:figuremechanism}
 \end{figure}
+
+\clearpage
 
@@
 	\includegraphics[width=1.0\linewidth]{gfx/figure_snapshot}
-	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them $ad$   $arbitrium$.
+	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them \textit{ad arbitrium}.
 The initial state (orange node) of the single compartment model in the previous example is saved using the \texttt{snapshot} method.
@@
 After evolving the model for some time (green node), the conductance profile and voltage dynamics returns to a state similar to the initial state (green plots).
-This configuration is now saved in a state called $final$ and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange).
-Another parameter is now changed (the Calcium target), and the model is integrated to reach a new state (blue node) where the voltage dynamics are different from the $initial$ state.
-In summary, any state can be bookmarked using a descriptive name using the \texttt{snapshot} method, and can be returned to using the \texttt{reset} method.}
+This configuration is now saved in a state called \textit{final} and the initial configuration is returned to using the "reset" method (backwards arrow from green to orange).
+Another parameter is now changed (the Calcium target), and the model is integrated to reach a new state (blue node) where the voltage dynamics are different from the \textit{initial} state.
+In summary, any state can be bookmarked using a descriptive name using the "snapshot" method, and can be returned to using the "reset"" method.}
 	\label{fig:figuresnapshot}
 \end{figure}
+
+\clearpage
 
@@
 	\caption{Simulating a network of conductance-based model neurons, coupled by voltage-dependent dynamic chemical synapses.
-A three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (\cite{prinzSimilarNetworkActivity2004}) (A).
-Each neuron is modeled with a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
+(A) A three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (\cite{prinzSimilarNetworkActivity2004}).
 Synapses can be one of two types (Cholinergic (dashed lines) or Glutamatergic (solid lines)) and have different kinetics.
-The code snippet shows how neurons can be created, wired together using synapses, and how the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse (B).
-Simulated voltage trace of a model network for the three compartments obtained from this simulation (C-E) .
-Time series activation variable of the Glutamatergic synapse between PY and LP (red connection in diagram) shows how the synapse becomes active every time the PY neuron spikes (F).}
+(B) Each neuron is modeled with a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances.
+The code snippet shows how neurons can be created and wired together using synapses.
+The model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse for the network.
+(C-E) Simulated voltage trace of a model network for the three compartments obtained from this simulation.
+(F) Time series of the synaptic activation variable for the Glutamatergic synapse between PY and LP (red connection in diagram), showing that the synapse becomes active every time the PY neuron spikes.}
 	\label{fig:figurenetwork}
@@
 	\caption{Manipulating neuron parameters in real time.
-Any set of parameter in the model can be manipulated; here, the maximal conductance of every conductance type in the model from Fig.
-\ref{fig:figurehh} is being manipulated using the code snippet shown here.
-The screenshot shows a GUI with sliders for every parameter of interest that is created by the \texttt{manipulate} method.
+(A) Any set of parameters in the model can be manipulated by the graphical interface.
+Here, we manipulate every maximal conductance from the model in Figure \ref{fig:figurehh} using the code snipped shown.
+The screenshot shows a GUI with sliders for every parameter of interest that is created by the "manipulate" method.
 These sliders can be linked to an arbitrary number of visualization functions.
-In this example, two visualization functions are used: the built in \texttt{plot} method (A) and a custom function that computes the firing-rate-$vs$.-injected current curve for this neuron (B).
-Both plots refresh themselves with every movement of any slider, allowing the user to build intuition about how every parameter controls the dynamical behavior of the model.
-A screen recording of this model being manipulated in real time is included in Supplementary Material}
+In this example, two visualization functions are used: (B) the built in "plot" method and (C) a custom function that computes the firing-rate-\vs-injected current curve for this neuron Both plots refresh themselves with every movement of any slider, allowing the user to build intuition about how every parameter controls the dynamical behavior of the model.
+A screen recording of this model being manipulated in real time is included in the Supplementary Material.}
 	\label{fig:figuremanipulate}
@@
 	\caption{Comparison of speed and accuracy of xolotl, NEURON and DynaSim.
-The top row shows simulations of a tonically-firing Hodgkin-Huxley model with three conductances with constant injected current (as in Fig.
-\ref{fig:figurehh}).
-The bottom row shows simulations of  a bursting stomatogastric ganglion neuron model with 8 conductances (as in Fig.
-\ref{fig:figurenetwork}).
-Ratio of run-time to simulation time (relative speed) as a function of simulation time step (A, B).
-Simulation error as a function of the step size (C, D).
-Relative speed of integration as a function of the simulation length (E, F).
-Relative speed of integration, normalised by system size, as a function of the number of compartments simulated simultaneously (G, H).
+The top row shows simulations of a tonically-firing Hodgkin-Huxley model with three conductances with constant injected current (\cf Figure \ref{fig:figurehh}).
+The bottom row shows simulations of a bursting stomatogastric ganglion neuron model with eight conductances (\cf Figure \ref{fig:figurenetwork}).
+(A-B) Ratio of run-time to simulation time (relative speed) as a function of simulation time step.
+(C-D) Simulation error as a function of the step size.
+(E-F) Relative speed of integration as a function of the simulation length.
+(G-H) Relative speed of integration, normalised by system size, as a function of the number of compartments simulated simultaneously.
 All benchmarks were performed on the same computer, and all simulators using fixed time-step integration methods.

### 2018-08-15 paper/paper.tex
minor changes

--- 2018-08-15 paper/paper.tex
@@
 \section*{Acknowledgments}
-The authors would like to thank Timothy O'Leary and Leandro Alonso for useful discussions, and Janis Li for help transcribing many conductance models.
+The authors would like to thank Timothy O'Leary and Leandro Alonso for useful discussions, Janis Li for help transcribing many conductance models, and members of the Marder Lab, especially Ekaterina Morozova, Jason Pipkin and Philipp Rosenbaum for careful reading of the manuscript.
 

### 2018-08-16 paper/paper.tex
minor changes to paper, continued to work on stability figure

@@
 %\def\corrAddress{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
-\def\corrAddress{Volen National Center for Complex Systems, Brandeis University, Waltham MA 02453}
+\def\corrAddress{Volen National Center for Complex Systems, Brandeis University, Waltham MA 02454}
 \def\corrEmail{\texttt{srinivasgs@brandeis.edu}}
@@
 
-\title[xolotl: neuron \& network simulator]{xolotl: An Intuitive and Approachable Neuron \& Network Simulator}
+\title[Xolotl: neuron and network simulator]{xolotl: An Intuitive and Approachable Neuron and Network Simulator for Research and Teaching}
 \author[\firstAuthorLast ]{\Authors}
@@
 Our approach builds on the tradition of uniting the speed of languages like C++ with the ease-of-use and feature-set of scientific programming languages like MATLAB.
-Our software allows for the creation and manipulation of hierarchical models with components that are named and searchable, permitting intuitive high-level programmatic control over all parameters.
-Our simulator's architecture allows for the interactive manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately.
-Our simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be extended to include arbitrary conductances, synapses, and mechanisms.
+This software allows for the creation and manipulation of hierarchical models with components that are named and searchable, permitting intuitive high-level programmatic control over all parameters.
+The simulator's architecture allows for the interactive manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately.
+xolotl is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be extended to include arbitrary conductances, synapses, and mechanisms.
 Several core features like bookmarking of parameters and automatic hashing of source code facilitate reproducible and auditable research.
-Finally, our simulator is written in a modular fashion, includes detailed tutorials and worked examples, and is freely available at https://github.com/sg-s/xolotl, enabling seamless integration into the workflow of other researchers.
+Its ease of use and rich visualization capabilites make it an attractive option in teaching environments.
+Finally, xolotl is written in a modular fashion, includes detailed tutorials and worked examples, and is freely available at https://github.com/sg-s/xolotl, enabling seamless integration into the workflow of other researchers.
 
@@
 Conductance-based models are a powerful biophysical simplification of an electrically excitable compartment in a neuron (\cite{hodgkinQuantitativeDescriptionMembrane1952}).
-Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small circuits (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}).
+Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in some circuits (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}).
 These models provide an approachable framework for understanding many salient principles of electrophysiology, since they explicitly model cell membranes and ion channels as electrical components in a circuit.
@@
 This approach permits considerable flexibility for simulating systems of differential equations.
-Since models need to be translated between the two languages, the hierarchical nature of neuron models is not naturally encapsulated by these tools, and the syntax can be verbose.
+Because models need to be translated between the two languages, the hierarchical nature of neuron models is not naturally encapsulated by these tools, and the syntax can be verbose.
 Neither approach facilitates the creation of tools that simultaneously maintain efficiency, ease-of-use, and clarity.
@@
 \paragraph{Automatic hashing and compiling.}
-Since every model requires a compiled binary to run, a potential stumbling block is the problem of unambiguously linking a model to a binary executable.
+Because every model requires a compiled binary to run, a potential stumbling block is the problem of unambiguously linking a model to a binary executable.
 xolotl solves this problem by hashing (\cite{rivestMD5MessageDigestAlgorithm1992}) the C++ header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier (its ``hash'').
@@
 
-Our focus on our simulator's ease-of-use and speed imposed some limitations on its feature set.
+Our focus on xolotl's ease-of-use and speed imposed some limitations on its feature set.
 
@@
 
-In this example, we demonstrate how to simulate the spiking activity of a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model and gain insight into the underlying biophysical mechanisms.
+In this example, we demonstrate how to simulate the spiking activity of a Hodgkin-Huxley-like model, and how the tools built into xolotl make it easy to set up and integrate the model and gain insight into the underlying biophysical mechanisms.
 This simple model consists of a single electrical compartment with three types of conductances (Figure \ref{fig:figurehh}A).
-This hierarchical organization of the neuron is mirrored in the structure of the model in our simulator: an object of type "compartment" is used to represent the cell body, and three objects of type "conductance" are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B).
-These models of ion channels were obtained from \cite{liuModelNeuronActivityDependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator.
+This hierarchical organization of the neuron is mirrored in the structure of the model in the simulator: an object of type "compartment" is used to represent the cell body, and three objects of type "conductance" are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B).
+These models of ion channels were obtained from (\cite{liuModelNeuronActivityDependent1998}) based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator.
 The code to set up this model is terse, idiomatic, relies on no special markup, and preserves the hierarchical nature of the model (Figure \ref{fig:figurehh}C).
@@
 
-The model can be integrated for various amplitudes of injected current to determine its f-I curve, or the relationship between injected current and its firing rate (\cite{kisperskyIncreaseSodiumConductance2012}).
+The model can be integrated for various amplitudes of injected current to determine its F-I (frequency current) curve, or the relationship between injected current and its firing rate (\cite{kisperskyIncreaseSodiumConductance2012}).
 Figure \ref{fig:figurehh}E shows the f-I curve of this model, obtained by repeated integration of the model.
@@
 First, the cell is clamped to a number of different voltages (Fig.
-\ref{fig:figureclamp}C) and the resultant clamping currents are measured by integrating the model (Fig.
+\ref{fig:figureclamp}C) and the resultant clamp currents are measured by integrating the model (Fig.
 \ref{fig:figureclamp}D).
@@
 
-By repeating the integration at a number of clamping voltages, we observe that the asymptotic clamping currents depend on the clamping voltage in a nonlinear manner (Fig.
+By repeating the integration at a number of clamp voltages, we observe that the asymptotic clamp currents depend on the clamp voltage in a nonlinear manner (Fig.
 \ref{fig:figureclamp}E), since the open probabilities of the channel are functions of the membrane voltage.
 Assuming the reversal potential is known, Eq.
-\eqref{eq:2} can be used to solve for the total conductance of the channel as a function of the clamping voltage (Fig.
+\eqref{eq:2} can be used to solve for the total conductance of the channel as a function of the clamp voltage (Fig.
 \ref{fig:figureclamp}F).
@@
 xolotl makes it possible to model and include arbitrary intracellular mechanisms.
-In our simulator, these intracellular mechanisms are represented by the "mechanism" object, and can be bound to compartments, conductances, and other object types.
+In xolotl, these intracellular mechanisms are represented by the "mechanism" object, and can be bound to compartments, conductances, and other object types.
 
@@
 Examining the voltage dynamics of the cell reveal that it transitions from quiescence to truncated bursts of action potentials to periodic bursting as this mechanism regulates the neuron's ion channel spectrum.
-In summary, our simulator can be used to construct neuron models with intracellular mechanisms such as Calcium influx and buffering, and homeostatic regulation.
+In summary, xolotl can be used to construct neuron models with intracellular mechanisms such as Calcium influx and buffering, and homeostatic regulation.
 
@@
 
-Our simulator has several types of synapses like Cholinergic, Glutamatergic, AMPA and NMDA built-in, and other synapse classes can be easily added using templates.
+xolotl has several types of synapses like Cholinergic, Glutamatergic, AMPA and NMDA built-in, and other synapse classes can be easily added using templates.
 Figure \ref{fig:figurenetwork} demonstrates the implementation of a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans (\cite{prinzSimilarNetworkActivity2004}).
@@
 
-Conductance-based neuron models are typically high dimensional and contain several parameters.
+Conductance-based neuron models are typically high dimensional and contain many parameters.
 Changing a single parameter monotonically can cause non-monotonic changes in behavior of the model, and certain dynamical features may only emerge when in specific non-convex regions of parameter space (\cite{golowasch2002failure}).
@@
 
-Our software is designed to streamline this process and allows for any parameter in any model to be manipulated using graphical sliders, with immediate, real-time feedback of its behavior.
-Any model in our simulator can be manipulated using
+Xoltol is designed to streamline this process and allows for any parameter in any model to be manipulated using graphical sliders, with immediate, real-time feedback of its behavior.
+Any model in the simulator can be manipulated using
 
@@
 \ref{fig:figurebenchmark}A-B), and were approximately linear in the region tested.
-Our simulator compared favorably to NEURON and DynaSim in this task.
+Xolotl compared favorably to NEURON and DynaSim in this task.
 We also measured the quality of the simulated output by comparing it to the simulated output at the smallest time step.
@@
 \ref{fig:figurebenchmark}C-D).
-Since our simulator sets up and runs the simulation in C++, it needs to transfer parameters and data to and from the underlying implementation.
+Since xolotl sets up and runs the simulation in C++, it needs to transfer parameters and data to and from the underlying implementation.
 To measure the performance cost of this overhead, we repeated these benchmarks on all three simulators at a fixed time step of $0.1$ ms and varied the length of time simulated for.
@@
 We set out to design a neuron and network simulator that could be useful in the classroom setting, especially for students of computational science, while also being powerful, fast, and extensible enough to be used for research.
-By using a novel architecture that permits the symbolic manipulation of C++ objects in a intuitive MATLAB interface, we demonstrated some of the features of our simulator using simulations of single-compartment models of Hodgkin-Huxley like neurons (Fig.
+By using a novel architecture that permits the symbolic manipulation of C++ objects in a intuitive MATLAB interface, we demonstrated some of the features of xolotl using simulations of single-compartment models of Hodgkin-Huxley like neurons (Fig.
 \ref{fig:figurehh}); voltage clamp experiments to recover activation functions of single channel types (Fig.
@@
 \ref{fig:figurenetwork}).
-We also illustrated how built-in features of our simulators make it easy to bookmark and jump between model configurations (Fig.
+We also illustrated how built-in features of the simulator make it easy to bookmark and jump between model configurations (Fig.
 \ref{fig:figuremechanism}, purple), and how parameters of the model can be changed using sliders and their effect can be viewed in real time (Fig.
@@
 
-Over the years, several simulators have been developed to integrate systems of coupled differential equations, those that model the spiking activity of neurons (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003, hinesNEURONSimulationEnvironment1997}).
+Over the years, several simulators have been developed to integrate systems of coupled differential equations, those that model the spiking activity of neurons (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003, hinesNEURONSimulationEnvironment1997, bower2003genesis}).
 A critical architectural choice in designing a simulator is how much ``scaffolding'' a user is provided to construct a model; whether a model is specified by equations or by components, or by some combination of the two.
@@
 
-Our simulator is a purely component-based simulator, and all equations need to be included in a C++ header file that specifies a object.
+Xolotl is a purely component-based simulator, and all equations need to be included in a C++ header file that specifies a object.
 Since our automatic type system binds MATLAB objects to the underlying C++ header files, objects can be inspected and parameters can be modified in the MATLAB workspace.
@@
 While this allows for powerful features such as support for units in NEURON, the user is required to learn a novel syntax and vocabulary, hindering ease of use.
-In general, the syntax for model specification in different simulator can be different.
-For example, DynaSim, BRIAN, XPP, and NEURON all use different, incompatible formalisms to represent equations, increasing the cognitive load of users using more than one simulator.
-In our software, we have elected not to specify a domain-specific ``middleware'' layer, and instead specify and implement models and equations in idiomatic C++.
-This greatly decreases the learning curve and allows users with a general familiarity with programming languages to quickly acquaint themselves even with the most technical parts of our simulator.
+In general, the syntax for model specification in different simulators can be different.
+For example, DynaSim, BRIAN, XPP, and NEURON all use different, incompatible formalisms to represent equations, increasing the cognitive load on users using more than one simulator.
+Here, we have elected not to specify a domain-specific ``middleware'' layer, and instead specify and implement models and equations in idiomatic C++.
+This greatly decreases the learning curve and allows users with a general familiarity with programming languages to quickly acquaint themselves even with the most technical parts of the simulator.
 
@@
 Our goal was to design software that would allow the user to verify for herself that the software was running as she intended, and to be able to reproduce results from others quickly and unambiguously.
-The primary design choice in our software that enables auditability and reproducibility is that every simulation is tied to an alphanumeric checksum, or hash, using the "MD5" algorithm (\cite{rivestMD5MessageDigestAlgorithm1992}).
+The primary design choice in our software that enables auditability and reproducibility is that every simulation is tied to an alphanumeric checksum, or hash, using the MD5 algorithm (\cite{rivestMD5MessageDigestAlgorithm1992}).
 The hash is computed from every C++ file that is part of the model, and any changes in any C++ file included in the model will trigger a recompilation of that model.
@@
 
-Since xolotl models are bonafide MATLAB objects, they are compatible with most of the powerful tools that exist within MATLAB.
+Because xolotl models are bonafide MATLAB objects, they are compatible with most of the powerful tools that exist within MATLAB.
 For example, it is possible to write simple scripts that run xolotl models in parallel using MATLAB's parallel processing toolbox, speeding up large simulations.
@@
 
-SG-S \& AH wrote the software and created the online user documentation. EM supervised the project. All authors wrote and reviewed the paper.
+SG-S \& AH wrote the software and created the online user documentation. EM provided general oversight. All authors wrote and reviewed the paper.
 
@@
 
-AH received funding from the National Institute on Drug Abuse (NIDA) through the undergraduate training grant in computational neuroscience (1R90DA033463-01).
+AH received funding from the National Institute on Drug Abuse (NIDA) through the undergraduate training grant in computational neuroscience (R90DA033463-01).
 
@@
 (A) Schematic representation of a single-compartment neuron with three populations of ion channels (colored rectangles).
-(B) In our simulator, the soma is represented using an object of class ``compartment'' and populations of ion channels are represented by ``conductance'' objects contained within the compartment object.
+(B) In xolotl, the soma is represented using an object of class ``compartment'' and populations of ion channels are represented by ``conductance'' objects contained within the compartment object.
 (C) The code snippet shown sets up this neuron model, injects current, integrates and plots the voltage, and displays activation functions, all in a few lines of code.
@@
 	\caption{Simulating a voltage-clamp experiment.
-The diagram shows a cell with delayed rectifier potassium conductance (from \cite{liuModelNeuronActivityDependent1998}) that is being recorded from in two-electrode voltage clamp (A).
+The diagram shows a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivityDependent1998}) that is being recorded from in two-electrode voltage clamp (A).
 The code snippet shown here sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it (B).
 Voltage steps that the cell is clamped to (C).
-Clamping currents as a function of time (D).
+Clamp currents as a function of time (D).
 Asymptotic clamped current {\em vs.} clamped voltage for this cell (E).
@@
 A single-compartment neuron model with 8 channel types (A).
-Since there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant (B), and the cell tonically spikes (C).
+Because there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant (B), and the cell tonically spikes (C).
 Intracellular Calcium buffering and influx through voltage gated Calcium channels (VGCCs) can be modeled using a simple differential equation (D).

### 2018-08-16 figures/benchmarks/figure_sup_stability.m

@@ -4,17 +4,16 @@
 
 clearvars -except z
 
+
+show_tol = logspace(-5,-2,5);
+n_tol = 20;
+
 redo = false;
 
 t_end = 30e3; % ms
 t_transient = 10e3; % ms
 dt = .1; % ms
 
-% make a vector of dt to vary
-max_dt = 1e2;
-K = 1:max_dt;
-all_dt = K(rem(max_dt,K) == 0);
-all_dt = all_dt/1e3;
 
 % fix pseudorandom number generation
 prng = 1984;
@@ -106,10 +105,226 @@ else
 	passingModels = cache(h);
 end
 
-
-
 % remove all non-passing models
 params = G(:, passingModels);
+
+show_this = 11;
+
+
+
+% make a curve of chosen time step vs. AbsTol for the ode solver
+all_tol = logspace(log10(show_tol(1)),log10(show_tol(end)),n_tol);
+min_time_step = NaN*all_tol;
+mean_time_step = NaN*all_tol;
+med_time_step = NaN*all_tol;
+burst_period = NaN*all_tol;
+firing_rate = NaN*all_tol;
+duty_cycle = NaN*all_tol;
+
+handles.summary_figure = figure('outerposition',[300 300 1200 600],'PaperUnits','points','PaperSize',[1200 600]); hold on
+
+handles.ode_time = subplot(2,2,1); hold on
+set(handles.ode_time,'XScale','log','YScale','log')
+handles.ode_time_min = plot(handles.ode_time,all_tol,NaN*all_tol,'g-o');
+handles.ode_time_med = plot(handles.ode_time,all_tol,NaN*all_tol,'r-o');
+handles.ode_time_mean = plot(handles.ode_time,all_tol,NaN*all_tol,'b-o');
+
+xlabel(handles.ode_time,'Tolerance of ODE solver')
+ylabel(handles.ode_time,'Chosen Timestep (ms)')
+
+
+handles.burst_period = subplot(2,2,2); hold on
+handles.firing_rate = subplot(2,2,3); hold on
+handles.duty_cycle = subplot(2,2,4); hold on
+
+
+
+T = t_end/1e3;
+time = vectorise(.1:.1:t_end);
+
+for i = 1:length(all_tol)
+	options = odeset('RelTol',all_tol(i));
[…]

### 2018-08-17 paper/paper.tex
added funding info
diff --git a/README.md b/README.md
index 4699fb8..9a34667 100644
--- a/README.md
+++ b/README.md
@@ -1,7 +1,19 @@
-# xolotl: An Intuitive and Approachable Neuron & Network Simulator in MATLAB
+# xolotl: An Intuitive and Approachable Neuron & Network Simulator

 This repository contains code, latex documents and supporting images that allows you to reproduce all figures and text in this paper.

+# Status
+
+This paper is now at the "preprint" stage, and has been submitted. You can read it on the arXiv [here]()
+
+
+# Structure
+
+The Latex source for the paper is in `paper/`

@@
 
-\title[Xolotl: neuron and network simulator]{xolotl: An Intuitive and Approachable Neuron and Network Simulator for Research and Teaching}
+\title[Xolotl: neuron and network simulator]{Xolotl: An Intuitive and Approachable Neuron and Network Simulator for Research and Teaching}
 \author[\firstAuthorLast ]{\Authors}
@@
 	\tiny
-	\keyFont{ \section{Keywords:} simulator, code:MATLAB, code:C++, conductance-based, neuron, network, neuron models}
+	\keyFont{ \section{Keywords:} code:MATLAB, code:C++, conductance-based, software, Hodgkin-Huxley}
 
@@
 
-AH received funding from the National Institute on Drug Abuse (NIDA) through the undergraduate training grant in computational neuroscience (R90DA033463-01).
+The authors acknowledge funding from the National Institute on Drug Abuse (NIDA) through the undergraduate training grant in computational neuroscience R90DA033463 (AH), T32NS007292-31 (SG-S) and R35 NS097343 (EM).
 

### 2018-08-17 paper/paper.tex
minor fixes to text

@@
 Work on the simulator continues in the open at a publicly accessible repository (https://github.com/sg-s/xolotl/), and the library of conductances, synapses and mechanisms that xolotl ships with grows continuously.
+While this simulator was intended as a research tool, the many worked examples that are built into it demonstrate how it could also be used as a teaching tool.
 
@@
 
-The authors acknowledge funding from the National Institute on Drug Abuse (NIDA) through the undergraduate training grant in computational neuroscience R90DA033463 (AH), T32NS007292-31 (SG-S) and R35 NS097343 (EM).
+The authors acknowledge funding from the National Institutes through R90DA033463 (AH), T32NS007292-31 (SG-S) and R35 NS097343 (EM).
 

### 2018-08-17 paper/paper.tex
minor changes to text

@@
 Several core features like bookmarking of parameters and automatic hashing of source code facilitate reproducible and auditable research.
-Its ease of use and rich visualization capabilites make it an attractive option in teaching environments.
+Its ease of use and rich visualization capabilities make it an attractive option in teaching environments.
 Finally, xolotl is written in a modular fashion, includes detailed tutorials and worked examples, and is freely available at https://github.com/sg-s/xolotl, enabling seamless integration into the workflow of other researchers.
@@
 
-Xoltol is designed to streamline this process and allows for any parameter in any model to be manipulated using graphical sliders, with immediate, real-time feedback of its behavior.
+Xolotl is designed to streamline this process and allows for any parameter in any model to be manipulated using graphical sliders, with immediate, real-time feedback of its behavior.
 Any model in the simulator can be manipulated using
@@
 Relative speed of integration as a function of the simulation length (E, F).
-Relative speed of integration, normalised by system size, as a function of the number of compartments simulated simultaneously (G, H).
+Relative speed of integration, normalized by system size, as a function of the number of compartments simulated simultaneously (G, H).
 All benchmarks were performed on the same computer, and all simulators using fixed time-step integration methods.

### 2018-08-17 paper/supplementary/supplementary.tex
getting close to the final stretch
minor changes to text

--- 2018-08-07 paper/supplementary/supplementary.tex
@@
+% !BIB TS-program = biber
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \title[Supplementary Material]{{\helveticaitalic{Supplementary Material}}:
-\\ \helvetica{Xolotl: An Intuitive and Approachable Neuron \& Network Simulator in MATLAB}} %Please insert the title of your article here
+\\ \helvetica{Xolotl: An Intuitive and Approachable Neuron and Network Simulator for Research and Teaching}} %Please insert the title of your article here
 
@@
 
-\section{Stability of Exponential Euler Method}
-
-PUT FIGURE HERE
+\cleardoublepage
 

### 2018-08-17 paper/paper.tex
minor typos in manuscript

@@
 Our approach builds on the tradition of uniting the speed of languages like C++ with the ease-of-use and feature-set of scientific programming languages like MATLAB.
-This software allows for the creation and manipulation of hierarchical models with components that are named and searchable, permitting intuitive high-level programmatic control over all parameters.
+Xolotl allows for the creation and manipulation of hierarchical models with components that are named and searchable, permitting intuitive high-level programmatic control over all parts of the model.
 The simulator's architecture allows for the interactive manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately.
-xolotl is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be extended to include arbitrary conductances, synapses, and mechanisms.
+Xolotl is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be extended to include arbitrary conductances, synapses, and mechanisms.
 Several core features like bookmarking of parameters and automatic hashing of source code facilitate reproducible and auditable research.
@@
 In this paper, we introduce xolotl, an implementation of the cpplab system specialized in integrating conductance-based neuron and network models.
-Models can be easily constructed from network building blocks in a few lines of MATLAB code using a hierarchical and intuitive syntax.
+Models can be easily constructed from components of different types in a few lines of MATLAB code using a hierarchical and intuitive syntax.
 Since models in the MATLAB workspace are automatically linked to models in the C++ implementation, configuring these objects in MATLAB transparently configures the underlying C++ objects.
-xolotl comes packaged with hundreds of components that can be used to assemble cells and networks;  has built-in visualization functions to inspect voltage time traces and activation functions; and a GUI for real-time manipulation of model parameters.
-The ease-of-use of the tools built into our software lends them to pedagogical applications and the rapid prototyping of models.
+Xolotl comes packaged with hundreds of components that can be used to assemble cells and networks;  has built-in visualization functions to inspect voltage time traces and activation functions; and a GUI for real-time manipulation of model parameters.
+Xolotl's ease of use makes it an attractive option for pedagogical applications, rapid prototyping of models, and primary research use.
 Our software aims to simplify the investigation of the dynamics of conductance-based network and neuron models, facilitate collaborative modeling, and is intended to complement other tools being developed in the computational neuroscience community.
@@
 
-xolotl is designed to be easy-to-use and richly featured while being fast enough to use in everyday research.
+Xolotl is designed to be easy-to-use and richly featured while being fast enough to use in everyday research.
 Our focus was on designing an approachable simulator of conductance-based neurons and networks of these neurons; simulating arbitrary dynamical systems is therefore beyond the scope of this software.
@@
 Our goal was to make xolotl completely usable entirely from within MATLAB.
-Models created using this simulator appear in the MATLAB workspace as native objects, are are thus fully scriptable, and are fully compatible with the large library of toolboxes that MATLAB provides, allowing the software to be used as a component of other packages and tools.
-All parameters of models and activation functions of any channel can be inspected at any point.
+Models created using this simulator appear in the MATLAB workspace as native objects, are thus fully scriptable, and are fully compatible with the large library of toolboxes that MATLAB provides, allowing the software to be used as a component of other packages and tools.
+All parameters of a model, and activation functions of any channel can be inspected at any point.
 We designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by calling functions that generate new C++ files on-the-fly.
-Finally, xolotl fully auditable by design, with several tools to verify model and parameter integrity and aid in reproducibility.
+Finally, xolotl is fully auditable by design, with several tools to verify model and parameter integrity and aid in reproducibility.
 
@@
 \paragraph{Object-oriented.}
-xolotl is designed to mirror the nested and hierarchical structure of networks and neurons.
+Xolotl is designed to mirror the nested and hierarchical structure of networks and neurons.
 Biological neuronal networks are made up of neurons that are connected to each other with synapses.
@@
 
-\paragraph{A rich library of network components.} xolotl comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and mechanisms) that can be used as building blocks to construct model neurons and networks.
-Compartments represent sections of membrane with a single membrane potential, Calcium level, and set of constituent components, representing either entire cells or parts of cells.
-Objects of type "conductance" represent populations of ion channels in a compartment that produce transmembrane currents.
-"synapse" objects connect two compartments together by introducing a current in the postsynaptic compartment that depends on the presynaptic compartment's membrane potential.
-Objects of type "mechanism" can represent any intracellular mechanism and can read and modify any other component in the cell, and can run arbitrary dynamical systems within them.
+\paragraph{A rich library of network components.} Xolotl comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and mechanisms) that can be used as building blocks to construct model neurons and networks.
+``Compartments'' represent sections of membrane with a single membrane potential, intracellular Calcium concentration, and set of constituent components; and can represent either entire cells or parts of cells.
+Objects of type ``conductance'' represent populations of ion channels in a compartment that produce transmembrane currents.
+``synapse'' objects connect two compartments together by introducing a current in the postsynaptic compartment that depends on the presynaptic compartment's membrane potential.
+Objects of type ``mechanism'' can represent any intracellular mechanism and can read and modify any other component in the cell, and can run arbitrary dynamical systems within them.
 
@@
 Because every model requires a compiled binary to run, a potential stumbling block is the problem of unambiguously linking a model to a binary executable.
-xolotl solves this problem by hashing (\cite{rivestMD5MessageDigestAlgorithm1992}) the C++ header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier (its ``hash'').
+Xolotl solves this problem by hashing (\cite{rivestMD5MessageDigestAlgorithm1992}) the C++ header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier (its ``hash'').
 Compiled binaries are named using this hash, ensuring both that the correct binary is run to integrate the model, and that compilation occurs only as needed.
@@
 
-\paragraph{Limited numerical integration strategies.} Most components in the software are integrated using the exponential Euler method, which has been shown to perform well in neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}).
+\paragraph{Limited numerical integration strategies.} Most components in the software are integrated using the exponential Euler method, which has been used in integrating neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}).
 However, it may be desirable to use other methods under certain conditions.
-It is possible to introduce new network components that implement other integration schemes, or to modify the integration schemes of existing components, but that requires writing new C++  code.
+It is possible to introduce new components that implement other integration schemes, or to modify the integration schemes of existing components, but that requires writing new C++  code.
 Currently, xolotl can only implement integration schemes with fixed step size.
@@
 \paragraph{Inefficient tools for handling large networks.} xolotl was designed to work with small but complex networks and models, where every compartment and component is named, rather than numbered.
-It is more therefore suited towards simulating small and heterogeneous networks rather than large but homogenous networks.
+It is more therefore suited towards simulating small, heterogeneous networks rather than large, homogenous networks.
 While the software can integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task, offering more natural frameworks for dealing with a large number of identical units.
-Similarly, xolotl is not optimized to solve coupled ODEs on complex branched morphologies, and other simulators like NEURON (\cite{hinesNEURONSimulationEnvironment1997}) are more suited for this task.
+Similarly, xolotl is not optimized to solve coupled ODEs on complex branched morphologies, that other simulators like NEURON (\cite{hinesNEURONSimulationEnvironment1997}) are specialized for.
 
@@
 This simple model consists of a single electrical compartment with three types of conductances (Figure \ref{fig:figurehh}A).
-This hierarchical organization of the neuron is mirrored in the structure of the model in the simulator: an object of type "compartment" is used to represent the cell body, and three objects of type "conductance" are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B).
+This hierarchical organization of the neuron is mirrored in the structure of the model in the simulator: an object of type ``compartment'' is used to represent the cell body, and three objects of type ``conductance'' are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B).
 These models of ion channels were obtained from (\cite{liuModelNeuronActivityDependent1998}) based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator.
@@
 
-The model can be integrated for various amplitudes of injected current to determine its F-I (frequency current) curve, or the relationship between injected current and its firing rate (\cite{kisperskyIncreaseSodiumConductance2012}).
-Figure \ref{fig:figurehh}E shows the f-I curve of this model, obtained by repeated integration of the model.
+The model can be integrated for various amplitudes of injected current to determine its F-I (frequency current) curve (\cite{kisperskyIncreaseSodiumConductance2012}).
+Figure \ref{fig:figurehh}E shows the F-I curve of this model, obtained by repeated integration of the model.
 Finally, the built-in \texttt{show} function can plot activation ($m$) and inactivation ($h$) curves and voltage-dependent timescales of any channel type in the simulator (Figure \ref{fig:figurehh}D-G).
@@
 Voltage clamping is an experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (\cite{dayanTheoreticalNeuroscience2001}).
-Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp (Figure \ref{fig:figureclamp}A).
+Under this paradigm, the membrane voltage is ``clamped'' or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp (Figure \ref{fig:figureclamp}A).
 By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized (\cite{coleIonicCurrentMeasurements1960, coleIonsPotentialsNerve1955, hodgkinEffectSodiumIons1949, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, turrigianoSelectiveRegulationCurrent1995}).
 
-xolotl can reproduce such a voltage clamp experiment \textit{in-silico}.
+Xolotl can reproduce such a voltage clamp experiment \textit{in-silico}.
 Figure \ref{fig:figureclamp}B illustrates how a simple model with a single compartment and a single ion channel type can be set up and clamped to a desired voltage.
@@
 \ref{fig:figureclamp}G-H).
-xolotl can therefore be used to describe graphically the theoretical underpinnings of ion channel characterization through voltage clamp and can serve as an effective pedagogical tool in computational neuroscience.
+Xolotl can therefore be used to describe graphically the theoretical underpinnings of ion channel characterization through voltage clamp and can serve as an effective pedagogical tool in computational neuroscience.
 
@@
 However, real neurons possess several dynamical features, arising from a variety of intracellular mechanisms.
-xolotl makes it possible to model and include arbitrary intracellular mechanisms.
-In xolotl, these intracellular mechanisms are represented by the "mechanism" object, and can be bound to compartments, conductances, and other object types.
+Xolotl makes it possible to model and include arbitrary intracellular mechanisms.
+In xolotl, these intracellular mechanisms are represented by the ``mechanism'' object, and can be bound to compartments, conductances, and other object types.
 
@@
 Switching back and forth points in parameter space and state space of a neuron model is a common occurrence in working with neuron models, and a significant fraction of a modeler's time is spent in a feedback loop of running simulations, viewing the output, changing parameters, and re-running simulations (\cite{deschutterConsumerGuideNeuronal1992}).
-xolotl makes it easy to bookmark configurations of a model and return to them at will.
+Xolotl makes it easy to bookmark configurations of a model and return to them at will.
 Internally, xolotl uses the "serialize" method to gather all parameters and dynamic variables into a vector of values that is passed to the underlying C++ implementation.
@@
 
-xolotl has several types of synapses like Cholinergic, Glutamatergic, AMPA and NMDA built-in, and other synapse classes can be easily added using templates.
+Xolotl has several types of synapses built-in, and other synapse classes can be easily added using templates.
 Figure \ref{fig:figurenetwork} demonstrates the implementation of a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans (\cite{prinzSimilarNetworkActivity2004}).
@@
 Conversely, when "PY" is hyperpolarized, the dynamical variable is close to 0.
-When "PY" spikes, IPSPs can be seen in the "LP" voltage trace (Figure \ref{fig:figurenetwork}D-E).
+In this model, when "PY" spikes, IPSPs can be seen in the "LP" voltage trace (Figure \ref{fig:figurenetwork}D-E).
 
@@
 In practice, this meant writing a script, running it, inspecting the output, changing parameters in the script, and repeating this process.
-It can be cumbersome, and every step in this process involves ``mode'' changes: between a text editor, viewing a graphical output, and the command line that can frustrate the researcher.
+It can be cumbersome, and every step in this process involves ``mode'' changes: switching between a text editor, viewing a graphical output, and the command line that can frustrate the researcher.
 
@@
 
-This feature is only possible due to our architectural decision to split the code base across two programming environments.
-A rich scientific programming environment like MATLAB makes it possible to generate user interface elements and to bind them to data in plots, while the sheer speed of compiled languages like C++ allow for the immediate, real-time feedback and updating of plots.
+This feature was only possible due to our architectural decision to split the code base across two programming environments.
+A rich scientific programming environment like MATLAB makes it possible to easily generate user interface elements and to bind them to data in plots, while the sheer speed of compiled languages like C++ allow for the immediate, real-time feedback and updating of plots.
 By default, any parameter in the model can be manipulated, including parameters in user-defined mechanisms that do not exist in the base simulator.
@@
 While our software is not designed for this task {\em per se}, we measured its performance as a function of the number of compartments simulated.
-xolotl can quickly create a number of identical copies of a compartment using the "replicate" method:
+Xolotl can quickly create a number of identical copies of a compartment using the "replicate" method:
 
@@
 We used the "replicate" method to create and run models with varying numbers of neurons (either Hodgkin-Huxley-like or bursting neurons) and measured the speed of all three simulators as a function of system size.
-Plotting the speed normalized by the system size vs.
-the system size, we observed that the speed of integration of xolotl is linear with system size (Figure \ref{fig:figurebenchmark}G-H, black lines), for up to 1000 single-compartment neurons (up to 13,000 ODEs).
+Plotting the speed normalized by the system size {\em vs.} the system size, we observed that the speed of integration of xolotl is linear with system size (Figure \ref{fig:figurebenchmark}G-H, black lines), for up to 1000 single-compartment neurons (up to 13,000 ODEs).
 Its performance compares favorably with that of NEURON and DynaSim as a function of system size (Figure \ref{fig:figurebenchmark}G-H).
@@
 
-We set out to design a neuron and network simulator that could be useful in the classroom setting, especially for students of computational science, while also being powerful, fast, and extensible enough to be used for research.
+We set out to design a neuron and network simulator that could be useful in the classroom setting, especially for students of computational neuroscience, while also being powerful, fast, and extensible enough to be used for research.
 By using a novel architecture that permits the symbolic manipulation of C++ objects in a intuitive MATLAB interface, we demonstrated some of the features of xolotl using simulations of single-compartment models of Hodgkin-Huxley like neurons (Fig.
@@
 Thus, a good user interface may contribute more to the overall efficiency of a project than pure computation speed.'' (\cite{deschutterConsumerGuideNeuronal1992}).
-xolotl is designed primarily with ease-of-use in mind.
+Xolotl is designed primarily with ease-of-use in mind.
 This includes the time it takes to install, setup, and learn how to use the software, the time to write and debug scripts, and the time to perform the simulations (\cite{rudolph2007much}).
@@
 
-Over the years, several simulators have been developed to integrate systems of coupled differential equations, those that model the spiking activity of neurons (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003, hinesNEURONSimulationEnvironment1997, bower2003genesis}).
-A critical architectural choice in designing a simulator is how much ``scaffolding'' a user is provided to construct a model; whether a model is specified by equations or by components, or by some combination of the two.
-In an equation-oriented architecture, the user starts with a blank slate and the primary method of specifying a model is to write out its differential equations.
+Over the years, several simulators have been developed to integrate systems of coupled differential equations that model the spiking activity of neurons (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003, hinesNEURONSimulationEnvironment1997, bower2003genesis}).
+A critical architectural choice in designing a simulator is how much ``scaffolding'' a user is provided with to construct a model; whether a model is specified by equations or by components, or by some combination of the two.
+In an equation-oriented architecture, the user starts with a blank slate and the primary method of specifying a model is to write out its differential equations (\cite{stimberg2014equation}).
 In contrast, in a component-oriented architecture, the primary method of specifying a model is to assemble it from pre-existing components, each of which include differential equations, parameters, and solvers.
-Both approaches have advantages and disadvantages that are illustrated below.
+Both approaches have advantages and disadvantages that are discussed below.
 
@@
 
-Another architectural choice in simulators is the syntax required for specifying models.
+Another architectural choice in designing simulators is the syntax required for specifying models.
 Equation-oriented simulators specify models as strings of equations, so must invent a new syntax to specify derivatives, variables, and other common elements in coupled differential equations.
@@
 For example, it is possible to write simple scripts that run xolotl models in parallel using MATLAB's parallel processing toolbox, speeding up large simulations.
-xolotl models are also compatible with the global optimization toolbox in MATLAB, allowing parameters in xolotl models to be optimized, enabling the creation of toolboxes that efficiently tune parameters in neuron models to satisfy arbitrary constraints (\cite{Achard:2006ia, Krichmar:2014jj, Keren:2005ja, VanGeit:2007ea, Druckmann:2008kg}).
+Xolotl models are also compatible with the global optimization toolbox in MATLAB, allowing parameters in xolotl models to be optimized, enabling the creation of toolboxes that efficiently tune parameters in neuron models to satisfy arbitrary constraints (\cite{Achard:2006ia, Krichmar:2014jj, Keren:2005ja, VanGeit:2007ea, Druckmann:2008kg}).
 

### 2018-08-17 paper/paper.tex
another minor typo

@@
 Its ease of use and rich visualization capabilities make it an attractive option in teaching environments.
-Finally, xolotl is written in a modular fashion, includes detailed tutorials and worked examples, and is freely available at https://github.com/sg-s/xolotl, enabling seamless integration into the workflow of other researchers.
+Finally, xolotl is written in a modular fashion, includes detailed tutorials and worked examples, and is freely available at https://github.com/sg-s/xolotl, enabling seamless integration into the workflows of other researchers.
 
@@
 Here, we will show how xolotl can be used to represent a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H).
-Since this mechanism affects each ion channel population individually, an object corresponding to this mechanism is added to each "conductance" object in the neuron (Figure \ref{fig:figuremechanism}I).
+Since this mechanism affects each ion channel population individually, an object corresponding to this mechanism is added to each conductance object in the neuron (Figure \ref{fig:figuremechanism}I).
 Setting all maximal conductance densities to some low value and integrating the model shows that the intracellular Calcium levels rise over time and approach the target Calcium concentration (Figure \ref{fig:figuremechanism}J), while all conductance densities increase and then remain bounded (Figure \ref{fig:figuremechanism}K).
@@
 
-Network models in xolotl consist of "compartment" objects that can be connected by "synapse" objects.
+Network models in xolotl consist of compartment objects that can be connected by synapse objects.
 Two compartments representing different neurons can be connected using synapses using the built-in "connect" method.

### 2018-08-17 paper/paper.tex
more typos fixed

@@
 
-	Conductance-based models of neurons are used extensively in computational neuroscience, and working with these models continues to be a challenge due to their high dimensionality and large number of parameters.
+	Conductance-based models of neurons are used extensively in computational neuroscience.
+Working with these models can be  challenging due to their high dimensionality and large number of parameters.
 Here, we present a neuron and network simulator built on a novel automatic type system that binds object-oriented code written in C++ to objects in MATLAB.
@@
 Objects of type ``mechanism'' can represent any intracellular mechanism and can read and modify any other component in the cell, and can run arbitrary dynamical systems within them.
+Parameters of any of these objects can be easily inspected and modified at any time, either manually or through a programmatic interface.
 
 \paragraph{Automatic type system.}
-To circumvent the tradeoff between high-performance but hard to use languages like C++ and richly-featured but potentially slow languages like MATLAB, we constructed an automatic type system that links object oriented code in C++ to object oriented code in MATLAB.
+To circumvent the tradeoff between high-performance but hard-to-use languages like C++ and richly-featured but potentially slow languages like MATLAB, we constructed an automatic type system that links object oriented code in C++ to object oriented code in MATLAB.
 This architecture lets us construct the core of the simulator in C++, leveraging features of C++ like pointers that are not readily available in MATLAB.
@@
 In this section, we illustrate how xolotl can be used to generate, inspect, and simulate a variety of models.
-These examples have been chosen to illustrate various features of xolotl, and are intended to serve as templates upon which researchers and educators can build.
+These examples have been chosen to demonstrate various features of xolotl, and are intended to serve as templates upon which researchers and educators can build.
 
@@
 
-A key intracellular mechanisms is the cytosolic buffering of Calcium and its influx through voltage-gated Calcium channels.
+A key intracellular mechanism is the cytosolic buffering of Calcium and its influx through voltage-gated Calcium channels.
 Figure \ref{fig:figuremechanism}A shows a model of a single-compartment model with 8 populations of ion channels (\cite{liuModelNeuronActivityDependent1998}).
@@
 \section*{Acknowledgments}
-The authors would like to thank Timothy O'Leary and Leandro Alonso for useful discussions, Janis Li for help transcribing many conductance models, and members of the Marder Lab, especially Ekaterina Morozova, Jason Pipkin and Philipp Rosenbaum for careful reading of the manuscript.
+The authors would like to thank Timothy O'Leary and Leandro Alonso for useful discussions, Janis Li for help transcribing many conductance models, and members of the Marder Lab, especially Sonal Kedia, Ekaterina Morozova, Jason Pipkin and Philipp Rosenbaum for careful reading of the manuscript.
 

### 2018-10-25 paper/paper.tex
whitespace

@@
 \usepackage{hyperref}
+
+\usepackage{tcolorbox}
+\tcbuselibrary{breakable}
+\usepackage[framemethod=tikz]{mdframed}
 
@@
 Parameters of any of these objects can be easily inspected and modified at any time, either manually or through a programmatic interface.
+\hl{Every compartment exists as a C++ file within the code repository, making it easy to browse these components and edit them.
+For example, mechanisms are stored under ``c++/mechanism''.
+}
 
@@
 
+\begin{tcolorbox}[breakable,notitle,boxrule=0pt,colback=yellow,colframe=yellow]
+
 \paragraph{Limited numerical integration strategies.} Most components in the software are integrated using the exponential Euler method, which has been used in integrating neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}).
+Single compartment models may also be integrated using the 4th order Runge-Kutta method.
+Unbranched multi-compartment models are automatically integrated using an implicit Crank-Nicholson method (\cite{dayanTheoreticalNeuroscience2001}).
 However, it may be desirable to use other methods under certain conditions.
@@
 Currently, xolotl can only implement integration schemes with fixed step size.
+\end{tcolorbox}
 
@@
 These examples have been chosen to demonstrate various features of xolotl, and are intended to serve as templates upon which researchers and educators can build.
+\hl{In every example, we assume that a new xolotl object has been initialized using}
+
+\begin{lstlisting}[style=Matlab-editor]
+x = xolotl;
+\end{lstlisting}
 
@@
 
+\begin{tcolorbox}[breakable,notitle,boxrule=0pt,colback=yellow,colframe=yellow]
+
 Network models in xolotl consist of compartment objects that can be connected by synapse objects.
 Two compartments representing different neurons can be connected using synapses using the built-in "connect" method.
-For example, to connect two single-compartment neurons called "LP" and "PY" using a chemical synapse of type "Cholinergic" with strength 30 nS, we can use
+For example, in the well-studied pyloric circuit in the crustacean stomatograstic ganglion, a cell called LP synapses onto PY cells using several synapses, that together have an effective combined maximal conductance of around 30 nS.
+This can be modelled in xolotl by using single compartments to represent the LP and PY cells, and connecting the LP cell to the PY cell using a chemical synapse of type "Glutamatergic" with strength 30 nS using
+
+\end{tcolorbox}
 
 \begin{lstlisting}[style=Matlab-editor]
-x.connect('LP','PY','Cholinergic','gbar',30); 
+x.connect('LP','PY','Glutamatergic','gbar',30);
 \end{lstlisting}
@@
 We have focussed on making our software as easy to use as possible, without sacrificing performance or extensibility.
-For example, the software and all dependencies can be installed using a single-line installer script from within the MATLAB command line.
-The installation includes worked example scripts that demonstrate various features of the simulator, that can be run without any configuration.
+For example, the installation includes worked example scripts that demonstrate various features of the simulator, that can be run without any configuration.
 We decided to built the the front-end interface to xolotl in MATLAB to facilitate interoperability with existing tools for time series analysis, optimization, and parallel computing.
 This allowed us to build rich tools for visualization and to interact with the simulation.
+
+\begin{tcolorbox}[breakable,notitle,boxrule=0pt,colback=yellow,colframe=yellow]
+
+\subsection{The cpplab architecture}
+
+A common theme in the architecture of many simulators and high-performance software is the union of user-facing code written in a scientific programming or user-friendly language and performance-critical code written in a fast, compiled language like C++.
+The reasons for splitting the codebase across two languages are often pragmatic: user interfaces and second-order tools are often easier to implement in richer programming environments, but performance requirements often mean that code needs to be written in a language like C++ or FORTRAN.
+While newer projects like Julia aim to merge ease of use and performance in a single language, a number of simulators and software projects embrace this division across languages.
+For example, DynaSim, a general purpose dynamical systems simulator, allows for models specified in MATLAB to be compiled into C and executed, speeding up execution (\cite{sherfeyDynaSimMATLABToolbox2018}).
+NEURON similarly presents a more use-friendly Python interface, while integration is carried out in compiled code (\cite{hinesNEURONSimulationEnvironment1997}).
+MATLAB's built in "codegen" command allows MATLAB functions to be automatically translated into C code and compiled, which can then be run instead of the MATLAB version.
+The SWIG project (\cite{beazley1996swig}) offers tools to connect programs written in C or C++ to a variety of scripting languages like Python and Ruby.
+
+\end{tcolorbox}
+
+\begin{tcolorbox}[breakable,notitle,boxrule=0pt,colback=yellow,colframe=yellow]
+
+Our approach in developing "cpplab", our bridge between C++ code and MATLAB, was conceptually different from many of these architectures.
+Unlike MATLAB's "codegen" function, which assumes the primacy of MATLAB code and makes C code that mirrors the MATLAB code, "cpplab" places C++ code first, and only aims to mimic the structure of objects specified in C++ header files, but not their functionality.
+Thus, cpplab is able to create a bonafide MATLAB object whose properties reflect that of the C++ object, but not its methods.
+This simpler task can be easily achieved without the construction of an intermediate interface file, as used in more powerful solutions like SWIG.
+For the purposes of a neuron and network simulator, this architecture is more robust since only simple parsing of the C++ code is needed.
+This architecture can be easily used in other applications, and it is trivial to remove some restrictions that the cpplab architecture currently imposes, like lack of support for vectors.
+
+\end{tcolorbox}
 
@@
 
-Xolotl is a purely component-based simulator, and all equations need to be included in a C++ header file that specifies a object.
+Xolotl is a purely component-based simulator, and all equations need to be included in a C++ header file that specifies an object.
 Since our automatic type system binds MATLAB objects to the underlying C++ header files, objects can be inspected and parameters can be modified in the MATLAB workspace.
@@
 This greatly decreases the learning curve and allows users with a general familiarity with programming languages to quickly acquaint themselves even with the most technical parts of the simulator.
+
+\begin{tcolorbox}[breakable,notitle,boxrule=0pt,colback=yellow,colframe=yellow]
+Finally, we walk through a typical workflow to compare xolotl to other neuron simulators.
+The first step is to initialize a xolotl object, e.g., using "x = xolotl".
+This is similar to importing the NEURON package from within Python.
+Compartments can be created and added to the object tree using "x.add(...)", and channels and mechanisms can be added to these compartments using further "add" commands.
+This workflow is similar to NEURON, where conductances are ``inserted'' into compartments.
+In both xolotl and NEURON, the tree of objects can be inspected and individual properties like maximal conductances or reversal potentials can be changed from the command line.
+The setup phase of the workflow is thus similar to that in NEURON, but differs from equation-oriented environments like Brian or DynaSim, where a set of equations has to be explicitly written out.
+xolotl differs from NEURON at this stage in that it is possible to ``click through'' the object tree all the way to view the underling C++ code for a particular component, and thus to view the ODE being solved, and the numerical method used to solve it.
+Since NEURON uses several intermediate files, that may have to be compiled, it is not trivial to identify and view the file that matches an object in the workspace.
+Another unique feature of xolotl is that every xolotl object is automatically and silently hashed in the background, and that this hash can be inspected at any time.
+This makes it easy to verify that the object constructed so far is what it should be.
+\end{tcolorbox}
 
@@
 Work on the simulator continues in the open at a publicly accessible repository (https://github.com/sg-s/xolotl/), and the library of conductances, synapses and mechanisms that xolotl ships with grows continuously.
-While this simulator was intended as a research tool, the many worked examples that are built into it demonstrate how it could also be used as a teaching tool.
+\hl{While this simulator was intended as a research tool, the many worked examples that are built into it demonstrate how it could also be used as a teaching tool.
+We intend to continue development of this software as we add new features and improve performance.
+We have developed an automatic testing and deployment framework that routinely tests the publicly available repository and packages a MATLAB toolbox for redistribution.
+As the user base grows, we intend to transition to finding maintainers who will ensure its continued availability and functionality.
+}
 
@@
 \section*{Acknowledgments}
-The authors would like to thank Timothy O'Leary and Leandro Alonso for useful discussions, Janis Li for help transcribing many conductance models, and members of the Marder Lab, especially Sonal Kedia, Ekaterina Morozova, Jason Pipkin and Philipp Rosenbaum for careful reading of the manuscript.
+\hl{The authors would like to thank Timothy O'Leary and Leandro Alonso for useful discussions, Janis Li for help transcribing many conductance models, and Stephen Van Hooser, Sonal Kedia, Ekaterina Morozova, Jason Pipkin and Philipp Rosenbaum for careful reading of the manuscript.}
 
@@
 (F-G) Steady-state gating functions for activation (m) and inactivation (h) gating variables.
-(H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables}
+(H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables.
+\hl{Every panel in this figure except E can be reproduced using built-in methods of xolotl}.}
 	\label{fig:figurehh}
@@
 Both plots refresh themselves with every movement of any slider, allowing the user to build intuition about how every parameter controls the dynamical behavior of the model.
-A screen recording of this model being manipulated in real time is included in Supplementary Material}
+A screen recording of this model being manipulated in real time is included in Supplementary Material.}
 	\label{fig:figuremanipulate}
@@
 NEURON was run using the Python wrapper.
-}
+\hl{Figures similar to this can be generated using the built-in benchmark() method}.}
 	\label{fig:figurebenchmark}

### 2018-10-26 paper/paper.tex
Fix path to mechanisms

@@
 Simulators designed in this way tend to perform fast computations with little overhead, but suffer from a steep learning curve.
-Wrapping these simulators in a more approachable language like Python or using graphical user interfaces (GUIs) mitigates these drawbacks (\cite{hinesNEURONPython2009, gratiy2018bionet}) at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
+Wrapping these simulators in a more approachable language like Python or using graphical user interfaces (GUIs) mitigates these drawbacks (\cite{hinesNEURONPython2009, gratiyBioNetPythonInterface2018}) at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}).
 In contrast, simulators designed from the ground up in popular scientific computing languages can be easier to use and benefit from interoperability with other commonly-used tools.
@@
 \hl{Every compartment exists as a C++ file within the code repository, making it easy to browse these components and edit them.
-For example, mechanisms are stored under ``c++/mechanism''.
+For example, mechanisms are stored under "xolotl/c++/mechanisms".
 }
@@
 Two compartments representing different neurons can be connected using synapses using the built-in "connect" method.
-For example, in the well-studied pyloric circuit in the crustacean stomatograstic ganglion, a cell called LP synapses onto PY cells using several synapses, that together have an effective combined maximal conductance of around 30 nS.
-This can be modelled in xolotl by using single compartments to represent the LP and PY cells, and connecting the LP cell to the PY cell using a chemical synapse of type "Glutamatergic" with strength 30 nS using
+For example, in the well-studied pyloric circuit in the crustacean stomatograstic ganglion \cite{marderUnderstandingCircuitDynamics2007}, a cell called LP synapses onto PY cells using many synapses, that together have an effective combined maximal conductance of around 30 nS.
+This can be modeled in xolotl by using single compartments to represent the LP and PY cells, and connecting the LP cell to the PY cell using a chemical synapse of type "Glutamatergic" with strength 30 nS using
 
@@
 Conductance-based neuron models are typically high dimensional and contain many parameters.
-Changing a single parameter monotonically can cause non-monotonic changes in behavior of the model, and certain dynamical features may only emerge when in specific non-convex regions of parameter space (\cite{golowasch2002failure}).
+Changing a single parameter monotonically can cause non-monotonic changes in behavior of the model, and certain dynamical features may only emerge when in specific non-convex regions of parameter space (\cite{golowaschFailureAveragingConstruction2002}).
 It is often challenging to build intuition about what effect a parameter has on the model under these conditions.
@@
 The reasons for splitting the codebase across two languages are often pragmatic: user interfaces and second-order tools are often easier to implement in richer programming environments, but performance requirements often mean that code needs to be written in a language like C++ or FORTRAN.
-While newer projects like Julia aim to merge ease of use and performance in a single language, a number of simulators and software projects embrace this division across languages.
+While newer projects like Julia (\cite{bezansonJuliaFreshApproach2017}) aim to merge ease of use and performance in a single language, a number of simulators and software projects embrace this division across languages.
 For example, DynaSim, a general purpose dynamical systems simulator, allows for models specified in MATLAB to be compiled into C and executed, speeding up execution (\cite{sherfeyDynaSimMATLABToolbox2018}).
-NEURON similarly presents a more use-friendly Python interface, while integration is carried out in compiled code (\cite{hinesNEURONSimulationEnvironment1997}).
+NEURON similarly presents a more user-friendly Python interface, while integration is carried out in compiled code (\cite{hinesNEURONSimulationEnvironment1997}).
 MATLAB's built in "codegen" command allows MATLAB functions to be automatically translated into C code and compiled, which can then be run instead of the MATLAB version.
-The SWIG project (\cite{beazley1996swig}) offers tools to connect programs written in C or C++ to a variety of scripting languages like Python and Ruby.
+The SWIG project (\cite{beazleySWIGEasyUse1996}) offers tools to connect programs written in C or C++ to a variety of scripting languages like Python and Ruby.
 
@@
 Thus, cpplab is able to create a bonafide MATLAB object whose properties reflect that of the C++ object, but not its methods.
-This simpler task can be easily achieved without the construction of an intermediate interface file, as used in more powerful solutions like SWIG.
+This task can be easily achieved without the construction of an intermediate interface file, as used in more powerful solutions like SWIG.
 For the purposes of a neuron and network simulator, this architecture is more robust since only simple parsing of the C++ code is needed.
-This architecture can be easily used in other applications, and it is trivial to remove some restrictions that the cpplab architecture currently imposes, like lack of support for vectors.
+This architecture can be easily used in other applications, and it is trivial to remove some restrictions that the cpplab architecture currently imposes, such as lack of support for vectors.
 
@@
 
-Over the years, several simulators have been developed to integrate systems of coupled differential equations that model the spiking activity of neurons (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003, hinesNEURONSimulationEnvironment1997, bower2003genesis}).
+Over the years, several simulators have been developed to integrate systems of coupled differential equations that model the spiking activity of neurons (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003, hinesNEURONSimulationEnvironment1997,bowerBookGENESIS2nd1998}).
 A critical architectural choice in designing a simulator is how much ``scaffolding'' a user is provided with to construct a model; whether a model is specified by equations or by components, or by some combination of the two.
-In an equation-oriented architecture, the user starts with a blank slate and the primary method of specifying a model is to write out its differential equations (\cite{stimberg2014equation}).
+In an equation-oriented architecture, the user starts with a blank slate and the primary method of specifying a model is to write out its differential equations (\cite{stimbergEquationorientedSpecificationNeural2014}).
 In contrast, in a component-oriented architecture, the primary method of specifying a model is to assemble it from pre-existing components, each of which include differential equations, parameters, and solvers.
@@
 BRIAN is an equation-oriented neuronal simulator meant to be used from within Python (\cite{goodmanBrianSimulator2009}).
-XPP is a general purpose dynamical system simulator that is equation-oriented (\cite{ermentrout2002simulating}).
+XPP is a general purpose dynamical system simulator that is equation-oriented (\cite{ermentroutSimulatingAnalyzingAnimating2002}).
 DynaSim is an equation-oriented simulator with some component-oriented capabilities.
@@
 Some component-oriented simulators like NEURON also specify their own syntax, or have invented their own language to specify components and write out equations.
-While this allows for powerful features such as support for units in NEURON, the user is required to learn a novel syntax and vocabulary, hindering ease of use.
+While this allows for powerful features such as support for units, the user is required to learn a novel syntax and vocabulary, hindering ease of use.
 In general, the syntax for model specification in different simulators can be different.
-For example, DynaSim, BRIAN, XPP, and NEURON all use different, incompatible formalisms to represent equations, increasing the cognitive load on users using more than one simulator.
+For example, DynaSim, BRIAN, XPP, and NEURON all use different, incompatible formalisms to represent equations, increasing the cognitive load on users working with more than one simulator.
 Here, we have elected not to specify a domain-specific ``middleware'' layer, and instead specify and implement models and equations in idiomatic C++.
@@
 Finally, we walk through a typical workflow to compare xolotl to other neuron simulators.
-The first step is to initialize a xolotl object, e.g., using "x = xolotl".
-This is similar to importing the NEURON package from within Python.
+First, initialize a xolotl object (e.g.
+"x = xolotl").
+This is similar to importing the NEURON or Brian packages from within Python.
 Compartments can be created and added to the object tree using "x.add(...)", and channels and mechanisms can be added to these compartments using further "add" commands.
@@
 In both xolotl and NEURON, the tree of objects can be inspected and individual properties like maximal conductances or reversal potentials can be changed from the command line.
-The setup phase of the workflow is thus similar to that in NEURON, but differs from equation-oriented environments like Brian or DynaSim, where a set of equations has to be explicitly written out.
-xolotl differs from NEURON at this stage in that it is possible to ``click through'' the object tree all the way to view the underling C++ code for a particular component, and thus to view the ODE being solved, and the numerical method used to solve it.
+The setup phase of the workflow is therefore similar to that in NEURON, but differs from equation-oriented environments like Brian or DynaSim, where a set of equations has to be explicitly written out.
+Unlike in NEURON, it is possible to ``click through'' the xolotl object tree all the way down to view the underling C++ code for a particular component, and thus to view the ODE being solved, and the numerical method used to solve it.
 Since NEURON uses several intermediate files, that may have to be compiled, it is not trivial to identify and view the file that matches an object in the workspace.
-Another unique feature of xolotl is that every xolotl object is automatically and silently hashed in the background, and that this hash can be inspected at any time.
-This makes it easy to verify that the object constructed so far is what it should be.
+While NEURON requires compilation of a custom binary version including all network components, xolotl automatically and silently hashes the network in the background to determine which files must be compiled and linked.
+This hash can be inspected at any time, making it easy to verify that the object constructed so far is what it should be.
 \end{tcolorbox}
